### 2020-05-04 ms/manuscript.Rmd
Copy text of MS to Rmd

---
title: "Untitled"
output: html_document
---

Intergenerational Niche Differentiation in Filmy Ferns (Hymenophyllaceae)^1^

Joel H. Nitta^2,5,6^, James E. Watkins, Jr.^3^, N. Michele Holbrook^2^, Ravahere Taputuarai^4^, Tristan W. Wang^2^, and Charles C. Davis^2^

Running Head: Intergenerational Niche Differentiation in Filmy Ferns

**PREMISE OF THE STUDY**: Desiccation tolerance was a key trait allowing plants to colonize land.
However, little is known about the transition from desiccation tolerant non-vascular plants to desiccation sensitive vascular ones.
Filmy ferns (Hymenophyllaceae) present a useful system to study how water-stress strategies differ between non-vascular and vascular stages within a single organism, as they have vascularized sporophytes and nonvascular gametophytes that are each capable of varying degrees of desiccation tolerance.

**METHODS**: We surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia).
We used chlorophyll fluorescence to measure desiccation tolerance and light responses.
We sequenced plastid *rbcL* and *trnHpsbA* to identify field-collected gametophytes, and conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.

**KEY RESULTS**: Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima than sporophytes.
Despite their lack of greater physiological tolerances, gametophytes of several species occurred over a wider elevational range than conspecific sporophytes.
Epiphytes were more tolerant of desiccation than terrestrial species.

**CONCLUSIONS**: Our results show that filmy fern gametophytes and sporophytes differ in their physiology and niche requirements, and point to the importance of microhabitat in shaping the evolution of water-use strategies in vascular plants.

**KEYWORDS**: chlorophyll fluorescence; desiccation tolerance; DNA barcoding; filmy ferns; French Polynesia; gametophyte; Hymenophyllaceae; Moorea; photosynthesis

^1^Manuscript received \_\_\_\_ ; revision accepted \_\_\_\_.

^2^Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, Cambridge, Massachusetts 02138 USA;

^3^Department of Biology, Colgate University, 13 Oak Drive, Hamilton, New York 13346 USA; and

^4^Association Te Rau Ati Ati a Tau a Hiti Noa Tu, Mahina, Tahiti, French Polynesia

^5^Current address: Department of Botany, National Museum of Nature and Science, 4-1-1 Amakubo, Tsukuba, Ibaraki 305-0005 Japan

^6^Author for correspondence (email: joelnitta\@gmail.com)

The movement of plants from water to land was one of the most important evolutionary events in the history of the earth, fundamentally altering the global carbon and water cycle and setting the stage for modern terrestrial ecosystems (Graham, 1993; Kenrick and Crane, 1997).
The transition from aquatic to terrestrial growth was complex, requiring the evolution of a suite of adaptations to survive in dry habitats bombarded with radiation (Waters, 2003).
The earliest land plants were poikilohydric, i.e., they lacked any mechanisms to regulate water relations (Raven, 1999).
Therefore, one particularly important early adaptation that enabled plants to survive on land was desiccation tolerance (DT), the ability to lose water until reaching equilibrium with atmospheric humidity, then recover upon rewetting (Mishler and Churchill, 1985).

Compared with the ubiquitous DT among the earliest land plants, the occurrence of DT varies across extant land plant lineages and life stages.
Many bryophytes have retained vegetative DT in the gametophyte stage (Oliver et al. 2005, and references therein).
In extant vascular plants (tracheophytes), DT occurs in certain non-vegetative stages (e.g., seeds and pollen), but has largely been lost in vegetative tissues (with some notable exceptions, the so-called 'resurrection plants').
Instead, tracheophytes avoid desiccation with adaptations including a waxy cuticle, stomata, and an extensive vascular system.
Correlated with these differences in functional strategies to cope with water stress is a major difference in life cycle: in most tracheophytes (including all seed plants) the long-lived, dominant phase of the life cycle is the sporophyte (diploid phase), whereas in bryophytes it is the gametophyte (haploid phase).
Little is known about the details of the transition from desiccation tolerance to sensitivity, nor how this took place in conjunction with the switch from a gametophyte to a sporophyte-dominated life cycle; yet, these two steps underpin the radiation of plants onto dry land.

Ferns represent an important, understudied group of tracheophytes for investigating the transition from desiccation tolerance to desiccation avoidance in land plants.
Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often *\<*1 cm), lack vascular tissue, and are poikilohydric.
Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT (Watkins et al., 2007; Pittermann et al., 2013).
Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.

Filmy ferns (Hymenophyllaceae) represent a particularly interesting case for studying the ecological and evolutionary significance of vegetative DT across life stages in ferns.
Filmy ferns are a large family (ca. 430 spp.; PPG I, 2016) of primarily tropical, leptosporangiate ferns named for their remarkably thin leaf laminae, which are usually only a single cell layer in thickness and lack cuticle or stomata (Fig. 1A--D).
Filmy fern sporophytes are an exception to the 'typical' fern sporophyte described above because they have reverted to poikilohydry: their thin laminae allow for passive, rapid uptake and loss of moisture, thereby maintaining nearly constant equilibrium with atmospheric humidity (Raven, 1999).
Many species of epiphytic filmy ferns have extremely reduced root systems (Schneider, 2000), relying completely upon absorption through leaf laminae for water (Shreve, 1911), although terrestrial species possess true roots and use ground water (Dubuisson et al., 2011).
As would be expected from their poikilohydric nature, DT of varying degrees has been reported in sporophytes of many filmy fern species (see references below).

Filmy fern sporophytes encompass a broad diversity of growth forms (terrestrial, saxicolous, hemi-epiphytic, and holoepiphytic) and morphologies, with rhizome and leaf sizes varying over an order of magnitude (Hennequin et al., 2008; Nitta and Epps, 2009; Dubuisson et al., 2013).
Filmy fern gametophytes are similarly diverse, with morphotypes including filamentous, ribbon, and a combination of these two types (Fig. 1E--H).
While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages (Watkins et al., 2007; Watkins and Cardelús, 2012; Pittermann et al., 2013), and may be expected to occur in filmy ferns as well.
The possibility for varying degrees of vegetative DT in both stages of the life cycle of filmy ferns makes them a potentially useful system to compare how DT is utilized between haploid, non-vascular and diploid, vascular life stages.
Furthermore, the diversity of growth forms and habitats in filmy ferns makes them an ideal group for a comparative study of water-stress strategies (i.e., desiccation tolerance vs. desiccation avoidance) and the correlation of these strategies with ecological niche.

The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
In a series of elegant experiments on Jamaican filmy ferns, Shreve (1911) demonstrated that water-use strategies differ between species occupying contrasting habitats (i.e., terrestrial, low elevation epiphyte, and high elevation epiphyte).
Proctor (2003, 2012) investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs. low DT and adaptation to low light levels in species occupying more sheltered habitats.
In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, Johnson et al. (2000) found that gametophytes of *Vandenboschia speciosa* (Willd.) G. Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
Parra et al. (2009) and Saldan˜a et al. (2014) investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
A recent study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly (Cea et al., 2014).

Despite this considerable interest in filmy fern ecophysiology, with the sole exception of Johnson et al. (2000), no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
Furthermore, to our knowledge, no physiological studies of filmy ferns have accounted for their phylogenetic relationships, and we lack an understanding of how physiological strategies have evolved in this clade.
In addition to DT, photosynthetic optima (adaptation to different light levels) are likely to be important in determining the niches of life stages and species of filmy ferns (Proctor, 2003; Parra et al., 2015; Flores-Bavestrello et al., 2016).
Terrestrial and epiphytic species experience different levels of photosynthetic radiation and desiccation, and light intensity varies with height along the host plant in epiphytes (Benzing, 1990; Watkins and Cardelús, 2009).
In a recent study (Testo et al., unpublished), the combined effect of desiccation and light stress was found to have a greater impact on filmy fern physiological performance than either of these stressors alone, suggesting the need to integrate studies of DT with photosynthesis to gain a full picture of physiological adaptation in filmy ferns.
Here, we leverage a recently developed DNA barcoding system (Nitta et al., 2016) to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes: 1.) Epiphytic filmy ferns have higher DT and photosynthetic optima than non-epiphytic species.
2.) Amongst epiphytic filmy ferns, those occurring at high elevation have higher DT and photosynthetic optima than those at low elevation.
3.) For filmy ferns with gametophytes that occur over a wider range than sporophytes, the gametophyte has higher DT and photosynthetic optima than the sporophyte.

MATERIALS AND METHODS

***Study system***---We selected the island of Moorea, French Polynesia (17*◦*32' S, 149*◦*50' W) for our survey of filmy fern ecophysiology (Fig. 2).
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca. 15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
At the time of the study, a total of 18 filmy fern species were known from Moorea (Murdock and Smith, 2003; Ranker et al., 2005; Nitta et al., 2011b; Florence, in press), including multiple terrestrial, saxicolous, and epiphytic species.
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages (Pryer et al., 2001), are represented by similar numbers of species.
We follow the taxonomic system of Ebihara et al. (2006).

***Field survey***---We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti (Nitta et al., 2016).
During this study, 17 sampling sites were established from ca. 200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt. Rotui (899 m), Mt. Mouaputa (880 m), and Mt. Tohiea (1206 m; Fig. 2).
At each site, sporophytes were sampled in a single 10 m *×* 10 m plot, and gametophytes sampled in two to three 50 cm *×* 50 cm subplots located within the sporophyte plot.
Epiphytes were sampled to a maximum height of ca. 2 m on their host plants.
Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family (e.g., filamentous, ribbon with round gemmae; Nayar and Kaur 1971).
This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
We characterized the elevational range of each species based on the results of the field survey; any species found with gametophytes occurring *\>*200 m above the maximum or below the minimum observed elevational range of sporophytes was considered 'widespread' (i.e., facultatively independent sensu Rumsey and Sheffield \[1996\]).
We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, i.e., below ca. 500 m), or high elevation epiphyte (mostly occurring in cloud forest, i.e., above ca. 500 m).

***Microclimate***---We characterized the microenvironment experienced by filmy ferns by logging relative humidity (RH) and temperature every 15 min from July 7, 2013 to July 5, 2014 at 16 of the sampling sites (Fig. 2).
To do this we used Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) with external sensors outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation.
To compare between terrestrial and epiphytic habitats, dataloggers were placed in pairs at each site (one 'epiphytic' datalogger at ca. 1.5 m on a tree, one 'terrestrial' datalogger at ca. 10 cm next to the tree).
During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
We excluded any days during which one or more dataloggers failed from the final dataset, and some sites are missing data completely for one or both dataloggers.
One terrestrial datalogger (Mt. Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013).
We used the previous year's data for this site instead.
The final dataset included 201 days of data from 15 sites (26 dataloggers).
We tested for differences in climatic variables between epiphytic and terrestrial habitats using analysis of covariance (ANCOVA) with each climatic variable in turn as the response variable and elevation as the explanatory variable with habitat (terrestrial vs. epiphytic) as a cofactor.

***Desiccation tolerance***---Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8--16 individuals (one individual = single whole frond including a ca. 2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment chlorophyll fluorescence (F~v~/F~m~) was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval (Testo and Watkins, 2013).
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22--25 *◦*C.
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (--282 MPa), Mg(NO~3~)~2~ (--86 MPa), and NaCl (--38 MPa).
Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at --86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Following the desiccation treatment, plants were rewetted using water from a local stream.
Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0*.*5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield (Watkins et al., 2007).

***Light responses***---Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400--700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
Next, plants were equilibrated to the actinic light for 30 s, and photosynthetic yield (∆F/F~m~) measured with a 0*.*8 s saturating pulse.
This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca. 500 µmol·m^--2^ ·s^--1^.
Response curves were fitted using the equation *y* = *A*(1 --*e^-kx^*), where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded (Proctor, 2012).
The PPFD at 95% saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
Light responses were measured in 8--16 field-collected sporophyte individuals for each species, and on single gametophyte individuals in the field, which were later identified to species using the DNA barcodes.

***DNA barcoding***---We used DNA barcoding to identify filmy fern gametophytes to species as described in Nitta et al. (2016).
Briefly, DNA extraction was performed using a modified CTAB protocol (Doyle and Doyle, 1987) or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following Nitta et al. (2016).
For each marker, a reference library was constructed from sporophyte sequences using the makeblastDB command in BLAST (Altschul et al., 1997).
Field-collected gametophytes were identified by a local BLAST query against the reference library.
Gametophytes matching *\>*99.5% with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched *\>*99.5% with multiple reference sequences were not identified and excluded from further analysis.

***Statistical analysis and phylogenetic comparative methods***---All statistical analyses were done in R v.3.2.1 (R Core Team, 2015).
Datasets and R scripts used in this study are available on Dryad (\#\#\#\#\#\#).

We derived the phylogenetic tree used in this study by extracting Hymenophyllaceae from the tree of Nitta et al. (2016) with the extract.clade function of the ape package in R (Paradis et al., 2004).
This tree is well-supported and generally in accord with previously published, more densely sampled phylogenies (Pryer et al., 2001; Dubuisson et al., 2003; Hennequin et al., 2006; Ebihara et al., 2007; Appendix S1, see Supplemental Data with the online version of this article).

We quantified the degree of phylogenetic signal in physiological traits using Blomberg's *K* (Blomberg et al., 2003) and Pagel's λ (Pagel, 1999; Freckleton et al., 2002) as implemented with the phylosig function in the phytools R package (Revell, 2012).
Both measures test the hypothesis that the trait of interest is evolving according to Brownian Motion (BM).
λ is a scaling parameter that ranges from zero to one: λ near zero indicates random distribution of trait values across the tree, and λ near one indicates evolution of traits along the phylogeny following BM.
For *K*, values near one indicate evolution of traits following BM; *K* *\>*1 indicates traits more conserved than expected under BM, and *K* *\<*1 indicates that traits have less phylogenetic signal than expected under BM.
We tested the significance of *K* by comparing the observed value against values from a null distribution of 1000 phylogenies with the traits randomly shuffled across the tips.
We tested the significance of λ with a log-likelihood test comparing the likelihood of the observed value of λ vs. λ = 0.
For all further analyses, we used methods that account for phylogenetic history for physiological traits that showed significant phylogenetic signal, and standard methods otherwise.

We used general linear mixed models (GLMMs) to investigate the effect of growth habit, range size, and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at --86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the MCMCglmm R package (Hadfield, 2010).
For each of the three response variables (DT, ETR~max~ , and PPFD~95%~), we used growth habit (epiphyte vs. non-epiphyte \[i.e., terrestrial or saxicolous\]) or range size (widespread vs. not widespread) and generation (sporophyte vs. gametophyte) and their interactions as fixed effects.
Two species were observed both as low-elevation epiphytes and saxicoles (Table 1), but we treated them as epiphytic to distinguish them from exclusively saxicolous species, which we regard as the more specialized growth habit.
We accounted for intraspecific variation and phylogeny (only for response variables with significant phylogenetic signal) by including species (within generation) and an inverse phylogenetic variance-covariance matrix as random effects, respectively (Hadfield, 2010).
We used an inverse-Gamma distribution for all priors, and ran analyses for 500 000 iterations, with burnin after 1000 iterations and thinning every 500 iterations.
Comparisons of DT, ETR~max~ , and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a *t*-test for these.

To test for a correlation between environmental water potential and DT, we first derived a single minimum water potential value for each species by taking the average of the minimum daily water potential at all sites where that species was present.
We used minimum values because these may represent a climatic extreme that limits the occurrence of species.
We used data from the 'epiphytic' dataloggers mounted at 1*.*5 m on trees for epiphytic species and data from the 'terrestrial' dataloggers on the ground for non-epiphytic species.
Data from the Mt. Rotui 830 m site were excluded because this site is much more exposed than the other high elevation sites, and appears as a significant outlier in the microclimatic data (see Results).
This may be due to its location; unlike other mountains on Moorea, Mt. Rotui is completely isolated with no connecting ridges to other peaks (Fig. 2).
Since DT showed significant phylogenetic signal (see Results), we used phylogenetic generalized least squares (PGLS) to test for a relationship between environmental water potential and DT while taking phylogeny into account, using species' means.
We ran PGLS using the caper R package (Orme et al., 2013).

RESULTS

***Field survey***---Twenty-two filmy fern taxa (including one new record and two species that were recognized as multiple varieties) were observed as sporophytes, and 18 as gametophytes on Moorea in total (Table 1).
*Hymenophyllum braithwaitei* (Ebihara & K. Iwats.), only observed near the peak of Mt. Tohiea (1206 m), is a new record for Moorea and French Polynesia.
It was previously known only from Vanuatu and New Caledonia (Ebihara et al., 2010).
Three taxa with distinct *rbcL* sequences, morphology, and elevational ranges could be distinguished within the *Crepidomanes minutum* (Blume) K. Iwats. species complex, but species status of these taxa remains uncertain (Nitta et al., 2011a).
We refer to these here as *C. minutum* var. 1, 2, and 3.
Similarly, two members of the *Abrodictyum asae-grayi* (Bosch) Ebihara & K. Iwats. complex (also including *Abrodictyum caudatum* (Brack.) Ebihara & K. Iwats. in our tree; Appendix S1) were observed on the basis of *rbcL* and morphology, but as taxonomic treatment is beyond the scope of the current study we refer to them as *A. asae-grayi* var. 1 and 2.
In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var. 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K. Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson.
All of these have sporophytes that are normally only found in cloud forest (ca. 500--1200 m), but gametophytes that are distributed to lower elevations.

***Microclimate***---Mean daily temperature was strongly correlated with elevation, and dropped ca. 6 *◦*C from the lowest surveyed site (201 m) to the highest (1206 m; Fig. 3). Although epiphytic and terrestrial habitats were similar in temperature, epiphytic habitats tended to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones (Fig. 3, Appendix S2).
This difference was greatest at low elevations, and smallest at high elevations.
One significant departure from these trends was observed at the Mt. Rotui 830 m site (indicated by arrows in Fig. 3).
This site is much more exposed than the other cloud forest sites (Fig. 2), and many of the plants on the N facing slope are stunted (J. Nitta, pers. obs.).
Daily variation in temperature and RH was much greater, and mean RH much lower, at this site than other high elevation sites.

***Phylogenetic signal***---Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* and λ both near 1, *P \<*0.05; sporophytes *K* = 0.78, *P* = 0.02 and λ = 0.96, *P* = 0.11), but no signal for PPFD~95%~ or ETR~max~ (Table 2).

***Desiccation tolerance***---Ability to recover from desiccation was tested in sporophytes of 15 species and gametophytes of 14 species, including 12 species with both life phases (Appendix S3).
A general correlation of DT with habitat was observed in sporophytes (Fig. 4).
Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats. and *C. apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 MPa (more stringent tests were not conducted on these species after confirming their lack of DT at this level).
Saxicolous species (*Crepidomanes kurzii* (Bedd.) Tagawa & K. Iwats., *Polyphlebium endlicherianum* (C. Presl) Ebihara & K. Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C. kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
All epiphytic species were capable of recovering from desiccation to a certain degree, but ability varied between species.
Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.* *pallidum* and *P.* *borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.

Gametophytes of low elevation epiphytes had the highest DT (ca. 80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from ca. 25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; Fig. 5).
DT of gametophytes did not significantly exceed that of sporophytes for any species except for saxicolous *P. endlicherianum* (Fig. 5; *t*-test, *P* = 0.012), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C. minutum* var. 2, *H.* *digitatum*, *H. polyanthos*, and *P. borbonicum* (Fig. 5; *t*-test, *P* = 0.0093, *\<*0.001, *\<*0.001, 0.012, respectively).

The GLMM showed that both generation and the interaction of generation and growth habit, but not growth habit alone, have a significant effect on DT while controlling for phylogeny (Table 3).
In contrast, range type (gametophytes widespread beyond sporophytes or not) has no significant effect on DT (Table 3).

***Photosynthetic optima***---Light responses were measured in sporophytes of 18 species and gametophytes of 11 species, including nine species with both life phases (Appendix S4).
Unlike DT, photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes are adapted for higher light levels, whereas gametophytes are adapted for lower light levels (Fig. 5).
Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca. 142 ± 46 µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at 51 ± 22 µmol·m^--2^ ·s^--1^.
Similarly, sporophytes attained higher rates of electron transport (ETR~max~; 8.1 ± 2.7 µmol·m^--2^ ·s^--1^) than gametophytes (3.9 ± 0.9 µmol·m^--2^ ·s^--1^).
The GLMMs showed a significant effect of generation, but neither range type nor growth habit, on PPFD~95%~ and ETR~max~ (Table 3).

***Environmental water potential and DT***---Desiccation tolerance tended to increase with decreasing environmental water potential for both gametophytes and sporophytes (Fig. 6).
These trends were significant when analyzed using PGLS for sporophytes (*P* = 0.036) but not gametophytes (*P* = 0.123; Appendix S5).

DISCUSSION

Vegetative DT was a key innovation linked to the rise of early land plants, but is largely limited to non-vascular groups such as bryophytes and algae in modern plants.
Filmy ferns have received considerable attention over the years for their remarkable reversion to vegetative DT in the sporophyte phase (e.g., Shreve, 1911; Proctor, 2003; Cea et al., 2014).
Yet, no previous study to our knowledge has evaluated parallel physiology of sporophytes and gametophytes of filmy ferns in relation to DT.
Here, we present the first comparisons of physiological parameters between generations within filmy ferns in a phylogenetic comparative framework.

***Filmy fern sporophytes have a wide range of DT***---We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below --200 MPa in some epiphytes.
Levels of similarly extreme DT have also been reported in other species of *Hymenophyllum* (Proctor, 2003, 2012) and a general correlation of habitat with DT in filmy ferns has been previously established (Shreve, 1911; Proctor, 2003, 2012; Parra et al., 2009; Saldan˜a et al., 2014; Testo et al., unpublished), but no other studies to our knowledge have documented a similar range of DT values in filmy ferns from a single site.
One may wonder then, what is the adaptive significance of being able to withstand drying at --200 MPa vs. --100 MPa when the ambient water potential is typically *\>*--5 MPa and only rarely, if ever, drops below −50 MPa.
The recent finding of a constitutive mechanism for DT in filmy ferns (Cea et al., 2014) supports the need for a relatively wide 'safety margin,' and similar (or even more extreme) levels of constitutive DT are well known in bryophytes (Oliver et al., 1993).
The conservative behavior of fern stomata, which lack a response to the plant stress hormone ABA, is well documented (Brodribb and McAdam, 2011; McAdam and Brodribb, 2012, 2013).
It is possible that filmy ferns also behave conservatively with respect to water-stress.
We suggest that the variation in DT observed here is relevant because it is correlated with habitat, and likely reflects overall physiological tolerance, not just artificial extremes.
For example, sporophytes of some high elevation epiphytes that we surveyed (*C. minutum* var. 2, *H. digitatum*, *H. multifidum*, *H. polyanthos*) seem to be able to better withstand intense vs. moderate desiccation, with higher recovery values at --282 and --86 MPa vs. --38 MPa when dry for 15 d.
This may be an adaptation to cloud forest conditions; although cloud forests are frequently moist due to constant fog, dry periods are intense during the times when there is no cloud cover.
The complete lack of DT in terrestrial species is not surprising, considering they have much more direct access to water than epiphytes and therefore substantially less selective pressure to maintain DT.
Since controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (Appendix S3), it is unlikely that the lack of tolerance seen in terrestrial species is due solely to removal from the soil.
Comparison of DT with the minimum water potential envelope of each species (Fig. 6) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with minimum water potential exceeding ca. --7 MPa.
In contrast, habitats with minimum water potential below --7 MPa are nearly all epiphytic, and the filmy fern sporophytes occurring there are all capable of DT.

***Filmy fern gametophytes are not more stress-tolerant than sporophytes***---Contrary to our hypotheses, gametophytes of filmy ferns are in general not more tolerant of the abiotic environment (light levels and water potential) than sporophytes.
This differs from other fern groups with desiccation tolerant gametophytes but desiccation sensitive sporophytes (Watkins et al., 2007).
In the filmy ferns we studied, DT levels of both sporophytes and gametophytes largely reflected habitat; in general, epiphytic species tended to have higher DT than terrestrial or saxicolous species, regardless of life stage.
Interestingly, this echoes much earlier field observations (not experimental data) by Holloway (1930) on filmy fern gametophytes in New Zealand: he reported apparent DT in gametophytes of *Hymenophyllum rarum* R. Br. and *H. villosum* Colenso, both high elevation epiphytes with desiccation tolerant sporophytes, and lack of DT in terrestrial filamentous gametophytes of *Polyphlebium colensoi* (Hook.) Ebihara & K. Iwats., *Abrodictyum strictum* (Menzies ex Hook. & Grev.) Ebihara & K. Iwats., and *A. elongatum* (A. Cunn.) Ebihara & K. Iwats. occupying moist, protected sites on the forest floor.
Watkins et al. (2007) also found that DT of fern gametophytes tended to match habitat, with terrestrial species having lower DT than epiphytes.
However, we found that within high elevation epiphytes, gametophytes had lower DT than sporophytes, and only one species (saxicolous *P. endlicherianum*) with gametophytes that had significantly greater DT than sporophytes (Fig. 5).
At high elevation sites (i.e., cloud forest), epiphytic ferns often occur in a dense matrix of bryophytes (Fig. 1B).
It is possible that this bryophyte cover helps to retain moisture during brief periods of drought and thus protect the epiphytic fern gametophytes that grow amongst them.
More detailed studies of the high elevation epiphytic environment and interactions between co-occurring gametophytes of ferns and bryophytes are needed to test this hypothesis.

Although we predicted that species with gametophytes that occur beyond the range of their sporophytes would have higher DT and be adapted for higher light levels, this was not the case.
This is perhaps best illustrated by *C. apiifolia* (Fig. 1D, E, H).
Sporophytes of this species are usually restricted to moist cloud forest sites from ca. 600--1200 m, but we observed gametophytes over a much wider range, down to 200 m.
A similar distribution pattern has been reported in congeneric *Callistopteris baldwinii* (D. C. Eaton) Copel. from Hawaii (Dassler and Farrar, 1997).
Since gametophytes of *C. apiifolia* are common on Moorea and have a unique morphology (Nitta et al., 2016), we were able to collect enough material of this species for additional tests of DT at higher humidity levels (48 h at 75% and 100% RH; results not shown).
However, it still failed to recover from even these gentle treatments.
This suggests that the widespread gametophytes of *C. apiifolia* do not rely on DT to attain their distribution, but instead must do so via some other mechanism.
Although we measured terrestrial microhabitats by positioning sensors a few cm above the ground, these sensors are still outside of the boundary layer.
It is likely that the gametophytes of *C. apiifolia* do not actually experience the conditions measured with our 'terrestrial' dataloggers, but instead grow within a much more consistently moist boundary layer.
Gametophytes of *C. apiifolia* produce gemmae (Fig. 1E, H), and often form dense clonal mats.
In the case that their microhabitats do become too dry, they may suffer loss at the edge of the population, then recover by clonal growth (Farrar et al., 2008).
Furthermore, the complex surface of clonal mats has been shown to increase the boundary layer and decrease evaporation rates in mosses (Rice et al., 2001), and may function similarly in *C. apiifolia*.
Other filmy fern species also produce gemmae, but they are much more frequent in *C. apiifolia*, and the large size of the clonal mats formed by *C. apiifolia* indicate that it may have a high growth rate (J. Nitta, pers. obs.).
It is possible that there is a tradeoff between growth rate and DT, such that species capable of rapid asexual growth like *C. apiifolia* are unable to tolerate desiccation, but slower growing epiphytes such as *Hymenophyllum* spp. can better tolerate it (Oliver et al., 2000).
Growth rates of tolerant and non-tolerant gametophytes including quantification of gemmae production should be investigated in these species to test this hypothesis.

***Gametophytes are adapted for lower light levels than sporophytes***---Unlike DT, photosynthetic parameters did not vary significantly between species from different habitats.
Rather, we observed a clear intergenerational difference, with gametophytes consistently adapted for lower light levels than sporophytes (Fig. 5, Table 3).
In the only other study to our knowledge that compared photosynthetic rates between gametophytes and sporophytes of a filmy fern, Johnson et al. (2000) also found that gametophytes of *V. speciosa* reached maximum ETR at lower levels than sporophytes: in gametophytes of this species, photosynthesis became saturated at a PPFD of 4--5 µmol·m^--2^ ·s^--1^ in the field or 30--50 µmol·m^--2^·s^--1^ in the lab vs. 30--50 µmol·m^--2^ ·s^--1^ for the sporophyte (lab measurement only).
Although we found similar differences in direction between gametophytes and sporophytes, the measurements of Johnson et al. (2000) are considerably lower than what we observed in the field for both Moorean filmy fern gametophytes (PPFD~95%~ ca. 40--60 µmol·m^--2^ ·s^--1^) and sporophytes (PPFD~95%~ ca. 130--160 µmol·m^--2^ ·s^--1^).
This makes sense in light of the fact that *V. speciosa* is an extremely deep-shade plant, with gametophytes occupying protected rock crevices that receive less than 1 µmol·m^--2^·s^--1^ PPFD for most of the day (Johnson et al., 2000); many of the gametophytes sampled in our study occur in more exposed conditions.
It is surprising, however, that we did not observe any differences between species in different habitats, given that light levels should vary between tree trunks and the forest floor, at least on a ca. 1 m scale.
It is possible that the actual light levels experienced by gametophytes growing in these habitats may not vary significantly if they occupy protected microsites such as bryophyte mats or crevices in rock or bark.

***Influence of phylogeny on traits***---Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates (Proctor, 2003, 2012; Parra et al., 2009; Saldan˜a et al., 2014); however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
We detected significant phylogenetic signal in DT, suggesting that phylogeny should be considered when interpreting differences in DT between species (Table 2).
In contrast, the lack of phylogenetic signal in photosynthetic traits, coupled with the clear intergenerational differences we observed, suggest that sporophytes and gametophytes are adapted for different light levels regardless of clade or ecological niche.
In a more densely sampled study of morphological evolution in the trichomanoid clade, Dubuisson et al. (2013) identified several morphological changes including reduced stele and root systems associated with the evolution of epiphytic growth within the HE subclade (comprising the genera *Polyphlebium*, *Didymoglossum*, *Crepidomanes*, and *Vandenboschia*).
It is possible that both reduced morphology and increased DT are parts of an integrated adaptive strategy for epiphytic growth within this subclade of filmy ferns, and have evolved together repeatedly.
However, Dubuisson et al. (2013) were unable to determine if these morphological changes had occurred independently, or if there was a single origin at the base of the HE clade followed by multiple losses.
Increased sampling of both morphological and physiological traits of additional trichomanoid taxa is needed to distinguish between these scenarios.

***Concluding remarks and future directions***---In filmy ferns, both sporophyte and gametophyte generations are poikilohydric; however, we have found here that this does not mean they necessarily share similar physiological optima.
Although filmy fern sporophytes are often of small stature, elevating a leaf even a few cm above the substrate surface means escape from the boundary layer and exposure to substantially greater evapotranspiration.
It is likely that even in this clade, wherein the physiological divide between sporophyte and gametophyte is probably the smallest amongst ferns, the two generations experience substantially different selective pressures.
Fossils indicate that some of the earliest vascular plants may have had sporophytes and gametophytes that were both branched and small (Kenrick and Crane, 1997), but it is unclear at what point selective pressures began to diverge to favor large, complex sporophytes and small, simple gametophytes.
The relevance of plant size to water relations is supported by a recent study showing increased DT in smaller size classes of the epiphytic fern *Asplenium auritum* Sw. (Testo and Watkins, 2012).
Future studies of DT in filmy ferns should include juvenile stages of the sporophyte, which are expected to experience similar microhabitats as gametophytes, to better determine the role of microhabitat vs. life stage (and plant size) on DT.
Finally, more detailed analyses of the strategies used by each generation to cope with desiccation stress, such as transcriptomics, are needed to better understand the ecology and evolution of filmy ferns as well as the transition from desiccation tolerance to avoidance in vascular plants.

ACKNOWLEDGEMENTS

Members of the Davis Lab and Jonathan B. Losos provided helpful discussion and comments on drafts.
Saad Amer, Tohei Theophilus, and Suzanne Vinnette assisted with fieldwork and experiments.
Evan Zeng helped with analysis of microclimatic data.
Weston Testo provided advice on settings for the DT test.
Staff at the University of California Berkeley's Richard B. Gump South Pacific Research Station, Moorea, French Polynesia, in particular Valentine Brotherson, Neil Davies, Hinano Teavai-Murphy, and Frank Murphy, provided logistical support for fieldwork.
Funding provided in part by the National Science Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to JHN and CCD), Setup Funds from Harvard University to CCD, American Society of Plant Taxonomists (Research Grant for Graduate Students to JHN), Garden Club of America (Award in Tropical Botany to JHN), Harvard University Herbaria (Fernald Fieldwork Fellowship to JHN), Society of Systematic Biologists (Graduate Student Research Award to JHN), and Systematics Association (Systematics Research Fund to JHN).

### 2020-10-07 ms/manuscript.Rmd
Update MS

@@
 output: html_document
+bibliography: references.bib
+editor_options: 
+  markdown: 
+    wrap: sentence
 ---
@@
 
-The movement of plants from water to land was one of the most important evolutionary events in the history of the earth, fundamentally altering the global carbon and water cycle and setting the stage for modern terrestrial ecosystems (Graham, 1993; Kenrick and Crane, 1997).
-The transition from aquatic to terrestrial growth was complex, requiring the evolution of a suite of adaptations to survive in dry habitats bombarded with radiation (Waters, 2003).
-The earliest land plants were poikilohydric, i.e., they lacked any mechanisms to regulate water relations (Raven, 1999).
-Therefore, one particularly important early adaptation that enabled plants to survive on land was desiccation tolerance (DT), the ability to lose water until reaching equilibrium with atmospheric humidity, then recover upon rewetting (Mishler and Churchill, 1985).
+The movement of plants from water to land was one of the most important evolutionary events in the history of the earth, fundamentally altering the global carbon and water cycle and setting the stage for modern terrestrial ecosystems [@Graham1993; @Kenrick1997a].
+The transition from aquatic to terrestrial growth was complex, requiring the evolution of a suite of adaptations to survive in dry habitats bombarded with radiation [@Waters2003].
+The earliest land plants were poikilohydric, i.e., they lacked any mechanisms to regulate water relations [@Raven1999].
+Therefore, one particularly important early adaptation that enabled plants to survive on land was desiccation tolerance (DT), the ability to lose water until reaching equilibrium with atmospheric humidity, then recover upon rewetting [@Mishler1985].
 
 Compared with the ubiquitous DT among the earliest land plants, the occurrence of DT varies across extant land plant lineages and life stages.
-Many bryophytes have retained vegetative DT in the gametophyte stage (Oliver et al. 2005, and references therein).
+Many bryophytes have retained vegetative DT in the gametophyte stage [@Oliver2005, and references therein].
 In extant vascular plants (tracheophytes), DT occurs in certain non-vegetative stages (e.g., seeds and pollen), but has largely been lost in vegetative tissues (with some notable exceptions, the so-called 'resurrection plants').
@@
 These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often *\<*1 cm), lack vascular tissue, and are poikilohydric.
-Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT (Watkins et al., 2007; Pittermann et al., 2013).
+Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013a].
 Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.
@@
 low DT and adaptation to low light levels in species occupying more sheltered habitats.
-In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, Johnson et al.
-(2000) found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
+In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, Johnson et al. (2000) found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
 Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
@@
 ***Study system***---We selected the island of Moorea, French Polynesia (17*◦*32' S, 149*◦*50' W) for our survey of filmy fern ecophysiology (Fig. 2).
-Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca. 15 km NW of its larger sister island, Tahiti.
+Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
+15 km NW of its larger sister island, Tahiti.
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
@@
 During this study, 17 sampling sites were established from ca.
-200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt.
-Rotui (899 m), Mt.
-Mouaputa (880 m), and Mt.
-Tohiea (1206 m; Fig.
-2).
+200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt. Rotui (899 m), Mt. Mouaputa (880 m), and Mt. Tohiea (1206 m; Fig. 2).
 At each site, sporophytes were sampled in a single 10 m *×* 10 m plot, and gametophytes sampled in two to three 50 cm *×* 50 cm subplots located within the sporophyte plot.
-Epiphytes were sampled to a maximum height of ca. 2 m on their host plants.
+Epiphytes were sampled to a maximum height of ca.
+2 m on their host plants.
 Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family (e.g., filamentous, ribbon with round gemmae; Nayar and Kaur 1971).
 This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
-We characterized the elevational range of each species based on the results of the field survey; any species found with gametophytes occurring *\>*200 m above the maximum or below the minimum observed elevational range of sporophytes was considered 'widespread' (i.e., facultatively independent sensu Rumsey and Sheffield \[1996\]).
+We characterized the elevational range of each species based on the results of the field survey; any species found with gametophytes occurring *\>*200 m above the maximum or below the minimum observed elevational range of sporophytes was considered 'widespread' (i.e., facultatively independent sensu Rumsey and Sheffield $$1996$$).
 We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, i.e., below ca.
@@
 Next, plants were equilibrated to the actinic light for 30 s, and photosynthetic yield (∆F/F~m~) measured with a 0*.*8 s saturating pulse.
-This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca. 500 µmol·m^--2^ ·s^--1^.
+This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca.
+500 µmol·m^--2^ ·s^--1^.
 Response curves were fitted using the equation *y* = *A*(1 --*e^-kx^*), where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded (Proctor, 2012).
@@
 For each of the three response variables (DT, ETR~max~ , and PPFD~95%~), we used growth habit (epiphyte vs.
-non-epiphyte \[i.e., terrestrial or saxicolous\]) or range size (widespread vs.
+non-epiphyte $$i.e., terrestrial or saxicolous$$) or range size (widespread vs.
 not widespread) and generation (sporophyte vs.
@@
 species complex, but species status of these taxa remains uncertain (Nitta et al., 2011a).
-We refer to these here as *C. minutum* var. 1, 2, and 3.
+We refer to these here as *C. minutum* var.
+1, 2, and 3.
 Similarly, two members of the *Abrodictyum asae-grayi* (Bosch) Ebihara & K.
@@
 Iwats.
-in our tree; Appendix S1) were observed on the basis of *rbcL* and morphology, but as taxonomic treatment is beyond the scope of the current study we refer to them as *A.
-asae-grayi* var.
+in our tree; Appendix S1) were observed on the basis of *rbcL* and morphology, but as taxonomic treatment is beyond the scope of the current study we refer to them as *A. asae-grayi* var.
 1 and 2.
-In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C.
-Presl) Copel., *Crepidomanes minutum* var.
+In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var.
 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
@@
 ***Microclimate***---Mean daily temperature was strongly correlated with elevation, and dropped ca.
-6 *◦*C from the lowest surveyed site (201 m) to the highest (1206 m; Fig.
-3).
+6 *◦*C from the lowest surveyed site (201 m) to the highest (1206 m; Fig. 3).
 Although epiphytic and terrestrial habitats were similar in temperature, epiphytic habitats tended to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones (Fig.
@@
 Saxicolous species (*Crepidomanes kurzii* (Bedd.) Tagawa & K.
-Iwats., *Polyphlebium endlicherianum* (C.
-Presl) Ebihara & K.
-Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C.
-kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
+Iwats., *Polyphlebium endlicherianum* (C. Presl) Ebihara & K.
+Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C. kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
 All epiphytic species were capable of recovering from desiccation to a certain degree, but ability varied between species.
-Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
+Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel.
+was able to recover from desiccation longer than 2 d.
 Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.* *pallidum* and *P.* *borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
-Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.
+Interestingly, *C. minutum* var.
+2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw.
+could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.
 
@@
 minutum* var.
-2, *H.* *digitatum*, *H.
-polyanthos*, and *P.
-borbonicum* (Fig.
-5; *t*-test, *P* = 0.0093, *\<*0.001, *\<*0.001, 0.012, respectively).
+2, *H.* *digitatum*, *H. polyanthos*, and *P. borbonicum* (Fig. 5; *t*-test, *P* = 0.0093, *\<*0.001, *\<*0.001, 0.012, respectively).
 
@@
 5).
-Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca. 142 ± 46 µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at 51 ± 22 µmol·m^--2^ ·s^--1^.
+Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
+142 ± 46 µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at 51 ± 22 µmol·m^--2^ ·s^--1^.
 Similarly, sporophytes attained higher rates of electron transport (ETR~max~; 8.1 ± 2.7 µmol·m^--2^ ·s^--1^) than gametophytes (3.9 ± 0.9 µmol·m^--2^ ·s^--1^).
@@
 Br.
-and *H.
-villosum* Colenso, both high elevation epiphytes with desiccation tolerant sporophytes, and lack of DT in terrestrial filamentous gametophytes of *Polyphlebium colensoi* (Hook.) Ebihara & K.
-Iwats., *Abrodictyum strictum* (Menzies ex Hook.
-& Grev.) Ebihara & K.
-Iwats., and *A.
-elongatum* (A.
-Cunn.) Ebihara & K.
+and *H. villosum* Colenso, both high elevation epiphytes with desiccation tolerant sporophytes, and lack of DT in terrestrial filamentous gametophytes of *Polyphlebium colensoi* (Hook.) Ebihara & K.
+Iwats., *Abrodictyum strictum* (Menzies ex Hook. & Grev.) Ebihara & K.
+Iwats., and *A. elongatum* (A. Cunn.) Ebihara & K.
 Iwats.
@@
 This is perhaps best illustrated by *C. apiifolia* (Fig. 1D, E, H).
-Sporophytes of this species are usually restricted to moist cloud forest sites from ca. 600--1200 m, but we observed gametophytes over a much wider range, down to 200 m.
-A similar distribution pattern has been reported in congeneric *Callistopteris baldwinii* (D. C. Eaton) Copel. from Hawaii (Dassler and Farrar, 1997).
+Sporophytes of this species are usually restricted to moist cloud forest sites from ca.
+600--1200 m, but we observed gametophytes over a much wider range, down to 200 m.
+A similar distribution pattern has been reported in congeneric *Callistopteris baldwinii* (D. C. Eaton) Copel.
+from Hawaii (Dassler and Farrar, 1997).
 Since gametophytes of *C.
@@
 Fossils indicate that some of the earliest vascular plants may have had sporophytes and gametophytes that were both branched and small (Kenrick and Crane, 1997), but it is unclear at what point selective pressures began to diverge to favor large, complex sporophytes and small, simple gametophytes.
-The relevance of plant size to water relations is supported by a recent study showing increased DT in smaller size classes of the epiphytic fern *Asplenium auritum* Sw. (Testo and Watkins, 2012).
+The relevance of plant size to water relations is supported by a recent study showing increased DT in smaller size classes of the epiphytic fern *Asplenium auritum* Sw.
+(Testo and Watkins, 2012).
 Future studies of DT in filmy ferns should include juvenile stages of the sporophyte, which are expected to experience similar microhabitats as gametophytes, to better determine the role of microhabitat vs.

### 2020-10-19 ms/manuscript.Rmd
Update manuscript
to follow JPR format

@@
 ---
-title: "Untitled"
-output: html_document
+title: ""
 bibliography: references.bib
+mainfont: Roboto-Regular.ttf
+fontsize: 12pt
+mainfontoptions: 
+  - BoldFont=Roboto-Bold.ttf
+  - ItalicFont=Roboto-Italic.ttf
+  - BoldItalicFont=Roboto-BoldItalic.ttf
+output: 
+  bookdown::pdf_document2:
+    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
+    toc: no
+    number_sections: no
+    fig_caption: yes
+    keep_tex: yes
+    latex_engine: xelatex
+    pandoc_args: [ "--csl", "journal-of-plant-research.csl"]
+    includes:
+      in_header: journal-of-plant-research.sty
+header-includes: 
+  \usepackage{float}
+  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
 editor_options: 
@@
 
-Intergenerational Niche Differentiation in Filmy Ferns (Hymenophyllaceae)^1^
+<!-- to move to the next line without starting a new paragraph, 
+add two spaces after last word, e.g., Atsushi Ebihara__ -->
+**Corresponding author**  
+Joel H. Nitta  
+Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan  
+Phone: +81-  
+Fax: +81-  
+joelnitta@gmail.com
 
-Joel H. Nitta^2,5,6^, James E. Watkins, Jr.^3^, N. Michele Holbrook^2^, Ravahere Taputuarai^4^, Tristan W. Wang^2^, and Charles C. Davis^2^
+**Membership of the Botanical Society of Japan**  
+YES (Joel H. Nitta)
 
-Running Head: Intergenerational Niche Differentiation in Filmy Ferns
+**Subject area**  
+(1)	Taxonomy/Phylogenetics/Evolutionary Biology
 
-**PREMISE OF THE STUDY**: Desiccation tolerance was a key trait allowing plants to colonize land.
+**Number of tables, black-and-white figures, and color figures.**  
+No tables / 2 black-and-white figure / 1 color figure
+
+<!-- See https://stackoverflow.com/questions/24672111/how-to-add-a-page-break-in-word-document-generated-by-rstudio-markdown/40474753
+for inserting page breaks when rendering to MS word -->
+#####
+\newpage
+
+# Intergenerational Niche Differentiation in Filmy Ferns (Hymenophyllaceae)
+
+Joel H. Nitta^1,4^, James E. Watkins, Jr.^2^, N. Michele Holbrook^1^, Ravahere Taputuarai^3^, Tristan W. Wang^1^, and Charles C. Davis^1^
+
+^1^Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, Cambridge, Massachusetts 02138 USA
+
+^2^Department of Biology, Colgate University, 13 Oak Drive, Hamilton, New York 13346 USA
+
+^3^Association Te Rau Ati Ati a Tau a Hiti Noa Tu, Mahina, Tahiti, French Polynesia
+
+^4^Current address: Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan
+
+Author for correspondence: Joel Nitta (joelnitta\@gmail.com)
+
+ORCID ID numbers  
+- Joel H. Nitta: \textnormal{https://orcid.org/0000-0003-4719-7472}  
+- N. Michele Holbrook: \textnormal{https://orcid.org/0000-0003-3325-5395}  
+- Charles C. Davis: \textnormal{https://orcid.org/0000-0001-8747-1101}
+
+### Abstract
+Desiccation tolerance was a key trait allowing plants to colonize land.
 However, little is known about the transition from desiccation tolerant non-vascular plants to desiccation sensitive vascular ones.
 Filmy ferns (Hymenophyllaceae) present a useful system to study how water-stress strategies differ between non-vascular and vascular stages within a single organism, as they have vascularized sporophytes and nonvascular gametophytes that are each capable of varying degrees of desiccation tolerance.
+We surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia) and used chlorophyll fluorescence to measure desiccation tolerance and light responses.
+We sequenced plastid *rbcL* and *trnHpsbA* to identify field-collected gametophytes, and conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.
+Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima than sporophytes.
+Epiphytes were more tolerant of desiccation than terrestrial species.
+Despite their lack of greater physiological tolerances, gametophytes of several species occurred over a wider elevational range than conspecific sporophytes.
+Our results show that filmy fern gametophytes and sporophytes differ in their physiology and niche requirements, and point to the importance of microhabitat in shaping the evolution of water-use strategies in vascular plants.
 
-**METHODS**: We surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia).
-We used chlorophyll fluorescence to measure desiccation tolerance and light responses.
-We sequenced plastid *rbcL* and *trnHpsbA* to identify field-collected gametophytes, and conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.
+**Keywords**: chlorophyll fluorescence; desiccation tolerance; DNA barcoding; French Polynesia; gametophyte; Hymenophyllaceae
 
-**KEY RESULTS**: Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima than sporophytes.
-Despite their lack of greater physiological tolerances, gametophytes of several species occurred over a wider elevational range than conspecific sporophytes.
-Epiphytes were more tolerant of desiccation than terrestrial species.
+#####
+\newpage
 
-**CONCLUSIONS**: Our results show that filmy fern gametophytes and sporophytes differ in their physiology and niche requirements, and point to the importance of microhabitat in shaping the evolution of water-use strategies in vascular plants.
-
-**KEYWORDS**: chlorophyll fluorescence; desiccation tolerance; DNA barcoding; filmy ferns; French Polynesia; gametophyte; Hymenophyllaceae; Moorea; photosynthesis
-
-^1^Manuscript received \_\_\_\_ ; revision accepted \_\_\_\_.
-
-^2^Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, Cambridge, Massachusetts 02138 USA;
-
-^3^Department of Biology, Colgate University, 13 Oak Drive, Hamilton, New York 13346 USA; and
-
-^4^Association Te Rau Ati Ati a Tau a Hiti Noa Tu, Mahina, Tahiti, French Polynesia
-
-^5^Current address: Department of Botany, National Museum of Nature and Science, 4-1-1 Amakubo, Tsukuba, Ibaraki 305-0005 Japan
-
-^6^Author for correspondence (email: joelnitta\@gmail.com)
+## Introduction
 
@@
 Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
-These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often *\<*1 cm), lack vascular tissue, and are poikilohydric.
-Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013a].
+These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often <1 cm), lack vascular tissue, and are poikilohydric.
+Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013].
 Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.
@@
 1A--D).
-Filmy fern sporophytes are an exception to the 'typical' fern sporophyte described above because they have reverted to poikilohydry: their thin laminae allow for passive, rapid uptake and loss of moisture, thereby maintaining nearly constant equilibrium with atmospheric humidity (Raven, 1999).
-Many species of epiphytic filmy ferns have extremely reduced root systems (Schneider, 2000), relying completely upon absorption through leaf laminae for water (Shreve, 1911), although terrestrial species possess true roots and use ground water (Dubuisson et al., 2011).
+Filmy fern sporophytes are an exception to the 'typical' fern sporophyte described above because they have reverted to poikilohydry: their thin laminae allow for passive, rapid uptake and loss of moisture, thereby maintaining nearly constant equilibrium with atmospheric humidity [@Raven1999].
+Many species of epiphytic filmy ferns have extremely reduced root systems [@Schneider2000], relying completely upon absorption through leaf laminae for water [@Shreve1911], although terrestrial species possess true roots and use ground water [@Dubuisson2011].
 As would be expected from their poikilohydric nature, DT of varying degrees has been reported in sporophytes of many filmy fern species (see references below).
 
-Filmy fern sporophytes encompass a broad diversity of growth forms (terrestrial, saxicolous, hemi-epiphytic, and holoepiphytic) and morphologies, with rhizome and leaf sizes varying over an order of magnitude (Hennequin et al., 2008; Nitta and Epps, 2009; Dubuisson et al., 2013).
+Filmy fern sporophytes encompass a broad diversity of growth forms (terrestrial, saxicolous, hemi-epiphytic, and holoepiphytic) and morphologies, with rhizome and leaf sizes varying over an order of magnitude [@Hennequin2008; @Nitta2009; @Dubuisson2013].
 Filmy fern gametophytes are similarly diverse, with morphotypes including filamentous, ribbon, and a combination of these two types (Fig. 1E--H).
-While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages (Watkins et al., 2007; Watkins and Cardelús, 2012; Pittermann et al., 2013), and may be expected to occur in filmy ferns as well.
+While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages [@WatkinsJr.2007a; @WatkinsJr.2012; @Pittermann2013], and may be expected to occur in filmy ferns as well.
 The possibility for varying degrees of vegetative DT in both stages of the life cycle of filmy ferns makes them a potentially useful system to compare how DT is utilized between haploid, non-vascular and diploid, vascular life stages.
@@
 The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
-In a series of elegant experiments on Jamaican filmy ferns, Shreve (1911) demonstrated that water-use strategies differ between species occupying contrasting habitats (i.e., terrestrial, low elevation epiphyte, and high elevation epiphyte).
-Proctor (2003, 2012) investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs.
+In a series of elegant experiments on Jamaican filmy ferns, [@Shreve1911] demonstrated that water-use strategies differ between species occupying contrasting habitats (i.e., terrestrial, low elevation epiphyte, and high elevation epiphyte).
+Proctor [-@Proctor2003; -@Proctor2012] investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs.
 low DT and adaptation to low light levels in species occupying more sheltered habitats.
-In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, Johnson et al. (2000) found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
+In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, @Johnson2000 found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
 Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
-Parra et al.
-(2009) and Saldan˜a et al.
-(2014) investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
+@Parra2009 and @Saldana2013 investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
 They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
-A recent study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly (Cea et al., 2014).
+A recent study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly [@GarcesCea2014].
 
-Despite this considerable interest in filmy fern ecophysiology, with the sole exception of Johnson et al.
-(2000), no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
+Despite this considerable interest in filmy fern ecophysiology, with the sole exception of @Johnson2000, no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
 Furthermore, to our knowledge, no physiological studies of filmy ferns have accounted for their phylogenetic relationships, and we lack an understanding of how physiological strategies have evolved in this clade.
-In addition to DT, photosynthetic optima (adaptation to different light levels) are likely to be important in determining the niches of life stages and species of filmy ferns (Proctor, 2003; Parra et al., 2015; Flores-Bavestrello et al., 2016).
-Terrestrial and epiphytic species experience different levels of photosynthetic radiation and desiccation, and light intensity varies with height along the host plant in epiphytes (Benzing, 1990; Watkins and Cardelús, 2009).
+In addition to DT, photosynthetic optima (adaptation to different light levels) are likely to be important in determining the niches of life stages and species of filmy ferns [@Proctor2003; @Parra2015; @Flores-Bavestrello2016].
+Terrestrial and epiphytic species experience different levels of photosynthetic radiation and desiccation, and light intensity varies with height along the host plant in epiphytes  [@Benzing1990; @WatkinsJr.2009].
 In a recent study (Testo et al., unpublished), the combined effect of desiccation and light stress was found to have a greater impact on filmy fern physiological performance than either of these stressors alone, suggesting the need to integrate studies of DT with photosynthesis to gain a full picture of physiological adaptation in filmy ferns.
-Here, we leverage a recently developed DNA barcoding system (Nitta et al., 2016) to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes: 1.) Epiphytic filmy ferns have higher DT and photosynthetic optima than non-epiphytic species.
+Here, we leverage a recently developed DNA barcoding system [@Nitta2017] to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes: 1.) Epiphytic filmy ferns have higher DT and photosynthetic optima than non-epiphytic species.
 2.) Amongst epiphytic filmy ferns, those occurring at high elevation have higher DT and photosynthetic optima than those at low elevation.
@@
 
-MATERIALS AND METHODS
+<!-- 
+## Materials and methods
 
-***Study system***---We selected the island of Moorea, French Polynesia (17*◦*32' S, 149*◦*50' W) for our survey of filmy fern ecophysiology (Fig. 2).
+### Study system
+
+We selected the island of Moorea, French Polynesia (17*◦*32' S, 149*◦*50' W) for our survey of filmy fern ecophysiology (Fig. 2).
 Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
@@
 
-***Field survey***---We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti (Nitta et al., 2016).
+### Field survey
+
+We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti (Nitta et al., 2016).
 During this study, 17 sampling sites were established from ca.
@@
 
-***Microclimate***---We characterized the microenvironment experienced by filmy ferns by logging relative humidity (RH) and temperature every 15 min from July 7, 2013 to July 5, 2014 at 16 of the sampling sites (Fig.
-2).
+### Microclimate
+
+We characterized the microenvironment experienced by filmy ferns by logging relative humidity (RH) and temperature every 15 min from July 7, 2013 to July 5, 2014 at 16 of the sampling sites (Fig. 2).
 To do this we used Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) with external sensors outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation.
@@
 
-***Desiccation tolerance***---Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
+### Desiccation tolerance
+
+Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
 For sporophytes, 8--16 individuals (one individual = single whole frond including a ca.
@@
 
-***Light responses***---Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400--700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
+### Light responses
+
+Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400--700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
 Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
@@
 
-***DNA barcoding***---We used DNA barcoding to identify filmy fern gametophytes to species as described in Nitta et al. (2016).
+### DNA barcoding
+
+We used DNA barcoding to identify filmy fern gametophytes to species as described in Nitta et al. (2016).
 Briefly, DNA extraction was performed using a modified CTAB protocol (Doyle and Doyle, 1987) or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
@@
 
-***Statistical analysis and phylogenetic comparative methods***---All statistical analyses were done in R v.3.2.1 (R Core Team, 2015).
+### Statistical analysis and phylogenetic comparative methods
+
+All statistical analyses were done in R v.3.2.1 (R Core Team, 2015).
 Datasets and R scripts used in this study are available on Dryad (\#\#\#\#\#\#).
@@
 Two species were observed both as low-elevation epiphytes and saxicoles (Table 1), but we treated them as epiphytic to distinguish them from exclusively saxicolous species, which we regard as the more specialized growth habit.
-We accounted for intraspecific variation and phylogeny (only for response variables with significant phylogenetic signal) by including species (within generation) and an inverse phylogenetic variance-covariance matrix as random effects, respectively (Hadfield, 2010).
+We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix only for response variables with significant phylogenetic signal (Hadfield, 2010).
 We used an inverse-Gamma distribution for all priors, and ran analyses for 500 000 iterations, with burnin after 1000 iterations and thinning every 500 iterations.
@@
 We used minimum values because these may represent a climatic extreme that limits the occurrence of species.
-We used data from the 'epiphytic' dataloggers mounted at 1*.*5 m on trees for epiphytic species and data from the 'terrestrial' dataloggers on the ground for non-epiphytic species.
+We used data from the 'epiphytic' dataloggers mounted at 1.5 m on trees for epiphytic species and data from the 'terrestrial' dataloggers on the ground for non-epiphytic species.
 Data from the Mt.
@@
 
-RESULTS
+## Results
 
-***Field survey***---Twenty-two filmy fern taxa (including one new record and two species that were recognized as multiple varieties) were observed as sporophytes, and 18 as gametophytes on Moorea in total (Table 1).
+### Field survey
+
+Twenty-two filmy fern taxa (including one new record and two species that were recognized as multiple varieties) were observed as sporophytes, and 18 as gametophytes on Moorea in total (Table 1).
 *Hymenophyllum braithwaitei* (Ebihara & K. Iwats.), only observed near the peak of Mt. Tohiea (1206 m), is a new record for Moorea and French Polynesia.
@@
 
-***Microclimate***---Mean daily temperature was strongly correlated with elevation, and dropped ca.
+### Microclimate
+
+Mean daily temperature was strongly correlated with elevation, and dropped ca.
 6 *◦*C from the lowest surveyed site (201 m) to the highest (1206 m; Fig. 3).
@@
 
-***Phylogenetic signal***---Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* and λ both near 1, *P \<*0.05; sporophytes *K* = 0.78, *P* = 0.02 and λ = 0.96, *P* = 0.11), but no signal for PPFD~95%~ or ETR~max~ (Table 2).
+### Phylogenetic signal
 
-***Desiccation tolerance***---Ability to recover from desiccation was tested in sporophytes of 15 species and gametophytes of 14 species, including 12 species with both life phases (Appendix S3).
+Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* and λ both near 1, *P \<*0.05; sporophytes *K* = 0.78, *P* = 0.02 and λ = 0.96, *P* = 0.11), but no signal for PPFD~95%~ or ETR~max~ (Table 2).
+
+### Desiccation tolerance
+
+Ability to recover from desiccation was tested in sporophytes of 15 species and gametophytes of 14 species, including 12 species with both life phases (Appendix S3).
 A general correlation of DT with habitat was observed in sporophytes (Fig. 4).
@@
 
-***Photosynthetic optima***---Light responses were measured in sporophytes of 18 species and gametophytes of 11 species, including nine species with both life phases (Appendix S4).
+### Photosynthetic optima
+
+Light responses were measured in sporophytes of 18 species and gametophytes of 11 species, including nine species with both life phases (Appendix S4).
 Unlike DT, photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes are adapted for higher light levels, whereas gametophytes are adapted for lower light levels (Fig.
@@
 
-***Environmental water potential and DT***---Desiccation tolerance tended to increase with decreasing environmental water potential for both gametophytes and sporophytes (Fig. 6).
+### Environmental water potential and DT
+
+Desiccation tolerance tended to increase with decreasing environmental water potential for both gametophytes and sporophytes (Fig. 6).
 These trends were significant when analyzed using PGLS for sporophytes (*P* = 0.036) but not gametophytes (*P* = 0.123; Appendix S5).
 
-DISCUSSION
+## Discussion
 
@@
 
-***Filmy fern sporophytes have a wide range of DT***---We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below --200 MPa in some epiphytes.
+### Filmy fern sporophytes have a wide range of DT
+
+We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below --200 MPa in some epiphytes.
 Levels of similarly extreme DT have also been reported in other species of *Hymenophyllum* (Proctor, 2003, 2012) and a general correlation of habitat with DT in filmy ferns has been previously established (Shreve, 1911; Proctor, 2003, 2012; Parra et al., 2009; Saldan˜a et al., 2014; Testo et al., unpublished), but no other studies to our knowledge have documented a similar range of DT values in filmy ferns from a single site.
@@
 
-***Filmy fern gametophytes are not more stress-tolerant than sporophytes***---Contrary to our hypotheses, gametophytes of filmy ferns are in general not more tolerant of the abiotic environment (light levels and water potential) than sporophytes.
+### Filmy fern gametophytes are not more stress-tolerant than sporophytes
+
+Contrary to our hypotheses, gametophytes of filmy ferns are in general not more tolerant of the abiotic environment (light levels and water potential) than sporophytes.
 This differs from other fern groups with desiccation tolerant gametophytes but desiccation sensitive sporophytes (Watkins et al., 2007).
@@
 
-***Gametophytes are adapted for lower light levels than sporophytes***---Unlike DT, photosynthetic parameters did not vary significantly between species from different habitats.
+### Gametophytes are adapted for lower light levels than sporophytes
+
+Unlike DT, photosynthetic parameters did not vary significantly between species from different habitats.
 Rather, we observed a clear intergenerational difference, with gametophytes consistently adapted for lower light levels than sporophytes (Fig. 5, Table 3).
@@
 
-***Influence of phylogeny on traits***---Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates (Proctor, 2003, 2012; Parra et al., 2009; Saldan˜a et al., 2014); however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
+### Influence of phylogeny on traits
+
+Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates (Proctor, 2003, 2012; Parra et al., 2009; Saldan˜a et al., 2014); however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
 We detected significant phylogenetic signal in DT, suggesting that phylogeny should be considered when interpreting differences in DT between species (Table 2).
@@
 
-***Concluding remarks and future directions***---In filmy ferns, both sporophyte and gametophyte generations are poikilohydric; however, we have found here that this does not mean they necessarily share similar physiological optima.
+### Concluding remarks and future directions
+
+In filmy ferns, both sporophyte and gametophyte generations are poikilohydric; however, we have found here that this does not mean they necessarily share similar physiological optima.
 Although filmy fern sporophytes are often of small stature, elevating a leaf even a few cm above the substrate surface means escape from the boundary layer and exposure to substantially greater evapotranspiration.
@@
 
-ACKNOWLEDGEMENTS
+Acknowledgments
 
@@
 Funding provided in part by the National Science Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to JHN and CCD), Setup Funds from Harvard University to CCD, American Society of Plant Taxonomists (Research Grant for Graduate Students to JHN), Garden Club of America (Award in Tropical Botany to JHN), Harvard University Herbaria (Fernald Fieldwork Fellowship to JHN), Society of Systematic Biologists (Graduate Student Research Award to JHN), and Systematics Association (Systematics Research Fund to JHN).
+
+-->
+
+## References
+<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
+
+<div id="refs"></div>

### 2020-10-19 ms/manuscript.Rmd
Update MS

@@
 
-<!-- 
 ## Materials and methods
@@
 
-We selected the island of Moorea, French Polynesia (17*◦*32' S, 149*◦*50' W) for our survey of filmy fern ecophysiology (Fig. 2).
-Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
-15 km NW of its larger sister island, Tahiti.
+We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology (Fig. 2).
+Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca. 15 km NW of its larger sister island, Tahiti.
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
-At the time of the study, a total of 18 filmy fern species were known from Moorea (Murdock and Smith, 2003; Ranker et al., 2005; Nitta et al., 2011b; Florence, in press), including multiple terrestrial, saxicolous, and epiphytic species.
-The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages (Pryer et al., 2001), are represented by similar numbers of species.
-We follow the taxonomic system of Ebihara et al. (2006).
+At the time of the study, a total of 18 filmy fern species were known from Moorea [@Murdock2003; @Ranker2005; @Nitta2011b; @Florence], including multiple terrestrial, saxicolous, and epiphytic species.
+The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by similar numbers of species.
+We follow the taxonomic system of @Ebihara2006.
 
@@
 
-We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti (Nitta et al., 2016).
-During this study, 17 sampling sites were established from ca.
-200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt. Rotui (899 m), Mt. Mouaputa (880 m), and Mt. Tohiea (1206 m; Fig. 2).
-At each site, sporophytes were sampled in a single 10 m *×* 10 m plot, and gametophytes sampled in two to three 50 cm *×* 50 cm subplots located within the sporophyte plot.
-Epiphytes were sampled to a maximum height of ca.
-2 m on their host plants.
-Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family (e.g., filamentous, ribbon with round gemmae; Nayar and Kaur 1971).
+We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti [@Nitta2017].
+During this study, 17 sampling sites were established from \circa 200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt.
+Rotui (899 m), Mt.
+Mouaputa (880 m), and Mt.
+Tohiea (1206 m; Fig.
+2).
+At each site, sporophytes were sampled in a single 10 m × 10 m plot, and gametophytes sampled in two to three 50 cm × 50 cm subplots located within the sporophyte plot.
+Epiphytes were sampled to a maximum height of \circa 2 m on their host plants.
+Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family [\eg, filamentous, ribbon with round gemmae; @Nayar1971].
 This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
-We characterized the elevational range of each species based on the results of the field survey; any species found with gametophytes occurring *\>*200 m above the maximum or below the minimum observed elevational range of sporophytes was considered 'widespread' (i.e., facultatively independent sensu Rumsey and Sheffield $$1996$$).
-We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, i.e., below ca.
-500 m), or high elevation epiphyte (mostly occurring in cloud forest, i.e., above ca.
-500 m).
+We characterized the elevational range of each species based on the results of the field survey; any species found with gametophytes occurring *\>*200 m above the maximum or below the minimum observed elevational range of sporophytes was considered 'widespread' [\ie, facultatively independent *sensu* @Rumsey1996].
+We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, \ie, below \circa 500 m), or high elevation epiphyte (mostly occurring in cloud forest, \ie, above \circa 500 m).
 
+<!-- 
 ### Microclimate

### 2020-10-20 ms/manuscript.Rmd
Read in climate data

@@
     wrap: sentence
+  chunk_output_type: console
 ---
 
+```{r setup}
+library(jntools)
+```
+
+```{r load-data}
+loadd(
+  list = c(
+    "climate"
+  )
+)
+```
+
+```{=html}
 <!-- to move to the next line without starting a new paragraph, 
 add two spaces after last word, e.g., Atsushi Ebihara__ -->
-**Corresponding author**  
-Joel H. Nitta  
-Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan  
-Phone: +81-  
-Fax: +81-  
-joelnitta@gmail.com
+```
+**Corresponding author**\
+Joel H.
+Nitta\
+Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan\
+Phone: +81-\
+Fax: +81-\
+[joelnitta\@gmail.com](mailto:joelnitta@gmail.com){.email}
 
-**Membership of the Botanical Society of Japan**  
+**Membership of the Botanical Society of Japan**\
 YES (Joel H. Nitta)
 
-**Subject area**  
+**Subject area**\
 (1) Taxonomy/Phylogenetics/Evolutionary Biology
 
-**Number of tables, black-and-white figures, and color figures.**  
+**Number of tables, black-and-white figures, and color figures.**\
 No tables / 2 black-and-white figure / 1 color figure
 
+```{=html}
 <!-- See https://stackoverflow.com/questions/24672111/how-to-add-a-page-break-in-word-document-generated-by-rstudio-markdown/40474753
 for inserting page breaks when rendering to MS word -->
+```
 #####
+
 \newpage
@@
 
-ORCID ID numbers  
-- Joel H. Nitta: \textnormal{https://orcid.org/0000-0003-4719-7472}  
-- N. Michele Holbrook: \textnormal{https://orcid.org/0000-0003-3325-5395}  
+ORCID ID numbers\
+- Joel H. Nitta: \textnormal{https://orcid.org/0000-0003-4719-7472}\
+- N
+. Michele Holbrook: \textnormal{https://orcid.org/0000-0003-3325-5395}\
 - Charles C. Davis: \textnormal{https://orcid.org/0000-0001-8747-1101}
@@
 ### Abstract
+
 Desiccation tolerance was a key trait allowing plants to colonize land.
@@
 #####
+
 \newpage
@@
 Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
-These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often <1 cm), lack vascular tissue, and are poikilohydric.
+These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often \<1 cm), lack vascular tissue, and are poikilohydric.
 Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013].
@@
 low DT and adaptation to low light levels in species occupying more sheltered habitats.
-In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, @Johnson2000 found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
+In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, [@Johnson2000] found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
 Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
-@Parra2009 and @Saldana2013 investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
+[@Parra2009] and [@Saldana2013] investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
 They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
@@
 
-Despite this considerable interest in filmy fern ecophysiology, with the sole exception of @Johnson2000, no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
+Despite this considerable interest in filmy fern ecophysiology, with the sole exception of [@Johnson2000], no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
 Furthermore, to our knowledge, no physiological studies of filmy ferns have accounted for their phylogenetic relationships, and we lack an understanding of how physiological strategies have evolved in this clade.
@@
 2.) Amongst epiphytic filmy ferns, those occurring at high elevation have higher DT and photosynthetic optima than those at low elevation.
-3.) For filmy ferns with gametophytes that occur over a wider range than sporophytes, the gametophyte has higher DT and photosynthetic optima than the sporophyte.
+3.) Gametophytes with wider ranges have higher DT than those with with narrow ranges.
 
@@
 We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology (Fig. 2).
-Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca. 15 km NW of its larger sister island, Tahiti.
+Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
+15 km NW of its larger sister island, Tahiti.
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
@@
 The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by similar numbers of species.
-We follow the taxonomic system of @Ebihara2006.
+We follow the taxonomic system of [@Ebihara2006].
 
@@
 
-<!-- 
 ### Microclimate
 
-We characterized the microenvironment experienced by filmy ferns by logging relative humidity (RH) and temperature every 15 min from July 7, 2013 to July 5, 2014 at 16 of the sampling sites (Fig. 2).
-To do this we used Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) with external sensors outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation.
-To compare between terrestrial and epiphytic habitats, dataloggers were placed in pairs at each site (one 'epiphytic' datalogger at ca.
+```{r microclimate-stats}
+climate_start <- climate %>% pull(date) %>% min()
+climate_end <- climate %>% pull(date) %>% max()
+climate_n_sites <- climate %>% pull(site) %>% n_distinct()
+climate_n_days <- climate %>% pull(date) %>% n_distinct()
+climate_n_loggers <- climate %>% count(site, habit) %>% nrow()
+```
+
+We used microclimate data collected by [@Nitta2020].
+This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA).
+The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at ca.
 1.5 m on a tree, one 'terrestrial' datalogger at ca.
@@
 During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
-We excluded any days during which one or more dataloggers failed from the final dataset, and some sites are missing data completely for one or both dataloggers.
+Any days during which one or more dataloggers failed are excluded from the dataset.
 One terrestrial datalogger (Mt.
-Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013).
-We used the previous year's data for this site instead.
-The final dataset included 201 days of data from 15 sites (26 dataloggers).
-We tested for differences in climatic variables between epiphytic and terrestrial habitats using analysis of covariance (ANCOVA) with each climatic variable in turn as the response variable and elevation as the explanatory variable with habitat (terrestrial vs.
-epiphytic) as a cofactor.
+Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013), so the previous year's data for this site was used instead.
+The dataset includes `r climate_n_days` days of data from `r climate_n_sites` sites (`r climate_n_loggers` dataloggers).
 
@@
 Pre-treatment chlorophyll fluorescence (F~v~/F~m~) was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
-Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval (Testo and Watkins, 2013).
+Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
 Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
-The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22--25 *◦*C.
-Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (--282 MPa), Mg(NO~3~)~2~ (--86 MPa), and NaCl (--38 MPa).
-Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at --86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
+The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22°C to 25°C.
+Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
+Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at -86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
 Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
 Following the desiccation treatment, plants were rewetted using water from a local stream.
-Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0*.*5 h, 24 h, and 48 h following rewetting.
-Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield (Watkins et al., 2007).
+Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0.5 h, 24 h, and 48 h following rewetting.
+Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield [@WatkinsJr.2007c].
 
@@
 
-Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400--700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
+Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400 nm to 700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
 Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
-Next, plants were equilibrated to the actinic light for 30 s, and photosynthetic yield (∆F/F~m~) measured with a 0*.*8 s saturating pulse.
+Next, plants were equilibrated to the actinic light for 30 s, and photosynthetic yield (∆F/F~m~) measured with a 0.8 s saturating pulse.
 This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca.
-500 µmol·m^--2^ ·s^--1^.
-Response curves were fitted using the equation *y* = *A*(1 --*e^-kx^*), where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded (Proctor, 2012).
+500 µmol·m^-2^ ·s^-1^.
+Response curves were fitted using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded [@Proctor2012].
 The PPFD at 95% saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
@@
 
-We used DNA barcoding to identify filmy fern gametophytes to species as described in Nitta et al. (2016).
-Briefly, DNA extraction was performed using a modified CTAB protocol (Doyle and Doyle, 1987) or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
-Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following Nitta et al. (2016).
-For each marker, a reference library was constructed from sporophyte sequences using the makeblastDB command in BLAST (Altschul et al., 1997).
+We used DNA barcoding to identify filmy fern gametophytes to species as described in @Nitta2017.
+Briefly, DNA extraction was performed using a modified CTAB protocol [@Doyle1987] or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
+Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following @Nitta2017.
+For each marker, a reference library was constructed from sporophyte sequences using the makeblastDB command in BLAST [@Altschul1997].
 Field-collected gametophytes were identified by a local BLAST query against the reference library.
-Gametophytes matching *\>*99.5% with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched *\>*99.5% with multiple reference sequences were not identified and excluded from further analysis.
+Gametophytes matching >99.5% with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched >99.5% with multiple reference sequences were not identified and excluded from further analysis.
 
@@
 
-All statistical analyses were done in R v.3.2.1 (R Core Team, 2015).
+All statistical analyses were done in R v.4.0.2 [@RCoreTeam2020a].
 Datasets and R scripts used in this study are available on Dryad (\#\#\#\#\#\#).
 
-We derived the phylogenetic tree used in this study by extracting Hymenophyllaceae from the tree of Nitta et al. (2016) with the extract.clade function of the ape package in R (Paradis et al., 2004).
-This tree is well-supported and generally in accord with previously published, more densely sampled phylogenies (Pryer et al., 2001; Dubuisson et al., 2003; Hennequin et al., 2006; Ebihara et al., 2007; Appendix S1, see Supplemental Data with the online version of this article).
+We derived the phylogenetic tree used in this study by extracting Hymenophyllaceae from the tree of @Nitta2017 with the extract.clade function of the ape package in R [@Paradis2004].
+This tree is well-supported and generally in accord with previously published, more densely sampled phylogenies [@Pryer2001a; @Dubuisson2003; @Hennequin2006; @Ebihara2007a].
 
-We quantified the degree of phylogenetic signal in physiological traits using Blomberg's *K* (Blomberg et al., 2003) and Pagel's λ (Pagel, 1999; Freckleton et al., 2002) as implemented with the phylosig function in the phytools R package (Revell, 2012).
+We quantified the degree of phylogenetic signal in physiological traits using Blomberg's *K* [@Blomberg2003] and Pagel's λ [@Pagel1999; @Freckleton2002] as implemented with the phylosig function in the phytools R package [@Revell2012].
 Both measures test the hypothesis that the trait of interest is evolving according to Brownian Motion (BM).
 λ is a scaling parameter that ranges from zero to one: λ near zero indicates random distribution of trait values across the tree, and λ near one indicates evolution of traits along the phylogeny following BM.
-For *K*, values near one indicate evolution of traits following BM; *K* *\>*1 indicates traits more conserved than expected under BM, and *K* *\<*1 indicates that traits have less phylogenetic signal than expected under BM.
+For *K*, values near one indicate evolution of traits following BM; *K* >1 indicates traits more conserved than expected under BM, and *K* <1 indicates that traits have less phylogenetic signal than expected under BM.
 We tested the significance of *K* by comparing the observed value against values from a null distribution of 1000 phylogenies with the traits randomly shuffled across the tips.
@@
 
-We used general linear mixed models (GLMMs) to investigate the effect of growth habit, range size, and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at --86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the MCMCglmm R package (Hadfield, 2010).
-For each of the three response variables (DT, ETR~max~ , and PPFD~95%~), we used growth habit (epiphyte vs.
-non-epiphyte $$i.e., terrestrial or saxicolous$$) or range size (widespread vs.
-not widespread) and generation (sporophyte vs.
+We used general linear mixed models (GLMMs) to investigate the effect of growth habit and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at -86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the MCMCglmm R package (Hadfield, 2010).
+For each of the three response variables (DT, ETR~max~ , and PPFD~95%~), we used growth habit (high-elevation epiphyte, low-elevation epiphyte, terrestrial or saxicolous) and generation (sporophyte vs.
 gametophyte) and their interactions as fixed effects.
-Two species were observed both as low-elevation epiphytes and saxicoles (Table 1), but we treated them as epiphytic to distinguish them from exclusively saxicolous species, which we regard as the more specialized growth habit.
-We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix only for response variables with significant phylogenetic signal (Hadfield, 2010).
-We used an inverse-Gamma distribution for all priors, and ran analyses for 500 000 iterations, with burnin after 1000 iterations and thinning every 500 iterations.
-Comparisons of DT, ETR~max~ , and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a *t*-test for these.
+We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix in the case the response variable had significant phylogenetic signal (Hadfield, 2010).
+We used an inverse-Gamma distribution for all priors, and ran analyses for 500,000 iterations, with burnin after 1,000 iterations and thinning every 500 iterations.
+Comparisons of DT, ETR~max~, and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a *t*-test for these.
 
@@
 Since DT showed significant phylogenetic signal (see Results), we used phylogenetic generalized least squares (PGLS) to test for a relationship between environmental water potential and DT while taking phylogeny into account, using species' means.
-We ran PGLS using the caper R package (Orme et al., 2013).
+
+To test whether gametophytes with wider ranges have higher DT, we conducted PGLS with DT as the response variable and range width (measured as the difference between maximum and minimum observed elevation) as the explanatory variable.
+We ran PGLS using the caper R package [@Orme2018].
+
+<!--
 
@@
 species complex, but species status of these taxa remains uncertain (Nitta et al., 2011a).
-We refer to these here as *C. minutum* var.
-1, 2, and 3.
+We refer to these here as *C. minutum* var. 1, 2, and 3.
 Similarly, two members of the *Abrodictyum asae-grayi* (Bosch) Ebihara & K.
@@
 Iwats.
-in our tree; Appendix S1) were observed on the basis of *rbcL* and morphology, but as taxonomic treatment is beyond the scope of the current study we refer to them as *A. asae-grayi* var.
+in our tree; Appendix S1) were observed on the basis of *rbcL* and morphology, but as taxonomic treatment is beyond the scope of the current study we refer to them as *A.
+asae-grayi* var.
 1 and 2.
-In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var.
+In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C.
+Presl) Copel., *Crepidomanes minutum* var.
 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
@@
 
--->
+--\>
 
 ## References
+
 <!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
 
-<div id="refs"></div>
+::: {#refs}
+:::

### 2020-10-20 R/plan.R

@@ -17,7 +17,7 @@ plan <- drake_plan(
     out3 = file_out("data/nitta_2017/all_plots.csv")),
 
   # -microclimate data
-  # hobo = readr::read_csv("data/filmy_hobo_data.csv"),
+  climate = readr::read_csv("data_raw/doi_10.5061_dryad.fqz612jps__v2/moorea_climate.csv"),
 
   # - site data
   moorea_sites = readr::read_csv(

### 2020-10-20 ms/manuscript.Rmd
Fix some Rmd weirdness induced by RStudio RMD editor

@@
 
-```{r setup}
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
 library(jntools)
@@
 
-```{=html}
 <!-- to move to the next line without starting a new paragraph, 
 add two spaces after last word, e.g., Atsushi Ebihara__ -->
-```
-**Corresponding author**\
-Joel H.
-Nitta\
-Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan\
-Phone: +81-\
-Fax: +81-\
-[joelnitta\@gmail.com](mailto:joelnitta@gmail.com){.email}
 
-**Membership of the Botanical Society of Japan**\
+**Corresponding author**  
+Joel H. Nitta  
+Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan  
+Phone: +81-  
+Fax: +81-  
+email: joelnitta\@gmail.com
+
+**Membership of the Botanical Society of Japan**  
 YES (Joel H. Nitta)
 
-**Subject area**\
+**Subject area**  
 (1) Taxonomy/Phylogenetics/Evolutionary Biology
 
-**Number of tables, black-and-white figures, and color figures.**\
+**Number of tables, black-and-white figures, and color figures.**  
 No tables / 2 black-and-white figure / 1 color figure
 
-```{=html}
 <!-- See https://stackoverflow.com/questions/24672111/how-to-add-a-page-break-in-word-document-generated-by-rstudio-markdown/40474753
 for inserting page breaks when rendering to MS word -->
-```
 #####
@@
 
-ORCID ID numbers\
-- Joel H. Nitta: \textnormal{https://orcid.org/0000-0003-4719-7472}\
-- N
-. Michele Holbrook: \textnormal{https://orcid.org/0000-0003-3325-5395}\
-- Charles C. Davis: \textnormal{https://orcid.org/0000-0001-8747-1101}
+ORCID ID numbers  
+Joel H. Nitta: \textnormal{https://orcid.org/0000-0003-4719-7472}  
+N. Michele Holbrook: \textnormal{https://orcid.org/0000-0003-3325-5395}  
+Charles C. Davis: \textnormal{https://orcid.org/0000-0001-8747-1101}
 
@@
 We surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia) and used chlorophyll fluorescence to measure desiccation tolerance and light responses.
-We sequenced plastid *rbcL* and *trnHpsbA* to identify field-collected gametophytes, and conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.
+We conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.
 Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima than sporophytes.
@@
 The transition from aquatic to terrestrial growth was complex, requiring the evolution of a suite of adaptations to survive in dry habitats bombarded with radiation [@Waters2003].
-The earliest land plants were poikilohydric, i.e., they lacked any mechanisms to regulate water relations [@Raven1999].
+The earliest land plants were poikilohydric, \ie, they lacked any mechanisms to regulate water relations [@Raven1999].
 Therefore, one particularly important early adaptation that enabled plants to survive on land was desiccation tolerance (DT), the ability to lose water until reaching equilibrium with atmospheric humidity, then recover upon rewetting [@Mishler1985].
@@
 Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
-These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often \<1 cm), lack vascular tissue, and are poikilohydric.
+These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often <1 cm), lack vascular tissue, and are poikilohydric.
 Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013].
@@
 The possibility for varying degrees of vegetative DT in both stages of the life cycle of filmy ferns makes them a potentially useful system to compare how DT is utilized between haploid, non-vascular and diploid, vascular life stages.
-Furthermore, the diversity of growth forms and habitats in filmy ferns makes them an ideal group for a comparative study of water-stress strategies (i.e., desiccation tolerance vs.
+Furthermore, the diversity of growth forms and habitats in filmy ferns makes them an ideal group for a comparative study of water-stress strategies (\ie, desiccation tolerance vs.
 desiccation avoidance) and the correlation of these strategies with ecological niche.
@@
 The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
-In a series of elegant experiments on Jamaican filmy ferns, [@Shreve1911] demonstrated that water-use strategies differ between species occupying contrasting habitats (i.e., terrestrial, low elevation epiphyte, and high elevation epiphyte).
+In a series of elegant experiments on Jamaican filmy ferns, @Shreve1911 demonstrated that water-use strategies differ between species occupying contrasting habitats (i.e., terrestrial, low elevation epiphyte, and high elevation epiphyte).
 Proctor [-@Proctor2003; -@Proctor2012] investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs.
 low DT and adaptation to low light levels in species occupying more sheltered habitats.
-In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, [@Johnson2000] found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
+In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, @Johnson2000 found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
 Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
-[@Parra2009] and [@Saldana2013] investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
+@Parra2009 and @Saldana2013 investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
 They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
@@
 
-Despite this considerable interest in filmy fern ecophysiology, with the sole exception of [@Johnson2000], no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
+Despite this considerable interest in filmy fern ecophysiology, with the sole exception of @Johnson2000, no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
 Furthermore, to our knowledge, no physiological studies of filmy ferns have accounted for their phylogenetic relationships, and we lack an understanding of how physiological strategies have evolved in this clade.
@@
 The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by similar numbers of species.
-We follow the taxonomic system of [@Ebihara2006].
+We follow the taxonomic system of @Ebihara2006.
 
@@
 We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti [@Nitta2017].
-During this study, 17 sampling sites were established from \circa 200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt.
+During this study, 17 sampling sites were established from \circa{ }200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt.
 Rotui (899 m), Mt.
@@
 At each site, sporophytes were sampled in a single 10 m × 10 m plot, and gametophytes sampled in two to three 50 cm × 50 cm subplots located within the sporophyte plot.
-Epiphytes were sampled to a maximum height of \circa 2 m on their host plants.
+Epiphytes were sampled to a maximum height of \circa{ }2 m on their host plants.
 Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family [\eg, filamentous, ribbon with round gemmae; @Nayar1971].
 This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
-We characterized the elevational range of each species based on the results of the field survey; any species found with gametophytes occurring *\>*200 m above the maximum or below the minimum observed elevational range of sporophytes was considered 'widespread' [\ie, facultatively independent *sensu* @Rumsey1996].
-We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, \ie, below \circa 500 m), or high elevation epiphyte (mostly occurring in cloud forest, \ie, above \circa 500 m).
+We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, \ie, below \circa{ }500 m), or high elevation epiphyte (mostly occurring in cloud forest, \ie, above \circa{ }500 m).
 
@@
 
-We used microclimate data collected by [@Nitta2020].
+We used microclimate data collected by @Nitta2020.
 This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA).
 The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at ca.
-1.5 m on a tree, one 'terrestrial' datalogger at ca.
-10 cm next to the tree).
+1.5 m on a tree, one 'terrestrial' datalogger at \circa{ }10 cm next to the tree).
 During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
@@
 This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca.
-500 µmol·m^-2^ ·s^-1^.
+500 µmol·m^-2^·s^-1^.
 Response curves were fitted using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded [@Proctor2012].
@@
 
-We used general linear mixed models (GLMMs) to investigate the effect of growth habit and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at -86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the MCMCglmm R package (Hadfield, 2010).
+We used general linear mixed models (GLMMs) to investigate the effect of growth habit and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at -86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the MCMCglmm R package [@Hadfield2010].
 For each of the three response variables (DT, ETR~max~ , and PPFD~95%~), we used growth habit (high-elevation epiphyte, low-elevation epiphyte, terrestrial or saxicolous) and generation (sporophyte vs.
 gametophyte) and their interactions as fixed effects.
-We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix in the case the response variable had significant phylogenetic signal (Hadfield, 2010).
+We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix in the case the response variable had significant phylogenetic signal [@Hadfield2010].
 We used an inverse-Gamma distribution for all priors, and ran analyses for 500,000 iterations, with burnin after 1,000 iterations and thinning every 500 iterations.
@@
 <!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
-
-::: {#refs}
-:::
+<div id="refs"></div>

### 2020-11-10 ms/manuscript.Rmd
Update MS

@@
 title: ""
-bibliography: references.bib
+bibliography: 
+  - references.bib
+  - references_other.yaml
 mainfont: Roboto-Regular.ttf
@@
   list = c(
-    "climate"
+    "climate",
+    "filmy_species",
+    "moorea_sites",
+    "phylosig",
+    "range_types"
   )
@@
 
-<!-- See https://stackoverflow.com/questions/24672111/how-to-add-a-page-break-in-word-document-generated-by-rstudio-markdown/40474753
-for inserting page breaks when rendering to MS word -->
+\clearpage
 #####
-
-\newpage
 
@@
 
+\clearpage
 #####
-
-\newpage
 
@@
 
+```{r study-system-stats}
+# Calculate number of all filmy fern species, trichomanoids, and hymenophylloids
+# (not considering numbered taxa)
+
+n_filmy_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
+  n_distinct()
+
+n_trich_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
+  magrittr::extract(., str_detect(., "^Hymeno", negate = TRUE)) %>%
+  n_distinct()
+
+n_hymeno_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
+  magrittr::extract(., str_detect(., "^Hymeno", negate = FALSE)) %>%
+  n_distinct()
+
+```
+
 We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology (Fig. 2).
@@
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
-At the time of the study, a total of 18 filmy fern species were known from Moorea [@Murdock2003; @Ranker2005; @Nitta2011b; @Florence], including multiple terrestrial, saxicolous, and epiphytic species.
-The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by similar numbers of species.
+A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence].
+The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
 We follow the taxonomic system of @Ebihara2006.
@@
 We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti [@Nitta2017].
-During this study, 17 sampling sites were established from \circa{ }200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt.
+During this study, `r n_distinct(moorea_sites$site)` sampling sites were established from \circa{ }200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt.
 Rotui (899 m), Mt.
@@
 
-<!--
-
 ## Results
@@
 
-Twenty-two filmy fern taxa (including one new record and two species that were recognized as multiple varieties) were observed as sporophytes, and 18 as gametophytes on Moorea in total (Table 1).
-*Hymenophyllum braithwaitei* (Ebihara & K. Iwats.), only observed near the peak of Mt. Tohiea (1206 m), is a new record for Moorea and French Polynesia.
-It was previously known only from Vanuatu and New Caledonia (Ebihara et al., 2010).
+In total, `r n_distinct(range_types$species)` filmy fern taxa (including two species that were recognized as multiple varieties) were observed on Moorea.
 Three taxa with distinct *rbcL* sequences, morphology, and elevational ranges could be distinguished within the *Crepidomanes minutum* (Blume) K.
@@
 1 and 2.
+
 In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C.
@@
 
+Sporophytes of two taxa, *C.
+minutum* var.
+3 and *Hymenophyllum braithwaitei* (Ebihara & K.
+Iwats.), did not occur in the study plots, but each was observed from a single population on Mt.
+Tohiea, at ca.
+1000 m and near the summit (1206 m), respectively.
+We include these observations for comparison of ranges between sporophytes and gametophytes.
+
 ### Microclimate
@@
 Mean daily temperature was strongly correlated with elevation, and dropped ca.
-6 *◦*C from the lowest surveyed site (201 m) to the highest (1206 m; Fig. 3).
+6 °C from the lowest surveyed site (201 m) to the highest (1206 m; Fig. 3).
 Although epiphytic and terrestrial habitats were similar in temperature, epiphytic habitats tended to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones (Fig.
@@
 
-Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* and λ both near 1, *P \<*0.05; sporophytes *K* = 0.78, *P* = 0.02 and λ = 0.96, *P* = 0.11), but no signal for PPFD~95%~ or ETR~max~ (Table 2).
+```{r phy-sig-stats}
+sporo_recover_k <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(kval) %>% jntools::round_t(2)
+sporo_recover_k_pval <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(k_pval) %>% jntools::round_t(3)
+sporo_recover_lambda <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(lambda) %>% jntools::round_t(2)
+sporo_recover_lambda_pval <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(lambda_pval) %>% jntools::round_t(3)
+
+# Make sure that all ETRmax and PAR have pval > 0.05
+phylosig %>% 
+  filter(trait != "recover_mean") %>%
+  verify(all(k_pval > 0.05)) %>%
+  verify(all(lambda_pval > 0.05), success_fun = success_logical, error_fun = error_stop)
+```
+
+Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* and λ both near 1, *P \<*0.05; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval` and λ = `r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`), but no signal for PPFD~95%~ or ETR~max~ (Table 2).
+
+<!--
 

### 2020-11-20 ms/manuscript.Rmd
Add t-test for responses between generations

@@
     "phylosig",
-    "range_types"
+    "range_types",
+    "recovery_species_means"
   )
@@
 
-<!--
-
 ### Desiccation tolerance
 
-Ability to recover from desiccation was tested in sporophytes of 15 species and gametophytes of 14 species, including 12 species with both life phases (Appendix S3).
+```{r dt-stats}
+# Count number of species included in DT test by generation
+n_dt_species_gameto <- recovery_species_means %>% 
+  count(species, generation) %>% 
+  count(generation) %>% 
+  filter(generation == "gameto") %>% 
+  pull(n) %>% english2
+
+n_dt_species_gameto <- recovery_species_means %>% 
+  count(species, generation) %>% 
+  count(generation) %>% 
+  filter(generation == "sporo") %>% 
+  pull(n) %>% english2
+
+n_dt_species_both <- recovery_species_means %>% 
+  count(species, generation) %>% 
+  group_by(species) %>% 
+  summarize(n_generation = n_distinct(generation), .groups = "drop") %>%
+  filter(n_generation == 2) %>%
+  pull(species) %>%
+  n_distinct() %>%
+  english2
+
+# Vector of species that I know *after* running the t-test have significant
+# differences and discuss in the MS.
+species_with_signif_dt_diff <- c("Hymenophyllum_polyanthos", "Hymenophyllum_digitatum", "Crepidomanes_minutum2", "Polyphlebium_endlicherianum", "Polyphlebium_borbonicum")
+
+t_test_results %>% 
+  filter(response == "recover", p_value < 0.05) %>%
+  # Make sure that species with signif diff in DT mentioned
+  # in MS match those in the actual analysis results
+  verify(setequal(.$species, species_with_signif_dt_diff)) %>%
+  select(species, estimate, p_value) %>%
+  arrange(p_value)
+
+```
+
+Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_gameto` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
 A general correlation of DT with habitat was observed in sporophytes (Fig. 4).
@@
 Saxicolous species (*Crepidomanes kurzii* (Bedd.) Tagawa & K.
-Iwats., *Polyphlebium endlicherianum* (C. Presl) Ebihara & K.
-Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C. kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
+Iwats., *Polyphlebium endlicherianum* (C.
+Presl) Ebihara & K.
+Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C.
+kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
 All epiphytic species were capable of recovering from desiccation to a certain degree, but ability varied between species.
-Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel.
-was able to recover from desiccation longer than 2 d.
+Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
 Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.* *pallidum* and *P.* *borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
-Interestingly, *C. minutum* var.
-2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw.
-could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.
+Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.
 
@@
 minutum* var.
-2, *H.* *digitatum*, *H. polyanthos*, and *P. borbonicum* (Fig. 5; *t*-test, *P* = 0.0093, *\<*0.001, *\<*0.001, 0.012, respectively).
+2, *H.* *digitatum*, *H.
+polyanthos*, and *P.
+borbonicum* (Fig.
+5; *t*-test, *P* = 0.0093, *\<*0.001, *\<*0.001, 0.012, respectively).
+
+<!--
 

### 2020-11-20 R/functions.R

@@ -101,19 +101,24 @@ calculate_mean_recovery <- function (data) {
   
 }
 
-calculate_mean_etr <- function (data) {
+calculate_indiv_etr <- function (data) {
   
-  individual_etr <-
-    data %>%
+  # Calculate maximum ETR for each individual
+  data %>%
     assert(not_na, ETR) %>%
     group_by(species, generation, individual) %>%
     summarize(
-      etr_max = max(ETR)
-    ) %>%
-    ungroup
+      etr_max = max(ETR),
+      .groups = "drop"
+    )
   
-  # Next calculate the mean value by species and generation
-  individual_etr %>%
+}
+
+calculate_mean_etr <- function (data) {
+  
+  # Calculate mean ETR max by species and generation
+  data %>%
+    assert(not_na, etr_max) %>%
     group_by(generation, species) %>%
     summarize(
       etr_mean = mean(etr_max),
@@ -124,11 +129,10 @@ calculate_mean_etr <- function (data) {
   
 }
 
-calculate_mean_par <- function (data) {
+calculate_indiv_par <- function (data) {
   
-  # First calculate critical PAR value for each individual
-  individual_par <-
-    data %>%
+  # Calculate critical PAR value for each individual
+  data %>%
     select(species, generation, individual, PAR, ETR) %>%
     nest(data = c(PAR, ETR)) %>%
     mutate(
@@ -149,8 +153,12 @@ calculate_mean_par <- function (data) {
     # calculate critical PAR (PAR where reach 95% of max ETR)
     mutate(par_critical = -log(0.05)/estimate)
   
+}
+
+calculate_mean_par <- function (data) {
+  
   # Next calculate the mean value by species and generation
-  individual_par %>%
+  data %>%
     group_by(generation, species) %>%
     summarize(
       par_mean = mean(par_critical),
@@ -183,6 +191,101 @@ combine_mean_phys_traits <- function (recovery_species_means, etr_species_means,
   
 }
 
+# t-test ----
+
+#' Run a t-test comparing response values between gametophytes and sporophytes.
+#' 
+#' Helper function for run_t_test()
+#'
+#' @param data Data with response variable (DT recovery, ETRmax or PAR95) for
+#' filmy fern sporophytes and gametophytes in wide format. Three columns: `species`,
+#' `gameto`, and `sporo`. `gameto` and `sporo` are each list-columns, where each
+#' element of the list is a character vector of response variables.
[…]

### 2020-11-20 R/plan.R

@@ -16,7 +16,7 @@ plan <- drake_plan(
     out2 = file_out("data/nitta_2017/treepl_Moorea_Tahiti.tre"),
     out3 = file_out("data/nitta_2017/all_plots.csv")),
 
-  # -microclimate data
+  # - microclimate data
   climate = readr::read_csv("data_raw/doi_10.5061_dryad.fqz612jps__v2/moorea_climate.csv"),
 
   # - site data
@@ -31,13 +31,13 @@ plan <- drake_plan(
 
   filmy_species = traits$species,
   
-  # -phylogenetic tree
+  # - phylogenetic tree
   filmy_phy = ape::read.tree(
     file_in("data/nitta_2017/treepl_Moorea_Tahiti.tre"
     )) %>%
     ape::keep.tip(filmy_species),
 
-  # -physiological data
+  # - physiological data
   recovery_data_raw = readr::read_csv("data/filmy_DT_data.csv"),
 
   # - community data
@@ -60,23 +60,37 @@ plan <- drake_plan(
   recovery_data = clean_recovery(recovery_data_raw),
 
   light_data = clean_light(light_data_raw),
+  
+  # - calculate PAR95 by individual
+  par_indiv = calculate_indiv_par(light_data),
+  
+  # - calculate ETRmax by individual
+  etr_indiv = calculate_indiv_etr(light_data),
 
   # Calculate means by species and generation ----
-  # -% recovery
+  # - DT recovery
   recovery_species_means = calculate_mean_recovery(recovery_data),
 
-  # -ETR max
-  etr_species_means = calculate_mean_etr(light_data),
+  # - ETR max
+  etr_species_means = calculate_mean_etr(etr_indiv),
 
-  # -PAR 95%
-  par_species_means = calculate_mean_par(light_data),
+  # - PAR95
+  par_species_means = calculate_mean_par(par_indiv),
 
-  # -Combine the means into a single df
+  # - Combine the means into a single df
   combined_species_means = combine_mean_phys_traits(
     recovery_species_means = recovery_species_means,
     etr_species_means = etr_species_means,
     par_species_means = par_species_means
   ),
+  
+  # t-test ----
+  
+  # Perform two-sided t-test on DT and light responses across generations
+  t_test_results = run_t_test(
+    recovery_data = recovery_data, 
+    par_indiv = par_indiv, 
+    etr_indiv = etr_indiv),
 
   # Phylogenetic signal ----
   phylosig = analyze_phylosig_by_generation(

### 2020-11-20 ms/manuscript.Rmd
Update MS
add function to print p-value from t-test results

@@
     "range_types",
-    "recovery_species_means"
+    "recovery_species_means",
+    "t_test_results"
   )
@@
 
-# Vector of species that I know *after* running the t-test have significant
-# differences and discuss in the MS.
-species_with_signif_dt_diff <- c("Hymenophyllum_polyanthos", "Hymenophyllum_digitatum", "Crepidomanes_minutum2", "Polyphlebium_endlicherianum", "Polyphlebium_borbonicum")
-
-t_test_results %>% 
-  filter(response == "recover", p_value < 0.05) %>%
-  # Make sure that species with signif diff in DT mentioned
-  # in MS match those in the actual analysis results
-  verify(setequal(.$species, species_with_signif_dt_diff)) %>%
-  select(species, estimate, p_value) %>%
-  arrange(p_value)
-
+# Make helper function to check on p-value significance and extract from t_test_results
+#
+#' @param species_select Name of species to select from t-test results
+#' @param response_select Name of response to select from t-test results
+#' @param data t-test results
+#' @param p_check Boolean: should the p-value be required to be less than 0.05?
+#'
+#' @return A p-value for the selected species and response
+ttest_pval <- function(species_select, response_select, data = t_test_results, p_check = TRUE) {
+  data %>% 
+    # Filter to appropriate species and response
+    filter(species == species_select, response == response_select) %>%
+    # Verify the p-value is significant (or override with `p_check = FALSE`)
+    assert(function (x) magrittr::is_less_than(x, 0.05) | !p_check, p_value) %>%
+    # Format the p-value
+    mutate(
+      p_value = case_when(
+        p_value < 0.001 ~ "<0.001",
+        TRUE ~ jntools::round_t(p_value, digits = 3))
+    ) %>% 
+    pull(p_value)
+}
 ```
@@
 endlicherianum* (Fig.
-5; *t*-test, *P* = 0.012), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C.
+5; *t*-test, *P* = `r ttest_pval("Polyphlebium_endlicherianum", "recover")`), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C.
 minutum* var.
@@
 borbonicum* (Fig.
-5; *t*-test, *P* = 0.0093, *\<*0.001, *\<*0.001, 0.012, respectively).
-
-<!--
+5; *t*-test, *P* = `r ttest_pval("Crepidomanes_minutum2", "recover")`, `r ttest_pval("Hymenophyllum_digitatum", "recover")`, `r ttest_pval("Hymenophyllum_polyanthos", "recover")`, and `r ttest_pval("Polyphlebium_borbonicum", "recover")`, respectively).
 
@@
 In contrast, range type (gametophytes widespread beyond sporophytes or not) has no significant effect on DT (Table 3).
+
+<!--
 

### 2020-11-20 R/plan.R
Fix location of zipped data folders

@@ -2,7 +2,7 @@
 
 plan <- drake_plan(
 
-  # Load raw data ----
+  # Unzip raw data ----
   
   # Unzip data from Nitta et al. 2017 Ecol. Monographs from Dryad
   # The dataset must be downloaded first by going to
@@ -16,8 +16,21 @@ plan <- drake_plan(
     out2 = file_out("data/nitta_2017/treepl_Moorea_Tahiti.tre"),
     out3 = file_out("data/nitta_2017/all_plots.csv")),
 
+  # Unzip data from Nitta et al. 2020 New Phyt. from Dryad
+  # The dataset must be downloaded first by going to
+  # https://datadryad.org/stash/dataset/doi:10.5061/dryad.fqz612jps, clicking on
+  # "Download dataset", and saving to the "data" folder in this project.
+  climate_unzipped = unzip(
+    zipfile = file_in("data/doi_10.5061_dryad.fqz612jps__v4.zip"), 
+    files = "moorea_climate.csv", 
+    exdir = file_out("data/nitta_2020/"), 
+    junkpaths = TRUE, 
+    overwrite = TRUE),
+  
+  # Load raw data ----
+  
   # - microclimate data
-  climate = readr::read_csv("data_raw/doi_10.5061_dryad.fqz612jps__v2/moorea_climate.csv"),
+  climate = readr::read_csv(file_in("data/nitta_2020/moorea_climate.csv")),
 
   # - site data
   moorea_sites = readr::read_csv(
@@ -25,7 +38,7 @@ plan <- drake_plan(
     col_types = "cnnn") %>%
     filter(str_detect(site, "Aorai", negate = TRUE)),
 
-  # -traits
+  # - traits
   traits = readr::read_csv("data/filmy_trait_data.csv") %>%
     rename(species = genus_species),
 

### 2020-11-20 R/packages.R
load jntools package

@@ -7,6 +7,7 @@ library(ape)
 library(assertr)
 library(assertthat)
 library(glue)
+library(jntools)
 
 conflict_prefer("filter", "dplyr")
 conflict_prefer("rename", "dplyr")

### 2020-11-20 ms/manuscript.Rmd
Add vapor pressure decifit (vpd) to PGLS analysis

@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
-library(jntools)
 ```
@@
     "range_types",
-    "recovery_species_means",
+    "combined_species_means",
     "t_test_results"
@@
 # Count number of species included in DT test by generation
-n_dt_species_gameto <- recovery_species_means %>% 
+n_dt_species_sporo <- combined_species_means %>% 
+  filter(!is.na(recover_mean)) %>% 
+  count(species, generation) %>% 
+  count(generation) %>% 
+  filter(generation == "sporo") %>% 
+  pull(n) %>% 
+  english2
+
+n_dt_species_gameto <- combined_species_means %>% 
+  filter(!is.na(recover_mean)) %>%
   count(species, generation) %>% 
@@
   filter(generation == "gameto") %>% 
-  pull(n) %>% english2
+  pull(n) %>% 
+  english2
 
-n_dt_species_gameto <- recovery_species_means %>% 
-  count(species, generation) %>% 
-  count(generation) %>% 
-  filter(generation == "sporo") %>% 
-  pull(n) %>% english2
-
-n_dt_species_both <- recovery_species_means %>% 
+n_dt_species_both <- combined_species_means %>% 
+  filter(!is.na(recover_mean)) %>% 
   count(species, generation) %>% 
@@
 
-Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_gameto` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
+Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
 A general correlation of DT with habitat was observed in sporophytes (Fig. 4).
@@
 
-<!--
-
 ### Photosynthetic optima
 
-Light responses were measured in sporophytes of 18 species and gametophytes of 11 species, including nine species with both life phases (Appendix S4).
+```{r photosyn-stats}
+# Count number of species included in light data by generation
+n_light_species_sporo <- combined_species_means %>% 
+  filter(!is.na(etr_mean), !is.na(par_mean)) %>%
+  count(species, generation) %>% 
+  count(generation) %>% 
+  filter(generation == "sporo") %>% 
+  pull(n) %>% 
+  english2
+
+n_light_species_gameto <- combined_species_means %>% 
+  filter(!is.na(etr_mean), !is.na(par_mean)) %>%
+  count(species, generation) %>% 
+  count(generation) %>% 
+  filter(generation == "gameto") %>% 
+  pull(n) %>% 
+  english2
+
+n_light_species_both <- combined_species_means %>% 
+  filter(!is.na(etr_mean), !is.na(par_mean)) %>% 
+  count(species, generation) %>% 
+  group_by(species) %>% 
+  summarize(n_generation = n_distinct(generation), .groups = "drop") %>%
+  filter(n_generation == 2) %>%
+  pull(species) %>%
+  n_distinct() %>%
+  english2
+
+# Calculate grand means by generation for light responses
+# (so mean is the mean of species means, and sd is sd of species means)
+light_overall_means <-
+combined_species_means %>%
+  group_by(generation) %>%
+  summarize(
+    etr = mean(etr_mean, na.rm = TRUE),
+    etr_sd = sd(etr_mean, na.rm = TRUE),
+    par = mean(par_mean, na.rm = TRUE),
+    par_sd = sd(par_mean, na.rm = TRUE),
+    .groups = "drop"
+  ) %>%
+  mutate(across(where(is.numeric), ~round_t(.x, 1))) %>%
+  mutate(
+    etr_pretty = glue("{etr} ± {etr_sd}"),
+    par_pretty = glue("{par} ± {par_sd}"),
+  )
+
+# Extract values by response and generation for citing in MS
+mean_par_sporo <- light_overall_means %>% filter(generation == "sporo") %>% pull(par_pretty)
+mean_par_gameto <- light_overall_means %>% filter(generation == "gameto") %>% pull(par_pretty)
+mean_etr_sporo <- light_overall_means %>% filter(generation == "sporo") %>% pull(etr_pretty)
+mean_etr_gameto <- light_overall_means %>% filter(generation == "gameto") %>% pull(etr_pretty)
+```
+
+Light responses were measured in sporophytes of `r n_light_species_sporo` species and gametophytes of `r n_light_species_gameto` species, including `r n_light_species_both` species with both life phases (Appendix S4).
 Unlike DT, photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes are adapted for higher light levels, whereas gametophytes are adapted for lower light levels (Fig.
@@
 Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
-142 ± 46 µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at 51 ± 22 µmol·m^--2^ ·s^--1^.
-Similarly, sporophytes attained higher rates of electron transport (ETR~max~; 8.1 ± 2.7 µmol·m^--2^ ·s^--1^) than gametophytes (3.9 ± 0.9 µmol·m^--2^ ·s^--1^).
+`r mean_par_sporo` µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at `r mean_par_gameto` µmol·m^--2^ ·s^--1^.
+Similarly, sporophytes attained higher rates of electron transport (ETR~max~; `r mean_etr_sporo` µmol·m^--2^ ·s^--1^) than gametophytes (`r mean_etr_gameto` µmol·m^--2^ ·s^--1^).
 The GLMMs showed a significant effect of generation, but neither range type nor growth habit, on PPFD~95%~ and ETR~max~ (Table 3).
 
-### Environmental water potential and DT
+### Relationships of water potential and range size with DT
 
@@
 These trends were significant when analyzed using PGLS for sporophytes (*P* = 0.036) but not gametophytes (*P* = 0.123; Appendix S5).
+
+No significant relationship was detected between range size of gametophytes and DT.
+
+<!--
 

### 2020-11-20 R/functions.R

@@ -513,37 +513,47 @@ summary_to_csv <- function (model) {
 }
 
 #' Run phylogenetic generalized least squares (PGLS)
-#' for gametophyte range size vs. desiccation tolerance
+#' for range size and VPD vs. desiccation tolerance
 #' in filmy ferns from Moorea
 #'
 #' @param range_data Dataframe with range size data
 #' @param combined_species_means Dataframe with desiccation tolerance data
 #' @param phy Phylogeny
 #'
-#' @return model object
+#' @return list of model objects:
+#' - sporo_vpd_model:
+#' - gameto_vpd_model:
+#' - sporo_range_model:
+#' - gameto_range_model:
 #' 
 run_pgls_range <- function (range_data, combined_species_means, phy) {
   
-  range_data <-
-    range_data %>% 
-    mutate(range_breadth = gameto_max_range - gameto_min_range) %>%
-    select(species, range_breadth) %>%
-    filter(!is.na(range_breadth))
-  
-  dt_range_data <-
+  # Make a dataframe combining vpd and recovery
+  range_recover_data <-
     combined_species_means %>%
-    filter(generation == "gameto") %>%
     filter(!is.na(recover_mean)) %>%
-    select(species, recover = recover_mean) %>%
-    left_join(range_data, by = "species") %>%
-    remove_missing() %>%
+    select(species, generation, recover = recover_mean) %>%
+    pivot_wider(names_from = "generation", values_from = "recover", names_prefix = "recover_") %>%
+    left_join(
+      select(range_data, species, 
+            sporo_mean_vpd, 
+            gameto_mean_vpd,
+            sporo_range_breadth,
+            gameto_range_breadth
+            ), 
+      by = "species") %>%
     as.data.frame()
   
-  dt_range_phy <- ape::keep.tip(phy, dt_range_data$species)
+  # Create a comparative.data object for caper combining the vpd and recovery data with the phylogeny
+  range_recover_data_comp <- caper::comparative.data(phy, range_recover_data, species, na.omit = FALSE)
   
-  dt_range_data_comp <- caper::comparative.data(dt_range_phy, dt_range_data, species)
-  
-  caper::pgls(recover ~ range_breadth, dt_range_data_comp)
+  # Run PGLS for each combination of generation x vpd or range breadth vs. recovery
+  res <- list(
[…]

### 2020-11-20 R/plan.R

@@ -61,7 +61,8 @@ plan <- drake_plan(
     community_matrix_raw = community_matrix_raw,
     filmy_species = filmy_species,
     phy = filmy_phy,
-    moorea_sites = moorea_sites
+    moorea_sites = moorea_sites,
+    climate = climate
   ),
 
   # FIXME: check number of measurements per individual; shouldn't have more than 10 each.

### 2020-11-24 ms/manuscript.Rmd
Update pgls analysis

@@
 
-Desiccation tolerance tended to increase with decreasing environmental water potential for both gametophytes and sporophytes (Fig. 6).
+Desiccation tolerance tended to increase with VPD for both gametophytes and sporophytes (Fig. 6).
 These trends were significant when analyzed using PGLS for sporophytes (*P* = 0.036) but not gametophytes (*P* = 0.123; Appendix S5).
@@
 
-<!--
-
 ## Discussion
@@
 Vegetative DT was a key innovation linked to the rise of early land plants, but is largely limited to non-vascular groups such as bryophytes and algae in modern plants.
-Filmy ferns have received considerable attention over the years for their remarkable reversion to vegetative DT in the sporophyte phase (e.g., Shreve, 1911; Proctor, 2003; Cea et al., 2014).
+Filmy ferns have received considerable attention over the years for their remarkable reversion to vegetative DT in the sporophyte phase [\eg, @Shreve1911; @Proctor2003; @GarcesCea2014].
 Yet, no previous study to our knowledge has evaluated parallel physiology of sporophytes and gametophytes of filmy ferns in relation to DT.
@@
 In contrast, habitats with minimum water potential below --7 MPa are nearly all epiphytic, and the filmy fern sporophytes occurring there are all capable of DT.
+
+<!--
 

### 2020-11-24 R/functions.R

@@ -191,24 +191,6 @@ combine_mean_phys_traits <- function (recovery_species_means, etr_species_means,
   
 }
 
-#' Calculate mean VPD by site x growth habit
-#'
-#' @param climate Climate data for sites on Moorea;
-#' relative humidity, temperature, and vapor pressure
-#' deficit (VPD) calculated from these once every 15 minutes
-#'
-#' @return Mean VPD by site and growth habit (epiphytic or terrestrial)
-#' 
-calculate_mean_vpd <- function(climate) {
-  climate %>%
-  assert(not_na, vpd) %>%
-  group_by(site, habit) %>%
-  summarize(
-    vpd = mean(vpd),
-    .groups = "drop"
-  )
-}
-
 #' Calculate mean VPD for filmy fern gametophytes
 #'
 #' Treats specimens collected from rock walls ("epipetric") as epiphytic
@@ -304,6 +286,34 @@ calculate_mean_vpd_sporo <- function (community_matrix_raw, filmy_species, moore
       sporo_vpd = mean(vpd),
       .groups = "drop"
     )
+}
+
+
+#' Make a dataframe combining range size, vpd, and recovery
+#' 
+#' For preparing a dataframe for analysis with caper::pgls()
+#'
+#' @param combined_species_means Dataframe with desiccation tolerance data
+#' @param env_range_data Dataframe with environmental (VPD) and range size data
+#'
+#' @return Datafrmae
+#'
+combine_env_env_range_recover <- function (combined_species_means, env_range_data) {
+
+  combined_species_means %>%
+  filter(!is.na(recover_mean)) %>%
+  select(species, generation, recover = recover_mean) %>%
+  pivot_wider(names_from = "generation", values_from = "recover", names_prefix = "recover_") %>%
+  left_join(
+    select(env_range_data, 
+           species, 
+           sporo_vpd, 
+           gameto_vpd,
+           sporo_range_breadth,
+           gameto_range_breadth
+    ), 
+    by = "species") %>%
+  as.data.frame()
 }
 
 # t-test ----
@@ -631,8 +641,8 @@ summary_to_csv <- function (model) {
 #' for range size and VPD vs. desiccation tolerance
 #' in filmy ferns from Moorea
 #'
-#' @param range_data Dataframe with range size data
-#' @param combined_species_means Dataframe with desiccation tolerance data
+#' @param env_range_recover_data Dataframe with range size, environmental (VPD), 
+#' and desiccation tolerance data
[…]

### 2020-11-24 R/plan.R

@@ -97,10 +97,12 @@ plan <- drake_plan(
   ),
   
   # - combine environmental and range data
-  env_range_data = left_join(
-    mean_vpd_gameto,
-    mean_vpd_sporo,
-    range_types,
+  env_range_data = purrr::reduce(
+    list(
+      mean_vpd_gameto,
+      mean_vpd_sporo,
+      range_types),
+    left_join,
     by = "species"
   ),
   
@@ -110,7 +112,7 @@ plan <- drake_plan(
   # - calculate ETRmax by individual
   etr_indiv = calculate_indiv_etr(light_data),
 
-  # Calculate means by species and generation ----
+  # Means by species and generation ----
   # - DT recovery
   recovery_species_means = calculate_mean_recovery(recovery_data),
 
@@ -142,7 +144,8 @@ plan <- drake_plan(
     traits_select = c("recover_mean", "etr_mean", "par_mean")
   ),
   
-  # Generalized linear mixed models ----
+  # GLMMS ----
+  # (Generalized Linear Mixed Models)
   
   # Uses phylogeny for DT only
   glmms = run_glmm(
@@ -150,12 +153,17 @@ plan <- drake_plan(
     traits = traits,
     phy = filmy_phy),
   
-  # Phylogenetic generalized least squares (PGLS) ----
+  # PGLS ----
+  # (Phylogenetic Generalized Least Squares)
   
   # Run PGLS for gametophyte range size vs. desiccation tolerance
-  range_dt_model = run_pgls_range(
-    range_data = range_types, 
+  env_range_recover_data = combine_env_env_range_recover(
     combined_species_means = combined_species_means, 
+    env_range_data = env_range_data
+  ),
+  
+  range_dt_model = run_pgls(
+    env_range_recover_data = env_range_recover_data, 
     phy = filmy_phy),
   
   # Render manuscript ----

### 2020-12-18 R/functions.R
Add functions to calculate VPD, run PGLS

@@ -442,6 +442,21 @@ combine_mean_phys_traits <- function (recovery_species_means, etr_species_means,
   
 }
 
+
+#' Calculate mean VPD from climate data
+#'
+#' @param climate Dataframe with climate data including
+#' relative humidity, temperature, and VPD calculated from these measured
+#' once every 15 minutes by site and growth habit (terrestrial or epiphytic)
+#'
+#' @return Tibble with mean VPD by site and growth habit 
+calculate_mean_vpd <- function (climate) {
+  climate %>% 
+    group_by(site, habit) %>% 
+    summarize(vpd = mean(vpd)) %>% 
+    assert(not_na, vpd)
+}
+
 #' Calculate mean VPD for filmy fern gametophytes
 #'
 #' Treats specimens collected from rock walls ("epipetric") as epiphytic
@@ -917,6 +932,39 @@ run_pgls <- function (env_range_recover_data, phy) {
     gameto_range_model = caper::pgls(recover_gameto ~ gameto_range_breadth, env_range_recover_data_comp)
     )
   
+}
+
+#' Extract fitted values from a PGLS model
+#'
+#' @param model The PGLS model (list of class "pgls")
+#'
+#' @return Tibble, with two columns: species and the predicted value
+#' 
+predict_pgls <- function(model) {
+  model$fitted %>%
+    as.data.frame() %>%
+    rownames_to_column("species") %>%
+    rename(predicted = V1) %>%
+    as_tibble()
+}
+
+#' Summarize statistics of a PGLS model
+#'
+#' @param model The PGLS model (list of class "pgls")
+#'
+#' @return Tibble, with various columns for model statistics (r_squared, p-value, etc)
+#' 
+tidy_pgls <- function (model) {
+  model_sum <- summary(model)
+  tibble(
+    sigma = model_sum$sigma,
+    df = model_sum$df %>% paste(collapse = ","),
+    r_squared = model_sum$r.squared,
+    f_value = model_sum[["fstatistic"]][["value"]],
+    adj_r_squared = model_sum$adj.r.squared[,1],
+    # Need to run anova() to get p-value
+    p_value = anova(model) %>% filter(!is.na(`F value`)) %>% pull(`Pr(>F)`)
+  )
 }
 
 # Etc ----

### 2020-12-18 R/plan.R

@@ -172,9 +172,11 @@ plan <- drake_plan(
     env_range_data = env_range_data
   ),
   
-  range_dt_model = run_pgls(
+  env_range_dt_model = run_pgls(
     env_range_recover_data = env_range_recover_data, 
     phy = filmy_phy),
+  
+  env_range_dt_model_summary = map_df(env_range_dt_model, tidy_pgls, .id = "model"),
   
   # Render manuscript ----
   

### 2020-12-18 ms/manuscript.Rmd
Update MS

@@
     "combined_species_means",
-    "t_test_results"
+    "t_test_results",
+    "env_range_dt_model_summary"
   )
@@
 Filmy ferns (Hymenophyllaceae) represent a particularly interesting case for studying the ecological and evolutionary significance of vegetative DT across life stages in ferns.
-Filmy ferns are a large family (ca.
-430 spp.; PPG I, 2016) of primarily tropical, leptosporangiate ferns named for their remarkably thin leaf laminae, which are usually only a single cell layer in thickness and lack cuticle or stomata (Fig.
+Filmy ferns are a large family (\circa{ }430 spp.; PPG I, 2016) of primarily tropical, leptosporangiate ferns named for their remarkably thin leaf laminae, which are usually only a single cell layer in thickness and lack cuticle or stomata (Fig.
 1A--D).
@@
 This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA).
-The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at ca.
-1.5 m on a tree, one 'terrestrial' datalogger at \circa{ }10 cm next to the tree).
+The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at \circa{ }1.5 m on a tree, one 'terrestrial' datalogger at \circa{ }10 cm next to the tree).
 During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
@@
 Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
-For sporophytes, 8--16 individuals (one individual = single whole frond including a ca.
-2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
+For sporophytes, 8--16 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
 We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
@@
 Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson.
-All of these have sporophytes that are normally only found in cloud forest (ca. 500--1200 m), but gametophytes that are distributed to lower elevations.
+All of these have sporophytes that are normally only found in cloud forest (\circa{ }500--1200 m), but gametophytes that are distributed to lower elevations.
 
@@
 Iwats.), did not occur in the study plots, but each was observed from a single population on Mt.
-Tohiea, at ca.
-1000 m and near the summit (1206 m), respectively.
+Tohiea, at \circa{ }1000 m and near the summit (1206 m), respectively.
 We include these observations for comparison of ranges between sporophytes and gametophytes.
@@
 
-Gametophytes of low elevation epiphytes had the highest DT (ca.
-80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from ca.
-25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; Fig.
+Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; Fig.
 5).
@@
 
+```{r pgls-stats}
+sporo_vpd_pval <- env_range_dt_model_summary %>% filter(model == "sporo_vpd_model") %>% pull(p_value) %>%
+  round_t(2)
+gameto_vpd_pval <- env_range_dt_model_summary %>% filter(model == "gameto_vpd_model") %>% pull(p_value) %>%
+  round_t(2)
+```
+
 Desiccation tolerance tended to increase with VPD for both gametophytes and sporophytes (Fig. 6).
-These trends were significant when analyzed using PGLS for sporophytes (*P* = 0.036) but not gametophytes (*P* = 0.123; Appendix S5).
+These trends were significant when analyzed using PGLS for gametophytes (*P* = `r gameto_vpd_pval`) but not sporophytes (*P* = `r sporo_vpd_pval`; Appendix S5).
 
@@
 
-We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below --200 MPa in some epiphytes.
-Levels of similarly extreme DT have also been reported in other species of *Hymenophyllum* (Proctor, 2003, 2012) and a general correlation of habitat with DT in filmy ferns has been previously established (Shreve, 1911; Proctor, 2003, 2012; Parra et al., 2009; Saldan˜a et al., 2014; Testo et al., unpublished), but no other studies to our knowledge have documented a similar range of DT values in filmy ferns from a single site.
-One may wonder then, what is the adaptive significance of being able to withstand drying at --200 MPa vs.
---100 MPa when the ambient water potential is typically *\>*--5 MPa and only rarely, if ever, drops below −50 MPa.
-The recent finding of a constitutive mechanism for DT in filmy ferns (Cea et al., 2014) supports the need for a relatively wide 'safety margin,' and similar (or even more extreme) levels of constitutive DT are well known in bryophytes (Oliver et al., 1993).
-The conservative behavior of fern stomata, which lack a response to the plant stress hormone ABA, is well documented (Brodribb and McAdam, 2011; McAdam and Brodribb, 2012, 2013).
+We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below -200 MPa in some epiphytes.
+Levels of similarly extreme DT have also been reported in other species of *Hymenophyllum* [@Proctor2003; @Proctor2012] and a general correlation of habitat with DT in filmy ferns has been previously established [@Shreve1911; @Proctor2003; @Proctor2012; @Parra2009; @Saldana2013], but no other studies to our knowledge have documented a similar range of DT values in filmy ferns from a single site.
+One may wonder then, what is the adaptive significance of being able to withstand drying at -200 MPa vs.
+-100 MPa when the ambient water potential is typically > -5 MPa and only rarely, if ever, drops below -50 MPa.
+The recent finding of a constitutive mechanism for DT in filmy ferns [@GarcesCea2014] supports the need for a relatively wide "safety margin," and similar (or even more extreme) levels of constitutive DT are well known in bryophytes [@Oliver1993].
+The conservative behavior of fern stomata, which lack a response to the plant stress hormone ABA, is well documented [@Brodribb2011; @Mcadam2012; @Mcadam2013].
 It is possible that filmy ferns also behave conservatively with respect to water-stress.
@@
 polyanthos*) seem to be able to better withstand intense vs.
-moderate desiccation, with higher recovery values at --282 and --86 MPa vs.
---38 MPa when dry for 15 d.
+moderate desiccation, with higher recovery values at -282 and -86 MPa vs.
+-38 MPa when dry for 15 d.
 This may be an adaptation to cloud forest conditions; although cloud forests are frequently moist due to constant fog, dry periods are intense during the times when there is no cloud cover.
@@
 Comparison of DT with the minimum water potential envelope of each species (Fig.
-6) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with minimum water potential exceeding ca.
---7 MPa.
-In contrast, habitats with minimum water potential below --7 MPa are nearly all epiphytic, and the filmy fern sporophytes occurring there are all capable of DT.
-
-<!--
+6) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with minimum water potential exceeding \circa{ }-7 MPa.
+In contrast, habitats with minimum water potential below -7 MPa are nearly all epiphytic, and the filmy fern sporophytes occurring there are all capable of DT.
 
@@
 Contrary to our hypotheses, gametophytes of filmy ferns are in general not more tolerant of the abiotic environment (light levels and water potential) than sporophytes.
-This differs from other fern groups with desiccation tolerant gametophytes but desiccation sensitive sporophytes (Watkins et al., 2007).
+This differs from other fern groups with desiccation tolerant gametophytes but desiccation sensitive sporophytes [@WatkinsJr.2007c].
 In the filmy ferns we studied, DT levels of both sporophytes and gametophytes largely reflected habitat; in general, epiphytic species tended to have higher DT than terrestrial or saxicolous species, regardless of life stage.
-Interestingly, this echoes much earlier field observations (not experimental data) by Holloway (1930) on filmy fern gametophytes in New Zealand: he reported apparent DT in gametophytes of *Hymenophyllum rarum* R.
+Interestingly, this echoes much earlier field observations (not experimental data) by @Holloway1930 on filmy fern gametophytes in New Zealand: he reported apparent DT in gametophytes of *Hymenophyllum rarum* R.
 Br.
-and *H. villosum* Colenso, both high elevation epiphytes with desiccation tolerant sporophytes, and lack of DT in terrestrial filamentous gametophytes of *Polyphlebium colensoi* (Hook.) Ebihara & K.
-Iwats., *Abrodictyum strictum* (Menzies ex Hook. & Grev.) Ebihara & K.
-Iwats., and *A. elongatum* (A. Cunn.) Ebihara & K.
+and *H.
+villosum* Colenso, both high elevation epiphytes with desiccation tolerant sporophytes, and lack of DT in terrestrial filamentous gametophytes of *Polyphlebium colensoi* (Hook.) Ebihara & K.
+Iwats., *Abrodictyum strictum* (Menzies ex Hook.
+& Grev.) Ebihara & K.
+Iwats., and *A.
+elongatum* (A.
+Cunn.) Ebihara & K.
 Iwats.
 occupying moist, protected sites on the forest floor.
-Watkins et al. (2007) also found that DT of fern gametophytes tended to match habitat, with terrestrial species having lower DT than epiphytes.
+@WatkinsJr.2007c also found that DT of fern gametophytes tended to match habitat, with terrestrial species having lower DT than epiphytes.
 However, we found that within high elevation epiphytes, gametophytes had lower DT than sporophytes, and only one species (saxicolous *P.
@@
 5).
-At high elevation sites (i.e., cloud forest), epiphytic ferns often occur in a dense matrix of bryophytes (Fig. 1B).
+At high elevation sites (\ie, cloud forest), epiphytic ferns often occur in a dense matrix of bryophytes (Fig. 1B).
 It is possible that this bryophyte cover helps to retain moisture during brief periods of drought and thus protect the epiphytic fern gametophytes that grow amongst them.
@@
 This is perhaps best illustrated by *C. apiifolia* (Fig. 1D, E, H).
-Sporophytes of this species are usually restricted to moist cloud forest sites from ca.
-600--1200 m, but we observed gametophytes over a much wider range, down to 200 m.
+Sporophytes of this species are usually restricted to moist cloud forest sites from \circa{ }600 m to 1200 m, but we observed gametophytes over a much wider range, down to 200 m.
 A similar distribution pattern has been reported in congeneric *Callistopteris baldwinii* (D. C. Eaton) Copel.
-from Hawaii (Dassler and Farrar, 1997).
+from Hawaii [@Dassler1997].
 Since gametophytes of *C.
-apiifolia* are common on Moorea and have a unique morphology (Nitta et al., 2016), we were able to collect enough material of this species for additional tests of DT at higher humidity levels (48 h at 75% and 100% RH; results not shown).
+apiifolia* are common on Moorea and have a unique morphology [@Nitta2017], we were able to collect enough material of this species for additional tests of DT at higher humidity levels (48 h at 75% and 100% RH; results not shown).
 However, it still failed to recover from even these gentle treatments.
@@
 In the case that their microhabitats do become too dry, they may suffer loss at the edge of the population, then recover by clonal growth (Farrar et al., 2008).
-Furthermore, the complex surface of clonal mats has been shown to increase the boundary layer and decrease evaporation rates in mosses (Rice et al., 2001), and may function similarly in *C.
-apiifolia*.
+Furthermore, the complex surface of clonal mats has been shown to increase the boundary layer and decrease evaporation rates in mosses [@Rice2001], and may function similarly in *C. apiifolia*.
 Other filmy fern species also produce gemmae, but they are much more frequent in *C.
@@
 apiifolia* are unable to tolerate desiccation, but slower growing epiphytes such as *Hymenophyllum* spp.
-can better tolerate it (Oliver et al., 2000).
+can better tolerate it [@Oliver2000].
 Growth rates of tolerant and non-tolerant gametophytes including quantification of gemmae production should be investigated in these species to test this hypothesis.
@@
 Rather, we observed a clear intergenerational difference, with gametophytes consistently adapted for lower light levels than sporophytes (Fig. 5, Table 3).
-In the only other study to our knowledge that compared photosynthetic rates between gametophytes and sporophytes of a filmy fern, Johnson et al.
-(2000) also found that gametophytes of *V.
-speciosa* reached maximum ETR at lower levels than sporophytes: in gametophytes of this species, photosynthesis became saturated at a PPFD of 4--5 µmol·m^--2^ ·s^--1^ in the field or 30--50 µmol·m^--2^·s^--1^ in the lab vs.
-30--50 µmol·m^--2^ ·s^--1^ for the sporophyte (lab measurement only).
-Although we found similar differences in direction between gametophytes and sporophytes, the measurements of Johnson et al.
-(2000) are considerably lower than what we observed in the field for both Moorean filmy fern gametophytes (PPFD~95%~ ca.
-40--60 µmol·m^--2^ ·s^--1^) and sporophytes (PPFD~95%~ ca.
-130--160 µmol·m^--2^ ·s^--1^).
+In the only other study to our knowledge that compared photosynthetic rates between gametophytes and sporophytes of a filmy fern, @Johnson2000 also found that gametophytes of *V.
+speciosa* reached maximum ETR at lower levels than sporophytes: in gametophytes of this species, photosynthesis became saturated at a PPFD of 4--5 µmol·m^-2^·s^-1^ in the field or 30--50 µmol·m^-2^·s^-1^ in the lab vs.
+30--50 µmol·m^-2^ ·s^-1^ for sporophytes (lab measurement only).
+Although we found similar differences in direction between gametophytes and sporophytes, the measurements of @Johnson2000 are considerably lower than what we observed in the field for both Moorean filmy fern gametophytes (PPFD~95%~ \circa{ }40--60 µmol·m^-2^·s^-1^) and sporophytes (PPFD~95%~ \circa{ }130--160 µmol·m^-2^·s^-1^).
 This makes sense in light of the fact that *V.
-speciosa* is an extremely deep-shade plant, with gametophytes occupying protected rock crevices that receive less than 1 µmol·m^--2^·s^--1^ PPFD for most of the day (Johnson et al., 2000); many of the gametophytes sampled in our study occur in more exposed conditions.
-It is surprising, however, that we did not observe any differences between species in different habitats, given that light levels should vary between tree trunks and the forest floor, at least on a ca.
-1 m scale.
+speciosa* is an extremely deep-shade plant, with gametophytes occupying protected rock crevices that receive less than 1 µmol·m^-2^·s^-1^ PPFD for most of the day [@Johnson2000]; many of the gametophytes sampled in our study occur in more exposed conditions.
+It is surprising, however, that we did not observe any differences between species in different habitats, given that light levels should vary between tree trunks and the forest floor, at least on a \circa{ }1 m scale.
 It is possible that the actual light levels experienced by gametophytes growing in these habitats may not vary significantly if they occupy protected microsites such as bryophyte mats or crevices in rock or bark.
@@
 
-Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates (Proctor, 2003, 2012; Parra et al., 2009; Saldan˜a et al., 2014); however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
+Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates [@Proctor2003; @Proctor2012; @Parra2009; @Saldana2013]; however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
 We detected significant phylogenetic signal in DT, suggesting that phylogeny should be considered when interpreting differences in DT between species (Table 2).
 In contrast, the lack of phylogenetic signal in photosynthetic traits, coupled with the clear intergenerational differences we observed, suggest that sporophytes and gametophytes are adapted for different light levels regardless of clade or ecological niche.
-In a more densely sampled study of morphological evolution in the trichomanoid clade, Dubuisson et al.
-(2013) identified several morphological changes including reduced stele and root systems associated with the evolution of epiphytic growth within the HE subclade (comprising the genera *Polyphlebium*, *Didymoglossum*, *Crepidomanes*, and *Vandenboschia*).
+In a more densely sampled study of morphological evolution in the trichomanoid clade, @Dubuisson2013 identified several morphological changes including reduced stele and root systems associated with the evolution of epiphytic growth within the HE subclade (comprising the genera *Polyphlebium*, *Didymoglossum*, *Crepidomanes*, and *Vandenboschia*).
 It is possible that both reduced morphology and increased DT are parts of an integrated adaptive strategy for epiphytic growth within this subclade of filmy ferns, and have evolved together repeatedly.
-However, Dubuisson et al.
-(2013) were unable to determine if these morphological changes had occurred independently, or if there was a single origin at the base of the HE clade followed by multiple losses.
+However, @Dubuisson2013 were unable to determine if these morphological changes had occurred independently, or if there was a single origin at the base of the HE clade followed by multiple losses.
 Increased sampling of both morphological and physiological traits of additional trichomanoid taxa is needed to distinguish between these scenarios.
@@
 It is likely that even in this clade, wherein the physiological divide between sporophyte and gametophyte is probably the smallest amongst ferns, the two generations experience substantially different selective pressures.
-Fossils indicate that some of the earliest vascular plants may have had sporophytes and gametophytes that were both branched and small (Kenrick and Crane, 1997), but it is unclear at what point selective pressures began to diverge to favor large, complex sporophytes and small, simple gametophytes.
-The relevance of plant size to water relations is supported by a recent study showing increased DT in smaller size classes of the epiphytic fern *Asplenium auritum* Sw.
-(Testo and Watkins, 2012).
+Fossils indicate that some of the earliest vascular plants may have had sporophytes and gametophytes that were both branched and small [@Kenrick1997a], but it is unclear at what point selective pressures began to diverge to favor large, complex sporophytes and small, simple gametophytes.
+The relevance of plant size to water relations is supported by a study showing increased DT in smaller size classes of the epiphytic fern *Asplenium auritum* Sw. [@Testo2012].
 Future studies of DT in filmy ferns should include juvenile stages of the sporophyte, which are expected to experience similar microhabitats as gametophytes, to better determine the role of microhabitat vs.
@@
 
---\>
-
 ## References
@@
 <!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
-<div id="refs"></div>

### 2020-12-19 ms/manuscript.Rmd
Fix variables
- don't abbreviate "sporophyte" and "gametophyte"
- add checks for gameto dt data

@@
 ```{r phy-sig-stats}
-sporo_recover_k <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(kval) %>% jntools::round_t(2)
-sporo_recover_k_pval <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(k_pval) %>% jntools::round_t(3)
-sporo_recover_lambda <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(lambda) %>% jntools::round_t(2)
-sporo_recover_lambda_pval <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(lambda_pval) %>% jntools::round_t(3)
+sporo_recover_k <- phylosig %>% filter(generation == "sporophyte", trait == "recovery_mean") %>% pull(kval) %>% round_t(2)
+sporo_recover_k_pval <- phylosig %>% filter(generation == "sporophyte", trait == "recovery_mean") %>% pull(k_pval) %>% round_t(3)
+sporo_recover_lambda <- phylosig %>% filter(generation == "sporophyte", trait == "recovery_mean") %>% pull(lambda) %>% round_t(2)
+sporo_recover_lambda_pval <- phylosig %>% filter(generation == "sporophyte", trait == "recovery_mean") %>% pull(lambda_pval) %>% round_t(3)
 
@@
 phylosig %>% 
-  filter(trait != "recover_mean") %>%
+  filter(trait != "recovery_mean") %>%
   verify(all(k_pval > 0.05)) %>%
@@
 n_dt_species_sporo <- combined_species_means %>% 
-  filter(!is.na(recover_mean)) %>% 
+  filter(!is.na(recovery_mean)) %>% 
   count(species, generation) %>% 
   count(generation) %>% 
-  filter(generation == "sporo") %>% 
+  filter(generation == "sporophyte") %>% 
   pull(n) %>% 
@@
 n_dt_species_gameto <- combined_species_means %>% 
-  filter(!is.na(recover_mean)) %>%
+  filter(!is.na(recovery_mean)) %>%
   count(species, generation) %>% 
   count(generation) %>% 
-  filter(generation == "gameto") %>% 
+  filter(generation == "gametophyte") %>% 
   pull(n) %>% 
@@
 n_dt_species_both <- combined_species_means %>% 
-  filter(!is.na(recover_mean)) %>% 
+  filter(!is.na(recovery_mean)) %>% 
   count(species, generation) %>% 
@@
 5).
-DT of gametophytes did not significantly exceed that of sporophytes for any species except for saxicolous *P.
+<!-- FIXME: check on these numbers DT of gametophytes did not significantly exceed that of sporophytes for any species except for saxicolous *P.
 endlicherianum* (Fig.
-5; *t*-test, *P* = `r ttest_pval("Polyphlebium_endlicherianum", "recover")`), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C.
+5; *t*-test, *P* = `r ttest_pval("Polyphlebium_endlicherianum", "recovery")`), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C.
 minutum* var.
@@
 borbonicum* (Fig.
-5; *t*-test, *P* = `r ttest_pval("Crepidomanes_minutum2", "recover")`, `r ttest_pval("Hymenophyllum_digitatum", "recover")`, `r ttest_pval("Hymenophyllum_polyanthos", "recover")`, and `r ttest_pval("Polyphlebium_borbonicum", "recover")`, respectively).
+5; *t*-test, *P* = `r ttest_pval("Crepidomanes_minutum2", "recovery")`, `r ttest_pval("Hymenophyllum_digitatum", "recovery")`, `r ttest_pval("Hymenophyllum_polyanthos", "recovery")`, and `r ttest_pval("Polyphlebium_borbonicum", "recovery")`, respectively).
+-->
 
@@
   count(generation) %>% 
-  filter(generation == "sporo") %>% 
+  filter(generation == "sporophyte") %>% 
   pull(n) %>% 
@@
   count(generation) %>% 
-  filter(generation == "gameto") %>% 
+  filter(generation == "gametophyte") %>% 
   pull(n) %>% 
@@
 # Extract values by response and generation for citing in MS
-mean_par_sporo <- light_overall_means %>% filter(generation == "sporo") %>% pull(par_pretty)
-mean_par_gameto <- light_overall_means %>% filter(generation == "gameto") %>% pull(par_pretty)
-mean_etr_sporo <- light_overall_means %>% filter(generation == "sporo") %>% pull(etr_pretty)
-mean_etr_gameto <- light_overall_means %>% filter(generation == "gameto") %>% pull(etr_pretty)
+mean_par_sporo <- light_overall_means %>% filter(generation == "sporophyte") %>% pull(par_pretty)
+mean_par_gameto <- light_overall_means %>% filter(generation == "gametophyte") %>% pull(par_pretty)
+mean_etr_sporo <- light_overall_means %>% filter(generation == "sporophyte") %>% pull(etr_pretty)
+mean_etr_gameto <- light_overall_means %>% filter(generation == "gametophyte") %>% pull(etr_pretty)
 ```

### 2020-12-19 R/clean_gameto_dt_data.R

@@ -329,7 +329,7 @@ gameto_dt_2013 <-
     
 ### Combine all datasets ----
 gameto_dt <- bind_rows(gameto_dt_2012, gameto_dt_2013) %>%
-  mutate(generation = "gameto")
+  mutate(generation = "gametophyte")
 
 #### Filter to only filmy ferns ----
 
@@ -355,7 +355,11 @@ gameto_dt %>%
     by = c(individual = "coll_num")
   ) %>%
   filter(family == "Hymenophyllaceae") %>%
-  select(-family)
+  select(-family) %>%
+  mutate(
+    species = str_replace_all(species, " ", "_"),
+    dry_time = factor(dry_time)) %>%
+  check_dt_data
 
 # Write to data/ ----
 write_csv(filmy_gameto_dt, "data/filmy_gameto_dt.csv")

### 2020-12-19 R/clean_sporo_dt_data.R

@@ -367,7 +367,7 @@ filmy_sporo_dt <- bind_rows(
   select(
     species, salt, dry_time, individual, dataset, contains("weight"), contains("yield")
   ) %>%
-  mutate(generation = "sporo") %>%
+  mutate(generation = "sporophyte") %>%
   check_dt_data
 
 # Write to data_raw/ ----

### 2020-12-19 R/functions.R

@@ -248,7 +248,25 @@ fix_gameto_dt_names <- function (data) {
 #' @return Dataframe
 #' 
 load_sporo_dt <- function (file) {
-  readr::read_csv(file) %>%
+  readr::read_csv(file, col_types = cols(
+    species = col_character(),
+    salt = col_character(),
+    dry_time = col_double(),
+    individual = col_character(),
+    dataset = col_character(),
+    weight_pre = col_double(),
+    weight_dry = col_double(),
+    weight_desiccated = col_double(),
+    weight_30min = col_double(),
+    weight_24hr = col_double(),
+    weight_48hr = col_double(),
+    yield_pre = col_double(),
+    yield_30min = col_double(),
+    yield_24hr = col_double(),
+    yield_48hr = col_double(),
+    yield_72hr = col_double(),
+    generation = col_character()
+  )) %>%
     mutate(
       salt = factor(salt, levels = c("Control", "H2O", "NaCl", "MgNO3", "LiCl")),
       dry_time = factor(dry_time, levels = c(2, 15))
@@ -272,7 +290,27 @@ load_sporo_dt <- function (file) {
 #' @return Dataframe
 #' 
 load_gameto_dt <- function (file) {
-  readr::read_csv(file) %>%
+  readr::read_csv(
+    file, 
+    col_types = cols(
+      individual = col_character(),
+      yield_24hr = col_double(),
+      yield_48hr = col_double(),
+      yield_72hr = col_double(),
+      yield_dry = col_double(),
+      yield_pre = col_double(),
+      salt = col_character(),
+      dry_time = col_double(),
+      yield_30min = col_double(),
+      time_pre = col_datetime(format = ""),
+      time_dry = col_datetime(format = ""),
+      time_30min = col_datetime(format = ""),
+      time_24hr = col_datetime(format = ""),
+      time_48hr = col_datetime(format = ""),
+      time_72hr = col_datetime(format = ""),
+      generation = col_character(),
+      species = col_character()
+    )) %>%
     mutate(
       salt = factor(salt, levels = c("Control", "H2O", "NaCl", "MgNO3", "LiCl")),
       dry_time = factor(dry_time, levels = c(2, 15))
@@ -330,12 +368,15 @@ unzip_nitta_2017 <- function (zipped_path, unzip_path, ...) {
 #' @return Dataframe
 calc_recovery <- function (data) {
   data %>%
[…]

### 2020-12-19 R/plan.R

@@ -61,24 +61,29 @@ plan <- drake_plan(
   # FIXME: check number of measurements per individual; shouldn't have more than 10 each.
   # Crepidomanes_minutum2 (2834), Crepidomanes_minutum2 (2998), and Hymenophyllum_braithwaitei (Hymenophyllum_sp1_6)
   # have too many
-  # light_data_raw = readr::read_csv("data/filmy_light_data.csv"),
+  light_data = read_csv("data/filmy_light_curves.csv"),
 
   # Raw specimen data from specimens spreadsheet
-  # specimens_raw = read_csv("data/specimens.csv"),
+  specimens_raw = read_csv("data/fern_specimens.csv"),
   
   # Process data ----
-  filmy_gameto_recovery = calc_recovery(filmy_gameto_dt),
   
-  filmy_sporo_recovery =
-    filmy_sporo_dt %>%
-    # Only include samples that should be presented in the main MS
-    filter(section == "main") %>%
+  # - calculate recovery during DT test per individual
+  gameto_recovery_indiv = filmy_gameto_dt %>%
+    # Filter out individuals with low pre-treatment yields
+    filter(yield_pre > 400) %>%
     calc_recovery(),
   
-  filmy_sporo_recovery_mean = calc_mean_recovery(filmy_sporo_recovery),
+  sporo_recovery_indiv = filmy_sporo_dt %>%
+    # Only include samples that should be presented in the main MS
+    filter(section == "main") %>%
+    # Filter out individuals with low pre-treatment yields
+    filter(yield_pre > 400) %>%
+    calc_recovery(),
   
-  filmy_gameto_recovery_mean = calc_mean_recovery(filmy_gameto_recovery),
+  recovery_indiv = bind_rows(gameto_recovery_indiv, sporo_recovery_indiv),
   
+  # - calculate mean VPD per datalogger
   mean_vpd = calculate_mean_vpd(climate),
   
   # - calculate mean VPD per species for gametophytes
@@ -123,9 +128,10 @@ plan <- drake_plan(
   etr_indiv = calculate_indiv_etr(light_data),
 
   # Means by species and generation ----
+  
   # - DT recovery
-  recovery_species_means = calculate_mean_recovery(recovery_data),
-
+  recovery_species_means = calc_mean_recovery(recovery_indiv),
+  
   # - ETR max
   etr_species_means = calculate_mean_etr(etr_indiv),
 
@@ -143,7 +149,7 @@ plan <- drake_plan(
   
[…]

### 2020-12-21 ms/manuscript.Rmd
Migrate from drake to targets

@@
 ```{r load-data}
-loadd(
-  list = c(
+tar_load(
+  names = c(
+    "refs",
+    "refs_other",
     "climate",
@@
 }
+
+# FIXME: double check on these p-values
+# ttest_pval("Polyphlebium_endlicherianum", "recovery")
+# ttest_pval("Crepidomanes_minutum2", "recovery")
+# ttest_pval("Hymenophyllum_digitatum", "recovery")
+# ttest_pval("Hymenophyllum_polyanthos", "recovery")
+# ttest_pval("Polyphlebium_borbonicum", "recovery")
 ```
@@
 
-Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; Fig.
-5).
-<!-- FIXME: check on these numbers DT of gametophytes did not significantly exceed that of sporophytes for any species except for saxicolous *P.
+DT of gametophytes did not significantly exceed that of sporophytes for any species except for saxicolous *P.
 endlicherianum* (Fig.
-5; *t*-test, *P* = `r ttest_pval("Polyphlebium_endlicherianum", "recovery")`), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C.
+5; *t*-test, *P* = ), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C.
 minutum* var.
@@
 borbonicum* (Fig.
-5; *t*-test, *P* = `r ttest_pval("Crepidomanes_minutum2", "recovery")`, `r ttest_pval("Hymenophyllum_digitatum", "recovery")`, `r ttest_pval("Hymenophyllum_polyanthos", "recovery")`, and `r ttest_pval("Polyphlebium_borbonicum", "recovery")`, respectively).
--->
+5; *t*-test, *P* = XXX, XXX, XXX, and XXX, respectively).
+
+Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; Fig.
+5).
 
@@
 
+## Figures
+
 ## References

### 2020-12-21 R/functions.R

@@ -1282,11 +1282,12 @@ render_tracked <- function (tracked_output, dep1 = NULL, dep2 = NULL, ...) {
 #' @param template Path to template docx file.
 #' @param wd Working directory to run conversion. Should be same as
 #' directory containing any files needed to render latex to pdf.
+#' @param ... Other arguments; not used by this function, but for tracking dependencies.
 #'
 #' @return List including STDOUT of pandoc; externally, the
 #' docx file will be rendered in `wd`.
 #' 
-latex2docx <- function (latex, docx, template = NULL, wd = getwd()) {
+latex2docx <- function (latex, docx, template = NULL, wd = getwd(), ...) {
   
   assertthat::assert_that(assertthat::is.readable(latex))
   

### 2020-12-21 R/packages.R

@@ -1,7 +1,8 @@
 # Load packages used in the analysis
 
 library(conflicted)
-library(drake)
+library(targets)
+library(tarchetypes)
 library(ape)
 library(assertr)
 library(assertthat)

### 2020-12-21 R/packages_dev.R

@@ -1,0 +1,1 @@
+library("visNetwork")

### 2020-12-21 _drake.R

@@ -1,7 +1,0 @@
-# Load packages, functions, and plan
-source("R/packages.R")
-source("R/functions.R")
-source("R/plan.R")
-
-# Set up drake configuration
-drake_config(plan, verbose = 1, seed = 0)

### 2020-12-21 _targets.R

@@ -1,70 +1,51 @@
-# Plan the analysis
+source("R/packages.R")
+source("R/functions.R")
 
-plan <- drake_plan(
-
-  # Unzip raw data ----
-  
-  # Unzip data from Nitta et al. 2017 Ecol. Monographs from Dryad
-  # The dataset must be downloaded first by going to
-  # https://datadryad.org/stash/dataset/doi:10.5061/dryad.df59g, clicking on
-  # "Download dataset", and saving to the "data" folder in this project.
-  nitta_2017_data = unzip_nitta_2017(
-    zipped_path = file_in("data/doi_10.5061_dryad.df59g__v1.zip"),
-    unzip_path = "data/nitta_2017",
-    # Track data files used as input in analyses
-    out1 = file_out("data/nitta_2017/sites.csv"),
-    out2 = file_out("data/nitta_2017/treepl_Moorea_Tahiti.tre"),
-    out3 = file_out("data/nitta_2017/all_plots.csv")),
-
-  # Unzip data from Nitta et al. 2020 New Phyt. from Dryad
-  # The dataset must be downloaded first by going to
-  # https://datadryad.org/stash/dataset/doi:10.5061/dryad.fqz612jps, clicking on
-  # "Download dataset", and saving to the "data" folder in this project.
-  climate_unzipped = unzip(
-    zipfile = file_in("data/doi_10.5061_dryad.fqz612jps__v4.zip"), 
-    files = "moorea_climate.csv", 
-    exdir = file_out("data/nitta_2020/"), 
-    junkpaths = TRUE, 
-    overwrite = TRUE),
+tar_plan(
   
   # Load data ----
   
   # - microclimate data
-  climate = readr::read_csv(file_in("data/nitta_2020/moorea_climate.csv")),
-
+  tar_file(climate_file, "data/nitta_2020/moorea_climate.csv"),
+  climate = read_csv(climate_file),
+  
   # - site data
-  moorea_sites = readr::read_csv(
-    file_in("data/nitta_2017/sites.csv"),
+  tar_file(moorea_sites_file, "data/nitta_2017/sites.csv"),
+  moorea_sites = read_csv(
+    moorea_sites_file,
     col_types = "cnnn") %>%
     filter(str_detect(site, "Aorai", negate = TRUE)),
-
+  
   # - growth habit
-  filmy_habit = readr::read_csv("data/filmy_growth_habit.csv"),
-
+  tar_file(filmy_habit_file, "data/filmy_growth_habit.csv"),
[…]

### 2021-01-04 ms/si.Rmd
Add SI

---
title: "SI"
bibliography: 
  - references.bib
  - references_other.yaml
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-plant-research.csl"]
    includes:
      in_header: journal-of-plant-research-si.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = TRUE)
# if running interactively, source _targets.R first
# source("_targets.R")
```

```{r load-data}
tar_load(
  names = c(
    "filmy_dt_chamber",
    "filmy_dt"
  )
)
```

```{r load-functions}
# Define functions specific to this file
plot_chamber_rh <- function (data, time_limits) {
  ggplot(data, aes(x = date_time, y = 0.01*rh)) +
    geom_line() +
    scale_x_datetime(limits = time_limits) +
    scale_y_continuous(label = scales::label_percent(accuracy = 1)) +
    labs(y = "Relative humidity") +
    theme(
      legend.position = "none",
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank())
}

plot_chamber_temp <- function (data, time_limits) {
  ggplot(data, aes(x = date_time, y = temp)) +
    geom_line() +
    scale_x_datetime(limits = time_limits) +
    scale_y_continuous(label = scales::label_number(suffix = "\u00b0C", accuracy = 1)) +
    labs(y = "Temperature") +
    theme(
      legend.position = "none",
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank())
}

plot_sporo_chamber_time <- function (data, time_limits) {
    ggplot(data, aes(x = species, ymin = pre, ymax = post, y = post, group = species_duration)) +
    geom_linerange(position = position_dodge(0.5)) +
    geom_point(position = position_dodge(0.5)) +
    scale_y_datetime(limits = time_limits) +
    coord_flip() +
    labs(x = "Species") +
    theme(
      axis.title.x = element_blank()
    )
}

plot_sporo_chamber <- function (sporo_times_data, filmy_dt_chamber_data, salt_select, k_select) {

  # Get mean times of each cluster of DT test samples for the selected drying salt
  sporo_clusters <- cluster_sporo_times(sporo_times_data, salt_select, k_select)

  # Calculate time range for the clustered data
  sporo_time_range <- c(min = min(sporo_clusters$pre), max = max(sporo_clusters$post))

  # Subset DT chamber data
  sporo_chamber_data <-
    filmy_dt_chamber_data %>% 
    filter(date_time >= sporo_time_range[["min"]], date_time <= sporo_time_range[["max"]], salt == salt_select)

  # RH plot
  rh_plot <- plot_chamber_rh(sporo_chamber_data, sporo_time_range)

  # Temp plot
  temp_plot <- plot_chamber_temp(sporo_chamber_data, sporo_time_range)

  # Dry times plot
  time_plot <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range)

  # Combine plots
  rh_plot + temp_plot + time_plot + plot_layout(ncol = 1)
}

#' Cluster times from a sporophyte desiccation tolerance (DT) test
#' 
#' Calculates mean pre- and post-DT test times by species and duration of DT test
#' (2 or 15 days), also adds a `species_duration` column to the output for plotting.
#'
#' @param sporo_time_data Dataframe with DT test data for sporophytes
#' @param salt_select Name of salt to subset data
#' @param k_select Number of groups to cluster. For example, if the DT
#' test was done in 3 groups, `k_select` = 3.
#'
#' @return Dataframe
cluster_sporo_times <- function (sporo_time_data, salt_select, k_select) {

  sporo_2d_clusters <- 
    sporo_time_data %>%
    filter(salt == salt_select, dry_time == "2") %>%
    add_clusters("pre", k_select) %>% # choose K by inspecting raw data first
    group_by(species, cluster, salt, dry_time) %>%
    summarize(pre = mean(pre), post = mean(post), n = n(), .groups = "drop") # %>%
    # select(-cluster)

  sporo_15d_clusters <- 
    sporo_time_data %>%
    filter(salt == salt_select, dry_time == "15") %>%
    add_clusters("pre", k_select) %>%
    group_by(species, cluster, salt, dry_time) %>%
    summarize(pre = mean(pre), post = mean(post), n = n(), .groups = "drop") # %>%
    # select(-cluster)

  bind_rows(sporo_2d_clusters, sporo_15d_clusters) %>%
    mutate(species_duration = paste(species, dry_time))

}

```

```{r gametophyte-dt-2013, fig.cap = ""}
# Calculate mean start/end times for gametophyte DT groups
gameto_mean_times <-
  filmy_dt %>%
  filter(generation == "gametophyte") %>%
  select(individual, contains("time_")) %>%
  pivot_longer(names_to = "condition", values_to = "date_time", -individual) %>%
  filter(!is.na(date_time)) %>%
  mutate(condition = str_remove_all(condition, "time_") %>%
           str_replace_all("30min", "post")) %>%
  filter(condition %in% c("pre", "post")) %>%
  add_count(individual) %>%
  filter(n > 1) %>%
  select(-n) %>%
  pivot_wider(names_from = "condition", values_from = "date_time") %>%
  # Cluster starts times into groups (here, I know there are 10 groups)
  add_clusters("pre", 10) %>%
  mutate(cluster = as.factor(cluster)) %>%
  group_by(cluster) %>%
  summarize(pre = mean(pre), post = mean(post), n = n(), .groups = "drop")

gameto_time_limits <- c(min = min(gameto_mean_times$pre), max = max(gameto_mean_times$post))

# DT start/end times plot
dt_times_gameto_2013_plot <- ggplot(
  gameto_mean_times, 
  aes(x = cluster, ymin = pre, ymax = post, y = post, group = cluster)) +
  geom_linerange(position = position_dodge(0.5)) +
  geom_point(position = position_dodge(0.5)) +
  scale_y_datetime(limits = gameto_time_limits) +
  coord_flip() +
  labs(x = "Group") +
  theme(
    axis.title.x = element_blank()
  )

# Subset DT chamber data
chamber_gameto_2013 <- filmy_dt_chamber %>% filter(generation == "gametophyte", year == "2013")

# RH plot
chamber_gameto_2013_rh_plot <-
  plot_chamber_rh(chamber_gameto_2013, gameto_time_limits)

# Temp plot
chamber_gameto_2013_temp_plot <-
  plot_chamber_temp(chamber_gameto_2013, gameto_time_limits)

chamber_gameto_2013_temp_plot + chamber_gameto_2013_rh_plot + dt_times_gameto_2013_plot + plot_layout(ncol = 1)

ggsave(here::here("results/S1_fig.pdf"))
```

**Figure S1**. Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes. MgNO3 was used to maintain constant humidity at \circa{ }58%. All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days by accident.

```{r sporophyte-dt-2013-prep}
# Extract pre- and post-DT treatment times for sporophytes
sporo_times <-
  filmy_dt %>%
  filter(generation == "sporophyte") %>%
  select(species:dataset, contains("time_")) %>%
  pivot_longer(names_to = "condition", values_to = "date_time", contains("time_")) %>%
  filter(!is.na(date_time)) %>%
  mutate(condition = str_remove_all(condition, "time_") %>%
           # Consider first recovery measurement (at 30 min) to be "post-DT treatment"
           str_replace_all("30min", "post")) %>%
  filter(condition %in% c("pre", "post")) %>%
  add_count(species, salt, dry_time, individual, dataset) %>%
  filter(n > 1) %>%
  select(-n) %>%
  pivot_wider(names_from = "condition", values_from = "date_time")
```

```{r sporophyte-dt-2013-LiCl, fig.cap = ""}

# Get mean times of each cluster of DT test samples for the selected drying salt
sporo_clusters_2013 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2013")), "LiCl", 3)
sporo_clusters_2014 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2014")), "LiCl", 1)

sporo_clusters <- bind_rows(sporo_clusters_2013, sporo_clusters_2014)

# Calculate time range for the clustered data
sporo_time_range_2013 <- c(min = min(sporo_clusters_2013$pre), max = max(sporo_clusters_2013$post))
sporo_time_range_2014 <- c(min = min(sporo_clusters_2014$pre), max = max(sporo_clusters_2014$post))

# Subset DT chamber data
licl_chamber_data_2013 <-
  filter(filmy_dt_chamber, 
         date_time >= sporo_time_range_2013[["min"]], date_time <= sporo_time_range_2013[["max"]], salt == "LiCl")
licl_chamber_data_2014 <-filter(
  filmy_dt_chamber, 
  date_time >= sporo_time_range_2014[["min"]], date_time <= sporo_time_range_2014[["max"]], salt == "LiCl")

# RH plot
rh_plot_2013 <- plot_chamber_rh(licl_chamber_data_2013, sporo_time_range_2013)
rh_plot_2014 <- plot_chamber_rh(licl_chamber_data_2014, sporo_time_range_2014)

# Temp plot
temp_plot_2013 <- plot_chamber_temp(licl_chamber_data_2013, sporo_time_range_2013)
temp_plot_2014 <- plot_chamber_temp(licl_chamber_data_2014, sporo_time_range_2014)

# Dry times plot
time_plot_2013 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2013)
time_plot_2014 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2014)

# Combine plots
temp_plot_2013 + temp_plot_2014 + 
  rh_plot_2013 + rh_plot_2014 + 
  time_plot_2013 + time_plot_2014 + 
  plot_layout(ncol = 2)

ggsave(here::here("results/S2_fig.pdf"))
```

**Figure S2**.

```{r sporophyte-dt-2013-MgNO3, fig.cap = ""}
plot_sporo_chamber(
  sporo_times_data = sporo_times %>% filter(species != "Callistopteris_apiifolia"), 
  filmy_dt_chamber_data = filter(filmy_dt_chamber, generation == "sporophyte"),
  salt_select = "MgNO3", 
  k_select = 3)

ggsave(here::here("results/S3_fig.pdf"))
```

**Figure S3**.

```{r sporophyte-dt-2013-NaCl, fig.cap = ""}
# Sporo NaCl plot: this includes three dataloggers with different SNs (and therefore chambers),
# but some data are missing
nacl_clusters_other <- cluster_sporo_times(
  filter(sporo_times, species != "Abrodictyum_dentatum"), 
  "NaCl", 4)

nacl_clusters_adent <- cluster_sporo_times(
  filter(sporo_times, species == "Abrodictyum_dentatum"), 
  "NaCl", 1)

nacl_clusters <- bind_rows(nacl_clusters_other, nacl_clusters_adent) %>%
  mutate(
  serial_no = case_when(
    species %in% c("Didymoglossum_tahitense", "Crepidomanes_kurzii", "Callistopteris_apiifolia") & dry_time == "2" ~ "10342561",
    species %in% c("Hymenophyllum_multifidum", "Hymenophyllum_digitatum") ~ "10140667",
    species %in% c("Hymenophyllum_polyanthos") ~ "10152774",
    species %in% c("Abrodictyum_dentatum") & dry_time == "2" ~ "missing",
    TRUE ~ "missing"
  )
)

nacl_time_range <- c(min = min(nacl_clusters$pre), max = max(nacl_clusters$post))

serial_no_cols <- brewer_pal(palette = "Dark2")(4)

# Time plot
nacl_time_plot <- ggplot(nacl_clusters, aes(x = species, ymin = pre, ymax = post, y = post, group = species_duration, color = serial_no)) +
  geom_linerange(position = position_dodge(0.5)) +
  geom_point(position = position_dodge(0.5)) +
  scale_y_datetime(limits = nacl_time_range) +
  scale_color_manual(values = c(
    "10342561" = serial_no_cols[[1]],
    "10140667" = serial_no_cols[[2]],
    "10152774" = serial_no_cols[[3]],
    "10342616" = serial_no_cols[[4]],
    "missing" = "grey20"
    )) +
  coord_flip() +
  labs(x = "Species") +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )

# Subset DT chamber data
nacl_chamber_sporo_data <-
  filmy_dt_chamber %>% 
  filter(
    salt == "NaCl", 
    date_time >= nacl_time_range[["min"]], date_time <= nacl_time_range[["max"]])

# RH plot
nacl_rh_plot <- ggplot(nacl_chamber_sporo_data, aes(x = date_time, y = 0.01*rh, color = serial_no)) +
  scale_color_manual(values = c(
    "10342561" = serial_no_cols[[1]],
    "10140667" = serial_no_cols[[2]],
    "10152774" = serial_no_cols[[3]],
    "10342616" = serial_no_cols[[4]],
    "missing" = "grey20"
  )) +
  geom_line() +
  scale_x_datetime(limits = nacl_time_range) +
  scale_y_continuous(label = scales::label_percent(accuracy = 1)) +
  labs(
    y = "Relative humidity",
    color = "Data logger\nserial no.") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank())

nacl_temp_plot <- ggplot(nacl_chamber_sporo_data, aes(x = date_time, y = temp, color = serial_no)) +
  geom_line() +
  scale_color_manual(values = c(
    "10342561" = serial_no_cols[[1]],
    "10140667" = serial_no_cols[[2]],
    "10152774" = serial_no_cols[[3]],
    "10342616" = serial_no_cols[[4]],
    "missing" = "grey20"
  )) +
  scale_x_datetime(limits = nacl_time_range) +
  scale_y_continuous(label = scales::label_number(suffix = "\u00b0C", accuracy = 1)) +
  labs(y = "Temperature") +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank())

nacl_temp_plot + nacl_rh_plot + nacl_time_plot + plot_layout(ncol = 1)

ggsave(here::here("results/S4_fig.pdf"))
```

**Figure S4**.

### 2021-01-04 R/functions.R

@@ -401,6 +401,22 @@ unzip_nitta_2017 <- function (zipped_path, unzip_path, ...) {
 
 # Data wrangling ----
 
+#' Cluster rows of a dataframe based on a single column
+#' 
+#' For example, start times in a dataframe of desiccation tolerance
+#' start and end times.
+#'
+#' @param data Data to cluster
+#' @param col Column to use for clustering
+#' @param k Number of clusters to generate
+#'
+#' @return Dataframe with column "cluster" added
+#' 
+add_clusters <- function(data, col, k) {
+  kmeans_uni_groups <- Ckmeans.1d.dp::Ckmeans.1d.dp(as.numeric(data[[col]]), k = k)
+  mutate(data, cluster = kmeans_uni_groups$cluster)
+}
+
 #' Calculate % recovery from desiccation in a desiccation tolerance (DT) experiment
 #'
 #' @param data Dataframe; data read in from DT experiment

### 2021-01-04 R/packages.R

@@ -11,8 +11,11 @@ library(jntools)
 library(janitor)
 library(here)
 library(lubridate)
+library(patchwork)
+library(scales)
 library(tidyverse)
 
+conflict_prefer("col_factor", "readr")
 conflict_prefer("filter", "dplyr")
 conflict_prefer("rename", "dplyr")
 conflict_prefer("gather", "tidyr")

### 2021-01-04 _targets.R

@@ -173,6 +173,9 @@ tar_plan(
     template = "ms/journal-of-plant-research.docx",
     wd = here::here("results"),
     depends = manuscript_pdf
-  )
+  ),
+  
+  # SI
+  tar_render(si_pdf, "ms/si.Rmd", output_dir = here::here("results"))
   
 )

### 2021-01-04 R/clean_chamber_data.R
Classify species in DT metadata

@@ -15,14 +15,14 @@ sporo_2012_chamber <- bind_rows(
   parse_logger_dat("data_raw/2012/filmyDT2_MgNO3.csv") %>% mutate(salt = "MgNO3"),
   parse_logger_dat("data_raw/2012/filmyDT2_NaCl.csv") %>% mutate(salt = "NaCl")
 ) %>%
-  mutate(generation = "sporophyte", year = 2012)
+  mutate(generation = "sporophyte", year = 2012, species = "mixed")
   
 # 2012 gametophytes
 gameto_2012_chamber <-
   bind_rows(
     parse_logger_dat("data_raw/2012/gameto1_7-14.csv"),
     parse_logger_dat("data_raw/2012/gameto2_7-24.csv")
-  ) %>% mutate(salt = "MgNO3", generation = "gametophyte", year = 2012)
+  ) %>% mutate(salt = "MgNO3", generation = "gametophyte", year = 2012, species = "mixed")
 
 # 2013 load metadata for hobo loggers
 hobo_metadata <- read_csv(
@@ -50,7 +50,7 @@ hobo_chamber_data <-
   unnest(data) %>%
   # Exclude data not marked "keep" in metadata
   filter(use_status == "keep") %>%
-  select(-use_status, -file, -note) %>%
+  select(-use_status, -note) %>%
   # modify generation for Hobo s/n 10140659, which was used partly for sporos, partly for gametos
   mutate(
     generation = case_when(
@@ -62,6 +62,6 @@ hobo_chamber_data <-
 
 # Combine and write out
 combined_chamber_data <- bind_rows(sporo_2012_chamber, gameto_2012_chamber, hobo_chamber_data) %>%
-  assert(not_na, date_time, temp, rh, salt, generation, year)
+  assert(not_na, date_time, temp, rh, salt, generation, year, species)
 
 write_csv(combined_chamber_data, "data/filmy_dt_chamber.csv")

### 2021-01-04 R/clean_dt_data.R
Manually add 2014 DT test times

@@ -753,7 +753,16 @@ filmy_dt_raw_yield_long %>%
     rename(minipam_2013_long, yield_pam = yield), 
     by = c("memory", "pam_set")) %>%
   select(species, salt, dry_time, individual, generation, condition, yield_pam, date_time, note, dataset) %>%
-  # Make sure no rows got dupliated during the join
+  # Manually specify times for 2014 data (as per excel sheet)
+  mutate(date_time = case_when(
+    dry_time == "15" & condition == "pre" & dataset == "2014" ~ as_datetime("2014-06-17 00:10:00"),
+    dry_time == "15" & condition == "30min" & dataset == "2014" ~ as_datetime("2014-07-02 18:00:00"),
+    dry_time == "2" & condition == "pre" & dataset == "2014" ~ as_datetime("2014-07-06 16:30:00"),
+    dry_time == "2" & condition == "30min" & dataset == "2014" ~ as_datetime("2014-07-08 17:00:00"),
+    TRUE ~ date_time
+  )
+  ) %>%
+  # Make sure no rows got duplicated during the join
   assert_rows(col_concat, is_uniq, species, salt, dry_time, individual, condition)
 
 filmy_dt_pam_yield <- bind_rows(sporo_dt_pam_yield, gameto_dt_pam_yield)

### 2021-01-04 ms/si.Rmd
Fix sporo LiCl plot for 2013, 2014
Remove MgNO3 and NaCl plots for now, still need to fix

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
 # if running interactively, source _targets.R first
@@
     scale_x_datetime(limits = time_limits) +
-    scale_y_continuous(label = scales::label_percent(accuracy = 1)) +
+    scale_y_continuous(label = label_percent(accuracy = 1)) +
     labs(y = "Relative humidity") +
@@
     scale_x_datetime(limits = time_limits) +
-    scale_y_continuous(label = scales::label_number(suffix = "\u00b0C", accuracy = 1)) +
+    scale_y_continuous(label = label_number(suffix = "\u00b0C", accuracy = 1)) +
     labs(y = "Temperature") +
@@
 plot_sporo_chamber_time <- function (data, time_limits) {
-    ggplot(data, aes(x = species, ymin = pre, ymax = post, y = post, group = species_duration)) +
+    ggplot(data, aes(x = species, ymin = pre, ymax = post, y = post, group = species_duration, linetype = dry_time)) +
     geom_linerange(position = position_dodge(0.5)) +
@@
     scale_y_datetime(limits = time_limits) +
+    scale_linetype_manual(values = c("15" = "solid", "2" = "solid")) +
     coord_flip() +
@@
     theme(
-      axis.title.x = element_blank()
+      axis.title.x = element_blank(),
+      legend.position = "none"
     )
@@
 ```{r sporophyte-dt-2013-LiCl, fig.cap = ""}
-
 # Get mean times of each cluster of DT test samples for the selected drying salt
@@
   filter(filmy_dt_chamber, 
-         date_time >= sporo_time_range_2013[["min"]], date_time <= sporo_time_range_2013[["max"]], salt == "LiCl")
-licl_chamber_data_2014 <-filter(
-  filmy_dt_chamber, 
-  date_time >= sporo_time_range_2014[["min"]], date_time <= sporo_time_range_2014[["max"]], salt == "LiCl")
+    date_time >= sporo_time_range_2013[["min"]], 
+    date_time <= sporo_time_range_2013[["max"]], 
+    salt == "LiCl")
+licl_chamber_data_2014 <- 
+  filter(filmy_dt_chamber, 
+    date_time >= sporo_time_range_2014[["min"]], 
+    date_time <= sporo_time_range_2014[["max"]], 
+    salt == "LiCl") %>%
+  # There are two sets of times: 2014-06-17 to 2014-07-02, and 2014-07-06 to 2014-07-08
+  # Add a group for these so ggplot doesn't connect lines betweeen them
+  mutate(
+    day = yday(date_time),
+    group = ifelse(day < 185, "a", "b"))
+
+licl_chamber_data <- bind_rows(licl_chamber_data_2013, licl_chamber_data_2014)
+licl_temp_range <- c(min(licl_chamber_data$temp), max(licl_chamber_data$temp))
+licl_rh_range <- c(min(licl_chamber_data$rh), max(licl_chamber_data$rh))*0.01
 
 # RH plot
-rh_plot_2013 <- plot_chamber_rh(licl_chamber_data_2013, sporo_time_range_2013)
-rh_plot_2014 <- plot_chamber_rh(licl_chamber_data_2014, sporo_time_range_2014)
+rh_plot_2013 <- plot_chamber_rh(licl_chamber_data_2013, sporo_time_range_2013) +
+  scale_y_continuous(limits = licl_rh_range, label = label_percent(accuracy = 1))
+
+rh_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = 0.01*rh, group = group)) +
+  geom_line() +
+  scale_x_datetime(limits = sporo_time_range_2014) +
+  blank_x_theme() +
+  blank_y_theme() +
+  theme(legend.position = "none")
 
 # Temp plot
-temp_plot_2013 <- plot_chamber_temp(licl_chamber_data_2013, sporo_time_range_2013)
-temp_plot_2014 <- plot_chamber_temp(licl_chamber_data_2014, sporo_time_range_2014)
+temp_plot_2013 <- plot_chamber_temp(licl_chamber_data_2013, sporo_time_range_2013) +
+  scale_y_continuous(limits = licl_temp_range, label = label_number(suffix = "\u00b0C", accuracy = 1))
+
+temp_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = temp, group = group)) +
+  geom_line() +
+  scale_x_datetime(limits = sporo_time_range_2014) +
+  blank_x_theme() +
+  blank_y_theme() +
+  theme(legend.position = "none")
 
 # Dry times plot
-time_plot_2013 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2013)
-time_plot_2014 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2014)
+time_plot_2013 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2013) +
+  labs(y = "2013") +
+  theme(axis.title.x = element_text()) 
+time_plot_2014 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2014)  +
+  labs(y = "2014")  +
+  theme(axis.title.x = element_text()) +
+  blank_y_theme()
 
@@
 **Figure S2**.
-
-```{r sporophyte-dt-2013-MgNO3, fig.cap = ""}
-plot_sporo_chamber(
-  sporo_times_data = sporo_times %>% filter(species != "Callistopteris_apiifolia"), 
-  filmy_dt_chamber_data = filter(filmy_dt_chamber, generation == "sporophyte"),
-  salt_select = "MgNO3", 
-  k_select = 3)
-
-ggsave(here::here("results/S3_fig.pdf"))
-```
-
-**Figure S3**.
-
-```{r sporophyte-dt-2013-NaCl, fig.cap = ""}
-# Sporo NaCl plot: this includes three dataloggers with different SNs (and therefore chambers),
-# but some data are missing
-nacl_clusters_other <- cluster_sporo_times(
-  filter(sporo_times, species != "Abrodictyum_dentatum"), 
-  "NaCl", 4)
-
-nacl_clusters_adent <- cluster_sporo_times(
-  filter(sporo_times, species == "Abrodictyum_dentatum"), 
-  "NaCl", 1)
-
-nacl_clusters <- bind_rows(nacl_clusters_other, nacl_clusters_adent) %>%
-  mutate(
-  serial_no = case_when(
-    species %in% c("Didymoglossum_tahitense", "Crepidomanes_kurzii", "Callistopteris_apiifolia") & dry_time == "2" ~ "10342561",
-    species %in% c("Hymenophyllum_multifidum", "Hymenophyllum_digitatum") ~ "10140667",
-    species %in% c("Hymenophyllum_polyanthos") ~ "10152774",
-    species %in% c("Abrodictyum_dentatum") & dry_time == "2" ~ "missing",
-    TRUE ~ "missing"
-  )
-)
-
-nacl_time_range <- c(min = min(nacl_clusters$pre), max = max(nacl_clusters$post))
-
-serial_no_cols <- brewer_pal(palette = "Dark2")(4)
-
-# Time plot
-nacl_time_plot <- ggplot(nacl_clusters, aes(x = species, ymin = pre, ymax = post, y = post, group = species_duration, color = serial_no)) +
-  geom_linerange(position = position_dodge(0.5)) +
-  geom_point(position = position_dodge(0.5)) +
-  scale_y_datetime(limits = nacl_time_range) +
-  scale_color_manual(values = c(
-    "10342561" = serial_no_cols[[1]],
-    "10140667" = serial_no_cols[[2]],
-    "10152774" = serial_no_cols[[3]],
-    "10342616" = serial_no_cols[[4]],
-    "missing" = "grey20"
-    )) +
-  coord_flip() +
-  labs(x = "Species") +
-  theme(
-    axis.title.x = element_blank(),
-    legend.position = "none"
-  )
-
-# Subset DT chamber data
-nacl_chamber_sporo_data <-
-  filmy_dt_chamber %>% 
-  filter(
-    salt == "NaCl", 
-    date_time >= nacl_time_range[["min"]], date_time <= nacl_time_range[["max"]])
-
-# RH plot
-nacl_rh_plot <- ggplot(nacl_chamber_sporo_data, aes(x = date_time, y = 0.01*rh, color = serial_no)) +
-  scale_color_manual(values = c(
-    "10342561" = serial_no_cols[[1]],
-    "10140667" = serial_no_cols[[2]],
-    "10152774" = serial_no_cols[[3]],
-    "10342616" = serial_no_cols[[4]],
-    "missing" = "grey20"
-  )) +
-  geom_line() +
-  scale_x_datetime(limits = nacl_time_range) +
-  scale_y_continuous(label = scales::label_percent(accuracy = 1)) +
-  labs(
-    y = "Relative humidity",
-    color = "Data logger\nserial no.") +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank())
-
-nacl_temp_plot <- ggplot(nacl_chamber_sporo_data, aes(x = date_time, y = temp, color = serial_no)) +
-  geom_line() +
-  scale_color_manual(values = c(
-    "10342561" = serial_no_cols[[1]],
-    "10140667" = serial_no_cols[[2]],
-    "10152774" = serial_no_cols[[3]],
-    "10342616" = serial_no_cols[[4]],
-    "missing" = "grey20"
-  )) +
-  scale_x_datetime(limits = nacl_time_range) +
-  scale_y_continuous(label = scales::label_number(suffix = "\u00b0C", accuracy = 1)) +
-  labs(y = "Temperature") +
-  theme(
-    legend.position = "none",
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank())
-
-nacl_temp_plot + nacl_rh_plot + nacl_time_plot + plot_layout(ncol = 1)
-
-ggsave(here::here("results/S4_fig.pdf"))
-```
-
-**Figure S4**.
+Conditions inside desiccation chamber and duration of desiccation treatment for sporophytes at most intense desiccation level.
+LiCl was used to maintain constant humidity at \circa{ }18%.

### 2021-01-05 ms/manuscript.Rmd
Add SI plots: rel water content, DT controls for sporophytes

@@
 Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
-For sporophytes, 8--16 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
+For sporophytes, 8 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
 We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
 All experiments were initiated within 48 h of collection.
-Pre-treatment chlorophyll fluorescence (F~v~/F~m~) was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
+Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
 Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
@@
 Following the desiccation treatment, plants were rewetted using water from a local stream.
-Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0.5 h, 24 h, and 48 h following rewetting.
-Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield [@WatkinsJr.2007c].
+Plants were allowed to rehydrate, and F~v~/F~m~ was again measured at 0.5 h, 24 h, and 48 h following rewetting.
+Ability to recover from desiccation was quantified by comparing pre- and post-treatment F~v~/F~m~ [@WatkinsJr.2007c].
 

### 2021-01-05 ms/si.Rmd
Add SI plots: rel water content, DT controls for sporophytes

@@
     fig_caption: yes
-    keep_tex: yes
+    keep_tex: no
     latex_engine: xelatex
@@
 
-```{r load-data}
+```{r load-data, cache = FALSE}
 tar_load(
@@
     "filmy_dt_chamber",
-    "filmy_dt"
+    "filmy_dt",
+    "filmy_habit",
+    "rel_water_indiv",
+    "rel_water_species_means"
   )
@@
 
-```{r load-functions}
+```{r load-functions, cache = FALSE}
 # Define functions specific to this file
@@
 
-```{r gametophyte-dt-2013, fig.cap = ""}
+```{r gametophyte-dt-2013, fig.height = 4.5, fig.width = 7, fig.cap = ""}
 # Calculate mean start/end times for gametophyte DT groups
@@
 
-ggsave(here::here("results/S1_fig.pdf"))
+# ggsave(here::here("results/S1_fig.png"))
 ```
@@
 Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes.
-MgNO3 was used to maintain constant humidity at \circa{ }58%.
-All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days by accident.
+MgNO~3~ was used to maintain constant humidity at \circa{ }58%.
+All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days unintentionally.
+
+\clearpage
 
@@
 
-```{r sporophyte-dt-2013-LiCl, fig.cap = ""}
+```{r sporophyte-dt-2013-LiCl, fig.height = 4.5, fig.width = 7, fig.cap = ""}
 # Get mean times of each cluster of DT test samples for the selected drying salt
@@
 
-ggsave(here::here("results/S2_fig.pdf"))
+# ggsave(here::here("results/S2_fig.png"))
 ```
@@
 LiCl was used to maintain constant humidity at \circa{ }18%.
+
+\clearpage
+
+```{r water-content-prep}
+# Define color-blind palette
+cbPalette <- c(grey = "#999999", gold = "#E69F00", skyblue = "#56B4E9", 
+               green = "#009E73", yellow = "#F0E442", blue = "#0072B2", 
+               red = "#D55E00", purple = "#CC79A7")
+
+# Set colors, shapes for desiccation intensity
+salt_cols <- c(
+  "NaCl" = cbPalette[["blue"]], 
+  "MgNO3" = cbPalette[["yellow"]], 
+  "LiCl" = cbPalette[["red"]], 
+  "Control" = "white")
+
+salt_shapes <- c(
+  "NaCl" = 21, # circle
+  "MgNO3" = 24, # triangle
+  "LiCl" = 23, # diamond
+  "Control" = 22 # square
+  )
+```
+
+```{r water-content-full, fig.height = 4.5, fig.width = 7, fig.cap = ""}
+# Plot only species with weights recorded during recovery
+species_with_rec <-
+  rel_water_species_means %>%
+  filter(rec_time %in% c("30min", "24hr", "48hr")) %>%
+  pull(species) %>%
+  unique()
+
+rel_water_species_means_with_rec <-
+  rel_water_species_means %>%
+  filter(species %in% species_with_rec)
+
+pd <- position_dodge(.5)
+
+# Define color-blind palette
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+# Set colors for desiccation intensity
+salt_cols <- c("NaCl" = cbPalette[7], "MgNO3" = cbPalette[5], "LiCl" = cbPalette[6], "Control" = "white")
+
+ggplot(rel_water_species_means_with_rec, aes(x = rec_time, y = rel_water, group = interaction(dry_time, salt))) +
+  geom_errorbar(
+    aes(ymin = rel_water - sd, ymax = rel_water + sd),
+    color = "grey50",
+    width = 0,
+    position = pd) +
+  geom_line(position = pd, aes(linetype = dry_time), color = "grey30") +
+  geom_point(aes(shape = salt, fill = salt), position = pd) +
+  scale_linetype_manual(name = "Desiccation Time",
+                        values = c("solid", "dashed"),
+                        breaks = c("15", "2"),
+                        labels = c("15 days", "2 days")) +
+  scale_shape_manual(name = "Desiccation Intensity",
+                     values = salt_shapes,
+                     breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                     labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
+  scale_fill_manual(name = "Desiccation Intensity",
+                    values = salt_cols, 
+                    breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                    labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) +
+  scale_y_continuous(label = scales::percent) +
+  facet_wrap(vars(species)) +
+  labs(x = "Condition",
+       y = "Relative water content") +
+  theme_bw() +
+  theme(
+    panel.grid.major.x = element_blank(),
+    legend.position = "bottom",
+    legend.box="vertical", 
+    legend.margin=margin()
+  )
+
+# ggsave(here::here("results/S_fig_water_content_full.png"))
+```
+
+**Figure SXX**.
+Relative water content (RWC) of sporophytes during desiccation tolerance test.
+RWC was not measured for all combinations of species and desiccation treatments; only species with complete data shown.
+
+\clearpage
+
+```{r water-content-desiccated, fig.height = 4.5, fig.width = 7, fig.cap = ""}
+# Plot relative water content in desiccated state for all species
+rel_water_species_means_desic <<-
+rel_water_species_means %>%
+  filter(rec_time == "desiccated") %>%
+  select(-rec_time)
+
+# Set dodge spacing, colors
+pd <- position_dodge(.7)
+
+ggplot(rel_water_species_means_desic, aes(x = rel_water, y = species, group = salt)) +
+  geom_errorbar(
+    aes(xmin = rel_water - sd, xmax = rel_water + sd),
+    color = "grey80",
+    width = 0.2,
+    position=position_dodge(width=0.5)) +
+  geom_point(aes(fill = salt, shape = salt), size = 1.25, position=position_dodge(width=0.5)) +
+  scale_x_continuous(label = scales::percent) +
+  scale_shape_manual(name = "Desiccation Intensity",
+                     values = salt_shapes,
+                     breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                     labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
+  scale_fill_manual(name = "Desiccation Intensity",
+                    values = salt_cols, 
+                    breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                    labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) +
+  labs(
+    x = "Relative Water Content",
+    y = "Species",
+    color = "Desiccation\nintensity") +
+  theme_bw() +
+  theme(
+    panel.grid.minor.y = element_blank(),
+    panel.grid.major.y = element_blank(),
+    legend.position = "bottom"
+  ) +
+  facet_grid(vars(dry_time), labeller = as_labeller(c(`2` = "2 days", `15` = "15 days")))
+
+# ggsave(here::here("results/S_fig_water_content_desic.pdf"))
+```
+
+**Figure SXX**.
+Relative water content (RWC) of sporophytes after desiccation by duration of desiccation treatment (2 or 15 days).
+RWC was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not RWC of zero).
+
+\clearpage
+
+```{r sporo-dt-controls, fig.height = 8, fig.width = 7, fig.cap = ""}
+# Plot sporophyte DT controls
+
+# Prepare data for plotting
+sporo_control_plot_data <-
+  # Start with DT data
+  filmy_dt %>%
+  filter(salt == "Control", generation == "sporophyte") %>%
+  # Filter out individuals with low pre-treatment yields
+  filter(!is.na(yield_pre), yield_pre > 400) %>%
+  # Calculate recovery by individual
+  verify(yield_pre > 0) %>%
+  assert(not_na, yield_pre) %>%
+  select(species, salt, dry_time, individual, dataset, contains("yield")) %>%
+  select(-yield_72hr, -contains("yield_dry")) %>%
+  pivot_longer(names_to = "rec_time", values_to = "yield_recover", matches("30|24|48")) %>%
+  mutate(
+    rec_time = str_remove_all(rec_time, "yield_"),
+    recovery = yield_recover / yield_pre) %>%
+  select(-contains("yield")) %>%
+  # Calculate mean recovery by species and dataset
+  mutate(rec_time = factor(rec_time, levels = c("30min", "24hr", "48hr"))) %>%
+  group_by(species, salt, dry_time, rec_time, dataset) %>%
+  summarize(
+    recovery_mean = mean(recovery, na.rm = TRUE),
+    sd = sd(recovery, na.rm = TRUE),
+    n = n(),
+    .groups = "drop") %>%
+  rename(recovery = recovery_mean) %>%
+  # Add growth habit data
+  left_join(filmy_habit, by = "species") %>%
+  assert(not_na, habit) %>%
+  mutate(
+    habit = str_replace_all(habit, "_", " ") %>% 
+      str_to_sentence() %>%
+      factor(levels = c("Terrestrial", "Saxicolous", "Low epiphyte", "High epiphyte"))) %>%
+  # Add dataset year
+  mutate(year = str_split(dataset, "_") %>% map_chr(1)) %>%
+  # Abbreviate species names to "G. species"
+  mutate(
+    species = str_replace_all(species, "[a-z]+_", ". "),
+    species = glue("*{species}*")) %>%
+  # Add year to species names with multiple datasets
+  add_count(species, name = "n_species") %>%
+  mutate(species = ifelse(n_species > 6, paste(species, year), species)) %>%
+  select(-n_species) %>%
+  # Only keep rows with recovery
+  filter(!is.na(recovery))
+
+# Make plot
+ggplot(sporo_control_plot_data, aes(x = rec_time, y = recovery, shape = dry_time, fill = salt)) +
+  geom_point(position = pd, size = 1.5, fill = "white") +
+  geom_errorbar(
+    aes(ymin = recovery - sd, ymax = recovery + sd), 
+    colour = "grey50", width = 0, 
+    position = pd) +
+  facet_wrap(vars(habit, species)) +
+  scale_shape_manual(name = "Desiccation Time",
+                        values = c(24, 21),
+                        breaks = c("15", "2"),
+                        labels = c("15 days", "2 days")) +
+  scale_y_continuous(label = label_percent(accuracy = 1)) +
+  scale_x_discrete(limits = c("30min","24hr","48hr"),
+                   labels = c("0.5", "24", "48")) +
+  xlab("Recovery Time (hr)") +
+  ylab("Recovery (%)") +
+  theme_bw() +
+  legend_theme +
+  theme(legend.position ="bottom",
+        panel.grid.minor = element_blank(), 
+        panel.grid.major.x = element_blank(),
+        strip.text = element_markdown()
+)
+```
+
+**Figure SXX**.
+Recovery of controls (samples kept on moist tissues) for desiccation tolerance test of filmy fern sporophytes from Moorea, French Polynesia.
+Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation treatment (see Methods).
+Recovery was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not recovery of 0%).

### 2021-01-05 R/functions.R

@@ -676,6 +676,48 @@ calculate_mean_vpd_sporo <- function (community_matrix_raw, filmy_species, moore
     )
 }
 
+#' Calculate relative water content for sporophytes at the individual level
+#'
+#' @param data Filmy fern desiccation tolerance test data
+#'
+#' @return Dataframe
+#' 
+calculate_indiv_water <- function (data) {
+  data %>%
+    filter(generation == "sporophyte") %>%
+    select(-generation) %>% 
+    select(species:dataset, contains("weight")) %>%
+    pivot_longer(names_to = "rec_time", values_to = "weight", weight_desiccated:weight_48hr) %>%
+    mutate(rel_water_content =  (weight - weight_dry) / (weight_pre - weight_dry)) %>%
+    filter(!is.na(rel_water_content)) %>%
+    mutate(
+      rec_time = str_remove_all(rec_time, "weight_") %>%
+        factor(levels = c("desiccated", "30min", "24hr", "48hr"))) %>%
+    # Exclude outliers (presumably due to measurement error)
+    # rel water content should be between 0 to 100%
+    filter(rel_water_content < 1.25) %>%
+    filter(rel_water_content > -0.25) %>%
+    select(-weight_pre, -weight_dry)
+}
+
+#' Calculate relative water content for sporophytes at the species level
+#'
+#' @param data Filmy fern relative water content data at the individual level
+#' from desiccation tolerance test
+#'
+#' @return Dataframe
+#' 
+calculate_mean_water <- function (data) {
+  
+  data %>%
+    group_by(salt, species, rec_time, dry_time) %>%
+    summarize(
+      rel_water = mean(rel_water_content),
+      sd = sd(rel_water_content),
+      .groups = "drop"
+    )
+  
+}
 
 #' Make a dataframe combining range size, vpd, and recovery
 #' 
@@ -1189,6 +1231,11 @@ match_comm_and_tree <- function (comm, phy, return = c("comm", "tree")) {
 
 # Plots ----
 
+# Format legend (include title, but get rid of boxes)
+legend_theme <- theme (legend.background = element_rect(colour = "transparent", fill = "transparent", size = 0.5),
+                       legend.key = element_rect(colour = "white", fill = "white", size = 0.5),
+                       legend.text = element_text(size = 12))
+
 make_sporo_dt_plot <- function  (recovery_species_means, traits) {
   # Define color-blind palette
[…]

### 2021-01-05 R/packages.R

@@ -13,6 +13,7 @@ library(here)
 library(lubridate)
 library(patchwork)
 library(scales)
+library(ggtext)
 library(tidyverse)
 
 conflict_prefer("col_factor", "readr")

### 2021-01-05 _targets.R

@@ -58,6 +58,9 @@ tar_plan(
     filter(!is.na(yield_pre), yield_pre > 400) %>%
     calc_recovery(),
   
+  # - calculate water content during DT test per individual
+  rel_water_indiv = calculate_indiv_water(filmy_dt),
+  
   # - calculate mean VPD per datalogger
   mean_vpd = calculate_mean_vpd(climate),
   
@@ -104,6 +107,9 @@ tar_plan(
   
   # - calculate mean DT recovery by species and generation
   recovery_species_means = calc_mean_recovery(recovery_indiv),
+  
+  # - calculate relative water content by species (sporophytes only)
+  rel_water_species_means = calculate_mean_water(rel_water_indiv),
   
   # - calculate mean ETR max by species and generation
   etr_species_means = calculate_mean_etr(etr_indiv),

### 2021-01-05 ms/si.Rmd
Add gametophytes to DT control plot

@@
 # Prepare data for plotting
-sporo_control_plot_data <-
+dt_control_plot_data <-
   # Start with DT data
   filmy_dt %>%
-  filter(salt == "Control", generation == "sporophyte") %>%
+  filter(salt == "Control") %>%
   # Filter out individuals with low pre-treatment yields
@@
   assert(not_na, yield_pre) %>%
-  select(species, salt, dry_time, individual, dataset, contains("yield")) %>%
+  select(species, salt, dry_time, individual, dataset, generation, contains("yield")) %>%
   select(-yield_72hr, -contains("yield_dry")) %>%
@@
   mutate(rec_time = factor(rec_time, levels = c("30min", "24hr", "48hr"))) %>%
-  group_by(species, salt, dry_time, rec_time, dataset) %>%
+  group_by(species, salt, dry_time, rec_time, generation, dataset) %>%
   summarize(
@@
   add_count(species, name = "n_species") %>%
-  mutate(species = ifelse(n_species > 6, paste(species, year), species)) %>%
+  mutate(species = ifelse(n_species > 6 & generation == "sporophyte", paste(species, year), species)) %>%
   select(-n_species) %>%
@@
 
-# Make plot
-ggplot(sporo_control_plot_data, aes(x = rec_time, y = recovery, shape = dry_time, fill = salt)) +
+# Make plot: sporophyte controls
+a <-
+  dt_control_plot_data %>% 
+  filter(generation == "sporophyte") %>%
+  ggplot(aes(x = rec_time, y = recovery, shape = dry_time, fill = salt)) +
   geom_point(position = pd, size = 1.5, fill = "white") +
@@
   legend_theme +
-  theme(legend.position ="bottom",
+  theme(
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(), 
+    panel.grid.major.x = element_blank(),
+    strip.text = element_markdown(),
+    axis.title.x = element_blank()
+  )
+
+# Make plot: gametophyte controls
+b <-
+  dt_control_plot_data %>% 
+  filter(generation == "gametophyte") %>% 
+  ggplot(aes(x = rec_time, y = recovery, shape = dry_time, fill = salt)) +
+  geom_point(position = pd, size = 1.5, fill = "white") +
+  geom_errorbar(
+    aes(ymin = recovery - sd, ymax = recovery + sd), 
+    colour = "grey50", width = 0, 
+    position = pd) +
+  facet_wrap(vars(habit, species)) +
+  scale_shape_manual(name = "Desiccation Time",
+                        values = c(24, 21),
+                        breaks = c("15", "2"),
+                        labels = c("15 days", "2 days")) +
+  scale_y_continuous(label = label_percent(accuracy = 1)) +
+  scale_x_discrete(limits = c("30min","24hr","48hr"),
+                   labels = c("0.5", "24", "48")) +
+  xlab("Recovery Time (hr)") +
+  ylab("Recovery (%)") +
+  theme_bw() +
+  theme(
         panel.grid.minor = element_blank(), 
@@
         strip.text = element_markdown()
-)
+) +
+  # Use guides to make the legend go away
+  # otherwise it will be duplicated by patchwork
+  # https://github.com/thomasp85/patchwork/issues/170#issuecomment-641310103
+  guides(shape = "none")
+
+a + b + plot_layout(guides = "collect", ncol = 1, heights = c(85, 15)) + plot_annotation(tag_levels = "a") &
+  theme(legend.position='bottom')
 ```
@@
 **Figure SXX**.
-Recovery of controls (samples kept on moist tissues) for desiccation tolerance test of filmy fern sporophytes from Moorea, French Polynesia.
+Recovery of controls (samples kept on moist tissues) for desiccation tolerance test of filmy fern sporophytes (**a**) and gametophytes (**b**) from Moorea, French Polynesia.
 Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation treatment (see Methods).
 Recovery was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not recovery of 0%).
+Gametophytes only included a single individual per species thus lack error bars.

### 2021-01-06 ms/manuscript.Rmd
Use bare variable names for tar_load
or dependencies won't get tracked
https://github.com/wlandau/tarchetypes/issues/26

@@
 ```{r load-data}
-tar_load(
-  names = c(
-    "refs",
-    "refs_other",
-    "climate",
-    "filmy_species",
-    "moorea_sites",
-    "phylosig",
-    "range_types",
-    "combined_species_means",
-    "t_test_results",
-    "env_range_dt_model_summary"
-  )
-)
+tar_load(refs)
+tar_load(refs_other)
+tar_load(climate)
+tar_load(filmy_species)
+tar_load(moorea_sites)
+tar_load(phylosig)
+tar_load(range_types)
+tar_load(combined_species_means)
+tar_load(t_test_results)
+tar_load(env_range_dt_model_summary)
+tar_load(recovery_species_means)
+tar_load(filmy_habit)
 ```

### 2021-01-06 ms/si.Rmd
Use bare variable names for tar_load
or dependencies won't get tracked
https://github.com/wlandau/tarchetypes/issues/26

@@
 ```{r load-data, cache = FALSE}
-tar_load(
-  names = c(
-    "filmy_dt_chamber",
-    "filmy_dt",
-    "filmy_habit",
-    "rel_water_indiv",
-    "rel_water_species_means"
-  )
-)
+tar_load(filmy_dt_chamber)
+tar_load(filmy_dt)
+tar_load(filmy_habit)
+tar_load(rel_water_indiv)
+tar_load(rel_water_species_means)
 ```

### 2021-01-06 _targets.R
Move ms output to results/ms

@@ -170,18 +170,18 @@ tar_plan(
   tar_file(refs_other, "ms/references_other.yaml"),
 
   # Render PDF
-  tar_render(manuscript_pdf, "ms/manuscript.Rmd", output_dir = here::here("results")),
+  tar_render(manuscript_pdf, "ms/manuscript.Rmd", output_dir = here::here("results/ms")),
   
   # Render MS Word
   ms_docx = latex2docx(
-    latex = "results/manuscript.tex",
-    docx = "results/manuscript.docx",
+    latex = "results/ms/manuscript.tex",
+    docx = "results/ms/manuscript.docx",
     template = "ms/journal-of-plant-research.docx",
     wd = here::here("results"),
     depends = manuscript_pdf
   ),
   
   # SI
-  tar_render(si_pdf, "ms/si.Rmd", output_dir = here::here("results"))
+  tar_render(si_pdf, "ms/si.Rmd", output_dir = here::here("results/ms"))
   
 )

### 2021-01-06 ms/manuscript.Rmd
Add sporophyte DT figure

@@
 Mouaputa (880 m), and Mt.
-Tohiea (1206 m; Fig.
-2).
+Tohiea (1206 m).
 At each site, sporophytes were sampled in a single 10 m × 10 m plot, and gametophytes sampled in two to three 50 cm × 50 cm subplots located within the sporophyte plot.
 Epiphytes were sampled to a maximum height of \circa{ }2 m on their host plants.
-Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family [\eg, filamentous, ribbon with round gemmae; @Nayar1971].
-This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
 We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, \ie, below \circa{ }500 m), or high elevation epiphyte (mostly occurring in cloud forest, \ie, above \circa{ }500 m).
@@
 Gametophytes matching >99.5% with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched >99.5% with multiple reference sequences were not identified and excluded from further analysis.
+For this study, only gametophytes that could be identified as belonging to Hymenophyllaceae were used, and others were excluded from analysis.
 
@@
 
-## Figures
-
 ## References
@@
 <!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
+<div id="refs"></div>
+
+## Figure Captions
+
+**Fig. 1** Results of sporophyte desiccation tolerance test on filmy fern sporophytes from Moorea, French Polynesia.
+Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation at three different intensities for either 2 or 15 days (see Methods).
+Recovery was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not recovery of 0%).
+Error bars indicate standard deviation.
+Growth habit indicated in parentheses after each species: T, terrestrial; S, saxicolous; LE, low elevation epiphyte; HE, high elevation epiphyte
+
+## Figures
+
+```{r sporo-dt, fig.width = 6.85, fig.height = 7.5, fig.cap = ""}
+### Set up plotting variables ###
+
+# Define color-blind palette
+cbPalette <- c(grey = "#999999", gold = "#E69F00", skyblue = "#56B4E9", 
+               green = "#009E73", yellow = "#F0E442", blue = "#0072B2", 
+               red = "#D55E00", purple = "#CC79A7")
+
+# Set colors, shapes for desiccation intensity
+salt_cols <- c(
+  "NaCl" = cbPalette[["blue"]], 
+  "MgNO3" = cbPalette[["yellow"]], 
+  "LiCl" = cbPalette[["red"]])
+
+salt_shapes <- c(
+  "NaCl" = 21, # circle
+  "MgNO3" = 24, # triangle
+  "LiCl" = 23 # diamond
+)
+
+# Set offset of datapoints (dodge)
+pd <- position_dodge(.3)
+
+### Format data for plotting ###
+plot_data <-
+  recovery_species_means %>%
+  # Use only sporophytes
+  filter(generation == "sporophyte") %>%
+  # Add growth habit data
+  left_join(filmy_habit, by = "species") %>%
+  assert(not_na, habit) %>%
+  # Abbreviate growth habit, set as factor and specify levels
+  mutate(
+    habit = case_when(
+      habit == "High_Epiphyte" ~ "HE",
+      habit == "Low_Epiphyte" ~ "LE",
+      habit == "Saxicolous" ~ "S",
+      habit == "Terrestrial" ~ "T"
+    ) %>% factor(., levels = c("T", "S", "LE", "HE"))) %>%
+  # Format labels:
+  # - abbreviate species names to "G. species", make italic, add growth habit
+  mutate(
+    label = str_replace_all(species, "[a-z]+_", ". "),
+    label = glue("*{label}*"),
+    label = glue("{label} ({habit})")
+  ) %>%
+  # - order by alphabetical species within growth habit
+  arrange(habit, species) %>%
+  mutate(
+    order = 1:nrow(.),
+    # fct_reorder must reorder based on a numeric vector
+    label = fct_reorder(label, order)) %>%
+  # Tweak variable names
+  rename(recovery = recovery_mean, sd = recovery_sd, n = recovery_n)
+
+### Make plot ###
+ggplot(data = plot_data, aes(x = rec_time, y = recovery, group = interaction(dry_time, salt), shape = salt, fill = salt)) +
+  geom_line(position = pd, aes(linetype = dry_time)) +
+  geom_errorbar(
+    aes(ymin = recovery - sd, ymax = recovery + sd), 
+    colour = "grey50", width = 0, 
+    position = pd) +
+  geom_point(position = pd, size = 1.5) +
+  scale_linetype_manual(name = "Desiccation Time",
+                        values = c("solid", "dashed"),
+                        breaks = c("15", "2"),
+                        labels = c("15 days", "2 days")) +
+  scale_shape_manual(name = "Desiccation Intensity",
+                     values = salt_shapes,
+                     breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                     labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
+  scale_fill_manual(name = "Desiccation Intensity",
+                    values = salt_cols, 
+                    breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                    labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) +
+  scale_y_continuous(label = label_percent(accuracy = 1), limits = c(-0.1,1.1), expand = c(0,0), breaks=c(0,.2,.4,.6,.8,1.0)) +
+  scale_x_discrete(limits = c("30min","24hr","48hr"),
+                   labels = c("0.5", "24", "48")) +
+  xlab("Recovery Time (hr)") +
+  ylab("Recovery (%)") +
+  theme_bw() +
+  theme(
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(), 
+    panel.grid.major.x = element_blank(),
+    # Use markdown to format facet strip labels
+    strip.text = element_markdown(),
+    # Wrap legend parts into multiple lines
+    legend.box = "vertical", 
+    legend.margin = margin()
+  ) + 
+  facet_wrap(~label)
+
+# For JPR:
+# 39 mm, 84 mm, 129 mm, or 174 mm wide
+# not higher than 234 mm
+ggsave(here::here("results/ms/fig_1_sporo_dt.eps"), width = 6.85, height = 7.5)
+```
+
+**Fig. 1**

### 2021-01-06 ms/manuscript.Rmd
Add captions using captioner

@@
 Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
-A general correlation of DT with habitat was observed in sporophytes (Fig. 4).
+A general correlation of DT with habitat was observed in sporophytes (`r figure("sporo-dt")`).
 Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats.
@@
 
-Acknowledgments
+**Acknowledgments**
 
@@
 ## Figure Captions
+<!-- JPR requirements: Figure captions begin with the term Fig.
+in bold type, followed by the figure number, also in boldtype.
+No punctuation is to be included after the number, nor is any punctuation to be placed at the end of the caption.
+-->
 
-**Fig. 1** Results of sporophyte desiccation tolerance test on filmy fern sporophytes from Moorea, French Polynesia.
+**`r figure("sporo-dt")`** Results of sporophyte desiccation tolerance test on filmy fern sporophytes from Moorea, French Polynesia.
 Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation at three different intensities for either 2 or 15 days (see Methods).
@@
 
-**Fig. 1**
+**`r figure("sporo-dt")`**

### 2021-01-06 _targets.R

@@ -1,5 +1,6 @@
 source("R/packages.R")
 source("R/functions.R")
+source("ms/captions.R")
 
 tar_plan(
   

### 2021-01-06 ms/captions.R

@@ -1,0 +1,30 @@
+# Set up captions
+
+# Follow Journal of Plant Research (JPR) style:
+# - All figures are to be numbered using Arabic numerals.
+# - Figures should always be cited in text in consecutive numerical order.
+# - Figure parts should be denoted by lowercase letters (a, b, c, etc.).
+# - If an appendix appears in your article and it contains one or more figures, 
+#   continue the consecutive numbering of the main text. 
+# - Do not number the appendix figures, "A1, A2, A3, etc." 
+# - Figures in online appendices (Electronic Supplementary Material) should, 
+#   however, be numbered separately
+
+# - Figures
+figure_full <- captioner::captioner(prefix = "Fig.", suffix = "")
+figure_full(name = "sporo-dt", caption = "sporo-dt")
+
+# - Tables
+table_full <- captioner::captioner(prefix = "Table")
+
+# - SI figures
+s_figure_full <- captioner::captioner(prefix = "Fig. S", suffix = "")
+
+# - SI tables
+s_table_full <- captioner::captioner(prefix = "S", auto_space = FALSE, suffix = " Table. ")
+
+# Make short versions of citation functions (just the number)
+figure <- pryr::partial(figure_full, display = "cite")
+table <- pryr::partial(table_full, display = "cite")
+s_figure <- pryr::partial(s_figure_full, display = "cite")
+s_table <- pryr::partial(s_table_full, display = "cite")

### 2021-01-07 ms/manuscript.Rmd
Split dt figure into two parts

--- 2021-01-06 ms/manuscript.Rmd
@@
 <!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
+<div id="refs"></div>
+
+## Figure Captions
+<!-- JPR requirements: Figure captions begin with the term Fig.
+in bold type, followed by the figure number, also in boldtype.
+No punctuation is to be included after the number, nor is any punctuation to be placed at the end of the caption.
+-->
+
+**`r figure("sporo-dt")`** Results of sporophyte desiccation tolerance test on filmy fern sporophytes from Moorea, French Polynesia.
+Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation at three different intensities for either 2 or 15 days (see Methods).
+Recovery was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not recovery of 0%).
+Error bars indicate standard deviation.
+Growth habit indicated in parentheses after each species: T, terrestrial; S, saxicolous; LE, low elevation epiphyte; HE, high elevation epiphyte
+
+## Figures
+
+```{r dt-plot-setup}
+### Format data for plotting ###
+
+# Format labels for number of samples per species/generation
+n_labels <-
+recovery_species_means %>%
+  group_by(species, generation) %>%
+  mutate(n_count = n_distinct(recovery_n)) %>%
+  mutate(
+    n_label = case_when(
+      # no label if exactly 8
+      n_count == 1 & recovery_n == 8 ~ "",
+      # single number if no. samples is NOT 8 and consistent across treatments
+      n_count == 1 & recovery_n != 8 ~ as.character(glue(", *n* = {recovery_n}")),
+      # range if no. samples is NOT 8 and varies across treatments
+      n_count > 1 ~ as.character(glue(", *n* = {min(recovery_n)}-{max(recovery_n)}"))
+    )) %>%
+  select(species, generation, n_label) %>%
+  unique() %>%
+  ungroup()
+
+dt_plot_data <-
+  recovery_species_means %>%
+  # Add growth habit data
+  left_join(filmy_habit, by = "species") %>%
+  assert(not_na, habit) %>%
+  # Add labels for number of samples
+  left_join(n_labels, by = c("species", "generation")) %>%
+  # Abbreviate growth habit, set as factor and specify levels
+  mutate(
+    habit = case_when(
+      habit == "High_Epiphyte" ~ "HE",
+      habit == "Low_Epiphyte" ~ "LE",
+      habit == "Saxicolous" ~ "S",
+      habit == "Terrestrial" ~ "T"
+    ) %>% factor(., levels = c("T", "S", "LE", "HE"))) %>%
+  # Format labels:
+  # - abbreviate species names to "G. species", make italic, add growth habit
+  #   and sample number if not 8 
+  mutate(
+    label = str_replace_all(species, "[a-z]+_", ". "),
+    label = glue("*{label}* ({habit}{n_label})")
+  ) %>%
+  # - order by alphabetical species within growth habit
+  arrange(habit, species) %>%
+  mutate(
+    order = 1:nrow(.),
+    # fct_reorder must reorder based on a numeric vector
+    label = fct_reorder(label, order)) %>%
+  # Tweak variable names
+  rename(recovery = recovery_mean, sd = recovery_sd, n = recovery_n)
+
+### Set up plotting variables ###
+
+# Define color-blind palette
+cbPalette <- c(grey = "#999999", gold = "#E69F00", skyblue = "#56B4E9", 
+               green = "#009E73", yellow = "#F0E442", blue = "#0072B2", 
+               red = "#D55E00", purple = "#CC79A7")
+
+# Set colors, shapes for desiccation intensity
+salt_cols <- c(
+  "NaCl" = cbPalette[["blue"]], 
+  "MgNO3" = cbPalette[["yellow"]], 
+  "LiCl" = cbPalette[["red"]])
+
+salt_shapes <- c(
+  "NaCl" = 21, # circle
+  "MgNO3" = 24, # triangle
+  "LiCl" = 23 # diamond
+)
+
+# Set offset of datapoints (dodge)
+pd <- position_dodge(.3)
+
+### Define function for making dt plot ###
+make_dt_plot <- function (dt_plot_data, generation_select, plot_part) {
+  dt_plot_data %>% filter(generation == generation_select) %>%
+  ggplot(aes(x = rec_time, y = recovery, group = interaction(dry_time, salt), shape = salt, fill = salt)) +
+  geom_line(position = pd, aes(linetype = dry_time)) +
+  geom_errorbar(
+    aes(ymin = recovery - sd, ymax = recovery + sd), 
+    colour = "grey50", width = 0, 
+    position = pd) +
+  geom_point(position = pd, size = 1.5) +
+  scale_linetype_manual(name = "Desiccation Time",
+                        values = c("solid", "dashed"),
+                        breaks = c("15", "2"),
+                        labels = c("15 days", "2 days")) +
+  scale_shape_manual(name = "Desiccation Intensity",
+                     values = salt_shapes,
+                     breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                     labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
+  scale_fill_manual(name = "Desiccation Intensity",
+                    values = salt_cols, 
+                    breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                    labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) +
+  scale_y_continuous(label = label_percent(accuracy = 1), limits = c(-0.1,1.2), expand = c(0,0), breaks=c(0,.25,.5,.75,1.0)) +
+  scale_x_discrete(limits = c("30min","24hr","48hr"),
+                   labels = c("0.5", "24", "48")) +
+  xlab("Recovery Time (hr)") +
+  ylab("Recovery (%)") +
+  labs(
+    x = "Recovery Time (hr)",
+    y = "Recovery (%)",
+    title = glue("**{plot_part}**")
+  ) +
+  theme_bw() +
+  theme(
+    legend.position = "bottom",
+    plot.title = element_markdown(),
+    panel.grid.minor = element_blank(), 
+    panel.grid.major.x = element_blank(),
+    # Use markdown to format facet strip labels
+    strip.text = element_markdown(),
+    # Wrap legend parts into multiple lines
+    legend.box = "vertical", 
+    legend.margin = margin()
+  ) + 
+  facet_wrap(~label, ncol = 4)
+}
+```
+
+```{r sporo-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = ""}
+make_dt_plot(dt_plot_data, "sporophyte", "a")
+ggsave(here::here("results/ms/fig_1a_sporo_dt.pdf"), width = 6.85, height = 7.5)
+```
+
+**`r figure("sporo-dt")`a**
+
+```{r gameto-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = ""}
+make_dt_plot(dt_plot_data, "gametophyte", "b")
+ggsave(here::here("results/ms/fig_1b_gameto_dt.pdf"), width = 6.85, height = 6.5)
+```
+
+**`r figure("sporo-dt")`b**

### 2021-01-08 ms/manuscript.Rmd
Set overall plot theme, fix facet labels in DT plot

@@
 
-**`r figure("sporo-dt")`** Results of sporophyte desiccation tolerance test on filmy fern sporophytes from Moorea, French Polynesia.
+**`r figure("sporo-dt")`** Results of desiccation tolerance test on filmy fern sporophytes (**a**) and gametophytes (**b**) from Moorea, French Polynesia.
 Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation at three different intensities for either 2 or 15 days (see Methods).
@@
 Error bars indicate standard deviation.
+*n* = 8 samples per species/treatment unless otherwise indicated.
 Growth habit indicated in parentheses after each species: T, terrestrial; S, saxicolous; LE, low elevation epiphyte; HE, high elevation epiphyte
@@
 ## Figures
+
+```{r plot-setup}
+original_theme <- theme_set(theme_bw(base_size = 10))
+```
 
@@
   ) +
-  theme_bw() +
   theme(
@@
 ```{r sporo-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = ""}
-make_dt_plot(dt_plot_data, "sporophyte", "a")
+dt_plot_data %>%
+  mutate(label = str_replace_all(
+    label, 
+    fixed("*C. bipunctatum* (LE, *n* = 10-12)"),
+    "*C. bipunctatum*<br>(LE, *n* = 10-12)")) %>%
+make_dt_plot("sporophyte", "a")
 ggsave(here::here("results/ms/fig_1a_sporo_dt.pdf"), width = 6.85, height = 7.5)
@@
 
-**`r figure("sporo-dt")`a**
+**`r figure("sporo-dt")`**
 
@@
 
-**`r figure("sporo-dt")`b**
+**`r figure("sporo-dt")`** continued

### 2021-01-08 R/process_refs.R
Update references

@@ -21,7 +21,7 @@ citations <-
   unique %>% sort
 
 # Read in entire reference library exported from Zotero with BetterBibTex
-# (note: this can be a soft-link to ~/Dropbox/references/main_library.bib)
+# (note: this can be a soft-link to ~/Dropbox/bibliography/main_library.bib)
 bib <- RefManageR::ReadBib("data_raw/main_library.bib", check = FALSE)
 
 # Convert reference library to dataframe

### 2021-01-08 ms/manuscript.Rmd
Add figure 1 (filmy fern growth habit)

@@
 
-\clearpage
-#####
+\newpage
 
@@
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
-A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence].
+A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (`r figure("filmy-habit")`).
 The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
@@
 
-# Make helper function to check on p-value significance and extract from t_test_results
-#
-#' @param species_select Name of species to select from t-test results
-#' @param response_select Name of response to select from t-test results
-#' @param data t-test results
-#' @param p_check Boolean: should the p-value be required to be less than 0.05?
-#'
-#' @return A p-value for the selected species and response
-ttest_pval <- function(species_select, response_select, data = t_test_results, p_check = TRUE) {
-  data %>% 
-    # Filter to appropriate species and response
-    filter(species == species_select, response == response_select) %>%
-    # Verify the p-value is significant (or override with `p_check = FALSE`)
-    assert(function (x) magrittr::is_less_than(x, 0.05) | !p_check, p_value) %>%
-    # Format the p-value
-    mutate(
-      p_value = case_when(
-        p_value < 0.001 ~ "<0.001",
-        TRUE ~ jntools::round_t(p_value, digits = 3))
-    ) %>% 
-    pull(p_value)
-}
-
 # FIXME: double check on these p-values
@@
 Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
-A general correlation of DT with habitat was observed in sporophytes (`r figure("sporo-dt")`).
+A general correlation of DT with habitat was observed in sporophytes (`r figure("filmy-dt")`).
 Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats.
@@
 
+\newpage
+
 ## Figure Captions
@@
 
-**`r figure("sporo-dt")`** Results of desiccation tolerance test on filmy fern sporophytes (**a**) and gametophytes (**b**) from Moorea, French Polynesia.
+**`r figure("filmy-habit")`** Examples of filmy ferns from French Polynesia.
+(**a**–**d**) sporophytes; (**e**–**h**) gametophytes.
+(**a**) *Crepidomanes humile*, a low elevation epiphyte. *Nitta 3392* (GH). 
+(**b**) *Hymenophyllum multifidum*, a high elevation epiphyte growing in association with leafy liverworts and mosses. *Nitta 1257*
+(GH). 
+(**c**) *Polyphlebium endlicherianum*, a saxicolous species found at low to middle elevations. *Nitta 629* (GH). 
+(**d**) *Callistopteris apiifolia*, a terrestrial species with sporophytes occurring only in humid cloud forests. *Nitta 3990* (GH).
+(**e**) *C. apiifolia* (ribbon morphotype). Arrow indicates gemmae (asexual propagules) at tip of lobe. *Nitta 2565* (GH). 
+(**f**) *Hymenophyllum polyanthos* (ribbon morphotype). *Nitta 3862* (GH).
+(**g**) Filamentous gametophytes of an unidentified species of Hymenophyllaceae. *Nitta 4036* (GH). 
+(**h**) *C. apiifolia*, showing detail of gemmifers (attachment points of gemmae, example indicated by arrow) and branched gemmae. *Nitta 3887* (GH). 
+Scalebars: (**a**–**c**) 1.0 cm; (**d**) 5.0 cm; (**e**) 5.0 mm; (**f**) 2.0 mm; (**g**, **h**) 1.0 mm. 
+Photographs by J. H. Nitta.
+
+**`r figure("filmy-dt")`** Results of desiccation tolerance test on filmy fern sporophytes (**a**) and gametophytes (**b**) from Moorea, French Polynesia.
 Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation at three different intensities for either 2 or 15 days (see Methods).
@@
 
-## Figures
+\newpage
 
@@
 original_theme <- theme_set(theme_bw(base_size = 10))
+# JPR requirements:
+# Figures should be 39 mm, 84 mm, 129 mm, or 174 mm wide and not higher than 234 mm
+# In inches, widths: 1.53, 3.31, 5.08, 6.85; height: 9.21
 ```
@@
 
-```{r sporo-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = ""}
+```{r filmies-fig, fig.width = 6.85, results = "asis"}
+knitr::include_graphics(here::here("data/figure_1.png"))
+```
+
+**`r figure("filmy-habit")`**
+
+\newpage
+
+```{r filmy-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = ""}
 dt_plot_data %>%
@@
 
-**`r figure("sporo-dt")`**
+**`r figure("filmy-dt")`**
+
+\newpage
 
@@
 
-**`r figure("sporo-dt")`** continued
+**`r figure("filmy-dt")`** (continued)
+
+\newpage

### 2021-01-08 R/functions.R

@@ -1447,8 +1447,30 @@ analyze_dist_pattern <- function(community_matrix_raw, filmy_species, phy, moore
   
 }
 
-# Manuscript rendering ----
+# Manuscript ----
 
+# Helper function to check on p-value significance and extract from t_test_results
+#
+#' @param species_select Name of species to select from t-test results
+#' @param response_select Name of response to select from t-test results
+#' @param data t-test results
+#' @param p_check Boolean: should the p-value be required to be less than 0.05?
+#'
+#' @return A p-value for the selected species and response
+ttest_pval <- function(species_select, response_select, data = t_test_results, p_check = TRUE) {
+  data %>% 
+    # Filter to appropriate species and response
+    filter(species == species_select, response == response_select) %>%
+    # Verify the p-value is significant (or override with `p_check = FALSE`)
+    assert(function (x) magrittr::is_less_than(x, 0.05) | !p_check, p_value) %>%
+    # Format the p-value
+    mutate(
+      p_value = case_when(
+        p_value < 0.001 ~ "<0.001",
+        TRUE ~ jntools::round_t(p_value, digits = 3))
+    ) %>% 
+    pull(p_value)
+}
 
 # Dummy function to track arbitary output from rmarkdown::render()
 render_tracked <- function (tracked_output, dep1 = NULL, dep2 = NULL, ...) {

### 2021-01-08 ms/captions.R

@@ -12,7 +12,8 @@
 
 # - Figures
 figure_full <- captioner::captioner(prefix = "Fig.", suffix = "")
-figure_full(name = "sporo-dt", caption = "sporo-dt")
+figure_full(name = "filmy-habit", caption = "filmy-habit")
+figure_full(name = "filmy-dt", caption = "filmy-dt")
 
 # - Tables
 table_full <- captioner::captioner(prefix = "Table")

### 2021-01-13 ms/manuscript.Rmd
Add combined phylogeny + trait data figure

@@
 tar_load(filmy_habit)
+tar_load(filmy_phy)
 ```
@@
 ```{r plot-setup}
-original_theme <- theme_set(theme_bw(base_size = 10))
 # JPR requirements:
@@
 # In inches, widths: 1.53, 3.31, 5.08, 6.85; height: 9.21
+
+# Set overall theme, saving default theme
+default_theme <- theme_set(theme_bw(base_size = 10))
+
+# Define color-blind palette
+cbPalette <- c(grey = "#999999", gold = "#E69F00", skyblue = "#56B4E9", 
+               green = "#009E73", yellow = "#F0E442", blue = "#0072B2", 
+               red = "#D55E00", purple = "#CC79A7")
+
+# Define function for formatting text in legends
+pretty_text <- function (x) {
+  x %>%
+  str_replace_all("_", " ") %>%
+    str_to_sentence()
+}
 ```
@@
 ### Set up plotting variables ###
-
-# Define color-blind palette
-cbPalette <- c(grey = "#999999", gold = "#E69F00", skyblue = "#56B4E9", 
-               green = "#009E73", yellow = "#F0E442", blue = "#0072B2", 
-               red = "#D55E00", purple = "#CC79A7")
 
@@
 \newpage
+
+```{r combined-plot, fig.width = 6.85, fig.height = 5, fig.cap = ""}
+# Copy phylogeny and species means data so they
+# can be manipulated for this plot
+phy <- filmy_phy
+traits <- combined_species_means
+
+# Set colors for growth habit, generation
+habit_colors <- c(
+  "Terrestrial" = cbPalette[["red"]], 
+  "Saxicolous" = cbPalette[["gold"]], 
+  "Low_Epiphyte" = cbPalette[["green"]],
+  "High_Epiphyte" = cbPalette[["blue"]])
+
+generation_colors <- c(
+  "gametophyte" = "grey70",
+  "sporophyte" = "grey30"
+)
+
+# Subset tree to only species with trait data
+phy <- drop.tip(phy, setdiff(phy$tip.label, traits$species))
+
+# Sort species by position in phylogeny
+traits <-
+  traits %>%
+  filter(species %in% phy$tip.label) %>%
+  mutate(species = factor(species, levels = jntools::get_tips_in_ape_plot_order(phy)))
+
+filmy_habit <-
+  filmy_habit %>%
+  filter(species %in% phy$tip.label)
+
+# Part a (left): tree
+a <- ggtree(phy) %<+% filmy_habit +
+  geom_tippoint(aes(color = habit), position = position_nudge(x = 7, y = 0), size = 4) +
+  scale_color_manual(values = habit_colors, labels = pretty_text) +
+  labs(color = "Growth habit")
+
+# Part b (middle): desiccation tolerance
+b <- 
+ggplot(traits, aes(y = species, x = recovery_mean, fill = generation)) +
+  geom_col(position = position_dodge(0.5, preserve = "single"), width = 0.5) +
+  geom_errorbarh(
+    aes(xmax = recovery_mean + recovery_sd, xmin = recovery_mean),
+    position = position_dodge(0.5, preserve = "single"),
+    height = 0,
+    color = "grey20") +
+  scale_x_continuous(
+    labels = percent, 
+    limits = c(0,1.2), 
+    expand = expansion(mult = c(0,0.025)), 
+    breaks=c(0,.25,.5,.75,1.0)) +
+  scale_fill_manual(values = generation_colors, labels = pretty_text) +
+  theme_bw(base_size = 10) +
+  labs(x = "Recovery", fill = "Generation") +
+  theme(
+    panel.grid.minor = element_blank(), 
+    panel.grid.major.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom"
+  )
+
+# Part c (right): light-levels
+c <-
+ggplot(traits, aes(y = species, x = etr_mean, fill = generation)) +
+  geom_col(position = position_dodge(0.5, preserve = "single"), width = 0.5) +
+  geom_errorbarh(
+    aes(xmax = etr_mean + etr_sd, xmin = etr_mean),
+    position = position_dodge(0.5, preserve = "single"),
+    height = 0,
+    color = "grey20") +
+  theme_bw(base_size = 10) +
+  scale_fill_manual(values = generation_colors, labels = pretty_text) +
+  scale_y_discrete(
+    position = "right",
+    # Format species names by converting underscore to space and italicizing
+    labels = function(x) str_replace_all(x, "_", " ") %>% paste0("*", ., "*")
+    ) +
+  scale_x_continuous(expand = expansion(mult = c(0,0.025))) +
+  labs(x = "ETR<sub>max</sub>", fill = "Generation") +
+  theme(
+    panel.grid.minor = element_blank(), 
+    panel.grid.major.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y.right = element_markdown(), # NOT axis.text.y
+    axis.title.x = element_markdown()
+  ) +
+  guides(fill = "none")
+
+# Combine plots
+a + b + c + plot_layout(nrow = 1, guides = "collect") &
+  theme(
+    legend.position = "bottom",
+    # Adjust invididual margins to remove whitespace between subplots
+    plot.margin = margin(t = 0.1, r = 0, b = 0.1, l = 0.04, unit = "in"))
+
+ggsave(here::here("results/ms/fig_combined.pdf"), width = 6.85, height = 5)
+```
+
+**`r figure("combined")`**

### 2021-01-13 R/packages.R

@@ -14,6 +14,7 @@ library(lubridate)
 library(patchwork)
 library(scales)
 library(ggtext)
+library(ggtree)
 library(tidyverse)
 
 conflict_prefer("col_factor", "readr")

### 2021-01-13 ms/captions.R

@@ -14,6 +14,7 @@
 figure_full <- captioner::captioner(prefix = "Fig.", suffix = "")
 figure_full(name = "filmy-habit", caption = "filmy-habit")
 figure_full(name = "filmy-dt", caption = "filmy-dt")
+figure_full(name = "combined", caption = "combined")
 
 # - Tables
 table_full <- captioner::captioner(prefix = "Table")

### 2021-01-25 ms/manuscript.Rmd
Use continuous range of gametos beyond sporos
instead of classifying into 'widespread' vs 'not widespread'

resolves issue https://github.com/joelnitta/moorea_filmies/issues/1

@@
 tar_load(phylosig)
-tar_load(range_types)
+tar_load(range_comparison)
 tar_load(combined_species_means)
@@
 
-In total, `r n_distinct(range_types$species)` filmy fern taxa (including two species that were recognized as multiple varieties) were observed on Moorea.
+In total, `r n_distinct(range_comparison$species)` filmy fern taxa (including two species that were recognized as multiple varieties) were observed on Moorea.
 Three taxa with distinct *rbcL* sequences, morphology, and elevational ranges could be distinguished within the *Crepidomanes minutum* (Blume) K.

### 2021-01-25 R/functions.R

@@ -816,8 +816,7 @@ combine_env_env_range_recover <- function (combined_species_means, env_range_dat
            species, 
            sporo_vpd, 
            gameto_vpd,
-           sporo_range_breadth,
-           gameto_range_breadth
+           gameto_beyond_sporo
     ), 
     by = "species") %>%
   as.data.frame()
@@ -1178,8 +1177,7 @@ run_pgls <- function (env_range_recover_data, phy) {
   res <- list(
     sporo_vpd_model = caper::pgls(recovery_sporophyte ~ sporo_vpd, env_range_recover_data_comp),
     gameto_vpd_model = caper::pgls(recovery_gametophyte ~ gameto_vpd, env_range_recover_data_comp),
-    sporo_range_model = caper::pgls(recovery_sporophyte ~ sporo_range_breadth, env_range_recover_data_comp),
-    gameto_range_model = caper::pgls(recovery_gametophyte ~ gameto_range_breadth, env_range_recover_data_comp)
+    gameto_range_model = caper::pgls(recovery_gametophyte ~ gameto_beyond_sporo, env_range_recover_data_comp)
     )
   
 }
@@ -1437,15 +1435,17 @@ analyze_dist_pattern <- function(community_matrix_raw, filmy_species, phy, moore
   ) %>%
     assert(is_uniq, species) %>%
     mutate(
-      range = case_when(
-        gameto_max_range > sporo_max_range + 200 ~ "widespread",
-        gameto_min_range < sporo_min_range - 200 ~ "widespread",
-        is.na(sporo_max_range) & is.na(sporo_min_range) ~ "widespread",
-        TRUE ~ "not_widespread"
-      )
+      # Calculate elevational range of gametophytes beyond sporophytes
+      gameto_beyond_sporo_min = sporo_min_range - gameto_min_range,
+      gameto_beyond_sporo_max = gameto_max_range - sporo_max_range,
+      # If observed sporo range exceeds observed gameto range, consider
+      # range of gameto beyond sporo to be zero
+      gameto_beyond_sporo_min = ifelse(gameto_beyond_sporo_min < 0, 0, gameto_beyond_sporo_min),
+      gameto_beyond_sporo_max = ifelse(gameto_beyond_sporo_max < 0, 0, gameto_beyond_sporo_max)
     ) %>%
-    mutate(gameto_range_breadth = gameto_max_range - gameto_min_range) %>%
[…]

### 2021-01-25 _targets.R

@@ -83,8 +83,8 @@ tar_plan(
     filmy_species = filmy_species
   ),
   
-  # - determine which species have widespread gametophytes
-  range_types = analyze_dist_pattern(
+  # - determine range of gametophytes beyond sporophytes
+  range_comparison = analyze_dist_pattern(
     community_matrix_raw = community_matrix_raw,
     filmy_species = filmy_species,
     phy = filmy_phy,
@@ -96,7 +96,7 @@ tar_plan(
     list(
       mean_vpd_gameto,
       mean_vpd_sporo,
-      range_types),
+      range_comparison),
     left_join,
     by = "species"
   ),
@@ -153,7 +153,7 @@ tar_plan(
   # PGLS ----
   # (Phylogenetic Generalized Least Squares)
   
-  # Run PGLS for gametophyte range size vs. desiccation tolerance
+  # Run PGLS for VPD and gametophyte range breadth vs. desiccation tolerance
   env_range_recover_data = combine_env_env_range_recover(
     combined_species_means = combined_species_means,
     env_range_data = env_range_data

### 2021-01-25 ms/manuscript.Rmd
Add dt vs environment figure

@@
 tar_load(filmy_phy)
+tar_load(mean_vpd_sporo)
+tar_load(mean_vpd_gameto)
+
 ```
@@
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
-A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (`r figure("filmy-habit")`).
+A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (`r figure("habit")`).
 The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
@@
 The dataset includes `r climate_n_days` days of data from `r climate_n_sites` sites (`r climate_n_loggers` dataloggers).
+Mean daily temperature on Moorea is strongly correlated with elevation, dropping \circa{ }6 °C from the lowest surveyed site (201 m) to the highest (1206 m); temperature does not differ significantly between epiphytic and terrestrial habitats [@Nitta2020].
+However, epiphytic habitats tend to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones.
+This difference is greatest at low elevations, and smallest at high elevations.
 
@@
 
-In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C.
+In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes  including *Callistopteris apiifolia* (C.
 Presl) Copel., *Crepidomanes minutum* var.
@@
 We include these observations for comparison of ranges between sporophytes and gametophytes.
-
-### Microclimate
-
-Mean daily temperature was strongly correlated with elevation, and dropped ca.
-6 °C from the lowest surveyed site (201 m) to the highest (1206 m; Fig. 3).
-Although epiphytic and terrestrial habitats were similar in temperature, epiphytic habitats tended to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones (Fig.
-3, Appendix S2).
-This difference was greatest at low elevations, and smallest at high elevations.
-One significant departure from these trends was observed at the Mt. Rotui 830 m site (indicated by arrows in Fig. 3).
-This site is much more exposed than the other cloud forest sites (Fig. 2), and many of the plants on the N facing slope are stunted (J. Nitta, pers. obs.).
-Daily variation in temperature and RH was much greater, and mean RH much lower, at this site than other high elevation sites.
 
@@
 Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
-A general correlation of DT with habitat was observed in sporophytes (`r figure("filmy-dt")`).
+A general correlation of DT with habitat was observed in sporophytes (`r figure("dt")`**a**).
 Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats.
 and *C.
-apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 MPa (more stringent tests were not conducted on these species after confirming their lack of DT at this level).
+apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 MPa (more stringent tests were not conducted on *A.
+dentatum* after confirming its lack of DT at this level).
 Saxicolous species (*Crepidomanes kurzii* (Bedd.) Tagawa & K.
@@
 
+Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; `r figure("dt")`**b**).
+
 DT of gametophytes did not significantly exceed that of sporophytes for any species except for saxicolous *P.
@@
 5; *t*-test, *P* = XXX, XXX, XXX, and XXX, respectively).
-
-Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; Fig.
-5).
 
@@
 
+## Tables
+
+**`r table("glmm")`**
+
 ## Figure Captions
@@
 
-**`r figure("filmy-habit")`** Examples of filmy ferns from French Polynesia.
+**`r figure("habit")`** Examples of filmy ferns from French Polynesia.
 (**a**–**d**) sporophytes; (**e**–**h**) gametophytes.
@@
 
-**`r figure("filmy-dt")`** Results of desiccation tolerance test on filmy fern sporophytes (**a**) and gametophytes (**b**) from Moorea, French Polynesia.
+**`r figure("dt")`** Desiccation tolerance of filmy fern sporophytes (**a**) and gametophytes (**b**) from Moorea, French Polynesia.
 Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation at three different intensities for either 2 or 15 days (see Methods).
@@
 Growth habit indicated in parentheses after each species: T, terrestrial; S, saxicolous; LE, low elevation epiphyte; HE, high elevation epiphyte
+
+**`r figure("combined")`** Phylogeny of filmy ferns from Moorea, French Polynesia with comparison of physiological parameters (mean ± 1 SD) between sporophytes (dark grey bars) and gametophytes (light grey bars).
+Phylogeny adapted from Nitta et al. (2016); only species with physiological data shown.
+Bootstrap support >50% from 100 ML bootstraps indicated at nodes.
+Bold letters indicate major clades (H = hymenophylloid clade, T = trichomanoid clade).
+Growth habit indicated by colored symbols (blue = low elevation epiphyte, green = high elevation epiphyte, yellow = saxicolous, red = terrestrial).
+**Crepidomanes bipunctatum** and **C. humile** were observed growing as both epiphytes and saxicoles, but are shown as epiphytes to distinguish them from exclusively saxicolous species.
+Asterisk indicates significant difference in means (**P** <0.05) between generations for a particular species (**t**-test); ns indicates no significant difference; lack of ns or asterisk means not enough observations available for that species for **t**-test.
+Left bar plot shows recovery (%) of chlorophyll fluorescence (Fv/Fm) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa). 
+Right bar plot shows maximum relative electron transport rate of photosystem II (ETR~max~). 
+The other photosynthetic parameter measured, photosynthetic photon flux density at 95% saturation of RETR (PPFD~95%~), was qualitatively very similar to ETR~max~ and not shown.
+Absence of a bar indicates measurements not available for gametophytes or sporophytes of that species.
+
+**`r figure("corr")`** Recovery of F~v~/F~m~ (%) following 2 d desiccation at –86 MPa plotted against environmental vapor pressure deficit (VPD; kPa) (**a**, **b**) or range of gametophytes beyond sporophytes (m) (**c**) for gametophytes (**a**, **c**) or sporophytes (**b**).
+Species means ± 1 SD are shown. 
+Growth habit indicated by colored symbols (blue = low elevation epiphyte, green = high elevation epiphyte, yellow = saxicolous, red = terrestrial). 
+**Crepidomanes bipunctatum** and **C. humile** were observed growing as both epiphytes and saxicoles, but were treated as epiphytes to distinguish them from exclusively saxicolous species.
+**Abrodictyum dentatum** sporophytes were desiccated at –38 MPa instead of –86 MPa.
+Range of gametophytes beyond sporophytes calculated as the total elevation that gametophytes exceeded sporophytes at either end of the elevational gradient.
+Relationships between recovery and VPD or range of gametophytes beyond sporophytes were analyzed with phylogenetic generalized least squares (PGLS); no significant relationships were detected.
 
@@
 # Set overall theme, saving default theme
-default_theme <- theme_set(theme_bw(base_size = 10))
+default_theme <- theme_set(
+  theme_bw(base_size = 10) +
+    theme(
+      panel.grid.minor = element_blank()
+    ))
 
@@
 
-**`r figure("filmy-habit")`**
+**`r figure("habit")`**
 
@@
 
-**`r figure("filmy-dt")`**
+**`r figure("dt")`**
 
@@
 
-**`r figure("filmy-dt")`** (continued)
+**`r figure("dt")`** (continued)
 
@@
 **`r figure("combined")`**
+
+```{r corr-plot, fig.width = 6.85, fig.height = 5, fig.cap = ""}
+# Combine recovery, VPD, and growth habit into a single dataframe (wide format)
+recovery_mean_data <- select(combined_species_means, species, generation, recovery_mean) %>%
+  pivot_wider(values_from = recovery_mean, names_from = generation) %>%
+  rename_with(~glue("recovery_{.}_mean"), .cols = -species)
+
+recovery_sd_data <- select(combined_species_means, species, generation, recovery_sd) %>%
+  pivot_wider(values_from = recovery_sd, names_from = generation) %>%
+  rename_with(~glue("recovery_{.}_sd"), .cols = -species)
+
+plot_data <-
+list(filmy_habit, recovery_mean_data, recovery_sd_data, mean_vpd_sporo, mean_vpd_gameto, range_comparison) %>%
+  reduce(left_join, by = "species")
+
+a <- ggplot(plot_data, aes(x = vpd_sporophyte_mean, y = recovery_sporophyte_mean, color = habit)) +
+  geom_linerange(aes(
+    ymin = recovery_sporophyte_mean - recovery_sporophyte_sd,
+    ymax = recovery_sporophyte_mean + recovery_sporophyte_sd),
+    alpha = 0.3) +
+  geom_linerange(aes(
+    xmin = vpd_sporophyte_mean - vpd_sporophyte_sd,
+    xmax = vpd_sporophyte_mean + vpd_sporophyte_sd),
+    alpha = 0.3) +
+  geom_point() +
+  scale_y_continuous(
+    labels = percent, 
+    limits = c(0,1.25)) +
+  scale_color_manual(values = habit_colors, labels = pretty_text) +
+  labs(
+    y = "Recovery",
+    x = "VPD (kPa)",
+    color = "Growth habit")
+
+b <- ggplot(plot_data, aes(x = vpd_gametophyte_mean, y = recovery_gametophyte_mean, color = habit)) +
+  geom_linerange(aes(
+    ymin = recovery_gametophyte_mean - recovery_gametophyte_sd,
+    ymax = recovery_gametophyte_mean + recovery_gametophyte_sd),
+    alpha = 0.3) +
+  geom_linerange(aes(
+    xmin = vpd_gametophyte_mean - vpd_gametophyte_sd,
+    xmax = vpd_gametophyte_mean + vpd_gametophyte_sd),
+    alpha = 0.3) +
+  geom_point() +
+  scale_y_continuous(
+    labels = percent, 
+    limits = c(0,1.25)) +
+  scale_color_manual(values = habit_colors, labels = pretty_text) +
+  labs(
+    y = "Recovery",
+    x = "VPD (kPa)",
+    color = "Growth habit") +
+  guides(color = "none")
+
+c <- ggplot(plot_data, aes(x = gameto_beyond_sporo, y = recovery_gametophyte_mean, color = habit)) +
+  geom_linerange(aes(
+    ymin = recovery_gametophyte_mean - recovery_gametophyte_sd,
+    ymax = recovery_gametophyte_mean + recovery_gametophyte_sd),
+    alpha = 0.3) +
+  geom_point() +
+  scale_color_manual(values = habit_colors, labels = pretty_text) +
+  scale_y_continuous(labels = percent) +
+  labs(
+    y = "Recovery",
+    x = "Elevation (m)",
+    color = "Growth habit") +
+  guides(color = "none")
+
+a + b + c + 
+  guide_area() + 
+  plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = "a")  &
+  theme(plot.tag = element_text(face = 'bold'))
+
+ggsave(here::here("results/ms/fig_corr.pdf"), width = 6.85, height = 5)
+```
+
+**`r figure("corr")`**

### 2021-01-25 R/functions.R

@@ -700,7 +700,9 @@ calculate_mean_vpd_gameto <- function(
     # there are several sites that are missing a datalogger,
     # so don't include na values when calculating mean
     summarize(
-      gameto_vpd = mean(vpd, na.rm = TRUE),
+      vpd_gametophyte_mean = mean(vpd, na.rm = TRUE),
+      vpd_gametophyte_sd = sd(vpd, na.rm = TRUE),
+      vpd_gametophyte_n = n(),
       .groups = "drop"
     )
   
@@ -748,7 +750,9 @@ calculate_mean_vpd_sporo <- function (community_matrix_raw, filmy_species, moore
     uncount(abundance) %>%
     group_by(species) %>%
     summarize(
-      sporo_vpd = mean(vpd),
+      vpd_sporophyte_mean = mean(vpd),
+      vpd_sporophyte_sd = sd(vpd),
+      vpd_sporophyte_n = n(),
       .groups = "drop"
     )
 }
@@ -814,8 +818,8 @@ combine_env_env_range_recover <- function (combined_species_means, env_range_dat
   left_join(
     select(env_range_data, 
            species, 
-           sporo_vpd, 
-           gameto_vpd,
+           vpd_sporophyte = vpd_sporophyte_mean, 
+           vpd_gametophyte = vpd_gametophyte_mean,
            gameto_beyond_sporo
     ), 
     by = "species") %>%
@@ -1175,8 +1179,8 @@ run_pgls <- function (env_range_recover_data, phy) {
   
   # Run PGLS for each combination of generation x vpd or range breadth vs. recovery
   res <- list(
-    sporo_vpd_model = caper::pgls(recovery_sporophyte ~ sporo_vpd, env_range_recover_data_comp),
-    gameto_vpd_model = caper::pgls(recovery_gametophyte ~ gameto_vpd, env_range_recover_data_comp),
+    sporo_vpd_model = caper::pgls(recovery_sporophyte ~ vpd_sporophyte, env_range_recover_data_comp),
+    gameto_vpd_model = caper::pgls(recovery_gametophyte ~ vpd_gametophyte, env_range_recover_data_comp),
     gameto_range_model = caper::pgls(recovery_gametophyte ~ gameto_beyond_sporo, env_range_recover_data_comp)
     )
   

### 2021-01-25 ms/captions.R

@@ -12,12 +12,14 @@
 
 # - Figures
 figure_full <- captioner::captioner(prefix = "Fig.", suffix = "")
-figure_full(name = "filmy-habit", caption = "filmy-habit")
-figure_full(name = "filmy-dt", caption = "filmy-dt")
+figure_full(name = "habit", caption = "habit")
+figure_full(name = "dt", caption = "dt")
 figure_full(name = "combined", caption = "combined")
+figure_full(name = "corr", caption = "corr")
 
 # - Tables
 table_full <- captioner::captioner(prefix = "Table")
+table_full(name = "glmm", caption = "glmm")
 
 # - SI figures
 s_figure_full <- captioner::captioner(prefix = "Fig. S", suffix = "")

### 2021-01-25 R/packages.R
Add packages

@@ -15,6 +15,7 @@ library(patchwork)
 library(scales)
 library(ggtext)
 library(ggtree)
+library(kableExtra)
 library(tidyverse)
 
 conflict_prefer("col_factor", "readr")

### 2021-01-25 R/packages_dev.R

@@ -1,2 +1,3 @@
 library("visNetwork")
 library("tflow")
+library("pdftools")

### 2021-01-26 install_latex.R
Add 'caption' package to latex packages to install
needed for GT

@@ -12,6 +12,7 @@ latex_packages <- c(
   "bigintcalc",
   "bitset",
   "booktabs",
+  "caption",
   "colortbl",
   "environ",
   "etexcmds",

### 2021-01-26 ms/manuscript.Rmd
Add tables
still need to decide between GT and kableExtra

@@
 tar_load(mean_vpd_gameto)
-
+tar_load(glmm_summary)
+tar_load(glmm_params)
+tar_load(env_range_dt_model_summary)
 ```
@@
 
-Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* and λ both near 1, *P \<*0.05; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval` and λ = `r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`), but no signal for PPFD~95%~ or ETR~max~ (Table 2).
+Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* and λ both near 1, *P \<*0.05; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval` and λ = `r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`), but no signal for PPFD~95%~ or ETR~max~ (`r table("physig")`).
 
@@
   english2
+
+# Extract effect of generation on DT from GLMM
+sporo_dt_effect_mean <-
+glmm_params %>% 
+  filter(response == "recovery") %>%
+  verify(predictor == "generationsporophyte") %>%
+  pull(post.mean) %>%
+  scales::percent(accuracy = 0.1)
+
+sporo_dt_effect_lower <-
+glmm_params %>% 
+  filter(response == "recovery") %>%
+  verify(predictor == "generationsporophyte") %>%
+  pull(`l-95% CI`) %>%
+  scales::percent(accuracy = 0.1)
+
+sporo_dt_effect_upper <-
+glmm_params %>% 
+  filter(response == "recovery") %>%
+  verify(predictor == "generationsporophyte") %>%
+  pull(`u-95% CI`) %>%
+  scales::percent(accuracy = 0.1)
 
@@
 
-The GLMM showed that both generation and the interaction of generation and growth habit, but not growth habit alone, have a significant effect on DT while controlling for phylogeny (Table 3).
-In contrast, range type (gametophytes widespread beyond sporophytes or not) has no significant effect on DT (Table 3).
+The GLMM including the effect of generation only was selected as the best-fitting model for DT (`r table("glmm")`) while taking phylogeny into account.
+Sporophytes had `r sporo_dt_effect_mean` greater DT (95% CI `r sporo_dt_effect_lower` to `r sporo_dt_effect_upper`) relative to gametophytes (`r table("glmm-pars")`).
 
@@
 
-**`r table("glmm")`**
+**`r table("physig")`** Results of tests for phylogenetic signal (Blomberg’s *K* and Pagel’s λ) in desiccation tolerance (DT) and photosynthetic parameters of filmy ferns from Moorea, French Polynesia by generation
+
+```{r table-physig, results = "asis"}
+# Format phylogenetic signal results as a pretty table
+phylosig %>%
+  mutate(k_pval_stars = case_when(
+    k_pval < 0.001 ~ "***",
+    k_pval < 0.01 ~ "**",
+    k_pval < 0.05 ~ "*",
+    TRUE ~ ""
+  )) %>%
+  mutate(lambda_pval_stars = case_when(
+    lambda_pval < 0.001 ~ "***",
+    lambda_pval < 0.01 ~ "**",
+    lambda_pval < 0.05 ~ "*",
+    TRUE ~ ""
+  )) %>%
+  mutate(
+    kval = ifelse(kval < 0.001, format(kval, digits = 3), round_t(kval, 3)) %>% paste0(., k_pval_stars),
+    lambda = ifelse(lambda < 0.001, format(lambda, digits = 3), round_t(lambda, 3)) %>% paste0(., lambda_pval_stars),
+    trait = str_replace_all(trait, "recovery_mean", "Recovery") %>%
+      str_replace_all("etr_mean", "ETR~max~") %>%
+      str_replace_all("par_mean", "PPFD~95%~"),
+    generation = str_to_sentence(generation)
+  ) %>%
+  select(generation, trait, kval, lambda) %>%
+  group_by(generation) %>%
+  gt(rowname_col = "species") %>%
+  cols_label(kval = md("*K*"), lambda = "λ", trait = "") %>%
+  as_latex()
+```
+
+Recovery, recovery of F~v~/F~m~ (%) following 2 d desiccation at –86 MPa; ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached.
+
+Asterisks indicate significance of phylogenetic signal: \*\*, \pval{ }< 0.01, \*, \pval{ }< 0.05.
+
+\newpage
+
+**`r table("glmm")`** Deviance information criterion (DIC) and delta DIC for general linear mixed models (GLMMs) of photosynthetic parameters and desiccation tolerance in response to generation and growth habit in filmy ferns from Moorea, French Polynesia
+
+```{r glmm-summary, results = "asis"}
+# Format GLMM summary for printing
+glmm_summary %>%
+  mutate(
+    fixed_effects = case_when(
+      fixed_effects == "both" ~ "Generation + habit",
+      fixed_effects == "intersect" ~ "Generation + habit + Generation:habit",
+      TRUE ~ str_to_sentence(fixed_effects)
+    ),
+    response = str_replace_all(response, "recovery", "Recovery") %>%
+      str_replace_all("etr", "ETR~max~") %>%
+      str_replace_all("par", "PPFD~95%~")
+  ) %>%
+  group_by(response) %>%
+  gt() %>%
+  cols_label(
+    fixed_effects = "Fixed effects", 
+    response = "Response", 
+    delta_DIC = "delta DIC") %>%
+  tab_style(
+    style = cell_text(weight = "bold"),
+    locations = cells_body(
+      columns = everything(),
+      rows = delta_DIC == 0)) %>%
+  fmt_number(
+    columns = vars(DIC, delta_DIC)
+  ) %>%
+  as_latex()
+```
+
+ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached; Recovery, recovery of F~v~/F~m~ (%) following 2 d desiccation at –86 MPa
+
+Best model (that with the lowest DIC) for each response variable in bold
+
+\newpage
+
+**`r table("glmm-pars")`** Predictor effects, lower and upper 95% confidence interval (CI), effective sample size, and significance of model parameters for the best general linear mixed models (GLMMs) of photosynthetic parameters and desiccation tolerance in response to generation and growth habit in filmy ferns from Moorea, French Polynesia
+
+```{r glmm-params, results = "asis"}
+# Format summary of best GLMM parameters for printing
+glmm_params %>%
+  mutate(
+    fixed_effects = str_to_sentence(fixed_effects),
+    response = str_replace_all(response, "recovery", "Recovery") %>%
+      str_replace_all("etr", "ETR~max~") %>%
+      str_replace_all("par", "PPFD~95%~"),
+    predictor = str_replace_all(predictor, "generationsporophyte", "Sporophyte") %>%
+      str_replace_all("habitLow_Epiphyte", "Low Epiphyte") %>%
+      str_replace_all("habitSaxicolous", "Saxicolous") %>%
+      str_replace_all("habitTerrestrial", "Terrestrial")
+  ) %>%
+  group_by(response) %>%
+  gt() %>%
+  cols_label(
+    fixed_effects = "Fixed effects", 
+    predictor = "Predictor", 
+    post.mean = "Mean effect",
+    `l-95% CI` = "Lower 95% CI",
+    `u-95% CI` = "Upper 95% CI",
+    eff.samp = "Ef. sample size",
+    pMCMC = md("*P* MCMC")) %>%
+  tab_style(
+    style = cell_text(weight = "bold"),
+    locations = cells_body(
+      columns = everything(),
+      rows = pMCMC < 0.05)) %>%
+  fmt_number(columns = vars(post.mean, `l-95% CI`, `u-95% CI`)) %>%
+  fmt_number(columns = vars(eff.samp), decimals = 0) %>%
+  fmt_number(columns = vars(pMCMC), decimals = 3, drop_trailing_zeros = TRUE) %>%
+as_latex()
+```
+
+ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached; Recovery, recovery of F~v~/F~m~ (%) following 2 d desiccation at –86 MPa
+
+Effect of generation is relative to gametophytes; effect of growth habit is relative to high elevation epiphytes
+
+Significant predictors in bold
+
+\newpage
+
+**`r table("pglm")`** Phylogenetic generalized least squares (PLGS) models of desiccation tolerance as a function of environmental water potential (VPD) or the range of gametophytes beyond sporophytes (G beyond S) in filmy ferns from Moorea, French Polynesia
+
+```{r pglm, results = "asis"}
+env_range_dt_model_summary %>%
+  mutate(generation = case_when(
+    str_detect(model, "sporo") ~ "S",
+    str_detect(model, "gameto") ~ "G"
+  )) %>%
+  assert(not_na, generation) %>%
+  mutate(
+    predictor = case_when(
+      str_detect(model, "vpd") ~ "VPD",
+      str_detect(model, "range") ~ "G beyond S"
+    )
+  ) %>%
+  assert(not_na, predictor) %>%
+  select(generation, predictor, df, f_value, adj_r_squared, p_value) %>%
+  mutate(
+    f_value = round_t(f_value, 3),
+    p_value = round_t(p_value, 3),
+    adj_r_squared = round_t(adj_r_squared, 3)
+  ) %>%
+  kbl(col.names = c("Generation", "Predictor", "df", "$F$", "adj. $R^2$", "$P$"), escape = FALSE, booktabs = TRUE)
+```
+
+S, sporophyte; G, gametophyte
+
+\newpage
 
@@
 Growth habit indicated by colored symbols (blue = low elevation epiphyte, green = high elevation epiphyte, yellow = saxicolous, red = terrestrial).
-**Crepidomanes bipunctatum** and **C. humile** were observed growing as both epiphytes and saxicoles, but are shown as epiphytes to distinguish them from exclusively saxicolous species.
-Asterisk indicates significant difference in means (**P** <0.05) between generations for a particular species (**t**-test); ns indicates no significant difference; lack of ns or asterisk means not enough observations available for that species for **t**-test.
+*Crepidomanes bipunctatum* and *C. humile* were observed growing as both epiphytes and saxicoles, but are shown as epiphytes to distinguish them from exclusively saxicolous species.
+Asterisk indicates significant difference in means (*P* <0.05) between generations for a particular species (*t*-test); ns indicates no significant difference; lack of ns or asterisk means not enough observations available for that species for *t*-test.
 Left bar plot shows recovery (%) of chlorophyll fluorescence (Fv/Fm) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa). 
@@
 Growth habit indicated by colored symbols (blue = low elevation epiphyte, green = high elevation epiphyte, yellow = saxicolous, red = terrestrial). 
-**Crepidomanes bipunctatum** and **C. humile** were observed growing as both epiphytes and saxicoles, but were treated as epiphytes to distinguish them from exclusively saxicolous species.
-**Abrodictyum dentatum** sporophytes were desiccated at –38 MPa instead of –86 MPa.
+*Crepidomanes bipunctatum* and *C. humile* were observed growing as both epiphytes and saxicoles, but were treated as epiphytes to distinguish them from exclusively saxicolous species.
+*Abrodictyum dentatum* sporophytes were desiccated at –38 MPa instead of –86 MPa.
 Range of gametophytes beyond sporophytes calculated as the total elevation that gametophytes exceeded sporophytes at either end of the elevational gradient.

### 2021-01-26 R/functions.R

@@ -1219,6 +1219,60 @@ tidy_pgls <- function (model) {
   )
 }
 
+#' Summarize model statistics of a generalized linear mixed model (GLMM)
+#'
+#' @param model The model
+#'
+#' @return Tibble with DIC
+#' 
+summarize_glmm <- function(model) {
+  model_sum <- summary(model)
+  tibble(
+    DIC = model_sum$DIC
+  )
+}
+
+tidy_glmms <- function (model_df) {
+  model_df %>%
+    mutate(summary = map(model, summarize_glmm)) %>%
+    select(fixed_effects, response, summary) %>%
+    unnest(cols = "summary") %>%
+    group_by(response) %>%
+    mutate(
+      delta_DIC = DIC - min(DIC)) %>%
+    ungroup %>%
+    arrange(response, delta_DIC)
+}
+
+#' Summarize model parameters of a generalized linear mixed model (GLMM)
+#'
+#' @param model The model
+#'
+#' @return Tibble with estimated effect, confidence intervals, sample size, and probability
+#' 
+summarize_glmm_params <- function(model) {
+  model_sum <- summary(model)
+  model_sum$solutions %>% 
+    as.data.frame %>%
+    rownames_to_column("predictor") %>%
+    filter(!str_detect(predictor, "Intercept")) %>%
+    as_tibble()
+}
+
+tidy_best_glmm_params <- function (model_df) {
+  model_df %>%
+    mutate(param_summary = map(model, summarize_glmm_params)) %>%
+    mutate(summary = map(model, summarize_glmm)) %>%
+    select(fixed_effects, response, param_summary, summary) %>%
+    unnest(cols = c("param_summary", "summary")) %>%
+    group_by(response) %>%
+    mutate(
+      delta_DIC = DIC - min(DIC)) %>%
+    ungroup %>%
+    filter(delta_DIC == 0) %>%
+    select(-DIC, -delta_DIC)
+}
+
 # Etc ----
 
 #' Match trait data and tree

### 2021-01-26 R/packages.R

@@ -14,9 +14,10 @@ library(lubridate)
 library(patchwork)
 library(scales)
 library(ggtext)
-library(ggtree)
+suppressPackageStartupMessages(library(ggtree))
 library(kableExtra)
-library(tidyverse)
+library(gt)
+suppressPackageStartupMessages(library(tidyverse))
 
 conflict_prefer("col_factor", "readr")
 conflict_prefer("filter", "dplyr")

### 2021-01-26 _targets.R

@@ -150,6 +150,12 @@ tar_plan(
     traits = filmy_habit,
     phy = filmy_phy),
   
+  # Get DIC for each GLMM
+  glmm_summary = tidy_glmms(glmms),
+  
+  # Get parameters for best GLMMs only
+  glmm_params = tidy_best_glmm_params(glmms),
+  
   # PGLS ----
   # (Phylogenetic Generalized Least Squares)
   

### 2021-01-26 ms/captions.R

@@ -19,7 +19,9 @@ figure_full(name = "corr", caption = "corr")
 
 # - Tables
 table_full <- captioner::captioner(prefix = "Table")
+table_full(name = "physig", caption = "physig")
 table_full(name = "glmm", caption = "glmm")
+table_full(name = "glmm-pars", caption = "glmm-pars")
 
 # - SI figures
 s_figure_full <- captioner::captioner(prefix = "Fig. S", suffix = "")
@@ -27,8 +29,13 @@ s_figure_full <- captioner::captioner(prefix = "Fig. S", suffix = "")
 # - SI tables
 s_table_full <- captioner::captioner(prefix = "S", auto_space = FALSE, suffix = " Table. ")
 
-# Make short versions of citation functions (just the number)
+# Make short versions of citation functions (just the object type and number, e.g., "Fig. 1")
 figure <- pryr::partial(figure_full, display = "cite")
 table <- pryr::partial(table_full, display = "cite")
 s_figure <- pryr::partial(s_figure_full, display = "cite")
 s_table <- pryr::partial(s_table_full, display = "cite")
+
+# just the number (e.g., "1")
+table_num <- function (name) {table(name) %>% str_remove("Table ")}
+figure_num <- function (figure) {table(name) %>% str_remove("Fig. ")}
+

### 2021-01-27 ms/manuscript.Rmd
WIP: tables with kableExtra

@@
 Best model (that with the lowest DIC) for each response variable in bold
+
+\newpage
+
+```{r glmm-table-kableextra, results = "asis"}
+group_locs <-
+glmm_summary %>% mutate(row_num = 1:nrow(.)) %>% group_by(response) %>%
+  summarize(
+    start = min(row_num),
+    end = max(row_num),
+    .groups = "drop"
+  )
+
+glmm_summary %>%
+  mutate(
+    fixed_effects = case_when(
+      fixed_effects == "both" ~ "Generation + habit",
+      fixed_effects == "intersect" ~ "Generation + habit + Generation:habit",
+      TRUE ~ str_to_sentence(fixed_effects)
+    ),
+    response = str_replace_all(response, "recovery", "Recovery") %>%
+      str_replace_all("etr", "ETR~max~") %>%
+      str_replace_all("par", "PPFD~95%~"),
+    DIC = round_t(DIC, 2),
+    delta_DIC = round_t(delta_DIC, 2)
+  ) %>%
+  select(fixed_effects, response, DIC, delta_DIC) %>%
+  kbl(format = "latex", booktabs = TRUE, col.names = c("Fixed effects", "Response", "DIC", "delta DIC")) %>%
+  row_spec(which(glmm_summary$delta_DIC == 0), bold = TRUE) %>%
+  pack_rows(
+    "ETR\\textsubscript{max}", 
+    group_locs %>% filter(response == "etr") %>% pull(start),
+    group_locs %>% filter(response == "etr") %>% pull(end),
+    escape = FALSE)
+```
 

### 2021-01-27 ms/manuscript.Rmd
Add timescale to tree

@@
 ```{r combined-plot, fig.width = 6.85, fig.height = 5, fig.cap = ""}
-# Copy phylogeny and species means data so they
+# Copy data so they
 # can be manipulated for this plot
@@
 traits <- combined_species_means
+habit <- filmy_habit
 
@@
 
-filmy_habit <-
-  filmy_habit %>%
+# Filter growth habit data to only those in phylogeny
+habit <-
+  habit %>%
   filter(species %in% phy$tip.label)
@@
 # Part a (left): tree
-a <- ggtree(phy) %<+% filmy_habit +
+a <- ggtree(phy) %<+% habit +
   geom_tippoint(aes(color = habit), position = position_nudge(x = 7, y = 0), size = 4) +
   scale_color_manual(values = habit_colors, labels = pretty_text) +
-  labs(color = "Growth habit")
+  labs(color = "Growth habit") +
+  # There is no way modify the text in geom_treescale (to add units), so make the
+  # text disappear by setting fontsize = 0
+  geom_treescale(y = 12, x = 0, width = 40, fontsize = 0) +
+  # Manually place custom label above scale bar
+  annotate("text", x = 20, y = 12, label = "40 my", size = 2.5, vjust = -1)
 

### 2021-01-27 ms/manuscript.Rmd
Move pretty_text() to functions.R

@@
                red = "#D55E00", purple = "#CC79A7")
-
-# Define function for formatting text in legends
-pretty_text <- function (x) {
-  x %>%
-  str_replace_all("_", " ") %>%
-    str_to_sentence()
-}
 ```

### 2021-01-27 R/functions.R

@@ -1508,6 +1508,13 @@ analyze_dist_pattern <- function(community_matrix_raw, filmy_species, phy, moore
   
 }
 
+# Define function for formatting text in legends
+pretty_text <- function (x) {
+  x %>%
+    str_replace_all("_", " ") %>%
+    str_to_sentence()
+}
+
 # Manuscript ----
 
 # Helper function to check on p-value significance and extract from t_test_results

### 2021-01-27 ms/si.Rmd
Print tree with BS values in SI

--- 2021-01-05 ms/si.Rmd
@@
 ```{r load-data, cache = FALSE}
-tar_load(
-  names = c(
-    "filmy_dt_chamber",
-    "filmy_dt",
-    "filmy_habit",
-    "rel_water_indiv",
-    "rel_water_species_means"
-  )
-)
+tar_load(filmy_dt_chamber)
+tar_load(filmy_dt)
+tar_load(filmy_habit)
+tar_load(rel_water_indiv)
+tar_load(rel_water_species_means)
+tar_load(filmy_phy)
+tar_load(filmy_habit)
 ```
@@
 
+```{r plot-setup}
+# JPR requirements:
+# Figures should be 39 mm, 84 mm, 129 mm, or 174 mm wide and not higher than 234 mm
+# In inches, widths: 1.53, 3.31, 5.08, 6.85; height: 9.21
+
+# Set overall theme, saving default theme
+default_theme <- theme_set(
+  theme_bw(base_size = 10) +
+    theme(
+      panel.grid.minor = element_blank()
+    ))
+
+# Define color-blind palette
+cbPalette <- c(grey = "#999999", gold = "#E69F00", skyblue = "#56B4E9", 
+               green = "#009E73", yellow = "#F0E442", blue = "#0072B2", 
+               red = "#D55E00", purple = "#CC79A7")
+```
+
+```{r full-phylo, fig.height = 4.5, fig.width = 7, fig.cap = ""}
+# Copy phylogeny and species means data so they
+# can be manipulated for this plot
+phy <- filmy_phy
+habit <- filmy_habit
+
+# Set colors for growth habit, generation
+habit_colors <- c(
+  "Terrestrial" = cbPalette[["red"]], 
+  "Saxicolous" = cbPalette[["gold"]], 
+  "Low_Epiphyte" = cbPalette[["green"]],
+  "High_Epiphyte" = cbPalette[["blue"]])
+
+generation_colors <- c(
+  "gametophyte" = "grey70",
+  "sporophyte" = "grey30"
+)
+
+# Filter growth habit data to only species in phylogeny
+habit <-
+  habit %>%
+  filter(species %in% phy$tip.label)
+
+# Convert underscores to spaces
+# phy$tip.label <- str_replace_all(phy$tip.label, "_", "~")
+
+# Make a tibble of bootstrap values to map onto tree
+# - needs one column with "node", the other can be named as we like
+# - hard-code node ID: internal nodes start after tip nodes,
+#   and phy$node.label is in the same order as internal nodes
+bs_tibble <- tibble(
+  node=1:Nnode(phy) + Ntip(phy), 
+  # Don't print BS < 50%
+  bootstrap = ifelse(phy$node.label < 50, "", phy$node.label))
+
+# Join growth habit and BS data to ggtree object
+ggtree(phy) %<+% filmy_habit %<+% bs_tibble +
+  geom_text(aes(label=bootstrap), hjust=-.25, size = 3, color = "grey30") +
+  geom_tippoint(aes(color = habit), position = position_nudge(x = 0, y = 0), size = 4) +
+  # Add species labels
+  geom_tiplab(
+    # getting species names into italics is painful....
+    # need to use plotmath() parsing.
+    # to print a space, use `~`
+    # e.g., `italic("Hymenophyllum~polyanthos")`
+    aes(label=paste0('italic(', str_replace_all(label, "_", "~"), ')')), 
+    parse = TRUE,
+    hjust = -0.02) +
+  # There is no way modify the text in geom_treescale (to add units), so make the
+  # text disappear by setting fontsize = 0
+  geom_treescale(y = 20, x = 0, width = 40, fontsize = 0) +
+  # Manually place custom label above scale bar
+  annotate("text", x = 20, y = 20, label = "40 my", size = 3, vjust = -1) +
+  # Scale colors by growth habit
+  scale_color_manual(values = habit_colors, labels = pretty_text) +
+  labs(color = "Growth habit") +
+  # Need to remove clip so that species names print properly
+  coord_cartesian(clip = "off") +
+  theme(
+  plot.margin = margin(t = 0, r = 2.2, b = 0, l = 0, unit = "in"),
+  legend.position = "bottom"
+)
+```
+
+**Figure SXX**. Maximum likelihood ultrametric phylogenetic tree including all Moorean filmy fern species.
+Tree obtained by extracting Hymenophyllaceae from the tree of Nitta et al. (2016) with the extract.clade function of the ape package in R (Paradis et al., 2004).
+Bootstrap support values >50% shown at nodes.
+Growth habit indicated by colored dots. 
+*Crepidomanes bipunctatum* and *C. humile* were observed growing as both epiphytes and saxicoles, but are shown as epiphytes to distinguish them from exclusively saxicolous species.
+
 ```{r gametophyte-dt-2013, fig.height = 4.5, fig.width = 7, fig.cap = ""}
@@
 
-**Figure S1**.
+**Figure SXX**.
 Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes.
 MgNO~3~ was used to maintain constant humidity at \circa{ }58%.
-All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days unintentionally.
+All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days unintentionally
 
@@
 
-**Figure S2**.
+**Figure SXX**.
 Conditions inside desiccation chamber and duration of desiccation treatment for sporophytes at most intense desiccation level.

### 2021-01-27 R/functions.R

@@ -835,6 +835,37 @@ abbrev_sp <- function(data) {
     unite("species", genus, epithet)
 }
 
+#' Transfer bootstrap values from BS tree to time tree
+#' 
+#' The time tree lacks BS values, but the topology is the same.
+#'
+#' @param filmy_phy_bs Tree output from RAxML with bootstrap values at nodes
+#' @param filmy_phy Time tree without bootstrap values
+#'
+#' @return Time tree with bootstrap values
+#' 
+transfer_bs <- function(filmy_phy_bs, filmy_phy) {
+  
+  # Rotate nodes of BS tree so that the tip position matches the time tree
+  filmy_phy_bs <- rotateConstr(filmy_phy_bs, constraint = filmy_phy$tip.label)
+  
+  # This only changes how the phylogeny *looks*, not the underlying data.
+  # Write out the tree, then read it in again to change the actual order in the data.
+  temp_tree_file <- fs::path(tempdir(), "filmy_phy_bs.tre")
+  ape::write.tree(filmy_phy_bs, temp_tree_file)
+  filmy_phy_bs <- ape::read.tree(temp_tree_file)
+  fs::file_delete(temp_tree_file)
+  
+  # Make sure tips are now in the same order.
+  assert_that(isTRUE(all.equal(filmy_phy_bs$tip.label, filmy_phy$tip.label)))
+  
+  # Transfer BS support values to the time tree
+  filmy_phy$node.label <- filmy_phy_bs$node.label
+  
+  filmy_phy
+  
+}
+
 # t-test ----
 
 #' Run a t-test comparing response values between gametophytes and sporophytes.
@@ -1166,12 +1197,14 @@ summary_to_csv <- function (model) {
 #' @param phy Phylogeny
 #'
 #' @return list of model objects:
-#' - sporo_vpd_model:
-#' - gameto_vpd_model:
-#' - sporo_range_model:
-#' - gameto_range_model:
+#' - sporo_vpd_model
+#' - gameto_vpd_model
+#' - gameto_range_model
 #' 
 run_pgls <- function (env_range_recover_data, phy) {
+  
+  # Need to drop node labels for caper::comparative.data()
+  phy$node.label <- NULL
   
   # Create a comparative.data object for caper combining the range size, vpd, 
   # and recovery data with the phylogeny

### 2021-01-27 R/unzip_data.R

@@ -14,8 +14,8 @@ unzip_nitta_2017(
 # https://datadryad.org/stash/dataset/doi:10.5061/dryad.fqz612jps, clicking on
 # "Download dataset", and saving to the "data" folder in this project.
 unzip(
-  zipfile = file_in("data/doi_10.5061_dryad.fqz612jps__v4.zip"), 
+  zipfile = "data/doi_10.5061_dryad.fqz612jps__v4.zip", 
   files = "moorea_climate.csv", 
-  exdir = file_out("data/nitta_2020/"), 
+  exdir = "data/nitta_2020/", 
   junkpaths = TRUE, 
   overwrite = TRUE)

### 2021-01-27 _targets.R

@@ -24,15 +24,18 @@ tar_plan(
   # - vector of filmy fern species
   filmy_species = filmy_habit$species,
   
-  # - phylogenetic tree (time tree)
+  # - phylogenetic tree (time tree, lacks BS values at nodes)
   tar_file(filmy_phy_file, "data/nitta_2017/treepl_Moorea_Tahiti.tre"),
-  filmy_phy = ape::read.tree(filmy_phy_file) %>%
+  filmy_phy_no_bs = ape::read.tree(filmy_phy_file) %>%
     ape::keep.tip(filmy_species),
   
   # - phylogenetic tree (ML tree with BS values at nodes)
   tar_file(filmy_phy_bs_file, "data/nitta_2017/RAxML_bipartitions.all_broad.reduced"),
-  filmy_phy_bs = ape::read.tree(filmy_raxml_file) %>%
-    ape::keep.tip(filmy_phy$tip.label),
+  filmy_phy_bs = ape::read.tree(filmy_phy_bs_file) %>%
+    ape::keep.tip(filmy_phy_no_bs$tip.label),
+  
+  # - transfer bootstrap values from ML tree to time tree
+  filmy_phy = transfer_bs(filmy_phy_bs, filmy_phy_no_bs),
   
   # - desiccation tolerance yields
   tar_file(filmy_dt_file, "data/filmy_dt.csv"),

### 2021-01-29 R/functions.R
Refactor t-test
fixes incorrect lumping of individuals

@@ -399,15 +399,15 @@ load_filmy_dt <- function (file) {
 
 load_filmy_dt_chamber <- function (file) {
   readr::read_csv(file,
-  col_types = cols(
-    date_time = col_datetime(format = ""),
-    temp = col_double(),
-    rh = col_double(),
-    salt = col_character(),
-    generation = col_character(),
-    year = col_double(),
-    serial_no = col_character()
-  )
+                  col_types = cols(
+                    date_time = col_datetime(format = ""),
+                    temp = col_double(),
+                    rh = col_double(),
+                    salt = col_character(),
+                    generation = col_character(),
+                    year = col_double(),
+                    serial_no = col_character()
+                  )
   )
 }
 
@@ -550,99 +550,100 @@ process_community_matrix <- function (community_matrix_path, species_list, moore
     spread(site, abundance)
 }
 
-
-
-calculate_indiv_etr <- function (data) {
-  
-  # Calculate maximum ETR for each individual
+#' Fight light curve models to light curve data
+#'
+#' @param data Light curve data, with columns for 
+#' `coll_num`, `individual`, `species`, `generation`, `condition`, `date`
+#'
+#' @return Tibble with list-columns for `data`, `nls_mod` (model), 
+#' `k_stats` (model parameters), and `fitted` (fitted data points)
+#' 
+fit_lc_model <- function (data) {
   data %>%
-    assert(not_na, etr) %>%
-    group_by(species, generation, individual) %>%
-    summarize(
-      etr_max = max(etr),
-      .groups = "drop"
-    )
-  
-}
-
-calculate_mean_etr <- function (data) {
-  
-  # Calculate mean ETR max by species and generation
-  data %>%
-    assert(not_na, etr_max) %>%
-    group_by(generation, species) %>%
-    summarize(
-      etr_mean = mean(etr_max),
-      etr_sd = sd(etr_max),
-      etr_n = n()
-    ) %>%
-    ungroup()
-  
-}
-
-calculate_indiv_par <- function (data) {
-  
-  # Calculate critical PAR value for each individual
-  data %>%
-    select(species, generation, individual, par, etr) %>%
+    select(species, generation, individual, coll_num, condition, date, par, etr) %>%
+    # some sporophytes lack collection numbers, but
+    # we need to keep this column for differentiating between
[…]

### 2021-01-29 _targets.R

@@ -109,38 +109,33 @@ tar_plan(
     by = "species"
   ),
   
-  # - calculate PAR95 by individual
-  par_indiv = calculate_indiv_par(light_data),
+  # - fit light curve models to data
+  light_models = fit_lc_model(light_data),
   
-  # - calculate ETRmax by individual
-  etr_indiv = calculate_indiv_etr(light_data),
+  # - extract model parameters: 
+  # critical PAR, and ETR at 95% of estimated max value
+  filmy_lc_model_params = extract_lc_model_params(light_models),
   
+  # - calculate mean light curve parameters by species and generation
+  light_species_means = calculate_mean_light(filmy_lc_model_params),
+
   # - calculate mean DT recovery by species and generation
   recovery_species_means = calc_mean_recovery(recovery_indiv),
   
   # - calculate relative water content by species (sporophytes only)
   rel_water_species_means = calculate_mean_water(rel_water_indiv),
   
-  # - calculate mean ETR max by species and generation
-  etr_species_means = calculate_mean_etr(etr_indiv),
-  
-  # - calculate mean PAR95 by species and generation
-  par_species_means = calculate_mean_par(par_indiv),
-  
   # - combine the species means into a single df
   combined_species_means = combine_mean_phys_traits(
     recovery_species_means = recovery_species_means,
-    etr_species_means = etr_species_means,
-    par_species_means = par_species_means
+    light_species_means = light_species_means
   ),
   
   # t-test ----
-  
   # Perform two-sided t-test on DT and light responses across generations
   t_test_results = run_t_test(
-    recovery_data = recovery_indiv, 
-    par_indiv = par_indiv, 
-    etr_indiv = etr_indiv),
+    filmy_lc_model_params = filmy_lc_model_params, 
+    recovery_indiv = recovery_indiv),
   
   # Phylogenetic signal ----
   phylosig = analyze_phylosig_by_generation(

### 2021-01-29 ms/manuscript.Rmd
Add t-test stars to combined plot

@@
 ```{r combined-plot, fig.width = 6.85, fig.height = 5, fig.cap = ""}
-# Copy data so they
-# can be manipulated for this plot
+# Copy data so they can be manipulated for this plot
 phy <- filmy_phy
@@
 
-# Part a (left): tree
+# Part A (left): tree
 a <- ggtree(phy) %<+% habit +
-  geom_tippoint(aes(color = habit), position = position_nudge(x = 7, y = 0), size = 4) +
+  geom_tippoint(aes(color = habit), position = position_nudge(x = 14, y = 0), size = 3.5) +
   scale_color_manual(values = habit_colors, labels = pretty_text) +
@@
   # Manually place custom label above scale bar
-  annotate("text", x = 20, y = 12, label = "40 my", size = 2.5, vjust = -1)
-
-# Part b (middle): desiccation tolerance
-b <- 
-ggplot(traits, aes(y = species, x = recovery_mean, fill = generation)) +
-  geom_col(position = position_dodge(0.5, preserve = "single"), width = 0.5) +
+  annotate("text", x = 20, y = 12, label = "40 my", size = 2.5, vjust = -1) +
+  scale_x_continuous(expand = expansion(mult = c(0.01, 0.04))) +
+  theme_void(base_size = 10)
+
+# Part B (middle): desiccation tolerance
+# Make helper table of significant t-test results
+dt_t_test_data <-
+t_test_results %>%
+  filter(response == "recovery") %>%
+  mutate(star = case_when(
+    p_value < 0.001 ~ "***",
+    p_value < 0.01 ~ "**",
+    p_value < 0.05 ~ "*",
+    TRUE ~ "ns"
+    )) %>%
+  select(species, star) %>%
+  left_join(select(traits, species, generation, recovery_mean, recovery_sd), by = "species") %>%
+  rowwise %>%
+  mutate(recovery_top = sum(recovery_mean, recovery_sd)) %>%
+  group_by(species, star) %>%
+  summarize(recovery_top = max(recovery_top), .groups = "drop")
+
+# Set position dodge for bars
+pos <- position_dodge(
+  width = 0.5, # Bars right next to each other
+  preserve = "single" # Keep bar width constant
+  )
+
+b <- ggplot(traits, aes(y = species)) +
+  geom_col(
+    aes(x = recovery_mean, fill = generation),
+    position = pos,
+    width = 0.5) +
   geom_errorbarh(
-    aes(xmax = recovery_mean + recovery_sd, xmin = recovery_mean),
-    position = position_dodge(0.5, preserve = "single"),
+    aes(x = recovery_mean, xmax = recovery_mean + recovery_sd, xmin = recovery_mean, fill = generation),
+    position = pos,
     height = 0,
     color = "grey20") +
+  geom_text(
+    data = dt_t_test_data, 
+    aes(label = star, y = species, x = recovery_top),
+    angle = -90,
+    nudge_x = 0.01,
+    color = "grey20",
+    size = 3,
+    vjust = -0.5) +
   scale_x_continuous(
@@
   scale_fill_manual(values = generation_colors, labels = pretty_text) +
-  theme_bw(base_size = 10) +
+  # theme_bw(base_size = 10) +
   labs(x = "Recovery", fill = "Generation") +
@@
 
-# Part c (right): light-levels
-c <-
-ggplot(traits, aes(y = species, x = etr_mean, fill = generation)) +
-  geom_col(position = position_dodge(0.5, preserve = "single"), width = 0.5) +
+# Part C (right): light-levels
+
+# Make helper table of significant t-test results
+etr_t_test_data <-
+t_test_results %>%
+  filter(response == "etr") %>%
+  mutate(star = case_when(
+    p_value < 0.001 ~ "***",
+    p_value < 0.01 ~ "**",
+    p_value < 0.05 ~ "*",
+    TRUE ~ "ns"
+    )) %>%
+  select(species, star) %>%
+  left_join(select(traits, species, generation, etr_mean, etr_sd), by = "species") %>%
+  rowwise %>%
+  mutate(etr_top = sum(etr_mean, etr_sd)) %>%
+  group_by(species, star) %>%
+  summarize(etr_top = max(etr_top), .groups = "drop")
+
+c <- ggplot(traits, aes(y = species)) +
+  geom_col(
+    aes(x = etr_mean, fill = generation),
+    position = pos,
+    width = 0.5) +
   geom_errorbarh(
-    aes(xmax = etr_mean + etr_sd, xmin = etr_mean),
-    position = position_dodge(0.5, preserve = "single"),
+    aes(x = etr_mean, xmax = etr_mean + etr_sd, xmin = etr_mean, fill = generation),
+    position = pos,
     height = 0,
     color = "grey20") +
-  theme_bw(base_size = 10) +
+  geom_text(
+    data = etr_t_test_data, 
+    aes(label = star, y = species, x = etr_top),
+    angle = -90,
+    nudge_x = 0.01,
+    color = "grey20",
+    size = 3,
+    vjust = -0.5) +
   scale_fill_manual(values = generation_colors, labels = pretty_text) +
+  # theme_bw(base_size = 10) +
   scale_y_discrete(
@@
     # Adjust invididual margins to remove whitespace between subplots
-    plot.margin = margin(t = 0.1, r = 0, b = 0.1, l = 0.04, unit = "in"))
+    plot.margin = margin(t = 0.1, r = 0, b = 0.1, l = 0.04, unit = "in"),
+    # Wrap legend parts into multiple lines
+    legend.box = "vertical", 
+    legend.margin = margin())
 

### 2021-01-29 ms/manuscript.Rmd
Switch to kableExtra for (most) tables

@@
 **Number of tables, black-and-white figures, and color figures.**  
-No tables / 2 black-and-white figure / 1 color figure
+4 tables / 0 black-and-white figures / 4 color figures
 
@@
 sporo_vpd_pval <- env_range_dt_model_summary %>% filter(model == "sporo_vpd_model") %>% pull(p_value) %>%
-  round_t(2)
+  round_t(3)
 gameto_vpd_pval <- env_range_dt_model_summary %>% filter(model == "gameto_vpd_model") %>% pull(p_value) %>%
-  round_t(2)
-```
-
-Desiccation tolerance tended to increase with VPD for both gametophytes and sporophytes (Fig. 6).
-These trends were significant when analyzed using PGLS for gametophytes (*P* = `r gameto_vpd_pval`) but not sporophytes (*P* = `r sporo_vpd_pval`; Appendix S5).
-
-No significant relationship was detected between range size of gametophytes and DT.
+  round_t(3)
+```
+
+Desiccation tolerance tended to increase with VPD for sporophytes, but not gametophytes (`r figure("corr")`).
+These trends were marginally non-significant when analyzed using PGLS for sporophytes (*P* = `r sporo_vpd_pval`) and not significant for gametophytes (*P* = `r gameto_vpd_pval`; `r table("pgls")`).
+
+No significant relationship was detected between range size of gametophytes and DT (`r figure("corr")`, `r table("pgls")`).
 
@@
 
+```{r table-setup}
+options(knitr.table.format = "latex")
+```
+
 **`r table("physig")`** Results of tests for phylogenetic signal (Blomberg’s *K* and Pagel’s λ) in desiccation tolerance (DT) and photosynthetic parameters of filmy ferns from Moorea, French Polynesia by generation
 
-```{r table-physig, results = "asis"}
-# Format phylogenetic signal results as a pretty table
+```{r physig-table, results = "asis"}
+# Format phylogenetic signal results as a pretty table with kableExtra
+
+# Make helper table for formatting by groups
+phylosig_group_locs <-
+phylosig %>% mutate(row_num = 1:nrow(.)) %>% group_by(generation) %>%
+  summarize(
+    start = min(row_num),
+    end = max(row_num),
+    .groups = "drop"
+  )
+
 phylosig %>%
@@
     trait = str_replace_all(trait, "recovery_mean", "Recovery") %>%
-      str_replace_all("etr_mean", "ETR~max~") %>%
-      str_replace_all("par_mean", "PPFD~95%~"),
+      str_replace_all("etr_mean", "ETR\\\\textsubscript{max}") %>%
+      str_replace_all("par_mean", "PPFD\\\\textsubscript{95\\\\%}"),
     generation = str_to_sentence(generation)
   ) %>%
-  select(generation, trait, kval, lambda) %>%
-  group_by(generation) %>%
-  gt(rowname_col = "species") %>%
-  cols_label(kval = md("*K*"), lambda = "λ", trait = "") %>%
-  as_latex()
+  select(trait, kval, lambda) %>%
+  kbl(format = "latex", booktabs = TRUE, col.names = c("", "\\textit{K}", "λ"), escape = FALSE) %>%
+  pack_rows(
+    "Sporophytes", 
+    phylosig_group_locs %>% filter(generation == "sporophyte") %>% pull(start),
+    phylosig_group_locs %>% filter(generation == "sporophyte") %>% pull(end),
+    escape = FALSE) %>%
+  pack_rows(
+    "Gametophytes", 
+    phylosig_group_locs %>% filter(generation == "gametophyte") %>% pull(start),
+    phylosig_group_locs %>% filter(generation == "gametophyte") %>% pull(end),
+    escape = FALSE)
 ```
@@
 
-```{r glmm-summary, results = "asis"}
-# Format GLMM summary for printing
+```{r glmm-table, results = "asis"}
+glmm_summ_group_locs <-
+glmm_summary %>% mutate(row_num = 1:nrow(.)) %>% group_by(response) %>%
+  summarize(
+    start = min(row_num),
+    end = max(row_num),
+    .groups = "drop"
+  )
+
 glmm_summary %>%
@@
     ),
-    response = str_replace_all(response, "recovery", "Recovery") %>%
-      str_replace_all("etr", "ETR~max~") %>%
-      str_replace_all("par", "PPFD~95%~")
+    DIC = round_t(DIC, 2),
+    delta_DIC = round_t(delta_DIC, 2)
   ) %>%
-  group_by(response) %>%
-  gt() %>%
-  cols_label(
-    fixed_effects = "Fixed effects", 
-    response = "Response", 
-    delta_DIC = "delta DIC") %>%
-  tab_style(
-    style = cell_text(weight = "bold"),
-    locations = cells_body(
-      columns = everything(),
-      rows = delta_DIC == 0)) %>%
-  fmt_number(
-    columns = vars(DIC, delta_DIC)
-  ) %>%
-  as_latex()
+  select(fixed_effects, DIC, delta_DIC) %>%
+  kbl(format = "latex", booktabs = TRUE, col.names = c("Fixed effects", "DIC", "delta DIC"), escape = TRUE) %>%
+  row_spec(which(glmm_summary$delta_DIC == 0), bold = TRUE) %>%
+  pack_rows(
+    "ETR\\\\textsubscript{max}", 
+    glmm_summ_group_locs %>% filter(response == "etr") %>% pull(start),
+    glmm_summ_group_locs %>% filter(response == "etr") %>% pull(end),
+    escape = FALSE) %>%
+  pack_rows(
+    "PPFD\\\\textsubscript{95\\\\%}", 
+    glmm_summ_group_locs %>% filter(response == "par") %>% pull(start),
+    glmm_summ_group_locs %>% filter(response == "par") %>% pull(end),
+    escape = FALSE) %>%
+  pack_rows(
+    "Recovery", 
+    glmm_summ_group_locs %>% filter(response == "recovery") %>% pull(start),
+    glmm_summ_group_locs %>% filter(response == "recovery") %>% pull(end),
+    escape = FALSE)
 ```
@@
 
-```{r glmm-table-kableextra, results = "asis"}
-group_locs <-
-glmm_summary %>% mutate(row_num = 1:nrow(.)) %>% group_by(response) %>%
+**`r table("glmm-pars")`** Predictor effects, lower and upper 95% confidence interval (CI), effective sample size, and significance of model parameters for the best general linear mixed models (GLMMs) of photosynthetic parameters and desiccation tolerance in response to generation and growth habit in filmy ferns from Moorea, French Polynesia
+
+```{r glmm-params, results = "asis"}
+# Make helper table for formatting by groups
+glmm_params_group_locs <-
+glmm_params %>% mutate(row_num = 1:nrow(.)) %>% group_by(response) %>%
   summarize(
@@
 
-glmm_summary %>%
-  mutate(
-    fixed_effects = case_when(
-      fixed_effects == "both" ~ "Generation + habit",
-      fixed_effects == "intersect" ~ "Generation + habit + Generation:habit",
-      TRUE ~ str_to_sentence(fixed_effects)
-    ),
-    response = str_replace_all(response, "recovery", "Recovery") %>%
-      str_replace_all("etr", "ETR~max~") %>%
-      str_replace_all("par", "PPFD~95%~"),
-    DIC = round_t(DIC, 2),
-    delta_DIC = round_t(delta_DIC, 2)
-  ) %>%
-  select(fixed_effects, response, DIC, delta_DIC) %>%
-  kbl(format = "latex", booktabs = TRUE, col.names = c("Fixed effects", "Response", "DIC", "delta DIC")) %>%
-  row_spec(which(glmm_summary$delta_DIC == 0), bold = TRUE) %>%
-  pack_rows(
-    "ETR\\textsubscript{max}", 
-    group_locs %>% filter(response == "etr") %>% pull(start),
-    group_locs %>% filter(response == "etr") %>% pull(end),
-    escape = FALSE)
-```
-
-\newpage
-
-**`r table("glmm-pars")`** Predictor effects, lower and upper 95% confidence interval (CI), effective sample size, and significance of model parameters for the best general linear mixed models (GLMMs) of photosynthetic parameters and desiccation tolerance in response to generation and growth habit in filmy ferns from Moorea, French Polynesia
-
-```{r glmm-params, results = "asis"}
-# Format summary of best GLMM parameters for printing
+glmm_params_cols <-
+  c(
+    "Fixed effects" = "fixed_effects", 
+    "Predictor" = "predictor",
+    "Mean effect" = "post.mean",
+    "Lower 95\\% CI" = "l-95% CI",
+    "Upper 95\\% CI" = "u-95% CI",
+    "Ef. sample size" = "eff.samp",
+    "\\textit{P} MCMC" = "pMCMC"
+  )
+
 glmm_params %>%
@@
     fixed_effects = str_to_sentence(fixed_effects),
-    response = str_replace_all(response, "recovery", "Recovery") %>%
-      str_replace_all("etr", "ETR~max~") %>%
-      str_replace_all("par", "PPFD~95%~"),
     predictor = str_replace_all(predictor, "generationsporophyte", "Sporophyte") %>%
@@
   ) %>%
-  group_by(response) %>%
-  gt() %>%
-  cols_label(
-    fixed_effects = "Fixed effects", 
-    predictor = "Predictor", 
-    post.mean = "Mean effect",
-    `l-95% CI` = "Lower 95% CI",
-    `u-95% CI` = "Upper 95% CI",
-    eff.samp = "Ef. sample size",
-    pMCMC = md("*P* MCMC")) %>%
-  tab_style(
-    style = cell_text(weight = "bold"),
-    locations = cells_body(
-      columns = everything(),
-      rows = pMCMC < 0.05)) %>%
-  fmt_number(columns = vars(post.mean, `l-95% CI`, `u-95% CI`)) %>%
-  fmt_number(columns = vars(eff.samp), decimals = 0) %>%
-  fmt_number(columns = vars(pMCMC), decimals = 3, drop_trailing_zeros = TRUE) %>%
-as_latex()
+  select(all_of(glmm_params_cols)) %>%
+  kbl(format = "latex", booktabs = TRUE, escape = FALSE) %>%
+  kable_styling(latex_options = "scale_down") %>%
+  row_spec(which(glmm_params$pMCMC < 0.05), bold = TRUE) %>%
+  pack_rows(
+    "ETR\\\\textsubscript{max}", 
+    glmm_params_group_locs %>% filter(response == "etr") %>% pull(start),
+    glmm_params_group_locs %>% filter(response == "etr") %>% pull(end),
+    escape = FALSE) %>%
+  pack_rows(
+    "PPFD\\\\textsubscript{95\\\\%}", 
+    glmm_params_group_locs %>% filter(response == "par") %>% pull(start),
+    glmm_params_group_locs %>% filter(response == "par") %>% pull(end),
+    escape = FALSE) %>%
+  pack_rows(
+    "Recovery", 
+    glmm_params_group_locs %>% filter(response == "recovery") %>% pull(start),
+    glmm_params_group_locs %>% filter(response == "recovery") %>% pull(end),
+    escape = FALSE)
 ```
@@
 
-**`r table("pglm")`** Phylogenetic generalized least squares (PLGS) models of desiccation tolerance as a function of environmental water potential (VPD) or the range of gametophytes beyond sporophytes (G beyond S) in filmy ferns from Moorea, French Polynesia
-
-```{r pglm, results = "asis"}
+**`r table("pgls")`** Phylogenetic generalized least squares (PLGS) models of desiccation tolerance as a function of environmental water potential (VPD) or the range of gametophytes beyond sporophytes (G beyond S) in filmy ferns from Moorea, French Polynesia
+
+```{r pgls, results = "asis"}
 env_range_dt_model_summary %>%

### 2021-01-29 ms/captions.R

@@ -22,6 +22,7 @@ table_full <- captioner::captioner(prefix = "Table")
 table_full(name = "physig", caption = "physig")
 table_full(name = "glmm", caption = "glmm")
 table_full(name = "glmm-pars", caption = "glmm-pars")
+table_full(name = "pgls", caption = "pgls")
 
 # - SI figures
 s_figure_full <- captioner::captioner(prefix = "Fig. S", suffix = "")

### 2021-02-01 ms/manuscript.Rmd
Use kableExtra for all tables

@@
 ```{r glmm-table, results = "asis"}
+# Make helper table for formatting by groups
 glmm_summ_group_locs <-
@@
     .groups = "drop"
+  )
+
+# Format column names.
+# left side is new name (with latex formatting)
+# right side is original name
+glmm_summary_cols <-
+  c(
+    "Fixed effects" = "fixed_effects", 
+    "DIC" = "DIC",
+    "delta DIC" = "delta_DIC"
   )
@@
   ) %>%
-  select(fixed_effects, DIC, delta_DIC) %>%
-  kbl(format = "latex", booktabs = TRUE, col.names = c("Fixed effects", "DIC", "delta DIC"), escape = TRUE) %>%
+  select(all_of(glmm_summary_cols)) %>%
+  kbl(format = "latex", booktabs = TRUE, escape = TRUE) %>%
   row_spec(which(glmm_summary$delta_DIC == 0), bold = TRUE) %>%
@@
 
+# Format column names.
+# left side is new name (with latex formatting)
+# right side is original name
 glmm_params_cols <-
@@
 
+# Format cell values
 glmm_params %>%
   mutate(
-    fixed_effects = str_to_sentence(fixed_effects),
+    fixed_effects = case_when(
+      fixed_effects == "both" ~ "Generation + habit",
+      fixed_effects == "intersect" ~ "Generation + habit + Generation:habit",
+      TRUE ~ str_to_sentence(fixed_effects)
+    ),
     predictor = str_replace_all(predictor, "generationsporophyte", "Sporophyte") %>%
@@
       str_replace_all("habitSaxicolous", "Saxicolous") %>%
-      str_replace_all("habitTerrestrial", "Terrestrial")
+      str_replace_all("habitTerrestrial", "Terrestrial"),
+    # Round off long numeric values
+    across(c(post.mean, `l-95% CI`, `u-95% CI`, pMCMC), ~round_t(., 3)),
+    eff.samp = as.integer(eff.samp)
   ) %>%
+  # Format column names
   select(all_of(glmm_params_cols)) %>%
@@
   kable_styling(latex_options = "scale_down") %>%
+  # Make significant rows bold
   row_spec(which(glmm_params$pMCMC < 0.05), bold = TRUE) %>%
+  # Add group rows
   pack_rows(
@@
 ```{r pgls, results = "asis"}
+# Format column names.
+# left side is new name (with latex formatting)
+# right side is original name
+pgls_cols <-
+  c(
+    "Generation" = "generation", 
+    "Predictor" = "predictor",
+    "df" = "df",
+    "\\textit{F}" = "f_value",
+    "adj. \\textit{R}\\textsuperscript{2}" = "adj_r_squared",
+    "\\textit{P}" = "p_value"
+  )
+
+# Format cell values
 env_range_dt_model_summary %>%
@@
   assert(not_na, predictor) %>%
-  select(generation, predictor, df, f_value, adj_r_squared, p_value) %>%
   mutate(
@@
   ) %>%
-  kbl(col.names = c("Generation", "Predictor", "df", "$F$", "adj. $R^2$", "$P$"), escape = FALSE, booktabs = TRUE)
+  # Format column names
+  select(all_of(pgls_cols)) %>%
+  kbl(format = "latex", escape = FALSE, booktabs = TRUE)
 ```

### 2021-02-01 ms/si.Rmd
Source caption numbers in SI from MS
don't need separate captions.R

@@
 
-```{r setup, include=FALSE}
+```{r si-setup, include = FALSE, cache = FALSE, purl=FALSE}
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
@@
 
-```{r load-data, cache = FALSE}
+```{r captions}
+# Source captions from MS
+# so that caption numbers are numbered in order of appearance in MS
+
+# Run inline code when purling
+options(knitr.purl.inline = TRUE)
+
+# Make a temporary file to write out just inline R code from the SI Rmd
+temp_file <- tempfile()
+
+# Generate R script including inline R code from the SI Rmd
+# (add `purl = FALSE` to code chunks in SI Rmd so those don't run)
+knitr::purl(here::here("ms/manuscript.Rmd"), output = temp_file)
+
+# Trim this R script down to only functions that define figure and table captions
+read_lines(temp_file) %>%
+  magrittr::extract(
+    str_detect(., "^figure\\(|^table\\(|^s_figure\\(|^s_table\\(")
+  ) %>%
+  unique %>%
+  write_lines(temp_file)
+
+# Source the code. This will load the caption functions in the right order.
+source(temp_file)
+
+# Delete the temporary file
+fs::file_delete(temp_file)
+```
+
+```{r si-load-data, cache = FALSE}
 tar_load(filmy_dt_chamber)
@@
 
-```{r load-functions, cache = FALSE}
+```{r si-load-functions, cache = FALSE}
 # Define functions specific to this file
@@
 
-```{r plot-setup}
+```{r si-plot-setup}
 # JPR requirements:
@@
 
-**Figure SXX**. Maximum likelihood ultrametric phylogenetic tree including all Moorean filmy fern species.
+**`r s_figure("tree")`**. Maximum likelihood ultrametric phylogenetic tree including all Moorean filmy fern species.
 Tree obtained by extracting Hymenophyllaceae from the tree of Nitta et al. (2016) with the extract.clade function of the ape package in R (Paradis et al., 2004).
@@
 
-**Figure SXX**.
+**`r s_figure("chamber-g")`**.
 Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes.
@@
 
-**Figure SXX**.
+**`r s_figure("chamber-s")`**.
 Conditions inside desiccation chamber and duration of desiccation treatment for sporophytes at most intense desiccation level.
@@
 
-**Figure SXX**.
+**`r s_figure("water-s")`**.
 Relative water content (RWC) of sporophytes during desiccation tolerance test.
@@
 
-**Figure SXX**.
+**`r s_figure("water-all")`**.
 Relative water content (RWC) of sporophytes after desiccation by duration of desiccation treatment (2 or 15 days).
@@
 
-**Figure SXX**.
+**`r s_figure("dt-control")`**.
 Recovery of controls (samples kept on moist tissues) for desiccation tolerance test of filmy fern sporophytes (**a**) and gametophytes (**b**) from Moorea, French Polynesia.

### 2021-02-01 R/functions.R

@@ -1438,9 +1438,10 @@ match_comm_and_tree <- function (comm, phy, return = c("comm", "tree")) {
 # Plots ----
 
 # Format legend (include title, but get rid of boxes)
-legend_theme <- theme (legend.background = element_rect(colour = "transparent", fill = "transparent", size = 0.5),
-                       legend.key = element_rect(colour = "white", fill = "white", size = 0.5),
-                       legend.text = element_text(size = 12))
+legend_theme <- ggplot2::theme(
+  legend.background = ggplot2::element_rect(colour = "transparent", fill = "transparent", size = 0.5),
+  legend.key = ggplot2::element_rect(colour = "white", fill = "white", size = 0.5),
+  legend.text = ggplot2::element_text(size = 12))
 
 make_sporo_dt_plot <- function  (recovery_species_means, traits) {
   # Define color-blind palette
@@ -1664,3 +1665,33 @@ latex2docx <- function (latex, docx, template = NULL, lua_filter = NULL, wd = ge
   
 }
 
+# Captions ----
+
+# Follow Journal of Plant Research (JPR) style:
+# - All figures are to be numbered using Arabic numerals.
+# - Figures should always be cited in text in consecutive numerical order.
+# - Figure parts should be denoted by lowercase letters (a, b, c, etc.).
+# - If an appendix appears in your article and it contains one or more figures, 
+#   continue the consecutive numbering of the main text. 
+# - Do not number the appendix figures, "A1, A2, A3, etc." 
+# - Figures in online appendices (Electronic Supplementary Material) should, 
+#   however, be numbered separately
+
+# - First define the "full" version, which would include a caption
+# (except I never use the caption in the function, and instead replace with 'blank')
+figure_full <- captioner::captioner(prefix = "Fig.", suffix = "")
+table_full <- captioner::captioner(prefix = "Table")
+s_figure_full <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = "")
+s_table_full <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = "")
+
+# - Make a short function that prints only the object type and number, e.g., "Fig. 1"
+figure <- pryr::partial(figure_full, display = "cite", caption = "blank")
+table <- pryr::partial(table_full, display = "cite", caption = "blank")
+s_figure <- pryr::partial(s_figure_full, display = "cite", caption = "blank")
[…]

### 2021-02-01 _targets.R

@@ -1,7 +1,7 @@
 source("R/packages.R")
 source("R/functions.R")
-source("ms/captions.R")
 
+# Define workflow plan
 tar_plan(
   
   # Load data ----

### 2021-02-01 ms/captions.R

@@ -1,42 +1,0 @@
-# Set up captions
-
-# Follow Journal of Plant Research (JPR) style:
-# - All figures are to be numbered using Arabic numerals.
-# - Figures should always be cited in text in consecutive numerical order.
-# - Figure parts should be denoted by lowercase letters (a, b, c, etc.).
-# - If an appendix appears in your article and it contains one or more figures, 
-#   continue the consecutive numbering of the main text. 
-# - Do not number the appendix figures, "A1, A2, A3, etc." 
-# - Figures in online appendices (Electronic Supplementary Material) should, 
-#   however, be numbered separately
-
-# - Figures
-figure_full <- captioner::captioner(prefix = "Fig.", suffix = "")
-figure_full(name = "habit", caption = "habit")
-figure_full(name = "dt", caption = "dt")
-figure_full(name = "combined", caption = "combined")
-figure_full(name = "corr", caption = "corr")
-
-# - Tables
-table_full <- captioner::captioner(prefix = "Table")
-table_full(name = "physig", caption = "physig")
-table_full(name = "glmm", caption = "glmm")
-table_full(name = "glmm-pars", caption = "glmm-pars")
-table_full(name = "pgls", caption = "pgls")
-
-# - SI figures
-s_figure_full <- captioner::captioner(prefix = "Fig. S", suffix = "")
-
-# - SI tables
-s_table_full <- captioner::captioner(prefix = "S", auto_space = FALSE, suffix = " Table. ")
-
-# Make short versions of citation functions (just the object type and number, e.g., "Fig. 1")
-figure <- pryr::partial(figure_full, display = "cite")
-table <- pryr::partial(table_full, display = "cite")
-s_figure <- pryr::partial(s_figure_full, display = "cite")
-s_table <- pryr::partial(s_table_full, display = "cite")
-
-# just the number (e.g., "1")
-table_num <- function (name) {table(name) %>% str_remove("Table ")}
-figure_num <- function (figure) {table(name) %>% str_remove("Fig. ")}
-

### 2021-02-01 ms/manuscript.Rmd
Update MS

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
+```{r ms-setup, include = FALSE, cache = FALSE}
+knitr::opts_chunk$set(
+  echo = FALSE,     # Don't display the source code
+  message = FALSE,  # Don't preserve messages
+  warning = FALSE,  # Don't preserve warnings 
+  results = "hide", # Hide text output
+  cache = FALSE    # Don't use the cache
+  )
 ```
@@
 Many bryophytes have retained vegetative DT in the gametophyte stage [@Oliver2005, and references therein].
-In extant vascular plants (tracheophytes), DT occurs in certain non-vegetative stages (e.g., seeds and pollen), but has largely been lost in vegetative tissues (with some notable exceptions, the so-called 'resurrection plants').
+In extant vascular plants (tracheophytes), DT occurs in certain non-vegetative stages (\eg, seeds and pollen), but has largely been lost in vegetative tissues (with some notable exceptions, the so-called 'resurrection plants').
 Instead, tracheophytes avoid desiccation with adaptations including a waxy cuticle, stomata, and an extensive vascular system.
@@
 Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
-These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often <1 cm), lack vascular tissue, and are poikilohydric.
+These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (\ie, they regulate their internal water content), whereas fern gametophytes are much smaller (often < 1 cm), lack vascular tissue, and are poikilohydric.
 Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013].
@@
 Filmy ferns (Hymenophyllaceae) represent a particularly interesting case for studying the ecological and evolutionary significance of vegetative DT across life stages in ferns.
-Filmy ferns are a large family (\circa{ }430 spp.; PPG I, 2016) of primarily tropical, leptosporangiate ferns named for their remarkably thin leaf laminae, which are usually only a single cell layer in thickness and lack cuticle or stomata (Fig.
-1A--D).
+Filmy ferns are a large family [\circa{ }430 spp.; @PteridophytePhylogenyGroupI2016] of primarily tropical, leptosporangiate ferns named for their remarkably thin leaf laminae, which are usually only a single cell layer in thickness and lack cuticle or stomata (`r figure("habit")`a--d).
 Filmy fern sporophytes are an exception to the 'typical' fern sporophyte described above because they have reverted to poikilohydry: their thin laminae allow for passive, rapid uptake and loss of moisture, thereby maintaining nearly constant equilibrium with atmospheric humidity [@Raven1999].
@@
 Filmy fern sporophytes encompass a broad diversity of growth forms (terrestrial, saxicolous, hemi-epiphytic, and holoepiphytic) and morphologies, with rhizome and leaf sizes varying over an order of magnitude [@Hennequin2008; @Nitta2009; @Dubuisson2013].
-Filmy fern gametophytes are similarly diverse, with morphotypes including filamentous, ribbon, and a combination of these two types (Fig. 1E--H).
+Filmy fern gametophytes are similarly diverse, with morphotypes including filamentous, ribbon, and a combination of these two types (`r figure("habit")`e--h).
 While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages [@WatkinsJr.2007a; @WatkinsJr.2012; @Pittermann2013], and may be expected to occur in filmy ferns as well.
@@
 The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
-In a series of elegant experiments on Jamaican filmy ferns, @Shreve1911 demonstrated that water-use strategies differ between species occupying contrasting habitats (i.e., terrestrial, low elevation epiphyte, and high elevation epiphyte).
+In a series of elegant experiments on Jamaican filmy ferns, @Shreve1911 demonstrated that water-use strategies differ between species occupying contrasting habitats (\ie, terrestrial, low elevation epiphyte, and high elevation epiphyte).
 Proctor [-@Proctor2003; -@Proctor2012] investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs.
@@
 In a recent study (Testo et al., unpublished), the combined effect of desiccation and light stress was found to have a greater impact on filmy fern physiological performance than either of these stressors alone, suggesting the need to integrate studies of DT with photosynthesis to gain a full picture of physiological adaptation in filmy ferns.
-Here, we leverage a recently developed DNA barcoding system [@Nitta2017] to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes: 1.) Epiphytic filmy ferns have higher DT and photosynthetic optima than non-epiphytic species.
-2.) Amongst epiphytic filmy ferns, those occurring at high elevation have higher DT and photosynthetic optima than those at low elevation.
-3.) Gametophytes with wider ranges have higher DT than those with with narrow ranges.
+Here, we leverage a recently developed DNA barcoding system [@Nitta2017] to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes:
+1.) Epiphytic filmy ferns have higher DT and photosynthetic optima than terrestrial species.
+2.) Degree of DT is correlated with environmental water availability.
+3.) Gametophytes with wider ranges have higher DT than those with with narrower ranges.
 
@@
 
-We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology (Fig. 2).
-Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
-15 km NW of its larger sister island, Tahiti.
+We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology.
+Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located \circa{ }15 km NW of its larger sister island, Tahiti.
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
-A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (`r figure("habit")`).
+A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (Figs.
+`r figure_num("habit")`, `r s_figure_num("tree")`).
 The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
-We follow the taxonomic system of @Ebihara2006.
-
-### Field survey
+We follow the taxonomic system of @Ebihara2006 for Hymenophyllaceae.
+
+### Field survey and DNA barcoding
 
@@
 Tohiea (1206 m).
-At each site, sporophytes were sampled in a single 10 m × 10 m plot, and gametophytes sampled in two to three 50 cm × 50 cm subplots located within the sporophyte plot.
+At each site, sporophytes were sampled in a single 10 m × 10 m plot, and gametophytes sampled in two to three 50 cm × 50 cm plots located within the sporophyte plot.
+To estimate abundance of sporophytes, the 10 m × 10 m plot was subdivided into 25 2 m × 2 m subplots, and the presence of species in the sporophyte phase in each subplot was summed such that each species was assigned an abundance rank from 1 to 25.
+For the gametophyte plots, each 50 cm x 50 cm plot was subdivided into 25 5 x 5 cm squares, and a single gametophyte selected closest to the center of each square, to later be identified to species using DNA barcoding (see below).
 Epiphytes were sampled to a maximum height of \circa{ }2 m on their host plants.
-We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, \ie, below \circa{ }500 m), or high elevation epiphyte (mostly occurring in cloud forest, \ie, above \circa{ }500 m).
+For additional details on sampling procedure, see @Nitta2017.
+The growth habit of each species was scored based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, \ie, below \circa{ }500 m), or high elevation epiphyte (mostly occurring in cloud forest, \ie, above \circa{ }500 m).
+
+Unlike sporophytes, fern gametophytes are cryptic and cannot be identified to species based on morphology alone. 
+Therefore, we used DNA barcoding to identify filmy fern gametophytes to species as described in @Nitta2017.
+Briefly, DNA extraction was performed using a modified CTAB protocol [@Doyle1987] or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
+Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following @Nitta2017.
+For each marker, a reference library was constructed from sporophyte sequences using the 'makeblastDB' command in BLAST [@Altschul1997].
+Field-collected gametophytes were identified by a local BLAST query against the reference library.
+Gametophytes matching > 99.5 % with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched > 99.5 % with multiple reference sequences were not identified and excluded from further analysis.
+For this study, only gametophytes that could be identified as belonging to Hymenophyllaceae were used, and others were excluded from analysis.
 
@@
 We used microclimate data collected by @Nitta2020.
-This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA).
+This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) and vapor pressure deficit (VPD) calculated from these two values.
 The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at \circa{ }1.5 m on a tree, one 'terrestrial' datalogger at \circa{ }10 cm next to the tree).
@@
 However, epiphytic habitats tend to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones.
-This difference is greatest at low elevations, and smallest at high elevations.
+This difference is greatest at low elevations, and smallest at high elevations [@Nitta2020].
 
@@
 Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
-Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
+Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
 Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
-The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22°C to 25°C.
+The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22 °C to 25 °C (Figs. `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
 Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
-Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at -86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
-Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
-Following the desiccation treatment, plants were rewetted using water from a local stream.
+Following the desiccation treatment, plants were rewetted by wrapping them in wet tissues using water from a local stream.
 Plants were allowed to rehydrate, and F~v~/F~m~ was again measured at 0.5 h, 24 h, and 48 h following rewetting.
 Ability to recover from desiccation was quantified by comparing pre- and post-treatment F~v~/F~m~ [@WatkinsJr.2007c].
+
+We measured relative water content (RWC) in a subset of samples by recording the mass of samples prior to the DT test (fresh mass), at each step of the DT test (turgid mass), then after drying them overnight in a drying oven at 65 °C following the DT test (dry mass).
+Relative water content was calculated as \(RWC = \frac{TM - DM}{FM - DM} \times 100\), where *TM* is turgid mass, *DM* is dry mass, and *FM* is fresh mass.
+
+Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible.
+We therefore applied the same treatment for all gametophytes (2 d at -86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described above.
+Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
+Relative water content was not calculated for gametophytes, as these were too small to measure accurately with the balances available.
 
@@
 Response curves were fitted using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded [@Proctor2012].
-The PPFD at 95% saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
+The PPFD at 95 % saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
 Light responses were measured in 8--16 field-collected sporophyte individuals for each species, and on single gametophyte individuals in the field, which were later identified to species using the DNA barcodes.
 
-### DNA barcoding
-
-We used DNA barcoding to identify filmy fern gametophytes to species as described in @Nitta2017.
-Briefly, DNA extraction was performed using a modified CTAB protocol [@Doyle1987] or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
-Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following @Nitta2017.
-For each marker, a reference library was constructed from sporophyte sequences using the makeblastDB command in BLAST [@Altschul1997].
-Field-collected gametophytes were identified by a local BLAST query against the reference library.
-Gametophytes matching >99.5% with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched >99.5% with multiple reference sequences were not identified and excluded from further analysis.
-For this study, only gametophytes that could be identified as belonging to Hymenophyllaceae were used, and others were excluded from analysis.
-
-### Statistical analysis and phylogenetic comparative methods
+### Statistical analysis
 
 All statistical analyses were done in R v.4.0.2 [@RCoreTeam2020a].
-Datasets and R scripts used in this study are available on Dryad (\#\#\#\#\#\#).
-
-We derived the phylogenetic tree used in this study by extracting Hymenophyllaceae from the tree of @Nitta2017 with the extract.clade function of the ape package in R [@Paradis2004].
-This tree is well-supported and generally in accord with previously published, more densely sampled phylogenies [@Pryer2001a; @Dubuisson2003; @Hennequin2006; @Ebihara2007a].
-
-We quantified the degree of phylogenetic signal in physiological traits using Blomberg's *K* [@Blomberg2003] and Pagel's λ [@Pagel1999; @Freckleton2002] as implemented with the phylosig function in the phytools R package [@Revell2012].
+
+We derived the phylogenetic tree used in this study by extracting Hymenophyllaceae from the tree of @Nitta2017 with the \function{extract.clade} function of the \package{ape} package [@Paradis2004].
+This tree is in accord with previously published, more densely sampled phylogenies [@Pryer2001a; @Dubuisson2003; @Hennequin2006; @Ebihara2007a].
+
+We quantified the degree of phylogenetic signal in physiological traits using Blomberg's *K* [@Blomberg2003] and Pagel's λ [@Pagel1999; @Freckleton2002] as implemented with the \function{phylosig} function in the \package{phytools} package [@Revell2012].
 Both measures test the hypothesis that the trait of interest is evolving according to Brownian Motion (BM).
@@
 For *K*, values near one indicate evolution of traits following BM; *K* >1 indicates traits more conserved than expected under BM, and *K* <1 indicates that traits have less phylogenetic signal than expected under BM.
-We tested the significance of *K* by comparing the observed value against values from a null distribution of 1000 phylogenies with the traits randomly shuffled across the tips.
+We tested the significance of *K* by comparing the observed value against values from a null distribution of \num{1000} phylogenies with the traits randomly shuffled across the tips.
 We tested the significance of λ with a log-likelihood test comparing the likelihood of the observed value of λ vs. λ = 0.
@@
 
-We used general linear mixed models (GLMMs) to investigate the effect of growth habit and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at -86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the MCMCglmm R package [@Hadfield2010].
-For each of the three response variables (DT, ETR~max~ , and PPFD~95%~), we used growth habit (high-elevation epiphyte, low-elevation epiphyte, terrestrial or saxicolous) and generation (sporophyte vs.
-gametophyte) and their interactions as fixed effects.
+We used general linear mixed models (GLMMs) to investigate the effect of growth habit and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at -86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the \package{MCMCglmm} package [@Hadfield2010].
+For each of the three response variables (DT, ETR~max~, and PPFD~95%~), we constructed five models: a null model with no fixed effects, or models with either growth habit (high-elevation epiphyte, low-elevation epiphyte, terrestrial or saxicolous), generation (sporophyte vs.
+gametophyte), growth habit and generation, or growth habit, generation, and their interaction as fixed effects.
 We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix in the case the response variable had significant phylogenetic signal [@Hadfield2010].
 We used an inverse-Gamma distribution for all priors, and ran analyses for 500,000 iterations, with burnin after 1,000 iterations and thinning every 500 iterations.
-Comparisons of DT, ETR~max~, and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a *t*-test for these.
+We compared models using the Deviance Information Criterion (DIC) [@Spiegelhalter2002] and selected the best-fitting model as the one with the lowest DIC.
+
+Comparisons of DT, ETR~max~, and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a two-sided *t*-test for these.
 
@@
 Data from the Mt.
-Rotui 830 m site were excluded because this site is much more exposed than the other high elevation sites, and appears as a significant outlier in the microclimatic data (see Results).
-This may be due to its location; unlike other mountains on Moorea, Mt. Rotui is completely isolated with no connecting ridges to other peaks (Fig. 2).
-Since DT showed significant phylogenetic signal (see Results), we used phylogenetic generalized least squares (PGLS) to test for a relationship between environmental water potential and DT while taking phylogeny into account, using species' means.
-
-To test whether gametophytes with wider ranges have higher DT, we conducted PGLS with DT as the response variable and range width (measured as the difference between maximum and minimum observed elevation) as the explanatory variable.
-We ran PGLS using the caper R package [@Orme2018].
+Rotui 830 m site were excluded because this site is much more exposed than the other high elevation sites, and appears as a significant outlier in the microclimatic data [@Nitta2020].
+Since DT showed significant phylogenetic signal (see Results), we used phylogenetic generalized least squares (PGLS) [@Freckleton2002] to test for a relationship between environmental water potential and DT while taking phylogeny into account, using species' means.
+
+To test whether gametophytes with wider ranges have higher DT, we calculated the extent of gametophyte range beyond sporophytes ("G beyond S") as the total elevational range (m) gametophytes were observed growing beyond sporophytes at either end of the elevational gradient.
+We conducted PGLS with DT as the response variable and G beyond S as the explanatory variable with the \function{pgls} function in the \package{caper} [@Orme2018].
+
+Datasets used in this study are available on FigShare (\#\#\#\#\#\#). 
+Code used to analyze the data and generate figures and this manuscript are available on GitHub (https://github.com/joelnitta/moorea-filmies).
+A Docker image is available to run the code at https://hub.docker.com/r/joelnitta/moorea_filmies.
+The \package{targets} package [@Landau2021] was used to automate the workflow.
+The \package{ggplot2} [@Wickham2016], \package{ggtree}, [@Yu2017] and \package{patchwork} (https://github.com/thomasp85/patchwork) packages were used to generate the figures.
 
@@
 Iwats.
-species complex, but species status of these taxa remains uncertain (Nitta et al., 2011a).
+species complex, but species status of these taxa remains uncertain [@Nitta2011].
 We refer to these here as *C. minutum* var. 1, 2, and 3.
@@
 Iwats.
-in our tree; Appendix S1) were observed on the basis of *rbcL* and morphology, but as taxonomic treatment is beyond the scope of the current study we refer to them as *A.
+in our tree; `r s_figure("tree")`) were observed on the basis of *rbcL* and morphology, but as taxonomic treatment is beyond the scope of the current study we refer to them as *A.
 asae-grayi* var.
@@
 
-Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
-A general correlation of DT with habitat was observed in sporophytes (`r figure("dt")`**a**).
+Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases.
+A general correlation of DT with habitat was observed in sporophytes (`r figure("dt")`a).
 Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats.
 and *C.
-apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 MPa (more stringent tests were not conducted on *A.
+apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 NaCl (more stringent tests were not conducted on *A.
 dentatum* after confirming its lack of DT at this level).
@@
 
-Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; `r figure("dt")`**b**).
+Control samples generally did not show any decrease in F~v~/F~m~ throughout the duration of the experiment (`r s_figure("dt-control")`). 
+Relative water content reached \circa{ }0 % in the desiccated state (`r s_figure("water-all")`), and returned to \circa{ }100 % upon re-wetting in all species sampled (`r s_figure("water-s")`).
+
+Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; `r figure("dt")`b).
+
+Relative humidity in the growth chambers was high when samples were first inserted, then decreased to the target value maintained by each drying salt (Figs.
+`r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
+Brief fluctuations to higher humidities were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).
 
@@
 polyanthos*, and *P.
-borbonicum* (Fig.
-5; *t*-test, *P* = XXX, XXX, XXX, and XXX, respectively).
+borbonicum* (`r figure("combined")`; *t*-test, *P* = XXX, XXX, XXX, and XXX, respectively).
 
@@
 Since controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (Appendix S3), it is unlikely that the lack of tolerance seen in terrestrial species is due solely to removal from the soil.
-Comparison of DT with the minimum water potential envelope of each species (Fig.
-6) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with minimum water potential exceeding \circa{ }-7 MPa.
+Comparison of DT with the minimum water potential envelope of each species (`r figure("corr")`) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with minimum water potential exceeding \circa{ }-7 MPa.
 In contrast, habitats with minimum water potential below -7 MPa are nearly all epiphytic, and the filmy fern sporophytes occurring there are all capable of DT.
@@
 However, we found that within high elevation epiphytes, gametophytes had lower DT than sporophytes, and only one species (saxicolous *P.
-endlicherianum*) with gametophytes that had significantly greater DT than sporophytes (Fig.
-5).
+endlicherianum*) with gametophytes that had significantly greater DT than sporophytes (`r figure("compare")`).
 At high elevation sites (\ie, cloud forest), epiphytic ferns often occur in a dense matrix of bryophytes (Fig. 1B).
@@
 Although we predicted that species with gametophytes that occur beyond the range of their sporophytes would have higher DT and be adapted for higher light levels, this was not the case.
-This is perhaps best illustrated by *C. apiifolia* (Fig. 1D, E, H).
+This is perhaps best illustrated by *C. apiifolia* (`r figure("habit")`d, e, h).
 Sporophytes of this species are usually restricted to moist cloud forest sites from \circa{ }600 m to 1200 m, but we observed gametophytes over a much wider range, down to 200 m.
@@
 apiifolia* do not actually experience the conditions measured with our 'terrestrial' dataloggers, but instead grow within a much more consistently moist boundary layer.
-Gametophytes of *C. apiifolia* produce gemmae (Fig. 1E, H), and often form dense clonal mats.
-In the case that their microhabitats do become too dry, they may suffer loss at the edge of the population, then recover by clonal growth (Farrar et al., 2008).
+Gametophytes of *C. apiifolia* produce gemmae (`r figure("habit")`e, h), and often form dense clonal mats.
+In the case that their microhabitats do become too dry, they may suffer loss at the edge of the population, then recover by clonal growth [@Farrar2008].
 Furthermore, the complex surface of clonal mats has been shown to increase the boundary layer and decrease evaporation rates in mosses [@Rice2001], and may function similarly in *C. apiifolia*.
@@
 Unlike DT, photosynthetic parameters did not vary significantly between species from different habitats.
-Rather, we observed a clear intergenerational difference, with gametophytes consistently adapted for lower light levels than sporophytes (Fig. 5, Table 3).
+Rather, we observed a clear intergenerational difference, with gametophytes consistently adapted for lower light levels than sporophytes (`r figure("combined")`, `r table("glmm")`).
 In the only other study to our knowledge that compared photosynthetic rates between gametophytes and sporophytes of a filmy fern, @Johnson2000 also found that gametophytes of *V.
@@
 Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates [@Proctor2003; @Proctor2012; @Parra2009; @Saldana2013]; however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
-We detected significant phylogenetic signal in DT, suggesting that phylogeny should be considered when interpreting differences in DT between species (Table 2).
+We detected significant phylogenetic signal in DT, suggesting that phylogeny should be considered when interpreting differences in DT between species (`r table("physig")`).
 In contrast, the lack of phylogenetic signal in photosynthetic traits, coupled with the clear intergenerational differences we observed, suggest that sporophytes and gametophytes are adapted for different light levels regardless of clade or ecological niche.
@@
 
+**Electronic supplementary material**
+
+The online version of this article (DOI) contains supplementary material, which is available to authorized users.
+
 ## References
@@
 
-```{r table-setup}
+```{r table-setup, purl = FALSE}
 options(knitr.table.format = "latex")
@@
 
-```{r physig-table, results = "asis"}
+```{r physig-table, results = "asis", purl = FALSE}
 # Format phylogenetic signal results as a pretty table with kableExtra
@@
 
-```{r glmm-table, results = "asis"}
+```{r glmm-table, results = "asis", purl = FALSE}
 # Make helper table for formatting by groups
@@
 
-```{r glmm-params, results = "asis"}
+```{r glmm-params, results = "asis", purl = FALSE}
 # Make helper table for formatting by groups
@@
 
-```{r pgls, results = "asis"}
+```{r pgls, results = "asis", purl = FALSE}
 # Format column names.
@@
 
-```{r plot-setup}
+```{r plot-setup, purl = FALSE}
 # JPR requirements:
@@
 
-```{r dt-plot-setup}
+```{r dt-plot-setup, purl = FALSE}
 ### Format data for plotting ###
@@
 
-```{r filmies-fig, fig.width = 6.85, results = "asis"}
+```{r filmies-fig, fig.width = 6.85, results = "asis", purl = FALSE}
 knitr::include_graphics(here::here("data/figure_1.png"))
@@
 
-```{r filmy-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = ""}
+```{r filmy-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = "", purl = FALSE}
 dt_plot_data %>%
@@
 
-```{r gameto-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = ""}
+```{r gameto-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = "", purl = FALSE}
 make_dt_plot(dt_plot_data, "gametophyte", "b")
@@
 
-```{r combined-plot, fig.width = 6.85, fig.height = 5, fig.cap = ""}
+```{r combined-plot, fig.width = 6.85, fig.height = 5, fig.cap = "", purl = FALSE}
 # Copy data so they can be manipulated for this plot
@@
 
-```{r corr-plot, fig.width = 6.85, fig.height = 5, fig.cap = ""}
+```{r corr-plot, fig.width = 6.85, fig.height = 5, fig.cap = "", purl = FALSE}
 # Combine recovery, VPD, and growth habit into a single dataframe (wide format)

### 2021-02-01 ms/manuscript.Rmd
Update MS

@@
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
-A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (Figs.
-`r figure_num("habit")`, `r s_figure_num("tree")`).
+A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (\figs `r figure_num("habit")`, `r s_figure_num("tree")`).
 The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
@@
 Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
-The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22 °C to 25 °C (Figs. `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
+The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22 °C to 25 °C (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
 Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
@@
 
-# FIXME: double check on these p-values
-# ttest_pval("Polyphlebium_endlicherianum", "recovery")
-# ttest_pval("Crepidomanes_minutum2", "recovery")
-# ttest_pval("Hymenophyllum_digitatum", "recovery")
-# ttest_pval("Hymenophyllum_polyanthos", "recovery")
-# ttest_pval("Polyphlebium_borbonicum", "recovery")
+t_test_results
+
+# Extract list of species with significant p-values for gameto DT less than sporo
+dt_g_less_than_s <-
+t_test_results %>% 
+  filter(est_diff_sporo_gameto > 0, response == "recovery", p_value < 0.05) %>%
+  abbrev_sp() %>%
+  mutate(species = str_replace_all(species, "_", ". ")) %>%
+  pull(species) %>%
+  knitr::combine_words(before = "*")
+
+# Extract p-values for species with gameto DT less than sporo
+dt_g_less_than_s_pvals <-
+t_test_results %>%
+  filter(est_diff_sporo_gameto > 0, response == "recovery", p_value < 0.05) %>%
+  pull(species) %>%
+  map_chr(~ttest_pval(species_select = ., response_select = "recovery", data = t_test_results)) %>%
+  knitr::combine_words()
 ```
@@
 Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases.
+
 A general correlation of DT with habitat was observed in sporophytes (`r figure("dt")`a).
@@
 Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
-Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.* *pallidum* and *P.* *borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
+Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.
+pallidum* and *P.
+borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
 Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.
@@
 
+DT of gametophytes did not significantly exceed that of sporophytes for any species, and was significantly lower in `r dt_g_less_than_s` (`r figure("combined")`; *t*-test, *P* = `r dt_g_less_than_s_pvals`, respectively).
+
 Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; `r figure("dt")`b).
 
-Relative humidity in the growth chambers was high when samples were first inserted, then decreased to the target value maintained by each drying salt (Figs.
-`r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
+Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
 Brief fluctuations to higher humidities were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).
 
-DT of gametophytes did not significantly exceed that of sporophytes for any species except for saxicolous *P.
-endlicherianum* (Fig.
-5; *t*-test, *P* = ), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C.
-minutum* var.
-2, *H.* *digitatum*, *H.
-polyanthos*, and *P.
-borbonicum* (`r figure("combined")`; *t*-test, *P* = XXX, XXX, XXX, and XXX, respectively).
-
-The GLMM including the effect of generation only was selected as the best-fitting model for DT (`r table("glmm")`) while taking phylogeny into account.
+The GLMM including the effect of generation only was selected as the best-fitting model for DT (`r table("glmm")`). 
 Sporophytes had `r sporo_dt_effect_mean` greater DT (95% CI `r sporo_dt_effect_lower` to `r sporo_dt_effect_upper`) relative to gametophytes (`r table("glmm-pars")`).
@@
 
-Light responses were measured in sporophytes of `r n_light_species_sporo` species and gametophytes of `r n_light_species_gameto` species, including `r n_light_species_both` species with both life phases (Appendix S4).
-Unlike DT, photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes are adapted for higher light levels, whereas gametophytes are adapted for lower light levels (Fig.
-5).
+Light responses were measured in sporophytes of `r n_light_species_sporo` species and gametophytes of `r n_light_species_gameto` species, including `r n_light_species_both` species with both life phases.
+
+Photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (`r figure("combined")`).
 Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
@@
 Similarly, sporophytes attained higher rates of electron transport (ETR~max~; `r mean_etr_sporo` µmol·m^--2^ ·s^--1^) than gametophytes (`r mean_etr_gameto` µmol·m^--2^ ·s^--1^).
-The GLMMs showed a significant effect of generation, but neither range type nor growth habit, on PPFD~95%~ and ETR~max~ (Table 3).
+
+The GLMM including the effect of generation only was selected as the best-fitting model for ETR~max~, and growth habit + generation was selected for PPFD~95%~ (`r table("glmm")`).
+However, in the GLMM for PPFD~95%~, only the generation parameter was significant (`r table("glmm-pars")`).
 
@@
 Desiccation tolerance tended to increase with VPD for sporophytes, but not gametophytes (`r figure("corr")`).
-These trends were marginally non-significant when analyzed using PGLS for sporophytes (*P* = `r sporo_vpd_pval`) and not significant for gametophytes (*P* = `r gameto_vpd_pval`; `r table("pgls")`).
-
-No significant relationship was detected between range size of gametophytes and DT (`r figure("corr")`, `r table("pgls")`).
+This trend was marginally non-significant when analyzed using PGLS (*P* = `r sporo_vpd_pval`; `r table("pgls")`).
+
+No significant relationship was detected between range size of gametophytes beyond sporophytes and DT (`r figure("corr")`, `r table("pgls")`).
 

### 2021-02-01 R/functions.R

@@ -1590,7 +1590,7 @@ pretty_text <- function (x) {
 #' @param p_check Boolean: should the p-value be required to be less than 0.05?
 #'
 #' @return A p-value for the selected species and response
-ttest_pval <- function(species_select, response_select, data = t_test_results, p_check = TRUE) {
+ttest_pval <- function(species_select, response_select, data, p_check = TRUE) {
   data %>% 
     # Filter to appropriate species and response
     filter(species == species_select, response == response_select) %>%
@@ -1599,7 +1599,7 @@ ttest_pval <- function(species_select, response_select, data = t_test_results, p
     # Format the p-value
     mutate(
       p_value = case_when(
-        p_value < 0.001 ~ "<0.001",
+        p_value < 0.001 ~ "< 0.001",
         TRUE ~ jntools::round_t(p_value, digits = 3))
     ) %>% 
     pull(p_value)

### 2021-02-01 R/functions.R
Calculate mean of maximum daily VPD instead of overall mean

@@ -645,16 +645,19 @@ calculate_mean_light <- function (data) {
     )
 }
 
-#' Calculate mean VPD from climate data
+#' Calculate mean of daily env_range_dt_model VPD from climate data
 #'
 #' @param climate Dataframe with climate data including
 #' relative humidity, temperature, and VPD calculated from these measured
 #' once every 15 minutes by site and growth habit (terrestrial or epiphytic)
 #'
-#' @return Tibble with mean VPD by site and growth habit 
-calculate_mean_vpd <- function (climate) {
+#' @return Tibble with mean of daily env_range_dt_model VPD by site and growth habit 
+calculate_mean_max_vpd <- function (climate) {
   climate %>% 
-    group_by(site, habit) %>% 
+    group_by(site, habit, date) %>% 
+    summarize(vpd = max(vpd), .groups = "drop") %>% 
+    assert(not_na, vpd) %>%
+    group_by(site, habit) %>%
     summarize(vpd = mean(vpd), .groups = "drop") %>% 
     assert(not_na, vpd)
 }

### 2021-02-01 _targets.R

@@ -71,8 +71,8 @@ tar_plan(
   # - calculate water content during DT test per individual
   rel_water_indiv = calculate_indiv_water(filmy_dt),
   
-  # - calculate mean VPD per datalogger
-  mean_vpd = calculate_mean_vpd(climate),
+  # - calculate mean maximum daily VPD per datalogger
+  mean_vpd = calculate_mean_max_vpd(climate),
   
   # - calculate mean VPD per species for gametophytes
   mean_vpd_gameto = calculate_mean_vpd_gameto(

### 2021-02-01 ms/si.Rmd
Fix detecting all table or figure caption functions
when getting captions for SI from MS

@@
   magrittr::extract(
-    str_detect(., "^figure\\(|^table\\(|^s_figure\\(|^s_table\\(")
+    str_detect(., "^figure\\(|^table\\(|^s_figure\\(|^s_table\\(|^figure_num\\(|^table_num\\(|^s_figure_num\\(|^s_table_num\\(")
   ) %>%

### 2021-02-01 ms/si.Rmd
WIP: add SI figure showing range comparison between G and S

@@
 tar_load(filmy_habit)
+tar_load(range_comparison)
 ```
@@
                red = "#D55E00", purple = "#CC79A7")
+
+# Set colors for growth habit, generation
+habit_colors <- c(
+  "Terrestrial" = cbPalette[["red"]], 
+  "Saxicolous" = cbPalette[["gold"]], 
+  "Low_Epiphyte" = cbPalette[["green"]],
+  "High_Epiphyte" = cbPalette[["blue"]])
 ```
@@
 
-```{r gametophyte-dt-2013, fig.height = 4.5, fig.width = 7, fig.cap = ""}
-# Calculate mean start/end times for gametophyte DT groups
-gameto_mean_times <-
-  filmy_dt %>%
-  filter(generation == "gametophyte") %>%
-  select(individual, contains("time_")) %>%
-  pivot_longer(names_to = "condition", values_to = "date_time", -individual) %>%
-  filter(!is.na(date_time)) %>%
-  mutate(condition = str_remove_all(condition, "time_") %>%
-           str_replace_all("30min", "post")) %>%
-  filter(condition %in% c("pre", "post")) %>%
-  add_count(individual) %>%
-  filter(n > 1) %>%
-  select(-n) %>%
-  pivot_wider(names_from = "condition", values_from = "date_time") %>%
-  # Cluster starts times into groups (here, I know there are 10 groups)
-  add_clusters("pre", 10) %>%
-  mutate(cluster = as.factor(cluster)) %>%
-  group_by(cluster) %>%
-  summarize(pre = mean(pre), post = mean(post), n = n(), .groups = "drop")
-
-gameto_time_limits <- c(min = min(gameto_mean_times$pre), max = max(gameto_mean_times$post))
-
-# DT start/end times plot
-dt_times_gameto_2013_plot <- ggplot(
-  gameto_mean_times, 
-  aes(x = cluster, ymin = pre, ymax = post, y = post, group = cluster)) +
-  geom_linerange(position = position_dodge(0.5)) +
-  geom_point(position = position_dodge(0.5)) +
-  scale_y_datetime(limits = gameto_time_limits) +
-  coord_flip() +
-  labs(x = "Group") +
-  theme(
-    axis.title.x = element_blank()
-  )
-
-# Subset DT chamber data
-chamber_gameto_2013 <- filmy_dt_chamber %>% filter(generation == "gametophyte", year == "2013")
-
-# RH plot
-chamber_gameto_2013_rh_plot <-
-  plot_chamber_rh(chamber_gameto_2013, gameto_time_limits)
-
-# Temp plot
-chamber_gameto_2013_temp_plot <-
-  plot_chamber_temp(chamber_gameto_2013, gameto_time_limits)
-
-chamber_gameto_2013_temp_plot + chamber_gameto_2013_rh_plot + dt_times_gameto_2013_plot + plot_layout(ncol = 1)
-
-# ggsave(here::here("results/S1_fig.png"))
-```
-
-**`r s_figure("chamber-g")`**.
-Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes.
-MgNO~3~ was used to maintain constant humidity at \circa{ }58%.
-All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days unintentionally
-
-\clearpage
+\newpage
 
@@
 
-\clearpage
+\newpage
+
+```{r gametophyte-dt-2013, fig.height = 4.5, fig.width = 7, fig.cap = ""}
+# Calculate mean start/end times for gametophyte DT groups
+gameto_mean_times <-
+  filmy_dt %>%
+  filter(generation == "gametophyte") %>%
+  select(individual, contains("time_")) %>%
+  pivot_longer(names_to = "condition", values_to = "date_time", -individual) %>%
+  filter(!is.na(date_time)) %>%
+  mutate(condition = str_remove_all(condition, "time_") %>%
+           str_replace_all("30min", "post")) %>%
+  filter(condition %in% c("pre", "post")) %>%
+  add_count(individual) %>%
+  filter(n > 1) %>%
+  select(-n) %>%
+  pivot_wider(names_from = "condition", values_from = "date_time") %>%
+  # Cluster starts times into groups (here, I know there are 10 groups)
+  add_clusters("pre", 10) %>%
+  mutate(cluster = as.factor(cluster)) %>%
+  group_by(cluster) %>%
+  summarize(pre = mean(pre), post = mean(post), n = n(), .groups = "drop")
+
+gameto_time_limits <- c(min = min(gameto_mean_times$pre), max = max(gameto_mean_times$post))
+
+# DT start/end times plot
+dt_times_gameto_2013_plot <- ggplot(
+  gameto_mean_times, 
+  aes(x = cluster, ymin = pre, ymax = post, y = post, group = cluster)) +
+  geom_linerange(position = position_dodge(0.5)) +
+  geom_point(position = position_dodge(0.5)) +
+  scale_y_datetime(limits = gameto_time_limits) +
+  coord_flip() +
+  labs(x = "Group") +
+  theme(
+    axis.title.x = element_blank()
+  )
+
+# Subset DT chamber data
+chamber_gameto_2013 <- filmy_dt_chamber %>% filter(generation == "gametophyte", year == "2013")
+
+# RH plot
+chamber_gameto_2013_rh_plot <-
+  plot_chamber_rh(chamber_gameto_2013, gameto_time_limits)
+
+# Temp plot
+chamber_gameto_2013_temp_plot <-
+  plot_chamber_temp(chamber_gameto_2013, gameto_time_limits)
+
+chamber_gameto_2013_temp_plot + chamber_gameto_2013_rh_plot + dt_times_gameto_2013_plot + plot_layout(ncol = 1)
+
+# ggsave(here::here("results/S1_fig.png"))
+```
+
+**`r s_figure("chamber-g")`**.
+Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes.
+MgNO~3~ was used to maintain constant humidity at \circa{ }58%.
+All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days unintentionally
+
+\newpage
+
+```{r range-plot, fig.height = 4.5, fig.width = 7, fig.cap = ""}
+range_comparison %>%
+  left_join(filmy_habit, by = "species") %>%
+  mutate(species = factor(species) %>% fct_reorder(sporo_min_range)) %>%
+  ggplot(aes(y = species)) +
+  geom_linerange(aes(xmin = gameto_min_range, xmax = gameto_max_range)) +
+  geom_linerange(aes(xmin = sporo_min_range, xmax = sporo_max_range, color = habit), size = 3) +
+  geom_point(
+    data = filter(range_comparison, gameto_min_range < sporo_min_range), 
+    aes(x = gameto_min_range)) +
+  geom_point(
+    data = filter(range_comparison, gameto_max_range > sporo_max_range),
+    aes(x = gameto_max_range)) +
+  scale_color_manual(values = habit_colors, labels = pretty_text) +
+  labs(x = "Elevation (m)", y = "Species") +
+  theme(
+    legend.position = "bottom"
+  )
+```
+
+**`r s_figure("range")`**.
+Elevational range of sporophytes and gametophytes of filmy ferns on Moorea, French Polynesia.
+Observed elevational range width of sporophytes shown with colored bars; range of gametophytes occuring beyond sporophytes shown with lines and dots.
+Color indicates growth habit.
 
@@
 
-```{r water-content-full, fig.height = 4.5, fig.width = 7, fig.cap = ""}
-# Plot only species with weights recorded during recovery
-species_with_rec <-
-  rel_water_species_means %>%
-  filter(rec_time %in% c("30min", "24hr", "48hr")) %>%
-  pull(species) %>%
-  unique()
-
-rel_water_species_means_with_rec <-
-  rel_water_species_means %>%
-  filter(species %in% species_with_rec)
-
-pd <- position_dodge(.5)
-
-# Define color-blind palette
-cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
-# Set colors for desiccation intensity
-salt_cols <- c("NaCl" = cbPalette[7], "MgNO3" = cbPalette[5], "LiCl" = cbPalette[6], "Control" = "white")
-
-ggplot(rel_water_species_means_with_rec, aes(x = rec_time, y = rel_water, group = interaction(dry_time, salt))) +
-  geom_errorbar(
-    aes(ymin = rel_water - sd, ymax = rel_water + sd),
-    color = "grey50",
-    width = 0,
-    position = pd) +
-  geom_line(position = pd, aes(linetype = dry_time), color = "grey30") +
-  geom_point(aes(shape = salt, fill = salt), position = pd) +
-  scale_linetype_manual(name = "Desiccation Time",
-                        values = c("solid", "dashed"),
-                        breaks = c("15", "2"),
-                        labels = c("15 days", "2 days")) +
-  scale_shape_manual(name = "Desiccation Intensity",
-                     values = salt_shapes,
-                     breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
-                     labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
-  scale_fill_manual(name = "Desiccation Intensity",
-                    values = salt_cols, 
-                    breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
-                    labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) +
-  scale_y_continuous(label = scales::percent) +
-  facet_wrap(vars(species)) +
-  labs(x = "Condition",
-       y = "Relative water content") +
-  theme_bw() +
-  theme(
-    panel.grid.major.x = element_blank(),
-    legend.position = "bottom",
-    legend.box="vertical", 
-    legend.margin=margin()
-  )
-
-# ggsave(here::here("results/S_fig_water_content_full.png"))
-```
-
-**`r s_figure("water-s")`**.
-Relative water content (RWC) of sporophytes during desiccation tolerance test.
-RWC was not measured for all combinations of species and desiccation treatments; only species with complete data shown.
-
-\clearpage
-
-```{r water-content-desiccated, fig.height = 4.5, fig.width = 7, fig.cap = ""}
-# Plot relative water content in desiccated state for all species
-rel_water_species_means_desic <<-
-rel_water_species_means %>%
-  filter(rec_time == "desiccated") %>%
-  select(-rec_time)
+```{r sporo-dt-controls, fig.height = 8, fig.width = 7, fig.cap = ""}
+# Plot sporophyte DT controls
 
@@
 pd <- position_dodge(.7)
-
-ggplot(rel_water_species_means_desic, aes(x = rel_water, y = species, group = salt)) +
-  geom_errorbar(
-    aes(xmin = rel_water - sd, xmax = rel_water + sd),
-    color = "grey80",
-    width = 0.2,
-    position=position_dodge(width=0.5)) +
-  geom_point(aes(fill = salt, shape = salt), size = 1.25, position=position_dodge(width=0.5)) +
-  scale_x_continuous(label = scales::percent) +
-  scale_shape_manual(name = "Desiccation Intensity",
-                     values = salt_shapes,
-                     breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
-                     labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
-  scale_fill_manual(name = "Desiccation Intensity",
-                    values = salt_cols, 
-                    breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
-                    labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) +
-  labs(
-    x = "Relative Water Content",
-    y = "Species",
-    color = "Desiccation\nintensity") +
-  theme_bw() +
-  theme(
-    panel.grid.minor.y = element_blank(),
-    panel.grid.major.y = element_blank(),
-    legend.position = "bottom"
-  ) +
-  facet_grid(vars(dry_time), labeller = as_labeller(c(`2` = "2 days", `15` = "15 days")))
-
-# ggsave(here::here("results/S_fig_water_content_desic.pdf"))
-```
-
-**`r s_figure("water-all")`**.
-Relative water content (RWC) of sporophytes after desiccation by duration of desiccation treatment (2 or 15 days).
-RWC was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not RWC of zero).
-
-\clearpage
-
-```{r sporo-dt-controls, fig.height = 8, fig.width = 7, fig.cap = ""}
-# Plot sporophyte DT controls
 
@@
 Gametophytes only included a single individual per species thus lack error bars.
+
+\newpage
+
+```{r water-content-desiccated, fig.height = 4.5, fig.width = 7, fig.cap = ""}
+# Plot relative water content in desiccated state for all species
+rel_water_species_means_desic <<-
+rel_water_species_means %>%
+  filter(rec_time == "desiccated") %>%
+  select(-rec_time)
+
+# Set dodge spacing, colors
+pd <- position_dodge(.7)
+
+ggplot(rel_water_species_means_desic, aes(x = rel_water, y = species, group = salt)) +
+  geom_errorbar(
+    aes(xmin = rel_water - sd, xmax = rel_water + sd),
+    color = "grey80",
+    width = 0.2,
+    position=position_dodge(width=0.5)) +
+  geom_point(aes(fill = salt, shape = salt), size = 1.25, position=position_dodge(width=0.5)) +
+  scale_x_continuous(label = scales::percent) +
+  scale_shape_manual(name = "Desiccation Intensity",
+                     values = salt_shapes,
+                     breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                     labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
+  scale_fill_manual(name = "Desiccation Intensity",
+                    values = salt_cols, 
+                    breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                    labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) +
+  labs(
+    x = "Relative Water Content",
+    y = "Species",
+    color = "Desiccation\nintensity") +
+  theme_bw() +
+  theme(
+    panel.grid.minor.y = element_blank(),
+    panel.grid.major.y = element_blank(),
+    legend.position = "bottom"
+  ) +
+  facet_grid(vars(dry_time), labeller = as_labeller(c(`2` = "2 days", `15` = "15 days")))
+
+# ggsave(here::here("results/S_fig_water_content_desic.pdf"))
+```
+
+**`r s_figure("water-all")`**.
+Relative water content (RWC) of sporophytes after desiccation by duration of desiccation treatment (2 or 15 days).
+RWC was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not RWC of zero).
+
+\newpage
+
+```{r water-content-full, fig.height = 4.5, fig.width = 7, fig.cap = ""}
+# Plot only species with weights recorded during recovery
+species_with_rec <-
+  rel_water_species_means %>%
+  filter(rec_time %in% c("30min", "24hr", "48hr")) %>%
+  pull(species) %>%
+  unique()
+
+rel_water_species_means_with_rec <-
+  rel_water_species_means %>%
+  filter(species %in% species_with_rec)
+
+pd <- position_dodge(.5)
+
+# Define color-blind palette
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+# Set colors for desiccation intensity
+salt_cols <- c("NaCl" = cbPalette[7], "MgNO3" = cbPalette[5], "LiCl" = cbPalette[6], "Control" = "white")
+
+ggplot(rel_water_species_means_with_rec, aes(x = rec_time, y = rel_water, group = interaction(dry_time, salt))) +
+  geom_errorbar(
+    aes(ymin = rel_water - sd, ymax = rel_water + sd),
+    color = "grey50",
+    width = 0,
+    position = pd) +
+  geom_line(position = pd, aes(linetype = dry_time), color = "grey30") +
+  geom_point(aes(shape = salt, fill = salt), position = pd) +
+  scale_linetype_manual(name = "Desiccation Time",
+                        values = c("solid", "dashed"),
+                        breaks = c("15", "2"),
+                        labels = c("15 days", "2 days")) +
+  scale_shape_manual(name = "Desiccation Intensity",
+                     values = salt_shapes,
+                     breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                     labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
+  scale_fill_manual(name = "Desiccation Intensity",
+                    values = salt_cols, 
+                    breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
+                    labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) +
+  scale_y_continuous(label = scales::percent) +
+  facet_wrap(vars(species)) +
+  labs(x = "Condition",
+       y = "Relative water content") +
+  theme_bw() +
+  theme(
+    panel.grid.major.x = element_blank(),
+    legend.position = "bottom",
+    legend.box="vertical", 
+    legend.margin=margin()
+  )
+
+# ggsave(here::here("results/S_fig_water_content_full.png"))
+```
+
+**`r s_figure("water-s")`**.
+Relative water content (RWC) of sporophytes during desiccation tolerance test.
+RWC was not measured for all combinations of species and desiccation treatments; only species with complete data shown.

### 2021-02-02 R/functions.R
Tidy specimen data, filter to only filmy ferns from Moorea

@@ -515,6 +515,26 @@ process_community_matrix <- function (community_matrix_path, species_list, moore
     spread(site, abundance)
 }
 
+#' Tidy raw specimen data
+#' 
+#' and filter to only filmy ferns from Moorea
+#'
+#' @param specimens_raw Specimen collection data for ferns from Moorea and Tahiti
+#' @param filmy_species Character vector of species names of filmy ferns from Moorea and Tahiti
+#'
+#' @return Dataframe
+tidy_filmy_specimens <- function (specimens_raw, filmy_species) {
+  specimens_raw %>%
+    # only use needed columns
+    select(specimen, genus, specific_epithet, generation, county, site, locality, elevation, starts_with("gameto")) %>%
+    mutate(species = paste(genus, specific_epithet, sep = "_")) %>%
+    # subset to filmy ferns on Moorea
+    assert(not_na, county) %>%
+    assert(in_set("Moorea", "Tahiti"), county) %>%
+    filter(species %in% filmy_species, county == "Moorea") %>%
+    select(species, specimen, generation, site, locality, elevation, starts_with("gameto"))
+}
+
 #' Fight light curve models to light curve data
 #'
 #' @param data Light curve data, with columns for 
@@ -736,7 +756,7 @@ calculate_mean_max_vpd <- function (climate) {
 #'
 #' @param mean_vpd Mean Vapor Pressure Deficit (VPD) measured with dataloggers
 #' at each site
-#' @param specimens_raw Raw specimen collection data
+#' @param filmy_specimens Raw specimen collection data
 #' @param moorea_sites Dataframe of collection sites on Moorea
 #' @param filmy_species Vector of filmy fern species names
 #'
@@ -744,30 +764,22 @@ calculate_mean_max_vpd <- function (climate) {
 #' 
 calculate_mean_vpd_gameto <- function(
   mean_vpd,
-  specimens_raw,
+  filmy_specimens,
   moorea_sites,
   filmy_species
 ) {
   
-  gameto_filmy_specimens <-
-    specimens_raw %>%
-    # only use needed columns
-    select(starts_with("gameto"), specimen, genus, specific_epithet, generation, site) %>%
-    mutate(
-      species = paste(genus, specific_epithet, sep = "_"),
-      # **Important**: treat epipteric as epiphytic
-      gameto_habit = str_to_lower(gameto_habit) %>% str_replace_all("epipetric", "epiphytic")) %>%
+  filmy_specimens %>%
[…]

### 2021-02-02 _targets.R

@@ -60,6 +60,11 @@ tar_plan(
   
   # Process data ----
   
+  # - subset collection data to just filmy ferns on Moorea
+  filmy_specimens = tidy_filmy_specimens(
+    specimens_raw = specimens_raw,
+    filmy_species = filmy_species),
+  
   # - calculate recovery during DT test per individual
   recovery_indiv = filmy_dt %>%
     # Only include samples that should be presented in the main MS
@@ -77,7 +82,7 @@ tar_plan(
   # - calculate mean VPD per species for gametophytes
   mean_vpd_gameto = calculate_mean_vpd_gameto(
     mean_vpd = mean_vpd,
-    specimens_raw = specimens_raw,
+    filmy_specimens = filmy_specimens,
     moorea_sites = moorea_sites,
     filmy_species = filmy_species
   ),
@@ -95,6 +100,7 @@ tar_plan(
   range_comparison = analyze_dist_pattern(
     community_matrix_raw = community_matrix_raw,
     filmy_species = filmy_species,
+    filmy_specimens = filmy_specimens,
     phy = filmy_phy,
     moorea_sites = moorea_sites
   ),

### 2021-02-02 ms/si.Rmd
Add specimen data for sporophyte species not observed in plots when calculating range size

@@
 ```{r range-plot, fig.height = 4.5, fig.width = 7, fig.cap = ""}
+# Make plot showing range of gametophytes beyond sporophytes
+plot_data <-
 range_comparison %>%
   left_join(filmy_habit, by = "species") %>%
-  mutate(species = factor(species) %>% fct_reorder(sporo_min_range)) %>%
-  ggplot(aes(y = species)) +
-  geom_linerange(aes(xmin = gameto_min_range, xmax = gameto_max_range)) +
-  geom_linerange(aes(xmin = sporo_min_range, xmax = sporo_max_range, color = habit), size = 3) +
+  mutate(species = fct_reorder(species, sporo_mean_el)) %>%
+  mutate(
+    gameto_max_el = case_when(
+      is.na(gameto_max_el) ~ sporo_max_el,
+      sporo_max_el > gameto_max_el ~ sporo_max_el,
+      TRUE ~ gameto_max_el
+    ),
+    gameto_min_el = case_when(
+      is.na(gameto_min_el) ~ sporo_min_el,
+      sporo_min_el < gameto_min_el ~ sporo_min_el,
+      TRUE ~ gameto_min_el
+    )
+  )
+
+ggplot(plot_data, aes(y = species)) +
+  # Gameto range as line for those with different min/max elevation
+  geom_linerange(aes(xmin = gameto_min_el, xmax = gameto_max_el), color = "grey10") +
+  # Gameto range as point for those with single elevation, if different from sporo elevation
   geom_point(
-    data = filter(range_comparison, gameto_min_range < sporo_min_range), 
-    aes(x = gameto_min_range)) +
+    data = filter(
+      plot_data,
+      gameto_min_el == gameto_max_el, gameto_min_el != sporo_min_el, gameto_max_el != sporo_max_el),
+    aes(x = sporo_mean_el), 
+    fill = "grey10"
+  ) +
+  # Sporo range as fat line for those with different min/max elevation
+  geom_linerange(
+    data = filter(plot_data, sporo_min_el != sporo_max_el),
+    aes(xmin = sporo_min_el, xmax = sporo_max_el, color = habit), 
+    size = 3, fill = "grey10") +
+  # Sporo range as point for those with single elevation
   geom_point(
-    data = filter(range_comparison, gameto_max_range > sporo_max_range),
-    aes(x = gameto_max_range)) +
+    data = filter(plot_data, sporo_min_el == sporo_max_el),
+    aes(x = sporo_mean_el, color = habit)
+  ) +
+  # Diamond for gameto min el below sporo min el
+  geom_point(
+    data = filter(range_comparison, gameto_min_el < sporo_min_el), 
+    aes(x = gameto_min_el),
+    shape = 23, fill = "grey10") +
+  # Diamond for gameto min el above sporo max el
+  geom_point(
+    data = filter(range_comparison, gameto_max_el > sporo_max_el),
+    aes(x = gameto_max_el),
+    shape = 23, fill = "grey10") +
   scale_color_manual(values = habit_colors, labels = pretty_text) +
-  labs(x = "Elevation (m)", y = "Species") +
+  scale_y_discrete(labels = pretty_text) +
+  labs(x = "Elevation (m)", y = "Species", color = "Growth habit") +
   theme(
-    legend.position = "bottom"
+    legend.position = "bottom",
+    axis.text.y = element_text(face = "italic")
   )
@@
 **`r s_figure("range")`**.
-Elevational range of sporophytes and gametophytes of filmy ferns on Moorea, French Polynesia.
-Observed elevational range width of sporophytes shown with colored bars; range of gametophytes occuring beyond sporophytes shown with lines and dots.
+Elevational range (m) of sporophytes and gametophytes of filmy ferns on Moorea, French Polynesia.
+Elevational range width shown with colored bars (sporophytes) or black lines (gametophytes).
+Species observed only at a single elevation (not a range) shown with points.
+Extent of range of gametophytes beyond sporophytes shown with black diamonds.
 Color indicates growth habit.
+Presence of gametophytes assumed at sites where sporophytes were observed, even if gametophytes of that species were not collected from that site.
 

### 2021-02-02 R/functions.R

@@ -611,16 +611,17 @@ extract_lc_model_params <- function (data) {
 #'
 #' @param raw_community_matrix Community matrix of Nitta et al 2017
 #' @param filmy_species Character vector of filmy fern species names
+#' @param filmy_specimens Dataframe of specimen voucher data for filmy ferns from Moorea
 #' @param phy Phylogeny including filmy ferns of Moorea and Tahiti
 #' @param moorea_sites Site data for plots on Moorea including elevation
 #' 
 #' @return Dataframe
 #' 
-analyze_dist_pattern <- function(community_matrix_raw, filmy_species, phy, moorea_sites) {
+analyze_dist_pattern <- function(community_matrix_raw, filmy_species, filmy_specimens, phy, moorea_sites) {
   
   # Convert raw community data to long format by sporophyte and gametophyte
   # min and max elevational range.
-  range_long <-
+  range_long_plots_only <-
     community_matrix_raw %>%
     rename(site = X1) %>% 
     gather(species, abundance, -site) %>%
@@ -637,39 +638,75 @@ analyze_dist_pattern <- function(community_matrix_raw, filmy_species, phy, moore
     assert(not_na, el) %>%
     group_by(species, generation) %>%
     summarize(
-      min_range = min(el),
-      max_range = max(el),
+      min_el = min(el),
+      max_el = max(el),
+      mean_el = mean(el),
+      n = n(),
       .groups = "drop"
     )
   
-  # Compare ranges to see which species have gametophytes beyond sporophytes
+  # Add observations of sporophytes that occurred outside of the plots
+  # for species that were missing from plot data
+  sporos_missing_from_plot_data <-
+    anti_join(filmy_specimens, range_long_plots_only, by = c("generation", "species")) %>%
+    filter(generation == "sporophyte") %>%
+    rename(el = elevation) %>%
+    group_by(species, generation) %>%
+    summarize(
+      min_el = min(el),
+      max_el = max(el),
+      mean_el = mean(el),
+      n = n(),
+      .groups = "drop"
+    )
+  
+  range_long <- bind_rows(range_long_plots_only, sporos_missing_from_plot_data)
+  
+  # Calculate range of gametophytes beyond sporophytes
+  
+  # Convert data to wide format
   full_join(
     range_long %>%
       gather(variable, value, -species, -generation) %>%
[…]

### 2021-02-02 R/packages_dev.R
Add "breakerofchains" package

@@ -1,3 +1,4 @@
 library("visNetwork")
 library("tflow")
 library("pdftools")
+library("breakerofchains")

### 2021-02-02 R/clean_lc_data.R
Add light_id to light data
corresponds to each specimen

@@ -301,11 +301,13 @@ filmy_all_lc_plot_filtered <-
 ggsave(plot = filmy_all_lc_plot_filtered, file = "data_raw/intermediates/all_light_curves_filtered.pdf", height = 40, width = 40)
 
 # INTERACTIVE STEP:
+# To run, uncomment the lines below
 # In RStudio, go through data by increasing row selected one at a time,
 # select outliers, and save each as CSV
-filmy_lc_models_filtered %>%
-  filter(str_detect(id, "2479_1_2013-06-10")) %>%
-  select_lc_points()
+
+# filmy_lc_models_filtered %>%
+#   filter(str_detect(id, "2479_1_2013-06-10")) %>%
+#   select_lc_points()
   
 # Read in outliers, flag them in the filtered data
 
@@ -340,7 +342,18 @@ filmy_lc_data_filtered_outliers_flagged <-
   # make sure join didn't duplicate any values
   # (number of rows should be the same before and after)
   verify(nrow(.) == nrow(filmy_lc_data_filtered)) %>%
-  mutate(outlier = replace_na(outlier, FALSE))
+  mutate(outlier = replace_na(outlier, FALSE)) %>%
+  # Add a light-curve ID number for each sample:
+  # unique to combination of: species, generation, individual, coll_num, condition, date
+  # - replace NA values in collection number with "sn" (no number)
+  mutate(coll_num = replace_na(coll_num, "sn")) %>%
+  assert(not_na, species, generation, individual, coll_num, condition, date, yield, par, etr, outlier) %>%
+  group_by(species, generation, individual, coll_num, condition, date) %>%
+  mutate(light_id = cur_group_id()) %>%
+  ungroup() %>%
+  assert(not_na, light_id) %>%
+  # Don't need p-value any more after flagging outliers
+  select(-p.value)
 
 # Write out final cleaned dataset ----
 write_csv(filmy_lc_data_filtered_outliers_flagged, "data/filmy_light_curves.csv")

### 2021-02-02 R/functions.R

@@ -428,7 +428,8 @@ load_filmy_lc <- function (file) {
     sporo_site = col_character(),
     condition = col_character(),
     date = col_date(format = ""),
-    outlier = col_logical()
+    outlier = col_logical(),
+    light_id = col_integer()
   )
   )
 }
@@ -545,16 +546,10 @@ tidy_filmy_specimens <- function (specimens_raw, filmy_species) {
 #' 
 fit_lc_model <- function (data) {
   data %>%
-    select(species, generation, individual, coll_num, condition, date, par, etr) %>%
-    # some sporophytes lack collection numbers, but
-    # we need to keep this column for differentiating between
-    # multiple individuals within gametophyte collections
-    mutate(coll_num = replace_na(coll_num, "nc")) %>%
+    select(species, generation, individual, coll_num, condition, date, par, etr, light_id) %>%
     assert(not_na, everything()) %>%
     # Group into measurements by individual
-    # (`individual` column only applies *within* a species/generation, 
-    # so there are many `1`, `2`, etc)
-    group_by(coll_num, individual, species, generation, condition, date) %>%
+    group_by(light_id) %>%
     # Make sure the number of measurements per individual is
     # within the expected range (5-9, after removing outliers)
     add_count() %>%
@@ -577,7 +572,8 @@ fit_lc_model <- function (data) {
       k_stats = map(nls_mod, broom::tidy)
     ) %>%
     # Extract fitted data points
-    mutate(fitted = map(nls_mod, ~fitted(.) %>% tibble(etr_fit = .)))
+    mutate(fitted = map(nls_mod, ~fitted(.) %>% tibble(etr_fit = .))) %>%
+    ungroup()
 }
 
 #' Extract maximum ETR and PPFD95% from light curve models
@@ -593,7 +589,7 @@ extract_lc_model_params <- function (data) {
     # Unroll fitted light curve points
     unnest(cols = fitted) %>%
     # Get maximum estimated etr fit for each individual
-    group_by(coll_num, individual, species, generation, condition, date) %>%
+    group_by(light_id) %>%
     mutate(etr = max(etr_fit)) %>%
     ungroup %>%
     select(-etr_fit) %>%

### 2021-02-02 ms/manuscript.Rmd
WIP: add SI figure showing range comparison between G and S
Add light curves to SI

@@
 Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following @Nitta2017.
-For each marker, a reference library was constructed from sporophyte sequences using the 'makeblastDB' command in BLAST [@Altschul1997].
+For each marker, a reference library was constructed from sporophyte sequences using the \function{makeblastDB} command in BLAST [@Altschul1997].
 Field-collected gametophytes were identified by a local BLAST query against the reference library.
@@
 The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22 °C to 25 °C (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
-Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
+Salts used for desiccation and their corresponding mean water potentials and approximate relative humidity are as follows: LiCl (-282 MPa, 18 % RH), Mg(NO~3~)~2~ (-86 MPa, 58 % RH), and NaCl (-38 MPa, 80 % RH).
 Following the desiccation treatment, plants were rewetted by wrapping them in wet tissues using water from a local stream.
@@
 500 µmol·m^-2^·s^-1^.
-Response curves were fitted using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded [@Proctor2012].
+Response curves were fitted using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (RETR), *x* is photosynthetic photon flux density (PPFD), *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded [@Proctor2012].
 The PPFD at 95 % saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
@@
 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
-Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson.
+Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson (`r s_figure("range")`).
 All of these have sporophytes that are normally only found in cloud forest (\circa{ }500--1200 m), but gametophytes that are distributed to lower elevations.
@@
 
-Photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (`r figure("combined")`).
+Photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (\figs `r figure_num("combined")`, `r s_figure_num("light-curves")`).
 Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.

### 2021-02-02 ms/si.Rmd
Add light curves to SI

@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
-# if running interactively, source _targets.R first
-# source("_targets.R")
+# if running interactively, source packages and functions first
+# source("R/packages.R")
+# source("R/functions.R")
 ```
@@
 tar_load(range_comparison)
+tar_load(light_data_all)
+tar_load(light_models)
+tar_load(filmy_lc_model_params)
+tar_load(filmy_lc_model_fitted_data)
 ```
@@
 
+# Define function to plot facetted light curve
+make_facet_lc_plot <- function(light_data, lc_model_fitted_data, lc_model_params) {
+
+  cbPalette <- c(grey = "#999999", gold = "#E69F00", skyblue = "#56B4E9", 
+               green = "#009E73", yellow = "#F0E442", blue = "#0072B2", 
+               red = "#D55E00", purple = "#CC79A7")
+
+  ggplot(mapping = aes(x = par)) +
+    geom_point(
+      data = light_data, 
+      aes(y = etr, alpha = outlier)) +
+    geom_line(
+      data = lc_model_fitted_data, 
+      aes(y = etr_fit)) +
+    geom_hline(
+      data = lc_model_params, 
+      aes(yintercept = etr), color = cbPalette[["blue"]]) +
+    geom_vline(
+      data = lc_model_params, 
+      aes(xintercept = par), color = cbPalette[["red"]]) +
+    scale_alpha_manual(values = c("TRUE" = 0.2, "FALSE" = 1.0)) +
+    labs(
+      y = "ETR<sub>max</sub>", 
+      x = "PPFD (µmol·m<sup>-2</sup>·s<sup>-1</sup>)", 
+      alpha = "Outlier") +
+    theme(
+      axis.title.x = element_markdown(),
+      axis.title.y = element_markdown(),
+      strip.text = element_markdown(),
+      plot.title = element_markdown(),
+      legend.position = "none"
+    ) +
+    facet_wrap_paginate(
+      vars(str_abbrev_sp_pretty(species), light_id), 
+      ncol = 4, nrow = 6, page = 1,
+      scales = "free", labeller = label_value_2)
+}
+
 #' Cluster times from a sporophyte desiccation tolerance (DT) test
@@
 }
-
 ```
@@
 RWC was not measured for all combinations of species and desiccation treatments; only species with complete data shown.
+
+\newpage
+
+```{r sporo-lab-light-curves, fig.height = 9.5, fig.width = 7, fig.cap = ""}
+### Make full sets of light curve plots ###
+# - sporos in lab
+lc_plots_sporo_lab <- make_facet_lc_plot(
+  light_data = light_data_all %>% filter(generation == "sporophyte", condition == "lab"),
+  lc_model_fitted_data = filmy_lc_model_fitted_data %>% filter(generation == "sporophyte", condition == "lab"),
+  lc_model_params = filmy_lc_model_params %>% filter(generation == "sporophyte", condition == "lab")
+)
+# - sporos in field
+lc_plots_sporo_field <- make_facet_lc_plot(
+  light_data = light_data_all %>% filter(generation == "sporophyte", condition == "field"),
+  lc_model_fitted_data = filmy_lc_model_fitted_data %>% filter(generation == "sporophyte", condition == "field"),
+  lc_model_params = filmy_lc_model_params %>% filter(generation == "sporophyte", condition == "field")
+)
+# - gametos
+lc_plots_gameto <- make_facet_lc_plot(
+  light_data = light_data_all %>% filter(generation == "gametophyte"),
+  lc_model_fitted_data = filmy_lc_model_fitted_data %>% filter(generation == "gametophyte"),
+  lc_model_params = filmy_lc_model_params %>% filter(generation == "gametophyte")
+)
+
+### Get the number of pages needed to print each facetted plot ###
+n_sporo_lab_lc_pages <- n_pages(lc_plots_sporo_lab)
+n_sporo_field_lc_pages <- n_pages(lc_plots_sporo_field)
+n_gameto_lc_pages <- n_pages(lc_plots_gameto)
+
+### Print sporos in the lab ###
+# first with title "a"
+lc_plots_sporo_lab +
+labs(title = "**a**") +
+facet_wrap_paginate(
+    vars(str_abbrev_sp_pretty(species), light_id),
+    ncol = 4, nrow = 6, page = 1,
+    scales = "free", labeller = label_value_2)
+
+# print the rest
+# with title "a (cont.)"
+for(i in 2:n_sporo_lab_lc_pages) {
+  p <- lc_plots_sporo_lab +
+    labs(title = "**a** (cont.)") +
+    facet_wrap_paginate(
+      vars(str_abbrev_sp_pretty(species), light_id),
+      ncol = 4, nrow = 6, page = i,
+      scales = "free", labeller = label_value_2)
+  print(p)
+}
+
+### Print sporos in the field ###
+# first with title "b"
+lc_plots_sporo_field +
+labs(title = "**b**") +
+facet_wrap_paginate(
+    vars(str_abbrev_sp_pretty(species), light_id),
+    ncol = 4, nrow = 6, page = 1,
+    scales = "free", labeller = label_value_2)
+
+# print the rest
+# with title "b (cont.)"
+for(i in 2:n_sporo_field_lc_pages) {
+  p <- lc_plots_sporo_field +
+    labs(title = "**b** (cont.)") +
+    facet_wrap_paginate(
+      vars(str_abbrev_sp_pretty(species), light_id),
+      ncol = 4, nrow = 6, page = i,
+      scales = "free", labeller = label_value_2)
+  print(p)
+}
+
+### Print the gametos ###
+# first with title "c"
+lc_plots_gameto +
+labs(title = "**c**") +
+facet_wrap_paginate(
+    vars(str_abbrev_sp_pretty(species), light_id),
+    ncol = 4, nrow = 6, page = 1,
+    scales = "free", labeller = label_value_2)
+
+# print the rest
+# with title "c (cont.)"
+for(i in 2:n_gameto_lc_pages) {
+  p <- lc_plots_gameto +
+    labs(title = "**c** (cont.)") +
+    facet_wrap_paginate(
+      vars(str_abbrev_sp_pretty(species), light_id),
+      ncol = 4, nrow = 6, page = i,
+      scales = "free", labeller = label_value_2)
+  print(p)
+}
+```
+
+**`r s_figure("light-curves")`**.
+Rapid light response curves of filmy ferns from Moorea, French Polynesia.
+**a**, sporophytes measured in the lab; **b**, sporophytes measured in the field; **c** gametophytes measured in the field (no gametophytes were measured in the lab).
+Title of each subplot indicates species followed by light curve ID number assigned to each sample.
+Light curves fit using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (ETR), *x* is photosynthetic photon flux density (PPFD), *A* is the asymptote of the curve, and *k* is a slope parameter.
+Outliers at high PPFD not included in the model indicated in light grey.
+Blue line indicates maximum modeled relative electron transport rate (ETR~max~).
+Red line indicates PPFD at 95 % saturation of ETR (PPFD~95%~).

### 2021-02-02 R/functions.R

@@ -997,6 +997,20 @@ transfer_bs <- function(filmy_phy_bs, filmy_phy) {
   
 }
 
+
+#' Extract fitted data points from light curve models
+#'
+#' @param light_models Tibble with light curve models.
+#' Ouput of fit_lc_model()
+#'
+#' @return Tibble
+#' 
+extract_fitted_lc_data <- function (light_models) {
+  light_models %>%
+  select(-nls_mod, -k_stats) %>%
+  unnest(cols = c(data, fitted))
+}
+
 # t-test ----
 
 #' Run a t-test comparing response values between gametophytes and sporophytes.
@@ -1564,6 +1578,28 @@ abbrev_sp <- function(data) {
     separate(species, c("genus", "epithet")) %>%
     mutate(genus = substr(genus, 1, 1)) %>%
     unite("species", genus, epithet)
+}
+
+# Convert genus to just first letter: character vector version
+str_abbrev_sp <- function(chr) {
+  tibble(species = chr) %>%
+    abbrev_sp %>%
+    pull(species)
+}
+
+# Convert genus to just first letter: character vector version, make italics
+# with period between genus and species
+str_abbrev_sp_pretty <- function(chr) {
+  tibble(species = chr) %>%
+    separate(species, c("genus", "epithet")) %>%
+    mutate(genus = substr(genus, 1, 1)) %>%
+    mutate(species = glue("*{genus}. {epithet}*")) %>%
+    pull(species)
+}
+
+# Facet labeller that doesn't print on multiple lines
+label_value_2 <- function(labels) {
+  label_value(labels = labels, multi_line = FALSE)
 }
 
 # Manuscript ----

### 2021-02-02 R/packages.R

@@ -14,8 +14,9 @@ library(lubridate)
 library(patchwork)
 library(scales)
 library(ggtext)
-suppressPackageStartupMessages(library(ggtree))
 library(kableExtra)
+library(ggforce)
+suppressPackageStartupMessages(library(ggtree))
 suppressPackageStartupMessages(library(tidyverse))
 
 conflict_prefer("col_factor", "readr")

### 2021-02-02 _targets.R

@@ -118,6 +118,9 @@ tar_plan(
   # - fit light curve models to data
   light_models = fit_lc_model(light_data),
   
+  # - extract fitted data points
+  filmy_lc_model_fitted_data = extract_fitted_lc_data(light_models),
+  
   # - extract model parameters: 
   # critical PAR, and ETR at 95% of estimated max value
   filmy_lc_model_params = extract_lc_model_params(light_models),

### 2021-02-03 ms/manuscript.Rmd
Update MS

@@
 
-# Intergenerational Niche Differentiation in Filmy Ferns (Hymenophyllaceae)
+# Intergenerational niche differentiation in filmy ferns (Hymenophyllaceae)
 
@@
 Here, we leverage a recently developed DNA barcoding system [@Nitta2017] to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes:
-1.) Epiphytic filmy ferns have higher DT and photosynthetic optima than terrestrial species.
-2.) Degree of DT is correlated with environmental water availability.
-3.) Gametophytes with wider ranges have higher DT than those with with narrower ranges.
+1.) Light responses and DT differ between species with different growth habits.
+2.) Light responses and DT differ between sporophytes and gametophytes.
+3.) DT increases with decreasing environmental water availability.
+4.) Gametophytes with wider ranges have higher DT than those with with narrower ranges.
 
@@
 
+gameto_recover_k <- phylosig %>% filter(generation == "gametophyte", trait == "recovery_mean") %>% pull(kval) %>% round_t(2)
+gameto_recover_k_pval <- phylosig %>% filter(generation == "gametophyte", trait == "recovery_mean") %>% pull(k_pval) %>% round_t(3)
+gameto_recover_lambda <- phylosig %>% filter(generation == "gametophyte", trait == "recovery_mean") %>% pull(lambda) %>% round_t(2)
+gameto_recover_lambda_pval <- phylosig %>% filter(generation == "gametophyte", trait == "recovery_mean") %>% pull(lambda_pval) %>% round_t(3)
+
 # Make sure that all ETRmax and PAR have pval > 0.05
@@
 
-Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* and λ both near 1, *P \<*0.05; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval` and λ = `r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`), but no signal for PPFD~95%~ or ETR~max~ (`r table("physig")`).
+Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* = `r gameto_recover_k`, *P* = `r gameto_recover_k_pval` and λ = `r gameto_recover_lambda`, *P* = `r gameto_recover_lambda_pval`; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval` and λ = `r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`), but no significant signal for PPFD~95%~ or ETR~max~ (`r table("physig")`).
 
@@
 
-Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
+Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (\figs{ }`r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
 Brief fluctuations to higher humidities were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).
@@
 The GLMM including the effect of generation only was selected as the best-fitting model for DT (`r table("glmm")`). 
-Sporophytes had `r sporo_dt_effect_mean` greater DT (95% CI `r sporo_dt_effect_lower` to `r sporo_dt_effect_upper`) relative to gametophytes (`r table("glmm-pars")`).
+Sporophytes had `r sporo_dt_effect_mean` greater DT (95 % CI `r sporo_dt_effect_lower` to `r sporo_dt_effect_upper`) relative to gametophytes (`r table("glmm-pars")`).
 
@@
 The complete lack of DT in terrestrial species is not surprising, considering they have much more direct access to water than epiphytes and therefore substantially less selective pressure to maintain DT.
-Since controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (Appendix S3), it is unlikely that the lack of tolerance seen in terrestrial species is due solely to removal from the soil.
+Since controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (`r s_figure("dt-control")`), it is unlikely that the lack of tolerance seen in terrestrial species is due solely to removal from the soil.
 Comparison of DT with the minimum water potential envelope of each species (`r figure("corr")`) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with minimum water potential exceeding \circa{ }-7 MPa.
@@
 @WatkinsJr.2007c also found that DT of fern gametophytes tended to match habitat, with terrestrial species having lower DT than epiphytes.
-However, we found that within high elevation epiphytes, gametophytes had lower DT than sporophytes, and only one species (saxicolous *P.
-endlicherianum*) with gametophytes that had significantly greater DT than sporophytes (`r figure("compare")`).
+However, we found that gametophytes had lower DT than sporophytes regardless of growth habit (`r figure("compare")`).
 At high elevation sites (\ie, cloud forest), epiphytic ferns often occur in a dense matrix of bryophytes (Fig. 1B).
@@
 
-Although we predicted that species with gametophytes that occur beyond the range of their sporophytes would have higher DT and be adapted for higher light levels, this was not the case.
+Although we predicted that species with gametophytes that occur far beyond the range of their sporophytes would have higher DT, this was not the case.
 This is perhaps best illustrated by *C. apiifolia* (`r figure("habit")`d, e, h).
@@
 
-Unlike DT, photosynthetic parameters did not vary significantly between species from different habitats.
+Photosynthetic parameters did not vary significantly between species from different habitats.
 Rather, we observed a clear intergenerational difference, with gametophytes consistently adapted for lower light levels than sporophytes (`r figure("combined")`, `r table("glmm")`).

### 2021-02-03 ms/si.Rmd
Update MS

@@
 ---
-title: "SI"
+title: ""
 bibliography: 
@@
 
+**Electronic supplementary materials**
+
+**Title:** Intergenerational niche differentiation in filmy ferns (Hymenophyllaceae)
+
+**Authors:** Nitta JH; Watkins JE; Holbrook NM; Taputuarai R; Wang T; Davis CC
+
+**Journal:** Journal of Plant Research
+
+**Corresponding author:** Joel H. Nitta  
+Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan  
+Phone: +81-  
+Fax: +81-  
+email: joelnitta\@gmail.com
+
+**Content:** \figs `r s_figure_num("tree")`--`r s_figure_num("light-curves")`
+
+**`r s_figure("tree")`**. Maximum likelihood ultrametric phylogenetic tree including all Moorean filmy fern species.
+
+**`r s_figure("chamber-s")`**. Conditions inside desiccation chamber and duration of desiccation treatment for sporophytes.
+
+**`r s_figure("chamber-g")`**. Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes.
+
+**`r s_figure("range")`**. Elevational range (m) of sporophytes and gametophytes of filmy ferns on Moorea, French Polynesia.
+
+**`r s_figure("dt-control")`**. Recovery of controls (samples kept on moist tissues) for desiccation tolerance test of filmy ferns from Moorea, French Polynesia.
+
+**`r s_figure("water-all")`**. Relative water content (RWC) of sporophytes after desiccation by duration of desiccation treatment (2 or 15 days).
+
+**`r s_figure("water-s")`**. Relative water content (RWC) of sporophytes during desiccation tolerance test.
+
+**`r s_figure("light-curves")`**. Rapid light response curves of filmy ferns from Moorea, French Polynesia.
+
+\newpage
+
 ```{r full-phylo, fig.height = 4.5, fig.width = 7, fig.cap = ""}
@@
 
-**`r s_figure("chamber-s")`**.
-Conditions inside desiccation chamber and duration of desiccation treatment for sporophytes at most intense desiccation level.
-LiCl was used to maintain constant humidity at \circa{ }18%.
+**`r s_figure("chamber-s")`**. Conditions inside desiccation chamber and duration of desiccation treatment for sporophytes. LiCl was used to maintain constant humidity at \circa{ }18%.
 

### 2021-02-08 R/clean_chamber_data.R
Drop data from preliminary period in 2012 dt chamber raw data

@@ -15,7 +15,11 @@ sporo_2012_chamber <- bind_rows(
   parse_logger_dat("data_raw/2012/filmyDT2_MgNO3.csv") %>% mutate(salt = "MgNO3"),
   parse_logger_dat("data_raw/2012/filmyDT2_NaCl.csv") %>% mutate(salt = "NaCl")
 ) %>%
-  mutate(generation = "sporophyte", year = 2012, species = "mixed")
+  mutate(generation = "sporophyte", year = 2012, species = "mixed") %>%
+  # It appears that data before 2012-07-31 18:10:00 UTC was prelminary (NaCl datalogger
+  # stabilizes to a different RH than 18%, 58%, or 80%). 
+  # Consider the period after this to be the correct data
+  filter(date_time > "2012-07-31 18:10:00 UTC")
   
 # 2012 gametophytes
 gameto_2012_chamber <-
@@ -45,7 +49,7 @@ hobo_chamber_data <-
     )
   ) %>%
   # Add metadata
-  left_join(hobo_metadata, by = "file") %>%
+  left_join(select(hobo_metadata, -serial_no), by = "file") %>%
   mutate(data = map(file, read_hobo)) %>%
   unnest(data) %>%
   # Exclude data not marked "keep" in metadata

### 2021-02-08 ms/si.Rmd
Add NaCl sporo chamber plot

@@
     scale_y_datetime(limits = time_limits) +
+    scale_x_discrete(labels = str_abbrev_sp_pretty) +
     scale_linetype_manual(values = c("15" = "solid", "2" = "solid")) +
@@
       axis.title.x = element_blank(),
+      axis.text.y = element_markdown(),
       legend.position = "none"
@@
 
-```{r sporophyte-dt-2013-prep}
+```{r sporophyte-dt-prep}
 # Extract pre- and post-DT treatment times for sporophytes
@@
   add_count(species, salt, dry_time, individual, dataset) %>%
-  filter(n > 1) %>%
+  filter(n > 1) %>%  # only keep those with both pre- and post- times
   select(-n) %>%
@@
 # Get mean times of each cluster of DT test samples for the selected drying salt
-sporo_clusters_2013 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2013")), "LiCl", 3)
-sporo_clusters_2014 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2014")), "LiCl", 1)
-
-sporo_clusters <- bind_rows(sporo_clusters_2013, sporo_clusters_2014)
+licl_clusters_2013 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2013")), "LiCl", 3)
+licl_clusters_2014 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2014")), "LiCl", 1)
+
+licl_clusters <- bind_rows(licl_clusters_2013, licl_clusters_2014)
 
 # Calculate time range for the clustered data
-sporo_time_range_2013 <- c(min = min(sporo_clusters_2013$pre), max = max(sporo_clusters_2013$post))
-sporo_time_range_2014 <- c(min = min(sporo_clusters_2014$pre), max = max(sporo_clusters_2014$post))
+licl_time_range_2013 <- c(min = min(licl_clusters_2013$pre), max = max(licl_clusters_2013$post))
+licl_time_range_2014 <- c(min = min(licl_clusters_2014$pre), max = max(licl_clusters_2014$post))
 
@@
   filter(filmy_dt_chamber, 
-    date_time >= sporo_time_range_2013[["min"]], 
-    date_time <= sporo_time_range_2013[["max"]], 
+    date_time >= licl_time_range_2013[["min"]], 
+    date_time <= licl_time_range_2013[["max"]], 
     salt == "LiCl")
@@
   filter(filmy_dt_chamber, 
-    date_time >= sporo_time_range_2014[["min"]], 
-    date_time <= sporo_time_range_2014[["max"]], 
+    date_time >= licl_time_range_2014[["min"]], 
+    date_time <= licl_time_range_2014[["max"]], 
     salt == "LiCl") %>%
@@
 # RH plot
-rh_plot_2013 <- plot_chamber_rh(licl_chamber_data_2013, sporo_time_range_2013) +
+licl_rh_plot_2013 <- plot_chamber_rh(licl_chamber_data_2013, licl_time_range_2013) +
   scale_y_continuous(limits = licl_rh_range, label = label_percent(accuracy = 1))
 
-rh_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = 0.01*rh, group = group)) +
+licl_rh_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = 0.01*rh, group = group)) +
   geom_line() +
-  scale_x_datetime(limits = sporo_time_range_2014) +
+  scale_x_datetime(limits = licl_time_range_2014) +
   blank_x_theme() +
@@
 # Temp plot
-temp_plot_2013 <- plot_chamber_temp(licl_chamber_data_2013, sporo_time_range_2013) +
+licl_temp_plot_2013 <- plot_chamber_temp(licl_chamber_data_2013, licl_time_range_2013) +
   scale_y_continuous(limits = licl_temp_range, label = label_number(suffix = "\u00b0C", accuracy = 1))
 
-temp_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = temp, group = group)) +
+licl_temp_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = temp, group = group)) +
   geom_line() +
-  scale_x_datetime(limits = sporo_time_range_2014) +
+  scale_x_datetime(limits = licl_time_range_2014) +
   blank_x_theme() +
@@
 # Dry times plot
-time_plot_2013 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2013) +
+licl_time_plot_2013 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2013) +
   labs(y = "2013") +
   theme(axis.title.x = element_text()) 
-time_plot_2014 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2014)  +
+licl_time_plot_2014 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2014)  +
   labs(y = "2014")  +
@@
 # Combine plots
-temp_plot_2013 + temp_plot_2014 + 
-  rh_plot_2013 + rh_plot_2014 + 
-  time_plot_2013 + time_plot_2014 + 
-  plot_layout(ncol = 2)
-
-# ggsave(here::here("results/S2_fig.png"))
-```
-
-**`r s_figure("chamber-s")`**. Conditions inside desiccation chamber and duration of desiccation treatment for sporophytes. LiCl was used to maintain constant humidity at \circa{ }18%.
+licl_temp_plot_2013 + licl_temp_plot_2014 + 
+  licl_rh_plot_2013 + licl_rh_plot_2014 + 
+  licl_time_plot_2013 + licl_time_plot_2014 + 
+  plot_layout(ncol = 2) +
+  plot_annotation(title = "a")
+```
+
+```{r sporophyte-dt-2013-NaCl, fig.height = 7, fig.width = 7, fig.cap = ""}
+# Subset desiccation duration data
+# H_polyanthos shows n=1, but this is only because pre/post were available for only one sample
+# (actually had 8 samples total)
+nacl_clusters_2013 <- 
+  sporo_times %>%
+  filter(str_detect(dataset, "2013")) %>%
+  cluster_sporo_times("NaCl", 4) %>%
+  mutate(
+    serial_no = case_when(
+      species %in% c("Didymoglossum_tahitense", "Crepidomanes_kurzii", "Callistopteris_apiifolia") & dry_time == "2" ~ "missing",
+      species %in% c("Hymenophyllum_multifidum", "Hymenophyllum_digitatum") ~ "10140667",
+      species %in% c("Hymenophyllum_polyanthos") ~ "10152774",
+      species %in% c("Abrodictyum_dentatum") & dry_time == "2" ~ "missing",
+      TRUE ~ "missing"
+    ),
+    serial_no = as.factor(serial_no)
+  )
+
+nacl_clusters_2014 <-
+  sporo_times %>%
+  filter(str_detect(dataset, "2014")) %>%
+  cluster_sporo_times("NaCl", 1) %>%
+  mutate(
+    serial_no = case_when(
+      species %in% c("Vandenboschia_maxima") & dry_time == "2" ~ "10140661",
+      TRUE ~ "10152773"
+    ),
+    serial_no = as.factor(serial_no)
+  )
+
+# make levels of `serial_no` and `species` the same across datasets
+nacl_all_sn <- filmy_dt_chamber %>% 
+  filter(salt == "NaCl") %>% 
+  pull(serial_no) %>% 
+  unique %>% 
+  sort %>%
+  c("missing")
+
+nacl_all_species <- sporo_times %>% 
+  filter(salt == "NaCl") %>% 
+  pull(species) %>% 
+  unique %>% 
+  sort
+
+nacl_clusters_2013 <- mutate(
+  nacl_clusters_2013, 
+  # sort SN in numeric order
+  serial_no = fct_expand(serial_no, nacl_all_sn) %>% fct_inseq,
+  # sort species in reverse-alphabetical order (so 'A' is on top in plot)
+  species = fct_expand(species, nacl_all_species) %>% fct_relevel(function (x) sort(x, decreasing = TRUE))
+  )
+
+nacl_clusters_2014 <- mutate(
+  nacl_clusters_2014, 
+  serial_no = fct_expand(serial_no, nacl_all_sn) %>% fct_inseq,
+  species = fct_expand(species, nacl_all_species) %>% fct_relevel(function (x) sort(x, decreasing = TRUE))
+  )
+
+# Calculate time range for the clustered data
+# to standardize x-axis
+nacl_time_range_2013 <- c(min = min(nacl_clusters_2013$pre), max = max(nacl_clusters_2013$post))
+nacl_time_range_2014 <- c(min = min(nacl_clusters_2014$pre), max = max(nacl_clusters_2014$post))
+
+# Subset DT chamber data
+nacl_chamber_data_2012 <- 
+filmy_dt_chamber %>% 
+  filter(salt == "NaCl", date_time > "2012-07-31 18:10:00 UTC", date_time < "2013-01-01 00:10:00 UTC")
+
+nacl_time_range_2012 <- c(min = min(nacl_chamber_data_2012$date_time), max = max(nacl_chamber_data_2012$date_time))
+
+nacl_chamber_data_2013 <-
+  filter(filmy_dt_chamber, 
+         date_time >= nacl_time_range_2013[["min"]], 
+         date_time <= nacl_time_range_2013[["max"]], 
+         salt == "NaCl")
+
+nacl_chamber_data_2014 <- 
+  filter(filmy_dt_chamber, 
+         date_time >= nacl_time_range_2014[["min"]], 
+         date_time <= nacl_time_range_2014[["max"]], 
+         salt == "NaCl") %>%
+  mutate(
+    # There are two sets of times: 2014-06-17 to 2014-07-02, and 2014-07-06 to 2014-07-08
+    # Add a group for these so ggplot doesn't connect lines betweeen them
+    day = yday(date_time),
+    group = ifelse(day < 185, "a", "b")
+  )
+
+# Get overall range of temperature and RH for NaCl
+# to standardize y-axis
+nacl_temp_range <- filter(filmy_dt_chamber, salt == "NaCl") %>%
+  summarize(min = min(temp), max = max(temp)) %>%
+  pivot_longer(everything()) %>%
+  arrange(value) %>%
+  pull(value)
+
+nacl_rh_range <- filter(filmy_dt_chamber, salt == "NaCl") %>%
+  mutate(rh = rh * 0.01) %>%
+  summarize(min = min(rh), max = max(rh)) %>%
+  pivot_longer(everything()) %>%
+  arrange(value) %>%
+  pull(value)
+
+# Set up color palette for serial numbers
+serial_no_cols <- c(
+  brewer_pal(palette = "Dark2")(5),
+  "grey20") %>%
+  set_names(
+    bind_rows(nacl_chamber_data_2013, nacl_chamber_data_2014) %>%
+      filter(serial_no != "missing") %>%
+      pull(serial_no) %>%
+      unique %>%
+      sort %>%
+      c("missing")
+  )
+
+# Time plots
+nacl_time_plot_2013 <- 
+  ggplot(
+    nacl_clusters_2013, 
+    aes(
+      x = species, ymin = pre, ymax = post, y = post, 
+      group = species_duration, color = serial_no)
+  ) +
+  geom_linerange(position = position_dodge(0.5)) +
+  geom_point(position = position_dodge(0.5)) +
+  scale_y_datetime(limits = nacl_time_range_2013) +
+  # Print all 'species' on the y-axis using `drop = false`
+  scale_x_discrete(labels = str_abbrev_sp_pretty, drop = FALSE) +
+  # Print all levels of 'serial no' in the legend using `drop = false`
+  scale_color_manual(values = serial_no_cols, drop = FALSE) +
+  coord_flip() +
+  theme(
+    axis.title = element_blank(),
+    axis.text.y = element_blank())
+
+nacl_time_plot_2014 <- 
+  ggplot(
+    nacl_clusters_2014, 
+    aes(x = species, ymin = pre, ymax = post, y = post, 
+        group = species_duration, color = serial_no)
+  ) +
+  geom_linerange(position = position_dodge(0.5)) +
+  geom_point(position = position_dodge(0.5)) +
+  scale_y_datetime(limits = nacl_time_range_2014) +
+  scale_x_discrete(labels = str_abbrev_sp_pretty, drop = FALSE) +
+  scale_color_manual(values = serial_no_cols) +
+  coord_flip() +
+  labs(x = "Species") +
+  theme(
+    axis.title = element_blank(),
+    axis.text.y = element_markdown()) +
+  guides(color = "none")
+
+# RH plots
+nacl_rh_plot_2012 <- 
+  ggplot(
+    nacl_chamber_data_2012, 
+    aes(x = date_time, y = 0.01*rh)
+  ) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = nacl_time_range_2012) +
+  scale_y_continuous(
+    label = scales::label_percent(accuracy = 1), 
+    limits = nacl_rh_range) +
+  labs(
+    y = "Relative humidity",
+    color = "Data logger\nserial no.") +
+  guides(color = "none") +
+  theme(axis.title.x = element_blank()) +
+  blank_y_theme()
+
+nacl_rh_plot_2013 <- 
+  ggplot(
+    nacl_chamber_data_2013, 
+    aes(x = date_time, y = 0.01*rh, color = serial_no)
+  ) +
+  scale_color_manual(values = serial_no_cols) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = nacl_time_range_2013) +
+  scale_y_continuous(
+    label = scales::label_percent(accuracy = 1), 
+    limits = nacl_rh_range) +
+  labs(
+    y = "Relative humidity",
+    color = "Data logger\nserial no.") +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()) +
+  blank_y_theme() +
+  guides(color = "none")
+
+nacl_rh_plot_2014 <- 
+  ggplot(
+    nacl_chamber_data_2014, 
+    aes(x = date_time, y = 0.01*rh, color = serial_no, group = interaction(group, serial_no))
+  ) +
+  scale_color_manual(values = serial_no_cols) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = nacl_time_range_2014) +
+  scale_y_continuous(
+    label = scales::label_percent(accuracy = 1), 
+    limits = nacl_rh_range) +
+  labs(
+    y = "Relative humidity",
+    color = "Data logger\nserial no.") +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()) +
+  guides(color = "none")
+
+# Temp plots
+nacl_temp_plot_2012 <- 
+  ggplot(
+    nacl_chamber_data_2012, 
+    aes(x = date_time, y = temp)
+  ) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = nacl_time_range_2012) +
+  scale_y_continuous(
+    label = label_number(suffix = "\u00b0C", accuracy = 1), 
+    limits = nacl_temp_range) +
+  labs(
+    y = "Temperature",
+    color = "Data logger\nserial no.",
+    title = "2012") +
+  theme(
+    plot.title = element_text(hjust = 0.5),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()) +
+  blank_y_theme() +
+  guides(color = "none")
+
+nacl_temp_plot_2013 <- 
+  ggplot(
+    nacl_chamber_data_2013, 
+    aes(x = date_time, y = temp, color = serial_no)
+  ) +
+  geom_line(size = 0.25) +
+  scale_color_manual(values = serial_no_cols) +
+  scale_x_datetime(limits = nacl_time_range_2013) +
+  scale_y_continuous(
+    label = label_number(suffix = "\u00b0C", accuracy = 1), 
+    limits = nacl_temp_range) +
+  labs(
+    y = "Temperature",
+    color = "Data logger\nserial no.",
+    title = "2013") +
+  theme(
+    plot.title = element_text(hjust = 0.5),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()) +
+  blank_y_theme() +
+  guides(color = "none")
+
+nacl_temp_plot_2014 <- 
+  ggplot(
+    nacl_chamber_data_2014, 
+    aes(x = date_time, y = temp, color = serial_no, group = interaction(group, serial_no))
+  ) +
+  scale_color_manual(values = serial_no_cols) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = nacl_time_range_2014) +
+  scale_y_continuous(
+    label = label_number(suffix = "\u00b0C", accuracy = 1), 
+    limits = nacl_temp_range) +
+  labs(
+    y = "Temperature",
+    color = "Data logger\nserial no.",
+    title = "2014") +
+  theme(
+    plot.title = element_text(hjust = 0.5),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()) +
+  guides(color = "none")
+
+nacl_temp_plot_2014 + nacl_temp_plot_2013 + nacl_temp_plot_2012 +
+  nacl_rh_plot_2014 + nacl_rh_plot_2013 + nacl_rh_plot_2012 +
+  nacl_time_plot_2014 + nacl_time_plot_2013 + plot_spacer() +
+  plot_layout(ncol = 3, nrow = 3, guides = "collect") +
+  plot_annotation(title = "b") &
+  theme(legend.position = "bottom")
+```
+
+**`r s_figure("chamber-s")`**.
+Temperature, relative humidity (RH), and duration of desiccation treatment for sporophytes.
+Relative humidity was controlled by placing drying salts inside desiccation chambers: **a**, LiCl (-282 MPa, 18 % RH); **b**, Mg(NO~3~)~2~ (-86 MPa, 58 % RH); **c**, NaCl (-38 MPa, 80 % RH; **c**.
+Temperature and RH measured with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) in 2013 and 2014 and **--FIXME: add datalogger model--** dataloggers in 2012.
+For bottom-most panels (duration of desiccation treatment), line indicates duration and point indicates removal of sample from chamber.
+For **b** and **c**, multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color.
+"missing" indicates data were not collected on the correspondence between samples and data loggers (chambers).
 

### 2021-02-17 ms/si.Rmd
update SI

@@
   ggplot(data, aes(x = date_time, y = 0.01*rh)) +
-    geom_line() +
+    geom_line(size = 0.25) +
     scale_x_datetime(limits = time_limits) +
@@
       legend.position = "none",
-      axis.title.x = element_blank(),
-      axis.text.x = element_blank(),
-      axis.ticks.x = element_blank())
+      axis.title.x = element_blank())
 }
@@
   ggplot(data, aes(x = date_time, y = temp)) +
-    geom_line() +
+    geom_line(size = 0.25) +
     scale_x_datetime(limits = time_limits) +
@@
     theme(
-      axis.title.x = element_blank(),
+      axis.title = element_blank(),
       axis.text.y = element_markdown(),
@@
     )
-}
-
-plot_sporo_chamber <- function (sporo_times_data, filmy_dt_chamber_data, salt_select, k_select) {
-
-  # Get mean times of each cluster of DT test samples for the selected drying salt
-  sporo_clusters <- cluster_sporo_times(sporo_times_data, salt_select, k_select)
-
-  # Calculate time range for the clustered data
-  sporo_time_range <- c(min = min(sporo_clusters$pre), max = max(sporo_clusters$post))
-
-  # Subset DT chamber data
-  sporo_chamber_data <-
-    filmy_dt_chamber_data %>% 
-    filter(date_time >= sporo_time_range[["min"]], date_time <= sporo_time_range[["max"]], salt == salt_select)
-
-  # RH plot
-  rh_plot <- plot_chamber_rh(sporo_chamber_data, sporo_time_range)
-
-  # Temp plot
-  temp_plot <- plot_chamber_temp(sporo_chamber_data, sporo_time_range)
-
-  # Dry times plot
-  time_plot <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range)
-
-  # Combine plots
-  rh_plot + temp_plot + time_plot + plot_layout(ncol = 1)
 }
@@
 # Subset DT chamber data
+licl_chamber_data_2012 <- 
+  filter(filmy_dt_chamber, 
+         date_time > "2012-07-31 18:10:00 UTC", 
+         date_time < "2013-01-01 00:10:00 UTC",
+         salt == "LiCl")
+
+licl_time_range_2012 <- c(min = min(licl_chamber_data_2012$date_time), max = max(licl_chamber_data_2012$date_time))
+
 licl_chamber_data_2013 <-
@@
 
-licl_chamber_data <- bind_rows(licl_chamber_data_2013, licl_chamber_data_2014)
+licl_chamber_data <- bind_rows(licl_chamber_data_2012, licl_chamber_data_2013, licl_chamber_data_2014)
 licl_temp_range <- c(min(licl_chamber_data$temp), max(licl_chamber_data$temp))
@@
 
-# RH plot
+# Temp plots
+licl_temp_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = temp, group = group)) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = licl_time_range_2014) +
+  scale_y_continuous(
+    label = label_number(suffix = "\u00b0C", accuracy = 1), 
+    limits = licl_temp_range) +
+  labs(
+    y = "Temperature",
+    title = "2014") +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 11),
+    legend.position = "none") +
+  blank_x_theme()
+
+licl_temp_plot_2013 <- plot_chamber_temp(licl_chamber_data_2013, licl_time_range_2013) +
+  scale_y_continuous(limits = licl_temp_range) +
+  labs(title = "2013") +
+  theme(plot.title = element_text(hjust = 0.5, size = 11)) +
+  blank_y_theme() +
+  blank_x_theme()
+
+licl_temp_plot_2012 <- plot_chamber_temp(licl_chamber_data_2012, licl_time_range_2012) +
+  scale_y_continuous(limits = licl_temp_range) +
+  labs(title = "2012") +
+  theme(plot.title = element_text(hjust = 0.5, size = 11)) +
+  blank_y_theme() +
+  blank_x_theme()
+
+# RH plots
+licl_rh_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = 0.01*rh, group = group)) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = licl_time_range_2014) +
+  scale_y_continuous(
+    label = label_percent(accuracy = 1), 
+    limits = licl_rh_range) +
+  labs(y = "Relative humidity") +
+  theme(legend.position = "none") +
+  blank_x_theme()
+
 licl_rh_plot_2013 <- plot_chamber_rh(licl_chamber_data_2013, licl_time_range_2013) +
-  scale_y_continuous(limits = licl_rh_range, label = label_percent(accuracy = 1))
-
-licl_rh_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = 0.01*rh, group = group)) +
-  geom_line() +
-  scale_x_datetime(limits = licl_time_range_2014) +
-  blank_x_theme() +
+  scale_y_continuous(limits = licl_rh_range, label = label_percent(accuracy = 1)) +
   blank_y_theme() +
-  theme(legend.position = "none")
-
-# Temp plot
-licl_temp_plot_2013 <- plot_chamber_temp(licl_chamber_data_2013, licl_time_range_2013) +
-  scale_y_continuous(limits = licl_temp_range, label = label_number(suffix = "\u00b0C", accuracy = 1))
-
-licl_temp_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = temp, group = group)) +
-  geom_line() +
-  scale_x_datetime(limits = licl_time_range_2014) +
-  blank_x_theme() +
-  blank_y_theme() +
-  theme(legend.position = "none")
+  blank_x_theme()
+
+licl_rh_plot_2012 <- plot_chamber_rh(licl_chamber_data_2012, licl_time_range_2012) +
+  scale_y_continuous(limits = licl_rh_range, label = label_percent(accuracy = 1)) +
+  blank_y_theme()
 
@@
 licl_time_plot_2013 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2013) +
-  labs(y = "2013") +
-  theme(axis.title.x = element_text()) 
-licl_time_plot_2014 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2014)  +
-  labs(y = "2014")  +
-  theme(axis.title.x = element_text()) +
   blank_y_theme()
+licl_time_plot_2014 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2014)
 
 # Combine plots
-licl_temp_plot_2013 + licl_temp_plot_2014 + 
-  licl_rh_plot_2013 + licl_rh_plot_2014 + 
-  licl_time_plot_2013 + licl_time_plot_2014 + 
-  plot_layout(ncol = 2) +
+licl_temp_plot_2014 + licl_temp_plot_2013 + licl_temp_plot_2012 +
+  licl_rh_plot_2014 + licl_rh_plot_2013 + licl_rh_plot_2012 +
+  licl_time_plot_2014 + licl_time_plot_2013 + plot_spacer() +
+  plot_layout(ncol = 3, nrow = 3, guides = "collect") +
   plot_annotation(title = "a")
@@
   theme(
-    plot.title = element_text(hjust = 0.5),
+    plot.title = element_text(hjust = 0.5, size = 11),
     axis.title.x = element_blank(),

### 2021-02-19 ms/si.Rmd
Add MgNO3 chamber plot, put those plots in landscape mode
a few drawbacks... puts page numbers in weird place, adds blank page. May need to edit manually later.

@@
   \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
+  \usepackage{pdflscape}
+  \newcommand{\blandscape}{\begin{landscape}}
+  \newcommand{\elandscape}{\end{landscape}}
 editor_options: 
@@
 
-\newpage
+\blandscape
 
@@
 
-```{r sporophyte-dt-2013-LiCl, fig.height = 4.5, fig.width = 7, fig.cap = ""}
+```{r sporophyte-dt-LiCl, fig.height = 7, fig.width = 9.5, fig.cap = ""}
 # Get mean times of each cluster of DT test samples for the selected drying salt
@@
   plot_layout(ncol = 3, nrow = 3, guides = "collect") +
-  plot_annotation(title = "a")
-```
-
-```{r sporophyte-dt-2013-NaCl, fig.height = 7, fig.width = 7, fig.cap = ""}
+  plot_annotation(title = "a") &
+  theme(plot.tag = element_text(face = "bold"))
+```
+
+```{r sporophyte-dt-MgNO3, fig.height = 7, fig.width = 9.5, fig.cap = ""}
+# Get mean times of each cluster of DT test samples for the selected drying salt
+mgno3_clusters_2013 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2013")), "MgNO3", 3)
+mgno3_clusters_2014 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2014")), "MgNO3", 1)
+
+mgno3_clusters <- bind_rows(mgno3_clusters_2013, mgno3_clusters_2014)
+
+# Calculate time range for the clustered data
+mgno3_time_range_2013 <- c(min = min(mgno3_clusters_2013$pre), max = max(mgno3_clusters_2013$post))
+mgno3_time_range_2014 <- c(min = min(mgno3_clusters_2014$pre), max = max(mgno3_clusters_2014$post))
+
+# Subset DT chamber data
+mgno3_chamber_data_2012 <- 
+  filter(filmy_dt_chamber, 
+         date_time > "2012-07-31 18:10:00 UTC", 
+         date_time < "2013-01-01 00:10:00 UTC",
+         salt == "MgNO3")
+
+mgno3_time_range_2012 <- c(min = min(mgno3_chamber_data_2012$date_time), max = max(mgno3_chamber_data_2012$date_time))
+
+mgno3_chamber_data_2013 <-
+  filter(filmy_dt_chamber, 
+         serial_no == "10342564",
+         date_time >= mgno3_time_range_2013[["min"]], 
+         date_time <= mgno3_time_range_2013[["max"]], 
+         salt == "MgNO3")
+mgno3_chamber_data_2014 <- 
+  filter(filmy_dt_chamber, 
+         date_time >= mgno3_time_range_2014[["min"]], 
+         date_time <= mgno3_time_range_2014[["max"]], 
+         salt == "MgNO3") %>%
+  # There are two sets of times: 2014-06-17 to 2014-07-02, and 2014-07-06 to 2014-07-08
+  # Add a group for these so ggplot doesn't connect lines betweeen them
+  mutate(
+    day = yday(date_time),
+    group = ifelse(day < 185, "a", "b"))
+
+mgno3_chamber_data <- bind_rows(mgno3_chamber_data_2012, mgno3_chamber_data_2013, mgno3_chamber_data_2014)
+mgno3_temp_range <- c(min(mgno3_chamber_data$temp), max(mgno3_chamber_data$temp))
+mgno3_rh_range <- c(min(mgno3_chamber_data$rh), max(mgno3_chamber_data$rh))*0.01
+
+# Temp plots
+mgno3_temp_plot_2014 <- ggplot(mgno3_chamber_data_2014, aes(x = date_time, y = temp, group = group)) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = mgno3_time_range_2014) +
+  scale_y_continuous(
+    label = label_number(suffix = "\u00b0C", accuracy = 1), 
+    limits = mgno3_temp_range) +
+  labs(
+    y = "Temperature",
+    title = "2014") +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 11),
+    legend.position = "none") +
+  blank_x_theme()
+
+mgno3_temp_plot_2013 <- plot_chamber_temp(mgno3_chamber_data_2013, mgno3_time_range_2013) +
+  scale_y_continuous(limits = mgno3_temp_range) +
+  labs(title = "2013") +
+  theme(plot.title = element_text(hjust = 0.5, size = 11)) +
+  blank_y_theme() +
+  blank_x_theme()
+
+mgno3_temp_plot_2012 <- plot_chamber_temp(mgno3_chamber_data_2012, mgno3_time_range_2012) +
+  scale_y_continuous(limits = mgno3_temp_range) +
+  labs(title = "2012") +
+  theme(plot.title = element_text(hjust = 0.5, size = 11)) +
+  blank_y_theme() +
+  blank_x_theme()
+
+# RH plots
+mgno3_rh_plot_2014 <- ggplot(mgno3_chamber_data_2014, aes(x = date_time, y = 0.01*rh, group = group)) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = mgno3_time_range_2014) +
+  scale_y_continuous(
+    label = label_percent(accuracy = 1), 
+    limits = mgno3_rh_range) +
+  labs(y = "Relative humidity") +
+  theme(legend.position = "none") +
+  blank_x_theme()
+
+mgno3_rh_plot_2013 <- plot_chamber_rh(mgno3_chamber_data_2013, mgno3_time_range_2013) +
+  scale_y_continuous(limits = mgno3_rh_range, label = label_percent(accuracy = 1)) +
+  blank_y_theme() +
+  blank_x_theme()
+
+mgno3_rh_plot_2012 <- plot_chamber_rh(mgno3_chamber_data_2012, mgno3_time_range_2012) +
+  scale_y_continuous(limits = mgno3_rh_range, label = label_percent(accuracy = 1)) +
+  blank_y_theme()
+
+# Dry times plot
+mgno3_time_plot_2013 <- plot_sporo_chamber_time(mgno3_clusters, mgno3_time_range_2013) +
+  blank_y_theme()
+mgno3_time_plot_2014 <- plot_sporo_chamber_time(mgno3_clusters, mgno3_time_range_2014)
+
+# Combine plots
+mgno3_temp_plot_2014 + mgno3_temp_plot_2013 + mgno3_temp_plot_2012 +
+  mgno3_rh_plot_2014 + mgno3_rh_plot_2013 + mgno3_rh_plot_2012 +
+  mgno3_time_plot_2014 + mgno3_time_plot_2013 + plot_spacer() +
+  plot_layout(ncol = 3, nrow = 3, guides = "collect") +
+  plot_annotation(title = "b") &
+  theme(plot.tag = element_text(face = "bold"))
+```
+
+```{r sporophyte-dt-NaCl, fig.height = 7, fig.width = 9.5, fig.cap = ""}
 # Subset desiccation duration data
@@
   plot_layout(ncol = 3, nrow = 3, guides = "collect") +
-  plot_annotation(title = "b") &
-  theme(legend.position = "bottom")
-```
+  plot_annotation(title = "c") &
+  theme(
+    plot.tag = element_text(face = "bold"),
+    legend.position = "bottom")
+```
+
+\elandscape
+
+\newpage
 
@@
 For bottom-most panels (duration of desiccation treatment), line indicates duration and point indicates removal of sample from chamber.
-For **b** and **c**, multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color.
+For **c**, multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color.
 "missing" indicates data were not collected on the correspondence between samples and data loggers (chambers).

### 2021-02-19 install_latex.R

@@ -39,6 +39,7 @@ latex_packages <- c(
   "mdwtools",
   "multirow",
   "pdfescape",
+  "pdflscape",
   "pdftexcmds",
   "refcount",
   "rerunfilecheck",

### 2021-02-22 ms/si.Rmd
Fix landscape mode
needed to make image size slightly smaller to avoid extra blank page

@@
       in_header: journal-of-plant-research-si.sty
+# enable landscape mode selectively with pdflscape and newcommand 
+# https://stackoverflow.com/questions/25849814/rstudio-rmarkdown-both-portrait-and-landscape-layout-in-a-single-pdf
 header-includes: 
@@
 
-\blandscape
-
 ```{r sporophyte-dt-prep}
@@
 
-```{r sporophyte-dt-LiCl, fig.height = 7, fig.width = 9.5, fig.cap = ""}
+\blandscape
+
+```{r sporophyte-dt-LiCl, fig.height = 6.5, fig.width = 9, fig.cap = ""}
 # Get mean times of each cluster of DT test samples for the selected drying salt
@@
 
-```{r sporophyte-dt-MgNO3, fig.height = 7, fig.width = 9.5, fig.cap = ""}
+```{r sporophyte-dt-MgNO3, fig.height = 6.5, fig.width = 9, fig.cap = ""}
 # Get mean times of each cluster of DT test samples for the selected drying salt
@@
 
-```{r sporophyte-dt-NaCl, fig.height = 7, fig.width = 9.5, fig.cap = ""}
+```{r sporophyte-dt-NaCl, fig.height = 6.5, fig.width = 9, fig.cap = ""}
 # Subset desiccation duration data

### 2021-02-22 ms/si.Rmd
Fix mapping of colors in dry-down plot

@@
 
-# Define color-blind palette
-cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
 # Set colors for desiccation intensity
-salt_cols <- c("NaCl" = cbPalette[7], "MgNO3" = cbPalette[5], "LiCl" = cbPalette[6], "Control" = "white")
+salt_cols <- c(
+  "NaCl" = cbPalette[["blue"]], 
+  "MgNO3" = cbPalette[["yellow"]], 
+  "LiCl" = cbPalette[["red"]], 
+  "Control" = "white")
 

### 2021-02-22 ms/manuscript.Rmd
Update MS

@@
 tar_load(env_range_dt_model_summary)
+tar_load(filmy_dt_chamber)
 ```
@@
 Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
-Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
-Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
-The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22 °C to 25 °C (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
-Salts used for desiccation and their corresponding mean water potentials and approximate relative humidity are as follows: LiCl (-282 MPa, 18 % RH), Mg(NO~3~)~2~ (-86 MPa, 58 % RH), and NaCl (-38 MPa, 80 % RH).
+Any samples with pre-treatment F~v~/F~m~ < 0.4 were excluded from analysis, as typical values of F~v~/F~m~ in healthy plants tend to be \circa{ }0.8 [@Bjorkman1987].
+Samples were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
+Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 30 min (2012 field season) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014 field seasons).
+The desiccation chambers were kept in an air-conditioned room. 
+Temperature inside the chambers was mean `mean(filmy_dt_chamber$temp) %>% round(1)` ± SD `r sd(filmy_dt_chamber$temp) %>% round(2)` °C (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
+Salts used for desiccation and their corresponding mean water potentials and approximate relative humidity and VPD are as follows: LiCl (-282 MPa, 18 % RH, 2.45 kPa), Mg(NO~3~)~2~ (-86 MPa, 58 % RH, 1.25 kPa), and NaCl (-38 MPa, 80 % RH, 0.60 kPa).
 Following the desiccation treatment, plants were rewetted by wrapping them in wet tissues using water from a local stream.
@@
 The PPFD at 95 % saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
-Light responses were measured in 8--16 field-collected sporophyte individuals for each species, and on single gametophyte individuals in the field, which were later identified to species using the DNA barcodes.
+Light responses were measured in eight individuals per species in the lab for sporophytes, and on single individuals in the field for gametophytes.
+The gametophytes were later identified to species using the DNA barcodes.
 
@@
 
-To test for a correlation between environmental water potential and DT, we first derived a single minimum water potential value for each species by taking the average of the minimum daily water potential at all sites where that species was present.
-We used minimum values because these may represent a climatic extreme that limits the occurrence of species.
+To test for a correlation between environmental water availability and DT, we first derived a single maximum VPD for each species by taking the average of the maximum daily VPD at all sites where that species was present.
+We used maximum VPD (\ie, minimum humidity) values because these may represent a climatic extreme that limits the occurrence of species.
 We used data from the 'epiphytic' dataloggers mounted at 1.5 m on trees for epiphytic species and data from the 'terrestrial' dataloggers on the ground for non-epiphytic species.
@@
 Rotui 830 m site were excluded because this site is much more exposed than the other high elevation sites, and appears as a significant outlier in the microclimatic data [@Nitta2020].
-Since DT showed significant phylogenetic signal (see Results), we used phylogenetic generalized least squares (PGLS) [@Freckleton2002] to test for a relationship between environmental water potential and DT while taking phylogeny into account, using species' means.
+Since DT showed significant phylogenetic signal (see Results), we used phylogenetic generalized least squares (PGLS) [@Freckleton2002] to test for a relationship between VPD and DT while taking phylogeny into account, using species' means.
 
@@
 
-Photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (\figs `r figure_num("combined")`, `r s_figure_num("light-curves")`).
+Photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (\figs `r figure_num("combined")`, `r s_figure_num("light-curves")`, `r s_figure_num("traits-box")`).
 Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
@@
 
-### Relationships of water potential and range size with DT
+### Relationships of vapor pressure deficit and range size with DT
 
@@
 
-We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below -200 MPa in some epiphytes.
+```{r sporo-dt-disc-stats}
+# Calculate the avg daily maximum VPD tolerated by sax & terr sporophytes
+max_ter_vpd <- mean_vpd_sporo %>%
+  left_join(filmy_habit) %>%
+  filter(habit %in% c("Saxicolous", "Terrestrial")) %>%
+  slice_max(order_by = vpd_sporophyte_mean) %>%
+  pull(vpd_sporophyte_mean) %>%
+  round(1)
+```
+
+We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below -200 MPa (VPD > 2.45 kPa) in some epiphytes.
 Levels of similarly extreme DT have also been reported in other species of *Hymenophyllum* [@Proctor2003; @Proctor2012] and a general correlation of habitat with DT in filmy ferns has been previously established [@Shreve1911; @Proctor2003; @Proctor2012; @Parra2009; @Saldana2013], but no other studies to our knowledge have documented a similar range of DT values in filmy ferns from a single site.
-One may wonder then, what is the adaptive significance of being able to withstand drying at -200 MPa vs.
--100 MPa when the ambient water potential is typically > -5 MPa and only rarely, if ever, drops below -50 MPa.
+One may wonder then, what is the adaptive significance of being able to withstand drying at -282 MPa (\circa{ }2.45 kPa) vs.
+-86 MPa (\circa{ }1.25 kPa) when the ambient VPD is typically < (\circa{ }0.10 kPa) and only rarely exceeds > 0.50 kPa.
 The recent finding of a constitutive mechanism for DT in filmy ferns [@GarcesCea2014] supports the need for a relatively wide "safety margin," and similar (or even more extreme) levels of constitutive DT are well known in bryophytes [@Oliver1993].
@@
 Since controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (`r s_figure("dt-control")`), it is unlikely that the lack of tolerance seen in terrestrial species is due solely to removal from the soil.
-Comparison of DT with the minimum water potential envelope of each species (`r figure("corr")`) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with minimum water potential exceeding \circa{ }-7 MPa.
-In contrast, habitats with minimum water potential below -7 MPa are nearly all epiphytic, and the filmy fern sporophytes occurring there are all capable of DT.
+Comparison of DT with the average daily maximum VPD of each species (`r figure("corr")`) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with VPD exceeding \circa{ }`r max_ter_vpd` kPa.
 
@@
 This differs from other fern groups with desiccation tolerant gametophytes but desiccation sensitive sporophytes [@WatkinsJr.2007c].
-In the filmy ferns we studied, DT levels of both sporophytes and gametophytes largely reflected habitat; in general, epiphytic species tended to have higher DT than terrestrial or saxicolous species, regardless of life stage.
+In the filmy ferns we studied, DT levels of both sporophytes and gametophytes largely reflected habitat; in general, epiphytic species tended to have higher DT than terrestrial or saxicolous species, regardless of life stage (\figs `r figure_num("combined")`, `r s_figure_num("traits-box")`).
 Interestingly, this echoes much earlier field observations (not experimental data) by @Holloway1930 on filmy fern gametophytes in New Zealand: he reported apparent DT in gametophytes of *Hymenophyllum rarum* R.
@@
 @WatkinsJr.2007c also found that DT of fern gametophytes tended to match habitat, with terrestrial species having lower DT than epiphytes.
-However, we found that gametophytes had lower DT than sporophytes regardless of growth habit (`r figure("compare")`).
-At high elevation sites (\ie, cloud forest), epiphytic ferns often occur in a dense matrix of bryophytes (Fig. 1B).
+However, we found that gametophytes had lower DT than sporophytes regardless of growth habit (\figs `r figure_num("combined")`, `r s_figure_num("traits-box")`).
+At high elevation sites (\ie, cloud forest), epiphytic ferns often occur in a dense matrix of bryophytes (`r figure("habit")`b).
 It is possible that this bryophyte cover helps to retain moisture during brief periods of drought and thus protect the epiphytic fern gametophytes that grow amongst them.
@@
 Since gametophytes of *C.
-apiifolia* are common on Moorea and have a unique morphology [@Nitta2017], we were able to collect enough material of this species for additional tests of DT at higher humidity levels (48 h at 75% and 100% RH; results not shown).
-However, it still failed to recover from even these gentle treatments.
+apiifolia* are common on Moorea and have a unique morphology [@Nitta2017], we were able to collect enough material of this species for additional tests of DT at 100% RH for 2 d.
+However, it still failed to recover from even this gentle treatment (`r s_figure("capi-water")`.
 This suggests that the widespread gametophytes of *C. apiifolia* do not rely on DT to attain their distribution, but instead must do so via some other mechanism.
@@
 
-**`r table("pgls")`** Phylogenetic generalized least squares (PLGS) models of desiccation tolerance as a function of environmental water potential (VPD) or the range of gametophytes beyond sporophytes (G beyond S) in filmy ferns from Moorea, French Polynesia
+**`r table("pgls")`** Phylogenetic generalized least squares (PLGS) models of desiccation tolerance as a function of vapor pressure deficit (VPD) or the range of gametophytes beyond sporophytes (G beyond S) in filmy ferns from Moorea, French Polynesia
 
@@
 Asterisk indicates significant difference in means (*P* <0.05) between generations for a particular species (*t*-test); ns indicates no significant difference; lack of ns or asterisk means not enough observations available for that species for *t*-test.
-Left bar plot shows recovery (%) of chlorophyll fluorescence (Fv/Fm) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa). 
+Left bar plot shows recovery (%) of chlorophyll fluorescence (F~v~/F~m~) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa). 
 Right bar plot shows maximum relative electron transport rate of photosystem II (ETR~max~). 

### 2021-02-22 ms/si.Rmd
Update MS

@@
 tar_load(filmy_lc_model_fitted_data)
+tar_load(combined_species_means)
 ```
@@
     labs(
-      y = "ETR<sub>max</sub>", 
-      x = "PPFD (µmol·m<sup>-2</sup>·s<sup>-1</sup>)", 
+      y = "ETR<sub>max</sub> (µmol m<sup>-2</sup>·s<sup>-1</sup>)", 
+      x = "PPFD (µmol m<sup>-2</sup>·s<sup>-1</sup>)", 
       alpha = "Outlier") +
@@
   theme_bw(base_size = 10) +
-    theme(
-      panel.grid.minor = element_blank()
-    ))
+    theme(panel.grid.minor = element_blank()))
 
@@
 Relative humidity was controlled by placing drying salts inside desiccation chambers: **a**, LiCl (-282 MPa, 18 % RH); **b**, Mg(NO~3~)~2~ (-86 MPa, 58 % RH); **c**, NaCl (-38 MPa, 80 % RH; **c**.
-Temperature and RH measured with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) in 2013 and 2014 and **--FIXME: add datalogger model--** dataloggers in 2012.
-For bottom-most panels (duration of desiccation treatment), line indicates duration and point indicates removal of sample from chamber.
+Temperature and RH measured with Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 30 min (2012) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014).
+For bottom-most panels (duration of desiccation treatment), line indicates duration of treatment and point indicates removal of sample from chamber.
 For **c**, multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color.
@@
 Red line indicates PPFD at 95 % saturation of ETR (PPFD~95%~).
+
+\newpage
+
+```{r traits-box, fig.height = 5, fig.width = 7, fig.cap = ""}
+combined_species_means_habit <-
+  combined_species_means %>%
+  left_join(filmy_habit, by = "species") %>%
+  mutate(
+    habit = case_when(
+      habit == "High_Epiphyte" ~ "HE",
+      habit == "Low_Epiphyte" ~ "LE",
+      habit == "Saxicolous" ~ "S",
+      habit == "Terrestrial" ~ "T"
+    ) %>% factor(., levels = c("HE", "LE", "S", "T"))
+  )
+
+generation_colors <- c(
+  "gametophyte" = cbPalette[["skyblue"]],
+  "sporophyte" = cbPalette[["red"]]
+)
+
+plot_sp_mean <- function(data, var_name, var_label) {
+
+  ggplot(data, 
+         aes(x = habit, y = {{var_name}}, fill = generation)) +
+    geom_boxplot(outlier.size = 0) +
+    geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.2)) +
+    scale_fill_manual(values = generation_colors, labels = str_to_sentence) +
+    labs(
+      y = var_label,
+      x = "Growth Habit",
+      fill = "Generation"
+    )
+
+}
+
+a <- plot_sp_mean(combined_species_means_habit, etr_mean, "ETR<sub>max<sub>") +
+  theme(
+    axis.title.x = element_blank(),
+    axis.title.y = element_markdown())
+
+b <- plot_sp_mean(combined_species_means_habit, par_mean, "PPFD<sub>95%<sub>") +
+  theme(axis.title.y = element_markdown())
+
+c <- plot_sp_mean(combined_species_means_habit, recovery_mean, "Recovery") +
+  scale_y_continuous(labels = percent)
+
+a + b + c + guide_area() + plot_layout(ncol = 2, guides = 'collect') + plot_annotation(tag_levels = "a")
+```
+
+**`r s_figure("traits-box")`**.
+Box plots comparing physiological parameters between sporophytes (red) and gametophytes (blue) across growth habits.
+Each point represents one species; dark line is median, lower and upper edges correspond to first and third quartiles, and whiskers extend to the most extreme value no further than 1.5 times the inter-quartile range.
+**a**, ETR~max~ (µmol electrons m^--2^ ·s^--1^); **b** PPFD~95%~ (µmol photons m^--2^ ·s^--1^); **c**, recovery (%) of chlorophyll fluorescence (F~v~/F~m~) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa).
+*Crepidomanes bipunctatum* and *C.
+humile* were observed growing as both epiphytes and saxicoles, but are treated as epiphytes to distinguish them from exclusively saxicolous species.
+
+\newpage
+
+```{r capi-water, fig.height = 4, fig.width = 4, fig.cap = ""}
+# Calculate recovery from 100% RH treatment for Callistopteris apiifolia gametophytes
+cal_water_recovery <- filmy_dt %>%
+  # Only include samples that should be presented in the main MS
+  filter(salt == "H2O", species == "Callistopteris_apiifolia") %>%
+  # Filter out individuals with low pre-treatment yields
+  filter(!is.na(yield_pre), yield_pre > 400) %>%
+  calculate_recovery() %>%
+  calculate_mean_recovery() %>%
+  rename(recovery = recovery_mean, sd = recovery_sd) %>%
+  verify(length(unique(recovery_n)) == 1)
+
+cal_water_n <- cal_water_recovery$recovery_n %>% unique()
+
+# Plot recovery
+ggplot(cal_water_recovery, aes(x = rec_time, y = recovery, group = interaction(dry_time, salt))) +
+  geom_line() +
+  geom_errorbar(
+    aes(ymin = recovery - sd, ymax = recovery + sd), 
+    colour = "grey50", width = 0.1) +
+  geom_point(size = 1.5) +
+  scale_y_continuous(label = label_percent(accuracy = 1), limits = c(-0.1,1.2), expand = c(0,0), breaks=c(0,.25,.5,.75,1.0)) +
+  scale_x_discrete(limits = c("30min","24hr","48hr"),
+                   labels = c("0.5", "24", "48")) +
+  xlab("Recovery Time (hr)") +
+  ylab("Recovery (%)") +
+  labs(
+    x = "Recovery Time (hr)",
+    y = "Recovery (%)"
+  )
+```
+
+**`r s_figure("capi-water")`**.
+Recovery of *Callistopteris apiifolia* gametophytes from Moorea, French Polynesia from desiccation treatment at 100 % RH (water used in place of desiccation salt).
+*n* = `r cal_water_n`.

### 2021-02-22 R/functions.R

@@ -827,6 +827,8 @@ calculate_mean_light <- function (data) {
 }
 
 #' Calculate mean of daily env_range_dt_model VPD from climate data
+#' 
+#' Exclude outlier Rotui_800m_slope site
 #'
 #' @param climate Dataframe with climate data including
 #' relative humidity, temperature, and VPD calculated from these measured
@@ -835,6 +837,7 @@ calculate_mean_light <- function (data) {
 #' @return Tibble with mean of daily env_range_dt_model VPD by site and growth habit 
 calculate_mean_max_vpd <- function (climate) {
   climate %>% 
+    filter(site != "Rotui_800m_slope") %>%
     group_by(site, habit, date) %>% 
     summarize(vpd = max(vpd), .groups = "drop") %>% 
     assert(not_na, vpd) %>%
@@ -1319,6 +1322,8 @@ run_glmm <- function(combined_species_means, traits, phy) {
   
   # set number of iterations
   num <- 500000
+  n_burnin = 0.0002*num
+  n_thin = 0.0002*0.5*num
   
   # define prior list. Need one G for each random effect, one R for each fixed effect
   my_prior <- list(G=list(G1=list(V=1,nu=0.02)),
@@ -1350,7 +1355,7 @@ run_glmm <- function(combined_species_means, traits, phy) {
           family = "gaussian", 
           prior = my_prior, 
           data = as.data.frame(.y),
-          nitt = num, burnin = 1000, thin = 500,
+          nitt = num, burnin = n_burnin, thin = n_thin,
           verbose = FALSE)
       )
     )
@@ -1382,7 +1387,7 @@ run_glmm <- function(combined_species_means, traits, phy) {
           ginverse = list(species = dt_phy_inv$Ainv), 
           prior = my_prior, 
           data = as.data.frame(.y),
-          nitt = num, burnin = 1000, thin = 500,
+          nitt = num, burnin = n_burnin, thin = n_thin,
           verbose = FALSE)
       )
     )

### 2021-02-25 ms/manuscript.Rmd
Update MS, SI
- add ESM 2 (gameto dt batches)
- add time bars for gameto dt treatment, 2012 data

@@
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
-A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (\figs `r figure_num("habit")`, `r s_figure_num("tree")`).
+A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (\figs `r figure_num("habit")`, `r s_figure_num("tree")`; all supplemental figures available in Online Resource 1).
 The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
@@
 Samples were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
-Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 30 min (2012 field season) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014 field seasons).
+Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012 field season) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014 field seasons).
 The desiccation chambers were kept in an air-conditioned room. 
-Temperature inside the chambers was mean `mean(filmy_dt_chamber$temp) %>% round(1)` ± SD `r sd(filmy_dt_chamber$temp) %>% round(2)` °C (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
+Temperature inside the chambers was mean `mean(filmy_dt_chamber$temp) %>% round(1)` ± SD `r sd(filmy_dt_chamber$temp) %>% round(2)` °C (\figs `r s_figure_num("chamber")`).
 Salts used for desiccation and their corresponding mean water potentials and approximate relative humidity and VPD are as follows: LiCl (-282 MPa, 18 % RH, 2.45 kPa), Mg(NO~3~)~2~ (-86 MPa, 58 % RH, 1.25 kPa), and NaCl (-38 MPa, 80 % RH, 0.60 kPa).
@@
 Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible.
-We therefore applied the same treatment for all gametophytes (2 d at -86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described above.
+We therefore applied the same treatment for all gametophytes (2 d at -86 MPa) in batches of individuals (`r s_table("gameto-batches")`), and verified the species of each individual after the experiment using the DNA barcode approach described above.
 Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
@@
 
-Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (\figs{ }`r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
+Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (\figs{ }`r s_figure_num("chamber")`).
 Brief fluctuations to higher humidities were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).

### 2021-02-25 ms/si.Rmd
Update MS, SI
- add ESM 2 (gameto dt batches)
- add time bars for gameto dt treatment, 2012 data

@@
       in_header: journal-of-plant-research-si.sty
-# enable landscape mode selectively with pdflscape and newcommand 
 # https://stackoverflow.com/questions/25849814/rstudio-rmarkdown-both-portrait-and-landscape-layout-in-a-single-pdf
@@
 tar_load(combined_species_means)
+tar_load(gameto_indiv_times)
 ```
@@
 
-**Electronic supplementary materials**
+**Electronic supplementary materials: Online Resource 1**
 
@@
 
-**Content:** \figs `r s_figure_num("tree")`--`r s_figure_num("light-curves")`
-
-**`r s_figure("tree")`**. Maximum likelihood ultrametric phylogenetic tree including all Moorean filmy fern species.
-
-**`r s_figure("chamber-s")`**. Conditions inside desiccation chamber and duration of desiccation treatment for sporophytes.
-
-**`r s_figure("chamber-g")`**. Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes.
+**Content:** \figs `r s_figure_num("tree")`--`r s_figure_num("capi-water")`
+
+**`r s_figure("tree")`**. Maximum likelihood ultrametric phylogenetic tree including all filmy fern species from Moorea, French Polynesia.
+
+**`r s_figure("chamber")`**. Temperature, relative humidity (RH), and duration of desiccation treatment of filmy ferns from Moorea, French Polynesia.
 
@@
 **`r s_figure("light-curves")`**. Rapid light response curves of filmy ferns from Moorea, French Polynesia.
+
+**`r s_figure("traits-box")`**. Box plots comparing physiological parameters between sporophytes (red) and gametophytes (blue) across growth habits in filmy ferns from Moorea, French Polynesia.
+
+**`r s_figure("capi-water")`**. Recovery of *Callistopteris apiifolia* gametophytes from Moorea, French Polynesia from desiccation treatment at 100 % RH (water used in place of desiccation salt).
 
@@
 
-**`r s_figure("tree")`**. Maximum likelihood ultrametric phylogenetic tree including all Moorean filmy fern species.
+**`r s_figure("tree")`**. Maximum likelihood ultrametric phylogenetic tree including all filmy fern species from Moorea, French Polynesia.
 Tree obtained by extracting Hymenophyllaceae from the tree of Nitta et al. (2016) with the extract.clade function of the ape package in R (Paradis et al., 2004).
@@
   scale_color_manual(values = serial_no_cols, drop = FALSE) +
+  labs(x = "Species", color = "Serial no.") +
   coord_flip() +
@@
   coord_flip() +
-  labs(x = "Species") +
+  labs(x = "Species", color = "Serial no.") +
   theme(
@@
     y = "Relative humidity",
-    color = "Data logger\nserial no.") +
+    color = "Serial no.") +
   guides(color = "none") +
@@
     y = "Relative humidity",
-    color = "Data logger\nserial no.") +
+    color = "Serial no.") +
   theme(
@@
     y = "Temperature",
-    color = "Data logger\nserial no.",
+    color = "Serial no.",
     title = "2012") +
@@
     y = "Temperature",
-    color = "Data logger\nserial no.",
+    color = "Serial no.",
     title = "2013") +
@@
     y = "Temperature",
-    color = "Data logger\nserial no.",
+    color = "Serial no.",
     title = "2014") +
@@
 
-**`r s_figure("chamber-s")`**.
-Temperature, relative humidity (RH), and duration of desiccation treatment for sporophytes.
-Relative humidity was controlled by placing drying salts inside desiccation chambers: **a**, LiCl (-282 MPa, 18 % RH); **b**, Mg(NO~3~)~2~ (-86 MPa, 58 % RH); **c**, NaCl (-38 MPa, 80 % RH; **c**.
-Temperature and RH measured with Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 30 min (2012) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014).
-For bottom-most panels (duration of desiccation treatment), line indicates duration of treatment and point indicates removal of sample from chamber.
-For **c**, multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color.
-"missing" indicates data were not collected on the correspondence between samples and data loggers (chambers).
-
-\newpage
-
-```{r gametophyte-dt-2013, fig.height = 4.5, fig.width = 7, fig.cap = ""}
-# Calculate mean start/end times for gametophyte DT groups
+```{r gametophyte-dt, fig.height = 4.5, fig.width = 7, fig.cap = ""}
+# Combine and calculate mean start/end times by group (cluster)
 gameto_mean_times <-
-  filmy_dt %>%
-  filter(generation == "gametophyte") %>%
-  select(individual, contains("time_")) %>%
-  pivot_longer(names_to = "condition", values_to = "date_time", -individual) %>%
-  filter(!is.na(date_time)) %>%
-  mutate(condition = str_remove_all(condition, "time_") %>%
-           str_replace_all("30min", "post")) %>%
-  filter(condition %in% c("pre", "post")) %>%
-  add_count(individual) %>%
-  filter(n > 1) %>%
-  select(-n) %>%
-  pivot_wider(names_from = "condition", values_from = "date_time") %>%
-  # Cluster starts times into groups (here, I know there are 10 groups)
-  add_clusters("pre", 10) %>%
+  gameto_indiv_times %>%
   mutate(cluster = as.factor(cluster)) %>%
@@
 
-gameto_time_limits <- c(min = min(gameto_mean_times$pre), max = max(gameto_mean_times$post))
+# Set time limits so plots line up vertically
+gameto_times_2012 <- gameto_mean_times %>% filter(pre < "2013-01-01")
+gameto_times_2013 <- gameto_mean_times %>% filter(pre > "2013-01-01")
+gameto_time_limits_2013 <- c(min = min(gameto_times_2013$pre), max = max(gameto_times_2013$post))
+gameto_time_limits_2012 <- c(min = min(gameto_times_2012$pre), max = max(gameto_times_2012$post))
+
+# Subset DT chamber data
+chamber_gameto_2013 <- filmy_dt_chamber %>% filter(generation == "gametophyte", year == "2013")
+chamber_gameto_2012 <- filmy_dt_chamber %>% 
+  filter(
+    salt == "MgNO3", 
+    year == "2012",
+    generation == "gametophyte") %>%
+  mutate(group = case_when(
+    date_time < "2012-07-17 00:00:00" ~ "a",
+    date_time > "2012-07-17 00:00:00" ~ "b"
+  ))
 
@@
   geom_point(position = position_dodge(0.5)) +
-  scale_y_datetime(limits = gameto_time_limits) +
+  scale_y_datetime(limits = gameto_time_limits_2013) +
   coord_flip() +
-  labs(x = "Group") +
+  labs(x = "Batch") +
   theme(
@@
 
-# Subset DT chamber data
-chamber_gameto_2013 <- filmy_dt_chamber %>% filter(generation == "gametophyte", year == "2013")
+dt_times_gameto_2012_plot <- ggplot(
+  gameto_mean_times, 
+  aes(x = cluster, ymin = pre, ymax = post, y = post, group = cluster)) +
+  geom_linerange(position = position_dodge(0.5)) +
+  geom_point(position = position_dodge(0.5)) +
+  scale_y_datetime(limits = gameto_time_limits_2012) +
+  coord_flip() +
+  labs(x = "Batch") +
+  theme(
+    axis.title.x = element_blank()
+  ) +
+  blank_y_theme()
 
@@
 chamber_gameto_2013_rh_plot <-
-  plot_chamber_rh(chamber_gameto_2013, gameto_time_limits)
+  plot_chamber_rh(chamber_gameto_2013, gameto_time_limits_2013)
+
+chamber_gameto_2012_rh_plot <-
+  ggplot(chamber_gameto_2012, aes(x = date_time, y = 0.01*rh, group = group)) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = gameto_time_limits_2012) +
+  scale_y_continuous(label = label_percent(accuracy = 1)) +
+  labs(y = "Relative humidity") +
+  theme(
+    legend.position = "none",
+    axis.title.x = element_blank()) +
+  blank_y_theme()
 
@@
 chamber_gameto_2013_temp_plot <-
-  plot_chamber_temp(chamber_gameto_2013, gameto_time_limits)
-
-chamber_gameto_2013_temp_plot + chamber_gameto_2013_rh_plot + dt_times_gameto_2013_plot + plot_layout(ncol = 1)
-
-# ggsave(here::here("results/S1_fig.png"))
-```
-
-**`r s_figure("chamber-g")`**.
-Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes.
-MgNO~3~ was used to maintain constant humidity at \circa{ }58%.
-All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days unintentionally
+  plot_chamber_temp(chamber_gameto_2013, gameto_time_limits_2013) +
+  labs(title = "2013") +
+  theme(plot.title = element_text(hjust = 0.5, size = 11))
+
+chamber_gameto_2012_temp_plot <-
+  ggplot(chamber_gameto_2012, aes(x = date_time, y = temp, group = group)) +
+  geom_line(size = 0.25) +
+  scale_x_datetime(limits = gameto_time_limits_2012) +
+  scale_y_continuous(label = label_number(suffix = "\u00b0C", accuracy = 1)) +
+  labs(
+    y = "Temperature",
+    title = "2012") +
+  theme(
+    legend.position = "none",
+    plot.title = element_text(hjust = 0.5, size = 11),
+    axis.title.x = element_blank()) +
+  blank_y_theme()
+
+# Combine into final plot
+chamber_gameto_2013_temp_plot + chamber_gameto_2012_temp_plot +
+  chamber_gameto_2013_rh_plot + chamber_gameto_2012_rh_plot + 
+  dt_times_gameto_2013_plot + dt_times_gameto_2012_plot + 
+  plot_layout(ncol = 2, nrow = 3) +
+  plot_annotation(title = "d") &
+  theme(plot.tag = element_text(face = "bold"))
+```
+
+\newpage
+
+**`r s_figure("chamber")`**.
+Temperature, relative humidity (RH), and duration of desiccation treatment of filmy ferns from Moorea, French Polynesia.
+**a**--**c**, sporophytes; **d**, gametophytes.
+Relative humidity was controlled by placing drying salts inside desiccation chambers: **a**, LiCl (-282 MPa, 18 % RH); **b**, **d**, Mg(NO~3~)~2~ (-86 MPa, 58 % RH); **c**, NaCl (-38 MPa, 80 % RH).
+Temperature and RH measured with Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014).
+For **c**, multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color; "missing" indicates data were not collected on the correspondence between samples and data loggers (chambers).
+For bottom-most panels (duration of desiccation treatment), line indicates duration of treatment and point indicates removal of sample from chamber.
+Gametophytes were measured in batches; membership of individual gametophytes to each batch can be found in `r s_table("gameto-batches")` (Online Resource 2).
 
@@
 **`r s_figure("traits-box")`**.
-Box plots comparing physiological parameters between sporophytes (red) and gametophytes (blue) across growth habits.
+Box plots comparing physiological parameters between sporophytes (red) and gametophytes (blue) across growth habits in filmy ferns from Moorea, French Polynesia.
 Each point represents one species; dark line is median, lower and upper edges correspond to first and third quartiles, and whiskers extend to the most extreme value no further than 1.5 times the inter-quartile range.
@@
                    labels = c("0.5", "24", "48")) +
-  xlab("Recovery Time (hr)") +
-  ylab("Recovery (%)") +
   labs(
     x = "Recovery Time (hr)",
-    y = "Recovery (%)"
+    y = "Recovery"
   )

### 2021-02-25 R/functions.R

@@ -477,6 +477,24 @@ unzip_nitta_2017 <- function (zipped_path, unzip_path, ...) {
   
 }
 
+#' Load data on gametophyte DT test times from 2012
+#'
+#' @param file Path to data file with start and end times for gametophyte DT test in 2012
+#'
+#' @return Dataframe
+load_gameto_time_summary_2012 <- function(file) {
+  read_csv(
+    file,
+  col_types = cols(
+    date_time = col_datetime(format = ""),
+    event = col_character(),
+    group = col_character(),
+    cluster = col_double()
+  )) %>%
+  pivot_wider(values_from = "date_time", names_from = event) %>%
+  rename(pre = start, post = end)
+}
+
 # Data wrangling ----
 
 #' Cluster rows of a dataframe based on a single column
@@ -1072,6 +1090,116 @@ extract_fitted_lc_data <- function (light_models) {
   light_models %>%
   select(-nls_mod, -k_stats) %>%
   unnest(cols = c(data, fitted))
+}
+
+#' Format gametophyte DT data to include membership in treatment batches
+#'
+#' @param filmy_dt Cleaned DT data
+#' @param gameto_time_summary_2012 Data including start and end times of DT test for
+#' gametophytes in 2012
+#' @param filmy_specimens Filmy fern specimen data
+#'
+#' @return Dataframe
+#' 
+prepare_gameto_dt_groups <- function(filmy_dt, gameto_time_summary_2012, filmy_specimens) {
+  
+  # For gametophytes in 2012, individual level data is not available for start/end times,
+  # but there is data for start/end times of groups. So merge this by individual membership in groups.
+  gameto_times_2012 <-
+    filmy_dt %>%
+    filter(generation == "gametophyte") %>%
+    mutate(individual = parse_number(individual)) %>%
+    assert(not_na, individual) %>%
+    filter(individual < 1863 & individual > 984) %>% # lowest and highest collection numbers in 2012
+    mutate(
+      cluster = case_when(
+        individual < 1093 ~ 1, # epi/ter plots 1-3
+        individual >= 1093 ~ 2, # epi/ter plots 4-6
+        TRUE ~ NaN
+      )
+    ) %>%
+    select(individual, cluster) %>%
+    assert(not_na, everything()) %>%
+    assert(is_uniq, individual) %>%
+    left_join(gameto_time_summary_2012, by = "cluster") %>%
+    select(individual, pre, post, cluster) %>%
+    mutate(individual = as.character(individual))
+  
[…]

### 2021-02-25 _targets.R

@@ -59,6 +59,10 @@ tar_plan(
   tar_file(specimens_raw_file, "data/fern_specimens.csv"),
   specimens_raw = read_csv(specimens_raw_file),
   
+  # - gametophyte DT groups
+  tar_file(gameto_times_2012_file, "data/2012_gameto_dt_times.csv"),
+  gameto_time_summary_2012 = load_gameto_time_summary_2012(gameto_times_2012_file),
+  
   # Process data ----
   
   # - subset collection data to just filmy ferns on Moorea
@@ -141,6 +145,9 @@ tar_plan(
     light_species_means = light_species_means
   ),
   
+  # - make table of gametophyte DT treatment batches (groups)
+  gameto_indiv_times = prepare_gameto_dt_groups(filmy_dt, gameto_time_summary_2012, filmy_specimens),
+  
   # t-test ----
   # Perform two-sided t-test on DT and light responses across generations
   t_test_results = run_t_test(
@@ -203,7 +210,12 @@ tar_plan(
     depends = manuscript_pdf
   ),
   
-  # SI
-  tar_render(si_pdf, "ms/si.Rmd", output_dir = here::here("results/ms"))
+  # SI: figures (ESM 1)
+  tar_render(si_pdf, "ms/si.Rmd", output_file = "ESM_1.pdf", output_dir = here::here("results/si")),
+  
+  # SI: tables (ESM 2)
+  tar_file(
+    gameto_indiv_times_table,
+    write_gameto_times(gameto_indiv_times, "results/si/ESM_2.csv"))
   
 )

### 2021-02-26 ms/si.Rmd
Add 2012 time subplots to SI DT chamber figure

@@
 tar_load(gameto_indiv_times)
+tar_load(sporo_dt_times)
 ```
@@
     geom_line(size = 0.25) +
-    scale_x_datetime(limits = time_limits) +
+    scale_x_datetime(limits = time_limits, minor_breaks = "1 day") +
     scale_y_continuous(label = label_percent(accuracy = 1)) +
@@
     geom_line(size = 0.25) +
-    scale_x_datetime(limits = time_limits) +
+    scale_x_datetime(limits = time_limits, minor_breaks = "1 day") +
     scale_y_continuous(label = label_number(suffix = "\u00b0C", accuracy = 1)) +
@@
     geom_point(position = position_dodge(0.5)) +
-    scale_y_datetime(limits = time_limits) +
+    scale_y_datetime(limits = time_limits, minor_breaks = "1 day") +
     scale_x_discrete(labels = str_abbrev_sp_pretty) +
@@
 
-```{r sporophyte-dt-LiCl, fig.height = 6.5, fig.width = 9, fig.cap = ""}
+```{r change-theme-1}
+previous_theme <- theme_set(theme_bw(base_size = 10))
+```
+
+```{r sporophyte-dt-LiCl, fig.height = 6.6, fig.width = 9, fig.cap = ""}
 # Get mean times of each cluster of DT test samples for the selected drying salt
@@
 
-licl_clusters <- bind_rows(licl_clusters_2013, licl_clusters_2014)
+licl_clusters <- bind_rows(licl_clusters_2013, licl_clusters_2014) %>%
+  bind_rows(filter(sporo_dt_times, salt == "LiCl")) %>%
+  mutate(species_duration = paste(species, dry_time))
 
@@
 
-licl_time_range_2012 <- c(min = min(licl_chamber_data_2012$date_time), max = max(licl_chamber_data_2012$date_time))
+licl_time_range_2012 <- c(
+  # Manually entered data for sporo dt times LiCl was slightly earlier than 
+  # when started datalogger recording
+  min = filter(sporo_dt_times, salt == "LiCl") %>% pull(pre) %>% min, 
+  max = max(licl_chamber_data_2012$date_time)
+  )
 
@@
     legend.position = "none") +
+  blank_y_theme() +
   blank_x_theme()
@@
   theme(plot.title = element_text(hjust = 0.5, size = 11)) +
-  blank_y_theme() +
   blank_x_theme()
@@
   theme(legend.position = "none") +
-  blank_x_theme()
+  blank_x_theme() +
+  blank_y_theme()
 
@@
   scale_y_continuous(limits = licl_rh_range, label = label_percent(accuracy = 1)) +
-  blank_y_theme()
+  blank_x_theme()
 
 # Dry times plot
+licl_time_plot_2012 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2012) 
 licl_time_plot_2013 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2013) +
   blank_y_theme()
-licl_time_plot_2014 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2014)
+licl_time_plot_2014 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2014) +
+  blank_y_theme()
 
 # Combine plots
-licl_temp_plot_2014 + licl_temp_plot_2013 + licl_temp_plot_2012 +
-  licl_rh_plot_2014 + licl_rh_plot_2013 + licl_rh_plot_2012 +
-  licl_time_plot_2014 + licl_time_plot_2013 + plot_spacer() +
+licl_temp_plot_2012 + licl_temp_plot_2013 + licl_temp_plot_2014 +
+  licl_rh_plot_2012 + licl_rh_plot_2013 + licl_rh_plot_2014 +
+  licl_time_plot_2012 + licl_time_plot_2013 + licl_time_plot_2014 +
   plot_layout(ncol = 3, nrow = 3, guides = "collect") +
@@
 
-mgno3_clusters <- bind_rows(mgno3_clusters_2013, mgno3_clusters_2014)
+mgno3_clusters <- bind_rows(mgno3_clusters_2013, mgno3_clusters_2014) %>%
+  bind_rows(filter(sporo_dt_times, salt == "MgNO3")) %>%
+  mutate(species_duration = paste(species, dry_time))
 
@@
 
-mgno3_time_range_2012 <- c(min = min(mgno3_chamber_data_2012$date_time), max = max(mgno3_chamber_data_2012$date_time))
+mgno3_time_range_2012 <- c(
+  # Manually entered data for sporo dt times LiCl was slightly earlier than 
+  # when started datalogger recording
+  min = filter(sporo_dt_times, salt == "MgNO3") %>% pull(pre) %>% min, 
+  max = max(mgno3_chamber_data_2012$date_time)
+  )
 
@@
     legend.position = "none") +
-  blank_x_theme()
+  blank_x_theme() +
+  blank_y_theme()
 
@@
   geom_line(size = 0.25) +
-  scale_x_datetime(limits = mgno3_time_range_2014) +
+  scale_x_datetime(limits = mgno3_time_range_2014, minor_breaks = "1 day") +
   scale_y_continuous(
@@
   theme(legend.position = "none") +
+  blank_y_theme() +
   blank_x_theme()
@@
   scale_y_continuous(limits = mgno3_rh_range, label = label_percent(accuracy = 1)) +
-  blank_y_theme()
+  blank_x_theme()
 
 # Dry times plot
+mgno3_time_plot_2012 <- plot_sporo_chamber_time(mgno3_clusters, mgno3_time_range_2012)
 mgno3_time_plot_2013 <- plot_sporo_chamber_time(mgno3_clusters, mgno3_time_range_2013) +
   blank_y_theme()
-mgno3_time_plot_2014 <- plot_sporo_chamber_time(mgno3_clusters, mgno3_time_range_2014)
+mgno3_time_plot_2014 <- plot_sporo_chamber_time(mgno3_clusters, mgno3_time_range_2014) +
+  blank_y_theme()
 
 # Combine plots
-mgno3_temp_plot_2014 + mgno3_temp_plot_2013 + mgno3_temp_plot_2012 +
-  mgno3_rh_plot_2014 + mgno3_rh_plot_2013 + mgno3_rh_plot_2012 +
-  mgno3_time_plot_2014 + mgno3_time_plot_2013 + plot_spacer() +
+mgno3_temp_plot_2012 + mgno3_temp_plot_2013 + mgno3_temp_plot_2014 +
+  mgno3_rh_plot_2012 + mgno3_rh_plot_2013 + mgno3_rh_plot_2014 +
+  mgno3_time_plot_2012 + mgno3_time_plot_2013 + mgno3_time_plot_2014 +
   plot_layout(ncol = 3, nrow = 3, guides = "collect") +
@@
 \newpage
+
+```{r change-theme-2}
+theme_set(previous_theme)
+```
 

### 2021-02-26 R/functions.R

@@ -495,6 +495,27 @@ load_gameto_time_summary_2012 <- function(file) {
   rename(pre = start, post = end)
 }
 
+#' Load manually entered sporophyte DT times
+#'
+#' @param file Path to CSV file with manually entered sporophyte DT times
+#'
+#' @return Dataframe
+load_sporo_dt_times <- function(file) {
+  read_csv(
+    file, 
+    col_types = cols(
+      date_time = col_datetime(format = ""),
+      event = col_character(),
+      salt = col_character(),
+      dry_time = col_factor(),
+      species = col_character()
+    )) %>%
+    separate_rows(species, sep = ",") %>%
+    mutate(species = str_trim(species, "both")) %>%
+    pivot_wider(values_from = date_time, names_from = event) %>%
+    rename(pre = start, post = end)
+}
+
 # Data wrangling ----
 
 #' Cluster rows of a dataframe based on a single column

### 2021-02-26 _targets.R

@@ -63,6 +63,10 @@ tar_plan(
   tar_file(gameto_times_2012_file, "data/2012_gameto_dt_times.csv"),
   gameto_time_summary_2012 = load_gameto_time_summary_2012(gameto_times_2012_file),
   
+  # - sporophyte DT times entered manually
+  tar_file(sporo_dt_times_file, "data/2012_sporo_dt_times.csv"),
+  sporo_dt_times = load_sporo_dt_times(sporo_dt_times_file),
+  
   # Process data ----
   
   # - subset collection data to just filmy ferns on Moorea

### 2021-03-01 ms/si.Rmd
Update SI

@@
 Relative humidity was controlled by placing drying salts inside desiccation chambers: **a**, LiCl (-282 MPa, 18 % RH); **b**, **d**, Mg(NO~3~)~2~ (-86 MPa, 58 % RH); **c**, NaCl (-38 MPa, 80 % RH).
-Temperature and RH measured with Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014).
+Temperature and RH measured with Track-It RH/Temp dataloggers logging every 10 min (2012) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014).
 For **c**, multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color; "missing" indicates data were not collected on the correspondence between samples and data loggers (chambers).

### 2021-03-01 ms/manuscript.Rmd
Update data cleaning, MS, SI
- Include C. apiifolia 2012 NaCl data
- Change C. minutum DT data to var 2 for both 2012 and 2014 (as confirmed by checking lab and field notes)
- Add times for all NaCl plots in SI

@@
 glmm_params %>% 
-  filter(response == "recovery") %>%
+  filter(response == "recovery", pMCMC < 0.05) %>%
   verify(predictor == "generationsporophyte") %>%
@@
   scales::percent(accuracy = 0.1)
-
-t_test_results
 
@@
 max_ter_vpd <- mean_vpd_sporo %>%
-  left_join(filmy_habit) %>%
+  left_join(filmy_habit, by = "species") %>%
   filter(habit %in% c("Saxicolous", "Terrestrial")) %>%

### 2021-03-01 ms/si.Rmd
Update data cleaning, MS, SI
- Include C. apiifolia 2012 NaCl data
- Change C. minutum DT data to var 2 for both 2012 and 2014 (as confirmed by checking lab and field notes)
- Add times for all NaCl plots in SI

@@
 
+# NaCl plots need separate funcs because they involve color by serial number
+make_nacl_time_plot <- function (data, serial_no_cols, time_limits) {
+  ggplot(
+    data, 
+    aes(
+      x = species, ymin = pre, ymax = post, y = post, 
+      group = species_duration, color = serial_no)
+  ) +
+  geom_linerange(position = position_dodge(0.5)) +
+  geom_point(position = position_dodge(0.5)) +
+  scale_y_datetime(limits = time_limits, minor_breaks = "1 day") +
+  # Print all 'species' on the y-axis using `drop = false`
+  scale_x_discrete(labels = str_abbrev_sp_pretty, drop = FALSE) +
+  # Print all levels of 'serial no' in the legend using `drop = false`
+  scale_color_manual(values = serial_no_cols, drop = FALSE) +
+  labs(x = "Species", color = "Serial no.") +
+  coord_flip() +
+  theme(
+    axis.title = element_blank(),
+    axis.text.y = element_markdown()
+    )
+}
+
+make_nacl_rh_plot <- function(data, serial_no_cols, time_range, rh_range, groups = FALSE) {
+
+  # If groups is TRUE, plot separate lines for different groups of data
+  p <- if(groups == TRUE) {
+    ggplot(
+      data, 
+      aes(x = date_time, y = 0.01*rh, color = serial_no, group = interaction(group, serial_no))
+    ) 
+  } else {
+    ggplot(
+      data, 
+      aes(x = date_time, y = 0.01*rh, color = serial_no)
+    )
+  }
+
+  p +
+  geom_line(size = 0.25) +
+  scale_color_manual(values = serial_no_cols) +
+  scale_x_datetime(limits = time_range, minor_breaks = "1 day") +
+  scale_y_continuous(
+    label = scales::label_percent(accuracy = 1), 
+    limits = rh_range) +
+  labs(
+    y = "Relative humidity",
+    color = "Serial no.") +
+  guides(color = "none") +
+  theme(axis.title.x = element_blank()) +
+  guides(color = "none")
+}
+
+make_nacl_temp_plot <- function(data, serial_no_cols, time_range, temp_range, title, groups = FALSE) {
+
+  # If groups is TRUE, plot separate lines for different groups of data
+  p <- if(groups == TRUE) {
+    ggplot(
+      data, 
+      aes(x = date_time, y = temp, color = serial_no, group = interaction(group, serial_no))
+    ) 
+  } else {
+    ggplot(
+      data, 
+      aes(x = date_time, y = temp, color = serial_no)
+    )
+  }
+
+  p +
+  geom_line(size = 0.25) +
+  scale_color_manual(values = serial_no_cols) +
+  scale_x_datetime(limits = time_range, minor_breaks = "1 day") +
+  scale_y_continuous(
+    label = label_number(suffix = "\u00b0C", accuracy = 1), 
+    limits = temp_range) +
+  labs(
+    y = "Temperature",
+    color = "Serial no.",
+    title = title) +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 11),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()) +
+  guides(color = "none")
+}
+
 # Define function to plot facetted light curve
@@
 # Subset desiccation duration data
+nacl_clusters_2012 <- filter(sporo_dt_times, salt == "NaCl") %>%
+  mutate(
+    serial_no = case_when(
+      species %in% c(
+        "Hymenophyllum_polyanthos", "Crepidomanes_humile", 
+        "Crepidomanes_bipunctatum", "Polyphlebium_endlicherianum", 
+        "Crepidomanes_minutum2") ~ "3200919",
+      species == "Callistopteris_apiifolia" ~ "10140667, 10140659",
+      TRUE ~ NA_character_
+    )
+  ) %>%
+  mutate(year = "2012", species_duration = paste(species, dry_time))
+
 # H_polyanthos shows n=1, but this is only because pre/post were available for only one sample
@@
   cluster_sporo_times("NaCl", 4) %>%
+  # Exclude A. dentatum, which is lacking chamber datalogger data
+  filter(species != "Abrodictyum_dentatum") %>%
   mutate(
@@
       species %in% c("Hymenophyllum_polyanthos") ~ "10152774",
-      species %in% c("Abrodictyum_dentatum") & dry_time == "2" ~ "missing",
       TRUE ~ "missing"
-    ),
-    serial_no = as.factor(serial_no)
-  )
+    )
+  ) %>%
+  mutate(year = "2013")
 
@@
       TRUE ~ "10152773"
-    ),
-    serial_no = as.factor(serial_no)
-  )
-
-# make levels of `serial_no` and `species` the same across datasets
-nacl_all_sn <- filmy_dt_chamber %>% 
-  filter(salt == "NaCl") %>% 
-  pull(serial_no) %>% 
-  unique %>% 
-  sort %>%
-  c("missing")
-
-nacl_all_species <- sporo_times %>% 
-  filter(salt == "NaCl") %>% 
-  pull(species) %>% 
-  unique %>% 
-  sort
-
-nacl_clusters_2013 <- mutate(
-  nacl_clusters_2013, 
-  # sort SN in numeric order
-  serial_no = fct_expand(serial_no, nacl_all_sn) %>% fct_inseq,
-  # sort species in reverse-alphabetical order (so 'A' is on top in plot)
-  species = fct_expand(species, nacl_all_species) %>% fct_relevel(function (x) sort(x, decreasing = TRUE))
-)
-
-nacl_clusters_2014 <- mutate(
-  nacl_clusters_2014, 
-  serial_no = fct_expand(serial_no, nacl_all_sn) %>% fct_inseq,
-  species = fct_expand(species, nacl_all_species) %>% fct_relevel(function (x) sort(x, decreasing = TRUE))
-)
+    )
+  ) %>%
+  mutate(year = "2014")
 
@@
 # to standardize x-axis
+nacl_time_range_2012 <- c(min = min(nacl_clusters_2012$pre), max = max(nacl_clusters_2012$post))
 nacl_time_range_2013 <- c(min = min(nacl_clusters_2013$pre), max = max(nacl_clusters_2013$post))
@@
 
-# Subset DT chamber data
 nacl_chamber_data_2012 <-
   filmy_dt_chamber %>%
-  filter(salt == "NaCl", date_time > "2012-07-31 18:10:00 UTC", date_time < "2013-01-01 00:10:00 UTC")
-
-nacl_time_range_2012 <- c(min = min(nacl_chamber_data_2012$date_time), max = max(nacl_chamber_data_2012$date_time))
+  mutate(year = as.character(year)) %>%
+  filter(
+    date_time >= nacl_time_range_2012[["min"]], 
+    date_time <= nacl_time_range_2012[["max"]], 
+    salt == "NaCl")
 
 nacl_chamber_data_2013 <-
-  filter(filmy_dt_chamber, 
+  filmy_dt_chamber %>%
+  mutate(year = as.character(year)) %>%
+  filter( 
     date_time >= nacl_time_range_2013[["min"]], 
@@
 nacl_chamber_data_2014 <- 
-  filter(filmy_dt_chamber, 
+  filmy_dt_chamber %>%
+  mutate(year = as.character(year)) %>%
+  filter( 
     date_time >= nacl_time_range_2014[["min"]], 
@@
 
+# Set up color palette for serial numbers
+serial_no_cols <- c(
+  brewer_pal(palette = "Dark2")(8),
+  "grey20") %>%
+  set_names(
+    bind_rows(
+      nacl_clusters_2012, nacl_clusters_2013, nacl_clusters_2014, 
+      nacl_chamber_data_2012, nacl_chamber_data_2013, nacl_chamber_data_2014) %>%
+      filter(serial_no != "missing", !is.na(serial_no)) %>%
+      pull(serial_no) %>%
+      unique %>%
+      sort %>%
+      c("missing")
+  )
+
+# Combine NaCl data so that factor levels are same across years
+nacl_chamber_data <- bind_rows(nacl_chamber_data_2012, nacl_chamber_data_2013, nacl_chamber_data_2014) %>%
+  mutate(
+    year = as.character(year),
+    serial_no = as.factor(serial_no) %>% 
+      fct_expand(names(serial_no_cols)) %>% 
+      fct_relevel(function (x) sort(x, decreasing = TRUE)) %>%
+      fct_relevel("10140667, 10140659", after = Inf) %>%
+      fct_relevel("missing", after = Inf))
+
+nacl_clusters <- bind_rows(nacl_clusters_2012, nacl_clusters_2013, nacl_clusters_2014) %>%
+  mutate(
+    species = as.factor(species) %>% fct_relevel(function (x) sort(x, decreasing = TRUE)),
+    serial_no = as.factor(serial_no) %>% 
+      fct_expand(names(serial_no_cols)) %>% 
+      fct_relevel(function (x) sort(x, decreasing = TRUE)) %>%
+      fct_relevel("10140667, 10140659", after = Inf) %>%
+      fct_relevel("missing", after = Inf))
+
 # Get overall range of temperature and RH for NaCl
 # to standardize y-axis
-nacl_temp_range <- filter(filmy_dt_chamber, salt == "NaCl") %>%
+nacl_temp_range <- nacl_chamber_data %>%
   summarize(min = min(temp), max = max(temp)) %>%
@@
 
-nacl_rh_range <- filter(filmy_dt_chamber, salt == "NaCl") %>%
+nacl_rh_range <- nacl_chamber_data %>%
   mutate(rh = rh * 0.01) %>%
@@
 
-# Set up color palette for serial numbers
-serial_no_cols <- c(
-  brewer_pal(palette = "Dark2")(5),
-  "grey20") %>%
-  set_names(
-    bind_rows(nacl_chamber_data_2013, nacl_chamber_data_2014) %>%
-      filter(serial_no != "missing") %>%
-      pull(serial_no) %>%
-      unique %>%
-      sort %>%
-      c("missing")
-  )
-
 # Time plots
-nacl_time_plot_2013 <- 
-  ggplot(
-    nacl_clusters_2013, 
-    aes(
-      x = species, ymin = pre, ymax = post, y = post, 
-      group = species_duration, color = serial_no)
-  ) +
-  geom_linerange(position = position_dodge(0.5)) +
-  geom_point(position = position_dodge(0.5)) +
-  scale_y_datetime(limits = nacl_time_range_2013) +
-  # Print all 'species' on the y-axis using `drop = false`
-  scale_x_discrete(labels = str_abbrev_sp_pretty, drop = FALSE) +
-  # Print all levels of 'serial no' in the legend using `drop = false`
-  scale_color_manual(values = serial_no_cols, drop = FALSE) +
-  labs(x = "Species", color = "Serial no.") +
-  coord_flip() +
-  theme(
-    axis.title = element_blank(),
-    axis.text.y = element_blank())
-
-nacl_time_plot_2014 <- 
-  ggplot(
-    nacl_clusters_2014, 
-    aes(x = species, ymin = pre, ymax = post, y = post, 
-        group = species_duration, color = serial_no)
-  ) +
-  geom_linerange(position = position_dodge(0.5)) +
-  geom_point(position = position_dodge(0.5)) +
-  scale_y_datetime(limits = nacl_time_range_2014) +
-  scale_x_discrete(labels = str_abbrev_sp_pretty, drop = FALSE) +
-  scale_color_manual(values = serial_no_cols) +
-  coord_flip() +
-  labs(x = "Species", color = "Serial no.") +
-  theme(
-    axis.title = element_blank(),
-    axis.text.y = element_markdown()) +
-  guides(color = "none")
-
-# RH plots
-nacl_rh_plot_2012 <- 
-  ggplot(
-    nacl_chamber_data_2012, 
-    aes(x = date_time, y = 0.01*rh)
-  ) +
-  geom_line(size = 0.25) +
-  scale_x_datetime(limits = nacl_time_range_2012) +
-  scale_y_continuous(
-    label = scales::label_percent(accuracy = 1), 
-    limits = nacl_rh_range) +
-  labs(
-    y = "Relative humidity",
-    color = "Serial no.") +
-  guides(color = "none") +
-  theme(axis.title.x = element_blank()) +
-  blank_y_theme()
-
-nacl_rh_plot_2013 <- 
-  ggplot(
-    nacl_chamber_data_2013, 
-    aes(x = date_time, y = 0.01*rh, color = serial_no)
-  ) +
-  scale_color_manual(values = serial_no_cols) +
-  geom_line(size = 0.25) +
-  scale_x_datetime(limits = nacl_time_range_2013) +
-  scale_y_continuous(
-    label = scales::label_percent(accuracy = 1), 
-    limits = nacl_rh_range) +
-  labs(
-    y = "Relative humidity",
-    color = "Serial no.") +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()) +
+nacl_time_plot_2012 <- make_nacl_time_plot(
+  nacl_clusters %>% filter(year == "2012"), serial_no_cols, nacl_time_range_2012)
+
+nacl_time_plot_2013 <- make_nacl_time_plot(
+  nacl_clusters %>% filter(year == "2013"), serial_no_cols, nacl_time_range_2013) +
   blank_y_theme() +
@@
 
-nacl_rh_plot_2014 <- 
-  ggplot(
-    nacl_chamber_data_2014, 
-    aes(x = date_time, y = 0.01*rh, color = serial_no, group = interaction(group, serial_no))
-  ) +
-  scale_color_manual(values = serial_no_cols) +
-  geom_line(size = 0.25) +
-  scale_x_datetime(limits = nacl_time_range_2014) +
-  scale_y_continuous(
-    label = scales::label_percent(accuracy = 1), 
-    limits = nacl_rh_range) +
-  labs(
-    y = "Relative humidity",
-    color = "Serial no.") +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()) +
-  guides(color = "none")
-
-# Temp plots
-nacl_temp_plot_2012 <- 
-  ggplot(
-    nacl_chamber_data_2012, 
-    aes(x = date_time, y = temp)
-  ) +
-  geom_line(size = 0.25) +
-  scale_x_datetime(limits = nacl_time_range_2012) +
-  scale_y_continuous(
-    label = label_number(suffix = "\u00b0C", accuracy = 1), 
-    limits = nacl_temp_range) +
-  labs(
-    y = "Temperature",
-    color = "Serial no.",
-    title = "2012") +
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 11),
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()) +
+nacl_time_plot_2014 <- make_nacl_time_plot(
+  nacl_clusters %>% filter(year == "2014"), serial_no_cols, nacl_time_range_2014) +
   blank_y_theme() +
@@
 
-nacl_temp_plot_2013 <- 
-  ggplot(
-    nacl_chamber_data_2013, 
-    aes(x = date_time, y = temp, color = serial_no)
+# RH plots
+nacl_rh_plot_2012 <- make_nacl_rh_plot(
+  nacl_chamber_data %>% filter(year == "2012"),
+  serial_no_cols,
+  nacl_time_range_2012, 
+  nacl_rh_range
+)
+
+nacl_rh_plot_2013 <- make_nacl_rh_plot(
+  nacl_chamber_data %>% filter(year == "2013"),
+  serial_no_cols,
+  nacl_time_range_2013, 
+  nacl_rh_range
 ) + 
-  geom_line(size = 0.25) +
-  scale_color_manual(values = serial_no_cols) +
-  scale_x_datetime(limits = nacl_time_range_2013) +
-  scale_y_continuous(
-    label = label_number(suffix = "\u00b0C", accuracy = 1), 
-    limits = nacl_temp_range) +
-  labs(
-    y = "Temperature",
-    color = "Serial no.",
+  blank_y_theme() +
+  blank_x_theme()
+
+nacl_rh_plot_2014 <- make_nacl_rh_plot(
+  nacl_chamber_data %>% filter(year == "2014"),
+  serial_no_cols,
+  nacl_time_range_2014, 
+  nacl_rh_range,
+  groups = TRUE
+) + 
+  blank_y_theme() +
+  blank_x_theme()
+
+# Temp plots
+nacl_temp_plot_2012 <- make_nacl_temp_plot(
+  nacl_chamber_data %>% filter(year == "2012"),
+  serial_no_cols,
+  nacl_time_range_2012, 
+  nacl_temp_range,
+  title = "2012") +
+  guides(color = "none")
+
+nacl_temp_plot_2013 <- make_nacl_temp_plot(
+  nacl_chamber_data %>% filter(year == "2013"),
+  serial_no_cols,
+  nacl_time_range_2013, 
+  nacl_temp_range,
   title = "2013") + 
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 11),
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()) +
   blank_y_theme() +
-  guides(color = "none")
-
-nacl_temp_plot_2014 <- 
-  ggplot(
-    nacl_chamber_data_2014, 
-    aes(x = date_time, y = temp, color = serial_no, group = interaction(group, serial_no))
-  ) +
-  scale_color_manual(values = serial_no_cols) +
-  geom_line(size = 0.25) +
-  scale_x_datetime(limits = nacl_time_range_2014) +
-  scale_y_continuous(
-    label = label_number(suffix = "\u00b0C", accuracy = 1), 
-    limits = nacl_temp_range) +
-  labs(
-    y = "Temperature",
-    color = "Serial no.",
-    title = "2014") +
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 11),
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()) +
-  guides(color = "none")
-
-nacl_temp_plot_2014 + nacl_temp_plot_2013 + nacl_temp_plot_2012 +
-  nacl_rh_plot_2014 + nacl_rh_plot_2013 + nacl_rh_plot_2012 +
-  nacl_time_plot_2014 + nacl_time_plot_2013 + plot_spacer() +
+  blank_x_theme()
+
+nacl_temp_plot_2014 <- make_nacl_temp_plot(
+  nacl_chamber_data %>% filter(year == "2014"),
+  serial_no_cols,
+  nacl_time_range_2014, 
+  nacl_temp_range,
+  title = "2014",
+  groups = TRUE) + 
+  blank_y_theme() +
+  blank_x_theme()
+
+# Combine plots
+nacl_temp_plot_2012 + nacl_temp_plot_2013 + nacl_temp_plot_2014 +
+  nacl_rh_plot_2012 + nacl_rh_plot_2013 + nacl_rh_plot_2014 +
+  nacl_time_plot_2012 + nacl_time_plot_2013 + nacl_time_plot_2014 +
   plot_layout(ncol = 3, nrow = 3, guides = "collect") +
@@
 Relative humidity was controlled by placing drying salts inside desiccation chambers: **a**, LiCl (-282 MPa, 18 % RH); **b**, **d**, Mg(NO~3~)~2~ (-86 MPa, 58 % RH); **c**, NaCl (-38 MPa, 80 % RH).
-Temperature and RH measured with Track-It RH/Temp dataloggers logging every 10 min (2012) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014).
-For **c**, multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color; "missing" indicates data were not collected on the correspondence between samples and data loggers (chambers).
+Temperature and RH measured with Track-It RH/Temp dataloggers logging every 10 min (2012 species other than *Callistopteris apiifolia*) or Hobo ProV2 dataloggers logging every 5 min (2012 *Callistopteris apiifolia*, 2013, 2014).
+For NaCl (**c**), multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color; "missing" indicates correspondence between samples and data loggers (chambers) unknown.
 For bottom-most panels (duration of desiccation treatment), line indicates duration of treatment and point indicates removal of sample from chamber.

### 2021-03-01 R/clean_chamber_data.R

@@ -6,9 +6,21 @@ source(here::here("R/packages.R"))
 # Load functions
 source("R/functions.R")
 
-# 2012 data were measured with Track-It RH/Temp dataloggers (no serial numbers) 
-# for most samples, and Hobo ProV2 dataloggers (with serial numbers) for
-# Callistopteris apiifolia only
+# Load metadata for data loggers
+logger_metadata <- read_csv(
+  "data_raw/intermediates/logger_metadata.csv", 
+  col_types = cols(
+    file = col_character(),
+    year = col_double(),
+    salt = col_character(),
+    use_status = col_character(),
+    generation = col_character(),
+    note = col_character(),
+    serial_no = col_character()
+  ))
+
+# 2012 data were measured with Track-It RH/Temp dataloggers for most samples,
+# and Hobo ProV2 dataloggers for Callistopteris apiifolia only
 # 2012 sporophytes
 sporo_2012_chamber <- bind_rows(
   parse_logger_dat("data_raw/2012/track-it/filmyDT2_LiCl.csv") %>% mutate(salt = "LiCl"),
@@ -19,26 +31,22 @@ sporo_2012_chamber <- bind_rows(
   # Data before 2012-07-31 18:10:00 UTC was prelminary (NaCl datalogger
   # was actually placed in MgCl2, which wasn't used in final analysis). 
   # Filter to the period after this.
-  filter(date_time > "2012-07-31 18:10:00 UTC")
+  filter(date_time > "2012-07-31 18:10:00 UTC") %>%
+  left_join(
+    select(logger_metadata, salt, generation, year, species, file, serial_no),
+    c("salt", "generation", "year", "species")
+  )
   
 # 2012 gametophytes
 gameto_2012_chamber <-
   bind_rows(
-    parse_logger_dat("data_raw/2012/gameto1_7-14.csv"),
-    parse_logger_dat("data_raw/2012/gameto2_7-24.csv")
-  ) %>% mutate(salt = "MgNO3", generation = "gametophyte", year = 2012, species = "mixed")
-
-# 2013 load metadata for hobo loggers
-hobo_metadata <- read_csv(
-"data_raw/intermediates/hobo_metadata.csv", 
-col_types = cols(
-  file = col_character(),
-  year = col_double(),
-  salt = col_character(),
-  use_status = col_character(),
-  generation = col_character(),
-  note = col_character()
-))
[…]

### 2021-03-01 R/clean_dt_data.R

@@ -35,19 +35,29 @@ minipam_2013_long <-
   arrange(date_time)
 
 # 2012 sporophyte DT data ----
+
+# Use the sheet "averages (new cmin)", which includes all raw data plus a single C minutum sample (2d treatment)
+# that was lacking in the "raw data" sheet, and extra columns for SD and mean (we won't use those though)
+
 ### Fix column names ###
 # In the original excel file, these are in two rows.
 moorea_filmy_dt_2012_raw_path <- "data_raw/2012/Moorea Filmy DT 2 8-26 (version 1).xlsx"
 
-headers_1 <- read_excel(moorea_filmy_dt_2012_raw_path) %>% colnames()
+headers_1 <- read_excel(
+  moorea_filmy_dt_2012_raw_path, sheet = "averages (new cmin)") %>% 
+  colnames()
 
-headers_2 <- read_excel(moorea_filmy_dt_2012_raw_path, skip = 1) %>% colnames()
+headers_2 <- read_excel(
+  moorea_filmy_dt_2012_raw_path, sheet = "averages (new cmin)", skip = 1) %>% 
+  colnames()
 
-dt_2012_head_1 <- read_excel(moorea_filmy_dt_2012_raw_path, skip = 2, col_names = headers_1) %>%
+dt_2012_head_1 <- read_excel(
+  moorea_filmy_dt_2012_raw_path, sheet = "averages (new cmin)", skip = 2, col_names = headers_1) %>%
   remove_empty(which = "cols", quiet = FALSE) %>% # delete empty columns
   clean_names()
 
-dt_2012_head_2 <- read_excel(moorea_filmy_dt_2012_raw_path, skip = 2, col_names = headers_2) %>%
+dt_2012_head_2 <- read_excel(
+  moorea_filmy_dt_2012_raw_path, sheet = "averages (new cmin)", skip = 2, col_names = headers_2) %>%
   remove_empty(which = "cols", quiet = FALSE) %>%
   clean_names()
 
@@ -57,12 +67,13 @@ headers <- tibble(
 )
 
 # Write out these column headers, then manually edit new column names
-# write.csv(headers, "moorea_filmy_dt_2012_headers.csv")
+# write.csv(headers, "moorea_filmy_dt_2012_headers_sheet-2.csv")
 
 # Read in file with new column names
-new_headers <- read_csv("data_raw/intermediates/moorea_filmy_dt_2012_headers.csv")
[…]

### 2021-03-01 R/functions.R

@@ -417,6 +417,7 @@ load_filmy_dt <- function (file) {
         species == "Callistopteris_apiifolia" & salt == "H2O" ~ "si",
         species == "Hymenophyllum_polyanthos" & dataset == "2013_4" ~ "si",
         species == "Polyphlebium_borbonicum" & dataset == "2013_2" ~ "si",
+        species == "Crepidomanes_minutum2" & dataset == "2014" ~ "si",
         salt == "Control" ~ "si",
         TRUE ~ "main"
       )

### 2021-03-01 ms/manuscript.Rmd
Fix mistakenly assigning P. endlicherianum 2012 data to P. borbonicum

@@
 Samples were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
-Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012 field season) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014 field seasons).
+Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012 field season for species other than *Callistopteris apiifolia*) or Hobo ProV2 dataloggers logging every 5 min (2012 *Callistopteris apiifolia*, 2013, 2014 field seasons).
 The desiccation chambers were kept in an air-conditioned room. 
@@
 We tested the significance of λ with a log-likelihood test comparing the likelihood of the observed value of λ vs. λ = 0.
-For all further analyses, we used methods that account for phylogenetic history for physiological traits that showed significant phylogenetic signal, and standard methods otherwise.
+For all further analyses, we used methods that account for phylogenetic history for physiological traits that showed significant phylogenetic signal in at least one of the two metrics, and standard methods otherwise.
 
@@
 
-Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* = `r gameto_recover_k`, *P* = `r gameto_recover_k_pval` and λ = `r gameto_recover_lambda`, *P* = `r gameto_recover_lambda_pval`; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval` and λ = `r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`), but no significant signal for PPFD~95%~ or ETR~max~ (`r table("physig")`).
+Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were mostly similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT when measured with *K* (gametophytes *K* = `r gameto_recover_k`, *P* = `r gameto_recover_k_pval`; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval`), but no significant signal in either metric for PPFD~95%~ or ETR~max~ (`r table("physig")`).
+Results for phylogenetic signal in DT as measured with λ differed slightly; although gametophytes showed significant signal (λ = `r gameto_recover_lambda`, *P* = `r gameto_recover_lambda_pval`) sporophytes did not (λ =`r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`).
 
@@
 glmm_params %>% 
-  filter(response == "recovery", pMCMC < 0.05) %>%
-  verify(predictor == "generationsporophyte") %>%
+  filter(response == "recovery", predictor == "generationsporophyte") %>%
+  verify(pMCMC < 0.05) %>%
   pull(post.mean) %>%
@@
 glmm_params %>% 
-  filter(response == "recovery", pMCMC < 0.05) %>%
-  verify(predictor == "generationsporophyte") %>%
+  filter(response == "recovery", predictor == "generationsporophyte") %>%
+  verify(pMCMC < 0.05) %>%
   pull(`l-95% CI`) %>%
@@
 glmm_params %>% 
-  filter(response == "recovery", pMCMC < 0.05) %>%
-  verify(predictor == "generationsporophyte") %>%
+  filter(response == "recovery", predictor == "generationsporophyte") %>%
+  verify(pMCMC < 0.05) %>%
   pull(`u-95% CI`) %>%
@@
 
-The GLMM including the effect of generation only was selected as the best-fitting model for DT (`r table("glmm")`). 
+The GLMM including the combined effects of generation plus growth habit and their interaction was selected as the best-fitting model for DT (`r table("glmm")`). 
 Sporophytes had `r sporo_dt_effect_mean` greater DT (95 % CI `r sporo_dt_effect_lower` to `r sporo_dt_effect_upper`) relative to gametophytes (`r table("glmm-pars")`).

### 2021-03-01 ms/si.Rmd
Fix mistakenly assigning P. endlicherianum 2012 data to P. borbonicum

@@
 Relative humidity was controlled by placing drying salts inside desiccation chambers: **a**, LiCl (-282 MPa, 18 % RH); **b**, **d**, Mg(NO~3~)~2~ (-86 MPa, 58 % RH); **c**, NaCl (-38 MPa, 80 % RH).
-Temperature and RH measured with Track-It RH/Temp dataloggers logging every 10 min (2012 species other than *Callistopteris apiifolia*) or Hobo ProV2 dataloggers logging every 5 min (2012 *Callistopteris apiifolia*, 2013, 2014).
-For NaCl (**c**), multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color; "missing" indicates correspondence between samples and data loggers (chambers) unknown.
+Temperature and RH measured with Track-It RH/Temp dataloggers logging every 10 min (2012 species other than *Callistopteris apiifolia*) or Hobo ProV2 dataloggers logging every 5 min (2012 *C.
+apiifolia*, 2013, 2014).
+For NaCl (**c**), multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color; "missing" indicates correspondence between samples and data loggers (chambers) unknown; *C.
+apiifolia* samples divided into two chambers.
 For bottom-most panels (duration of desiccation treatment), line indicates duration of treatment and point indicates removal of sample from chamber.
@@
 ```{r change-theme-2}
+# Change theme back after DT chamber plots
 theme_set(previous_theme)

### 2021-03-01 R/clean_dt_data.R

@@ -96,7 +96,7 @@ dt_2012_all <-
       species == "C min" ~ "Crepidomanes_minutum2",
       species == "Cal api" ~ "Callistopteris_apiifolia",
       species == "H pol" ~ "Hymenophyllum_polyanthos",
-      species == "P end" ~ "Polyphlebium_borbonicum",
+      species == "P end" ~ "Polyphlebium_endlicherianum",
       TRUE ~ NA_character_
     )) %>%
   # Convert dry time to factor (control will be NA)

### 2021-03-01 R/functions.R

@@ -416,7 +416,6 @@ load_filmy_dt <- function (file) {
         species == "Callistopteris_apiifolia" & dataset == "2012" ~ "si",
         species == "Callistopteris_apiifolia" & salt == "H2O" ~ "si",
         species == "Hymenophyllum_polyanthos" & dataset == "2013_4" ~ "si",
-        species == "Polyphlebium_borbonicum" & dataset == "2013_2" ~ "si",
         species == "Crepidomanes_minutum2" & dataset == "2014" ~ "si",
         salt == "Control" ~ "si",
         TRUE ~ "main"

### 2021-03-01 _targets.R
Fix broken join that was dropping C. bipunctatum

@@ -120,9 +120,9 @@ tar_plan(
       mean_vpd_gameto,
       mean_vpd_sporo,
       range_comparison),
-    left_join,
+    full_join,
     by = "species"
-  ),
+  )  %>% assert(is_uniq, species),
   
   # - fit light curve models to data
   light_models = fit_lc_model(light_data),

### 2021-03-01 ms/manuscript.Rmd
Update MS
- add PPFD95% to combined plot
- add trendline to correlation plot
- update phy sig. results

@@
 tar_load(glmm_params)
+tar_load(env_range_dt_model)
 tar_load(env_range_dt_model_summary)
@@
 
-Desiccation tolerance tended to increase with VPD for sporophytes, but not gametophytes (`r figure("corr")`).
-This trend was marginally non-significant when analyzed using PGLS (*P* = `r sporo_vpd_pval`; `r table("pgls")`).
+Desiccation tolerance increased significantly with VPD for sporophytes, but not gametophytes (`r figure("corr")`; sporophytes *P* = `r sporo_vpd_pval`; `r table("pgls")`).
 
@@
 One may wonder then, what is the adaptive significance of being able to withstand drying at -282 MPa (\circa{ }2.45 kPa) vs.
--86 MPa (\circa{ }1.25 kPa) when the ambient VPD is typically < (\circa{ }0.10 kPa) and only rarely exceeds > 0.50 kPa.
+-86 MPa (\circa{ }1.25 kPa) when the ambient VPD is typically < \circa{ }0.10 kPa and only rarely exceeds > 0.50 kPa.
 The recent finding of a constitutive mechanism for DT in filmy ferns [@GarcesCea2014] supports the need for a relatively wide "safety margin," and similar (or even more extreme) levels of constitutive DT are well known in bryophytes [@Oliver1993].
@@
 We suggest that the variation in DT observed here is relevant because it is correlated with habitat, and likely reflects overall physiological tolerance, not just artificial extremes.
-For example, sporophytes of some high elevation epiphytes that we surveyed (*C.
-minutum* var.
-2, *H.
-digitatum*, *H.
+For example, sporophytes of some high elevation epiphytes that we surveyed (*H.
 multifidum*, *H.
@@
 Since controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (`r s_figure("dt-control")`), it is unlikely that the lack of tolerance seen in terrestrial species is due solely to removal from the soil.
-Comparison of DT with the average daily maximum VPD of each species (`r figure("corr")`) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with VPD exceeding \circa{ }`r max_ter_vpd` kPa.
+The significant relationship between average daily maximum VPD and DT in sporophytes supports the specialization of sporophyte DT according to niche, and indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with VPD exceeding \circa{ }`r max_ter_vpd` kPa (`r figure("corr")`).
 
@@
 apiifolia* are common on Moorea and have a unique morphology [@Nitta2017], we were able to collect enough material of this species for additional tests of DT at 100% RH for 2 d.
-However, it still failed to recover from even this gentle treatment (`r s_figure("capi-water")`.
+However, it still failed to recover from even this gentle treatment (`r s_figure("capi-water")`).
 This suggests that the widespread gametophytes of *C. apiifolia* do not rely on DT to attain their distribution, but instead must do so via some other mechanism.
@@
 Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates [@Proctor2003; @Proctor2012; @Parra2009; @Saldana2013]; however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
-We detected significant phylogenetic signal in DT, suggesting that phylogeny should be considered when interpreting differences in DT between species (`r table("physig")`).
+We detected significant phylogenetic signal in DT with Blomberg's *K* (`r table("physig")`), suggesting that phylogeny should be considered when interpreting differences in DT between species.
 In contrast, the lack of phylogenetic signal in photosynthetic traits, coupled with the clear intergenerational differences we observed, suggest that sporophytes and gametophytes are adapted for different light levels regardless of clade or ecological niche.
@@
 Left bar plot shows recovery (%) of chlorophyll fluorescence (F~v~/F~m~) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa). 
-Right bar plot shows maximum relative electron transport rate of photosystem II (ETR~max~). 
-The other photosynthetic parameter measured, photosynthetic photon flux density at 95% saturation of RETR (PPFD~95%~), was qualitatively very similar to ETR~max~ and not shown.
+Middle bar plot shows maximum relative electron transport rate of photosystem II (ETR~max~) (µmol e-·m^-2^·s^-1^). 
+Right bar plot shows photosynthetic photon flux density at 95% saturation of RETR (PPFD~95%~) (µmol quanta·m^-2^·s^-1^).
 Absence of a bar indicates measurements not available for gametophytes or sporophytes of that species.
@@
 Range of gametophytes beyond sporophytes calculated as the total elevation that gametophytes exceeded sporophytes at either end of the elevational gradient.
-Relationships between recovery and VPD or range of gametophytes beyond sporophytes were analyzed with phylogenetic generalized least squares (PGLS); no significant relationships were detected.
+Relationships between recovery and VPD or range of gametophytes beyond sporophytes were analyzed with phylogenetic generalized least squares (PGLS); significant relationships (*P* < 0.05) shown with grey line.
 
@@
 a <- ggtree(phy) %<+% habit +
-  geom_tippoint(aes(color = habit), position = position_nudge(x = 14, y = 0), size = 3.5) +
+  geom_tippoint(aes(color = habit), position = position_nudge(x = 14, y = 0), size = 3.0) +
   scale_color_manual(values = habit_colors, labels = pretty_text) +
@@
   # text disappear by setting fontsize = 0
-  geom_treescale(y = 12, x = 0, width = 40, fontsize = 0) +
+  geom_treescale(y = 12, x = 2, width = 40, fontsize = 0) +
   # Manually place custom label above scale bar
-  annotate("text", x = 20, y = 12, label = "40 my", size = 2.5, vjust = -1) +
+  annotate("text", x = 22, y = 12, label = "40 my", size = 2.5, vjust = -1) +
   scale_x_continuous(expand = expansion(mult = c(0.01, 0.04))) +
@@
     color = "grey20",
-    size = 3,
+    size = 2.5,
     vjust = -0.5) +
@@
     limits = c(0,1.2), 
-    expand = expansion(mult = c(0,0.025)), 
+    expand = expansion(mult = c(0,0.030)), 
     breaks=c(0,.25,.5,.75,1.0)) +
@@
 
-# Make helper table of significant t-test results
+# Make helper table of significant t-test results: ETR
 etr_t_test_data <-
@@
   rowwise %>%
+  # `etr_top` is the highest possible value of ETR in the plot (mean + sd),
+  # used for placing the asterisks
   mutate(etr_top = sum(etr_mean, etr_sd)) %>%
@@
   summarize(etr_top = max(etr_top), .groups = "drop")
+
+# Make helper table of significant t-test results: PAR
+par_t_test_data <-
+t_test_results %>%
+  filter(response == "par") %>%
+  mutate(star = case_when(
+    p_value < 0.001 ~ "***",
+    p_value < 0.01 ~ "**",
+    p_value < 0.05 ~ "*",
+    TRUE ~ "ns"
+    )) %>%
+  select(species, star) %>%
+  left_join(select(traits, species, generation, par_mean, par_sd), by = "species") %>%
+  rowwise %>%
+  mutate(par_top = sum(par_mean, par_sd)) %>%
+  group_by(species, star) %>%
+  summarize(par_top = max(par_top), .groups = "drop")
 
@@
   geom_errorbarh(
-    aes(x = etr_mean, xmax = etr_mean + etr_sd, xmin = etr_mean, fill = generation),
+    # need `fill = generation` for position to work correctly,
+    # even though it produces a warning 'Ignoring unknown aesthetics: fill '
+    aes(xmax = etr_mean + etr_sd, xmin = etr_mean, fill = generation),
     position = pos,
@@
     axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.title.x = element_markdown()
+  ) +
+  guides(fill = "none")
+
+d <- ggplot(traits, aes(y = species)) +
+  geom_col(
+    aes(x = par_mean, fill = generation),
+    position = pos,
+    width = 0.5) +
+  geom_errorbarh(
+    # need `fill = generation` for position to work correctly,
+    # even though it produces a warning 'Ignoring unknown aesthetics: fill '
+    aes(xmax = par_mean + par_sd, xmin = par_mean, fill = generation),
+    position = pos,
+    height = 0,
+    color = "grey20") +
+  geom_text(
+    data = par_t_test_data, 
+    aes(label = star, y = species, x = par_top),
+    angle = -90,
+    nudge_x = 0.01,
+    color = "grey20",
+    size = 2.5,
+    vjust = -0.5) +
+  scale_fill_manual(values = generation_colors, labels = pretty_text) +
+  scale_y_discrete(
+    position = "right",
+    labels = str_abbrev_sp_pretty
+    ) +
+  scale_x_continuous(expand = expansion(mult = c(0,0.025))) +
+  labs(x = "PPFD<sub>95%</sub>", fill = "Generation") +
+  theme(
+    panel.grid.minor = element_blank(), 
+    panel.grid.major.y = element_blank(),
+    axis.title.y = element_blank(),
     axis.text.y.right = element_markdown(), # NOT axis.text.y
@@
 # Combine plots
-a + b + c + plot_layout(nrow = 1, guides = "collect") &
+a + b + c + d + plot_layout(nrow = 1, guides = "collect") &
   theme(
@@
 
+# Verify that model fit is significant
+vpd_sporophyte_pval <- env_range_dt_model_summary %>% filter(model == "sporo_vpd_model") %>% pull(p_value) %>%
+  round_t(3)
+
+assertthat::assert_that(vpd_sporophyte_pval < 0.05)
+
+vpd_sporophyte_adj_r_squared <- env_range_dt_model_summary %>% filter(model == "sporo_vpd_model") %>% pull(adj_r_squared) %>%
+  round_t(3)
+
+# Extract fitted points from model
+vpd_sporophyte_fitted <-
+  # PGLS model has fitted points by species name, so join with VPD data
+  env_range_dt_model$sporo_vpd_model$fitted %>%
+  as.data.frame() %>%
+  rownames_to_column("species") %>%
+  as_tibble() %>%
+  rename(recovery_sporophyte_mean = V1) %>%
+  left_join(select(
+    plot_data, species, vpd_sporophyte_mean
+  ), by = "species")
+
 a <- ggplot(plot_data, aes(x = vpd_sporophyte_mean, y = recovery_sporophyte_mean, color = habit)) +
+  geom_line(data = vpd_sporophyte_fitted, color = "dark grey") +
   geom_linerange(aes(
@@
   geom_point() +
+  annotate(
+    geom = "richtext", 
+    x = Inf, y = -Inf, 
+    label = glue("*P* = {vpd_sporophyte_pval}<br>*R*<sup>2</sup> = {vpd_sporophyte_adj_r_squared}"),
+    hjust = 1, vjust = 0, size = 3, label.colour = "transparent") +
   scale_y_continuous(

### 2021-03-01 ms/si.Rmd
Update SI - remove sporo lab light curves

@@
 
-```{r sporo-lab-light-curves, fig.height = 9.5, fig.width = 7, fig.cap = ""}
+```{r light-curves, fig.height = 9.5, fig.width = 7, fig.cap = ""}
 ### Make full sets of light curve plots ###
@@
 
-### Print sporos in the field ###
+### Print the gametos ###
 # first with title "b"
-lc_plots_sporo_field +
+lc_plots_gameto +
   labs(title = "**b**") +
@@
 # with title "b (cont.)"
-for(i in 2:n_sporo_field_lc_pages) {
-  p <- lc_plots_sporo_field +
+for(i in 2:n_gameto_lc_pages) {
+  p <- lc_plots_gameto +
     labs(title = "**b** (cont.)") +
@@
 }
-
-### Print the gametos ###
-# first with title "c"
-lc_plots_gameto +
-  labs(title = "**c**") +
-  facet_wrap_paginate(
-    vars(str_abbrev_sp_pretty(species), light_id),
-    ncol = 4, nrow = 6, page = 1,
-    scales = "free", labeller = label_value_2)
-
-# print the rest
-# with title "c (cont.)"
-for(i in 2:n_gameto_lc_pages) {
-  p <- lc_plots_gameto +
-    labs(title = "**c** (cont.)") +
-    facet_wrap_paginate(
-      vars(str_abbrev_sp_pretty(species), light_id),
-      ncol = 4, nrow = 6, page = i,
-      scales = "free", labeller = label_value_2)
-  print(p)
-}
 ```
@@
 Rapid light response curves of filmy ferns from Moorea, French Polynesia.
-**a**, sporophytes measured in the lab; **b**, sporophytes measured in the field; **c** gametophytes measured in the field (no gametophytes were measured in the lab).
+**a**, sporophytes; **b**, gametophytes.
 Title of each subplot indicates species followed by light curve ID number assigned to each sample.

### 2021-03-01 ms/si.Rmd
Update SI - remove gameto DT controls from plot

@@
 # Make plot: sporophyte controls
-a <-
 dt_control_plot_data %>% 
@@
   )
-
-# Make plot: gametophyte controls
-b <-
-  dt_control_plot_data %>% 
-  filter(generation == "gametophyte") %>% 
-  ggplot(aes(x = rec_time, y = recovery, shape = dry_time, fill = salt)) +
-  geom_point(position = pd, size = 1.5, fill = "white") +
-  geom_errorbar(
-    aes(ymin = recovery - sd, ymax = recovery + sd), 
-    colour = "grey50", width = 0, 
-    position = pd) +
-  facet_wrap(vars(habit, species)) +
-  scale_shape_manual(name = "Desiccation Time",
-                     values = c(24, 21),
-                     breaks = c("15", "2"),
-                     labels = c("15 days", "2 days")) +
-  scale_y_continuous(label = label_percent(accuracy = 1)) +
-  scale_x_discrete(limits = c("30min","24hr","48hr"),
-                   labels = c("0.5", "24", "48")) +
-  xlab("Recovery Time (hr)") +
-  ylab("Recovery (%)") +
-  theme_bw() +
-  theme(
-    panel.grid.minor = element_blank(), 
-    panel.grid.major.x = element_blank(),
-    strip.text = element_markdown()
-  ) +
-  # Use guides to make the legend go away
-  # otherwise it will be duplicated by patchwork
-  # https://github.com/thomasp85/patchwork/issues/170#issuecomment-641310103
-  guides(shape = "none")
-
-a + b + plot_layout(guides = "collect", ncol = 1, heights = c(85, 15)) + plot_annotation(tag_levels = "a") &
-  theme(legend.position='bottom')
 ```
@@
 **`r s_figure("dt-control")`**.
-Recovery of controls (samples kept on moist tissues) for desiccation tolerance test of filmy fern sporophytes (**a**) and gametophytes (**b**) from Moorea, French Polynesia.
+Recovery of controls (samples kept on moist tissues) for desiccation tolerance test of filmy fern sporophytes from Moorea, French Polynesia.
 Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation treatment (see Methods).
 Recovery was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not recovery of 0%).
-Gametophytes only included a single individual per species thus lack error bars.
 

### 2021-03-01 ms/manuscript.Rmd
Update MS: Add rel water means

@@
 tar_load(filmy_dt_chamber)
+tar_load(rel_water_species_means)
 ```
@@
   knitr::combine_words()
+
+# Rel. water content
+rel_water_table <- 
+rel_water_species_means %>%
+  filter(!is.na(rel_water)) %>%
+  group_by(salt, dry_time, rec_time) %>% 
+  summarize(mean = mean(rel_water), sd = sd(rel_water), n = n(), .groups = "drop") %>%
+  arrange(mean) %>%
+  mutate(across(c(mean, sd), ~scales::percent(., suffix = " %", accuracy = 0.1))) %>%
+  mutate(sum_val = glue("{mean} ± {sd} (*n* = {n})"))
+
+rel_water_table %>% filter(salt == "LiCl", dry_time == "15") %>% pull(sum_val)
 ```
@@
 
-Control samples generally did not show any decrease in F~v~/F~m~ throughout the duration of the experiment (`r s_figure("dt-control")`). 
-Relative water content reached \circa{ }0 % in the desiccated state (`r s_figure("water-all")`), and returned to \circa{ }100 % upon re-wetting in all species sampled (`r s_figure("water-s")`).
+Control samples generally did not show any decrease in F~v~/F~m~ throughout the duration of the experiment (`r s_figure("dt-control")`) and maintained RWC near 100% (\figs{ }`r s_figure_num("water-all")`, `r s_figure_num("water-s")`).
+
+Relative water content in the desiccated state ranged from `r rel_water_table %>% filter(salt == "LiCl", dry_time == "15", rec_time == "desiccated") %>% pull(sum_val)` (--282 MPa for 15 d) to `r rel_water_table %>% filter(salt == "NaCl", dry_time == "2", rec_time == "desiccated") %>% pull(sum_val)` (--38 MPa for 2 d), and returned to \circa{ }100 % upon re-wetting in all species sampled (`r s_figure("water-s")`).
 
@@
 
-Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (\figs{ }`r s_figure_num("chamber")`).
+Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (`r s_figure("chamber")`).
 Brief fluctuations to higher humidities were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).

### 2021-03-01 ms/si.Rmd
Update MS: Add rel water means

@@
 
-```{r sporophyte-dt-MgNO3, fig.height = 6.5, fig.width = 9, fig.cap = ""}
+```{r sporophyte-dt-MgNO3, fig.height = 6.6, fig.width = 9, fig.cap = ""}
 # Get mean times of each cluster of DT test samples for the selected drying salt
@@
 
-```{r sporophyte-dt-NaCl, fig.height = 6.5, fig.width = 9, fig.cap = ""}
+```{r sporophyte-dt-NaCl, fig.height = 6.6, fig.width = 9, fig.cap = ""}
 # Subset desiccation duration data

### 2021-03-09 R/clean_lc_data.R
Save raw light curve plots to results/

@@ -289,7 +289,7 @@ filmy_lc_model_params_filtered <-
   filter(p.value < 0.05)  %>%
   abbrev_sp()
 
-# - Make plots
+# - Optional: make plots to visualize raw light curves
 filmy_all_lc_plot_filtered <-
   ggplot(abbrev_sp(filmy_lc_model_fitted_data_filtered), aes(x = par)) +
   geom_point(aes(y = etr)) +
@@ -298,7 +298,7 @@ filmy_all_lc_plot_filtered <-
   geom_vline(data = filmy_lc_model_params_filtered, aes(xintercept = par_critical), color = "red") +
   facet_wrap(vars(generation, species, id), scales = "free")
 
-ggsave(plot = filmy_all_lc_plot_filtered, file = "data_raw/intermediates/all_light_curves_filtered.pdf", height = 40, width = 40)
+ggsave(plot = filmy_all_lc_plot_filtered, file = "results/all_light_curves_filtered.pdf", height = 40, width = 40)
 
 # INTERACTIVE STEP:
 # To run, uncomment the lines below

### 2021-03-09 R/process_refs.R
Move location of main_library.bib link

@@ -22,7 +22,7 @@ citations <-
 
 # Read in entire reference library exported from Zotero with BetterBibTex
 # (note: this can be a soft-link to ~/Dropbox/bibliography/main_library.bib)
-bib <- RefManageR::ReadBib("data_raw/main_library.bib", check = FALSE)
+bib <- RefManageR::ReadBib("main_library.bib", check = FALSE)
 
 # Convert reference library to dataframe
 bib_df <- as.data.frame(bib) %>% rownames_to_column("cite_key")

### 2021-03-09 ms/manuscript.Rmd
Update MS
- add links for reviewers to data and code

@@
 All experiments were initiated within 48 h of collection.
-Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
+Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany).
 Any samples with pre-treatment F~v~/F~m~ < 0.4 were excluded from analysis, as typical values of F~v~/F~m~ in healthy plants tend to be \circa{ }0.8 [@Bjorkman1987].
@@
 
-Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400 nm to 700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
+Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400 nm to 700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer as described by the manufacturer.
 Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
@@
 
-Datasets used in this study are available on FigShare (\#\#\#\#\#\#). 
+Datasets used in this study are available on FigShare (https://doi.org/10.6084/m9.figshare.14184572). 
 Code used to analyze the data and generate figures and this manuscript are available on GitHub (https://github.com/joelnitta/moorea-filmies).
@@
 The \package{ggplot2} [@Wickham2016], \package{ggtree}, [@Yu2017] and \package{patchwork} (https://github.com/thomasp85/patchwork) packages were used to generate the figures.
+
+\*\*\* NOTE TO REVIEWERS: The dataset and code will not be made public until the paper is accepted.
+While the paper is under review, these links can be used for the dataset (https://figshare.com/s/d6349abf01a3756a5aae) and code (https://www.dropbox.com/sh/h788g1h3sjcqj0j/AACwFWTmL6GxrbLkCfOrXZU9a?dl=0)  \*\*\*
 

### 2021-03-10 ms/manuscript.Rmd
Fix order of facet labels in DT recovery plot

@@
   ) %>%
+  # Add a line break only for C. bipunctatum
+  mutate(label = str_replace_all(
+    label, 
+    fixed("*C. bipunctatum* (LE, *n* = 10-12)"),
+    "*C. bipunctatum*<br>(LE, *n* = 10-12)")) %>%
   # - order by alphabetical species within growth habit
@@
 ```{r filmy-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = "", purl = FALSE}
-dt_plot_data %>%
-  mutate(label = str_replace_all(
-    label, 
-    fixed("*C. bipunctatum* (LE, *n* = 10-12)"),
-    "*C. bipunctatum*<br>(LE, *n* = 10-12)")) %>%
-make_dt_plot("sporophyte", "a")
+make_dt_plot(dt_plot_data, "sporophyte", "a")
 ggsave(here::here("results/ms/fig_1a_sporo_dt.pdf"), width = 6.85, height = 7.5)

### 2021-03-10 ms/manuscript.Rmd
Update MS
- add framing as suggest by CCD to intro

@@
 
-Joel H. Nitta^1,4^, James E. Watkins, Jr.^2^, N. Michele Holbrook^1^, Ravahere Taputuarai^3^, Tristan W. Wang^1^, and Charles C. Davis^1^
+Joel H. Nitta^1,4^, James E. Watkins, Jr.^2^, N. Michele Holbrook^1^, Tristan W. Wang^3^, and Charles C. Davis^1^
 
@@
 
-^3^Association Te Rau Ati Ati a Tau a Hiti Noa Tu, Mahina, Tahiti, French Polynesia
+^3^Section of Plant Pathology and Plant-Microbe Biology, Cornell University, Ithaca, New York 14853 USA
 
@@
 N. Michele Holbrook: \textnormal{https://orcid.org/0000-0003-3325-5395}  
+Tristan W. Wang: \textnormal{https://orcid.org/0000-0001-6889-2453}  
 Charles C. Davis: \textnormal{https://orcid.org/0000-0001-8747-1101}
@@
 
-Desiccation tolerance was a key trait allowing plants to colonize land.
-However, little is known about the transition from desiccation tolerant non-vascular plants to desiccation sensitive vascular ones.
-Filmy ferns (Hymenophyllaceae) present a useful system to study how water-stress strategies differ between non-vascular and vascular stages within a single organism, as they have vascularized sporophytes and nonvascular gametophytes that are each capable of varying degrees of desiccation tolerance.
-We surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia) and used chlorophyll fluorescence to measure desiccation tolerance and light responses.
+Desiccation tolerance was a key trait that allowed plants to colonize land.
+However, little is known about the transition from desiccation tolerant non-vascular plants to desiccation avoiding vascular ones.
+Filmy ferns (Hymenophyllaceae) represent a useful system to investigate how water-stress strategies differ between non-vascular and vascular stages within a single organism because they have vascularized sporophytes and nonvascular gametophytes that are each capable of varying degrees of desiccation tolerance.
+To explore this, we surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia) and used chlorophyll fluorescence to measure desiccation tolerance and light responses.
 We conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.
@@
 Despite their lack of greater physiological tolerances, gametophytes of several species occurred over a wider elevational range than conspecific sporophytes.
-Our results show that filmy fern gametophytes and sporophytes differ in their physiology and niche requirements, and point to the importance of microhabitat in shaping the evolution of water-use strategies in vascular plants.
+Our results demonstrate that filmy fern gametophytes and sporophytes differ in their physiology and niche requirements, and point to the importance of microhabitat in shaping the evolution of water-use strategies in vascular plants.
 
@@
 The movement of plants from water to land was one of the most important evolutionary events in the history of the earth, fundamentally altering the global carbon and water cycle and setting the stage for modern terrestrial ecosystems [@Graham1993; @Kenrick1997a].
-The transition from aquatic to terrestrial growth was complex, requiring the evolution of a suite of adaptations to survive in dry habitats bombarded with radiation [@Waters2003].
+The transition from aquatic to terrestrial growth was complex, requiring a suite of evolutionary adaptations to survive in dry habitats bombarded with radiation [@Waters2003].
 The earliest land plants were poikilohydric, \ie, they lacked any mechanisms to regulate water relations [@Raven1999].
@@
 Many bryophytes have retained vegetative DT in the gametophyte stage [@Oliver2005, and references therein].
-In extant vascular plants (tracheophytes), DT occurs in certain non-vegetative stages (\eg, seeds and pollen), but has largely been lost in vegetative tissues (with some notable exceptions, the so-called 'resurrection plants').
+In extant vascular plants (tracheophytes), DT occurs in certain non-vegetative stages (\eg, seeds and pollen), but has largely been lost in vegetative tissues [with some notable exceptions, the so-called 'resurrection plants'; @Hartung1998; @Kappen2007; @Scott2000].
 Instead, tracheophytes avoid desiccation with adaptations including a waxy cuticle, stomata, and an extensive vascular system.
 Correlated with these differences in functional strategies to cope with water stress is a major difference in life cycle: in most tracheophytes (including all seed plants) the long-lived, dominant phase of the life cycle is the sporophyte (diploid phase), whereas in bryophytes it is the gametophyte (haploid phase).
-Little is known about the details of the transition from desiccation tolerance to sensitivity, nor how this took place in conjunction with the switch from a gametophyte to a sporophyte-dominated life cycle; yet, these two steps underpin the radiation of plants onto dry land.
+Little is known about the details of the transition from desiccation tolerance to avoidance, nor how this took place in conjunction with the switch from a gametophyte to a sporophyte-dominated life cycle; yet, these two steps underpin the radiation of plants onto dry land [@Puttick2018; @Qiu2006; @Wickett2014].
 
@@
 These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (\ie, they regulate their internal water content), whereas fern gametophytes are much smaller (often < 1 cm), lack vascular tissue, and are poikilohydric.
-Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013].
+Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@Pittermann2013; @WatkinsJr.2007c].
 Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.
@@
 As would be expected from their poikilohydric nature, DT of varying degrees has been reported in sporophytes of many filmy fern species (see references below).
-
-Filmy fern sporophytes encompass a broad diversity of growth forms (terrestrial, saxicolous, hemi-epiphytic, and holoepiphytic) and morphologies, with rhizome and leaf sizes varying over an order of magnitude [@Hennequin2008; @Nitta2009; @Dubuisson2013].
+The sporophytes of most filmy ferns are relatively small, with many less than \circa{ }10 cm.
+Thus, the morphological and physiological gap between sporophytes and gametophytes is arguably smaller in filmy ferns compared to any other fern clade.
+This makes filmy ferns a potentially useful group for understanding the transition from a desiccation tolerant, gametophyte-dominated strategy to a desiccation avoiding, sporophyte-dominated one, since the sporophytes and gametophytes of ancestral tracheophytes were likely initially similar, then gradually diverged by a series of step-wise adaptations [@Bateman1998; @Ligrone2012a].
+
+Filmy fern sporophytes encompass a broad diversity of growth forms (terrestrial, saxicolous, hemi-epiphytic, and holoepiphytic) and morphologies, with rhizome and leaf sizes varying over an order of magnitude [@Dubuisson2013; @Hennequin2008; @Nitta2009].
 Filmy fern gametophytes are similarly diverse, with morphotypes including filamentous, ribbon, and a combination of these two types (`r figure("habit")`e--h).
-While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages [@WatkinsJr.2007a; @WatkinsJr.2012; @Pittermann2013], and may be expected to occur in filmy ferns as well.
+While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages [@Pittermann2013; @WatkinsJr.2007a; @WatkinsJr.2012], and may be expected to occur in filmy ferns as well.
 The possibility for varying degrees of vegetative DT in both stages of the life cycle of filmy ferns makes them a potentially useful system to compare how DT is utilized between haploid, non-vascular and diploid, vascular life stages.
@@
 Furthermore, to our knowledge, no physiological studies of filmy ferns have accounted for their phylogenetic relationships, and we lack an understanding of how physiological strategies have evolved in this clade.
-In addition to DT, photosynthetic optima (adaptation to different light levels) are likely to be important in determining the niches of life stages and species of filmy ferns [@Proctor2003; @Parra2015; @Flores-Bavestrello2016].
+In addition to DT, photosynthetic optima (adaptation to different light levels) are likely to be important in determining the niches of life stages and species of filmy ferns [@Flores-Bavestrello2016; @Parra2015; @Proctor2003].
 Terrestrial and epiphytic species experience different levels of photosynthetic radiation and desiccation, and light intensity varies with height along the host plant in epiphytes [@Benzing1990; @WatkinsJr.2009].
@@
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
-A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (\figs `r figure_num("habit")`, `r s_figure_num("tree")`; all supplemental figures available in Online Resource 1).
+A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Florence; @Murdock2003; @Nitta2011b; @Nitta2017; @Ranker2005] (\figs `r figure_num("habit")`, `r s_figure_num("tree")`; all supplemental figures available in Online Resource 1).
 The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
@@
 We derived the phylogenetic tree used in this study by extracting Hymenophyllaceae from the tree of @Nitta2017 with the \function{extract.clade} function of the \package{ape} package [@Paradis2004].
-This tree is in accord with previously published, more densely sampled phylogenies [@Pryer2001a; @Dubuisson2003; @Hennequin2006; @Ebihara2007a].
-
-We quantified the degree of phylogenetic signal in physiological traits using Blomberg's *K* [@Blomberg2003] and Pagel's λ [@Pagel1999; @Freckleton2002] as implemented with the \function{phylosig} function in the \package{phytools} package [@Revell2012].
+This tree is in accord with previously published, more densely sampled phylogenies [@Dubuisson2003; @Ebihara2007a; @Hennequin2006; @Pryer2001a].
+
+We quantified the degree of phylogenetic signal in physiological traits using Blomberg's *K* [@Blomberg2003] and Pagel's λ [@Freckleton2002; @Pagel1999] as implemented with the \function{phylosig} function in the \package{phytools} package [@Revell2012].
 Both measures test the hypothesis that the trait of interest is evolving according to Brownian Motion (BM).
@@
 
-Vegetative DT was a key innovation linked to the rise of early land plants, but is largely limited to non-vascular groups such as bryophytes and algae in modern plants.
-Filmy ferns have received considerable attention over the years for their remarkable reversion to vegetative DT in the sporophyte phase [\eg, @Shreve1911; @Proctor2003; @GarcesCea2014].
+Vegetative DT was a key innovation linked to the rise of early land plants, but is largely limited to non-vascular groups such as bryophytes and algae in extant plants.
+Filmy ferns have received considerable attention over the years for their remarkable reversion to vegetative DT in the sporophyte phase [\eg, @GarcesCea2014; @Proctor2003; @Shreve1911].
 Yet, no previous study to our knowledge has evaluated parallel physiology of sporophytes and gametophytes of filmy ferns in relation to DT.
@@
 We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below -200 MPa (VPD > 2.45 kPa) in some epiphytes.
-Levels of similarly extreme DT have also been reported in other species of *Hymenophyllum* [@Proctor2003; @Proctor2012] and a general correlation of habitat with DT in filmy ferns has been previously established [@Shreve1911; @Proctor2003; @Proctor2012; @Parra2009; @Saldana2013], but no other studies to our knowledge have documented a similar range of DT values in filmy ferns from a single site.
+Levels of similarly extreme DT have also been reported in other species of *Hymenophyllum* [@Proctor2003; @Proctor2012] and a general correlation of habitat with DT in filmy ferns has been previously established [@Parra2009; @Proctor2003; @Proctor2012; @Saldana2013; @Shreve1911], but no other studies to our knowledge have documented a similar range of DT values in filmy ferns from a single site.
 One may wonder then, what is the adaptive significance of being able to withstand drying at -282 MPa (\circa{ }2.45 kPa) vs.
@@
 This may be an adaptation to cloud forest conditions; although cloud forests are frequently moist due to constant fog, dry periods are intense during the times when there is no cloud cover.
-The complete lack of DT in terrestrial species is not surprising, considering they have much more direct access to water than epiphytes and therefore substantially less selective pressure to maintain DT.
+The complete lack of DT in terrestrial species is not surprising, considering they have much more direct access to water than epiphytes and therefore are under substantially less selective pressure to maintain DT.
 Since controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (`r s_figure("dt-control")`), it is unlikely that the lack of tolerance seen in terrestrial species is due solely to removal from the soil.
@@
 Contrary to our hypotheses, gametophytes of filmy ferns are in general not more tolerant of the abiotic environment (light levels and water potential) than sporophytes.
-This differs from other fern groups with desiccation tolerant gametophytes but desiccation sensitive sporophytes [@WatkinsJr.2007c].
+This differs from other fern groups with desiccation tolerant gametophytes but desiccation avoiding sporophytes [@WatkinsJr.2007c].
 In the filmy ferns we studied, DT levels of both sporophytes and gametophytes largely reflected habitat; in general, epiphytic species tended to have higher DT than terrestrial or saxicolous species, regardless of life stage (\figs `r figure_num("combined")`, `r s_figure_num("traits-box")`).
@@
 
-Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates [@Proctor2003; @Proctor2012; @Parra2009; @Saldana2013]; however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
+Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates [@Parra2009; @Proctor2003; @Proctor2012; @Saldana2013]; however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
 We detected significant phylogenetic signal in DT with Blomberg's *K* (`r table("physig")`), suggesting that phylogeny should be considered when interpreting differences in DT between species.
@@
 Members of the Davis Lab and Jonathan B. Losos provided helpful discussion and comments on drafts.
-Saad Amer, Tohei Theophilus, and Suzanne Vinnette assisted with fieldwork and experiments.
-Evan Zeng helped with analysis of microclimatic data.
+Saad Amer, Ravahere Taputuarai, Tohei Theophilus, and Suzanne Vinnette assisted with fieldwork and experiments.
 Weston Testo provided advice on settings for the DT test.

### 2021-03-11 ms/si.Rmd
Fix DT chamber data filtering for 2013 gametophytes

@@
 # Subset DT chamber data
-chamber_gameto_2013 <- filmy_dt_chamber %>% filter(generation == "gametophyte", year == "2013")
+# the datalogger in use changed part way through.
+# 10342564 was used in a combined chamber including both sporos and gametos
+chamber_gameto_2013_part_1 <- filmy_dt_chamber %>% filter(serial_no == "10342564", year == "2013")
+chamber_gameto_2013_part_2 <- filmy_dt_chamber %>% filter(serial_no == "10140659", year == "2013", generation == "gametophyte")
+
+# chamber_gameto_2013 <- filmy_dt_chamber %>% filter(generation == "gametophyte", year == "2013")
+chamber_gameto_2013 <- bind_rows(chamber_gameto_2013_part_1, chamber_gameto_2013_part_2)
 chamber_gameto_2012 <- filmy_dt_chamber %>% 

### 2021-03-11 _targets.R
Add pre-print version of PDF
- uses params to set table style

@@ -202,9 +202,25 @@ tar_plan(
   tar_file(refs_other, "ms/references_other.yaml"),
 
   # Render PDF
-  tar_render(manuscript_pdf, "ms/manuscript.Rmd", output_dir = here::here("results/ms")),
+  # - PDF for preprint
+  tar_render(
+    preprint_pdf, 
+    "ms/manuscript.Rmd", 
+    output_dir = here::here("results/ms"),
+    output_file = "moorea_filmies_preprint.pdf",
+    params = list(output_type = "preprint")),
+  
+  # - PDF for conversion to Word
+  tar_render(
+    manuscript_pdf, 
+    "ms/manuscript.Rmd", 
+    output_dir = here::here("results/ms")
+    ),
   
   # Render MS Word
+  # FIXME: 
+  # - change Florence to "in press" (in-line and refs)
+  # - remove extra space from bib entries
   ms_docx = latex2docx(
     latex = "results/ms/manuscript.tex",
     docx = "results/ms/manuscript.docx",

### 2021-03-11 ms/manuscript.Rmd
Update MS
- reframe intro
- use params for formatting table based on output type
- fix descriptions of results to match figures

@@
   \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
+params:
+  output_type: ms_for_word
 editor_options: 
@@
 **Subject area**  
-(1) Taxonomy/Phylogenetics/Evolutionary Biology
+(1) Ecology/Ecophysiology/Environmental Biology
 
@@
 
-Joel H. Nitta^1,4^, James E. Watkins, Jr.^2^, N. Michele Holbrook^1^, Tristan W. Wang^3^, and Charles C. Davis^1^
+Joel H. Nitta^1,4^, James E. Watkins, Jr.^2^, N. Michele Holbrook^1^, Tristan W. Wang^1,3^, and Charles C. Davis^1^
 
@@
 
-^3^Section of Plant Pathology and Plant-Microbe Biology, Cornell University, Ithaca, New York 14853 USA
+^3^Current address: Section of Plant Pathology and Plant-Microbe Biology, Cornell University, Ithaca, New York 14853 USA
 
@@
 
+```{r study-system-stats}
+# Calculate number of all filmy fern species, trichomanoids, and hymenophylloids
+# (not considering numbered taxa, ie, varieties)
+n_filmy_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
+  n_distinct()
+
+n_trich_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
+  magrittr::extract(., str_detect(., "^Hymeno", negate = TRUE)) %>%
+  n_distinct()
+
+n_hymeno_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
+  magrittr::extract(., str_detect(., "^Hymeno", negate = FALSE)) %>%
+  n_distinct()
+```
+
 Desiccation tolerance was a key trait that allowed plants to colonize land.
-However, little is known about the transition from desiccation tolerant non-vascular plants to desiccation avoiding vascular ones.
+However, little is known about the transition from desiccation tolerant non-vascular plants to desiccation sensitive vascular ones.
 Filmy ferns (Hymenophyllaceae) represent a useful system to investigate how water-stress strategies differ between non-vascular and vascular stages within a single organism because they have vascularized sporophytes and nonvascular gametophytes that are each capable of varying degrees of desiccation tolerance.
-To explore this, we surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia) and used chlorophyll fluorescence to measure desiccation tolerance and light responses.
+To explore this, we surveyed sporophytes and gametophytes of `r n_filmy_sp` species (`r n_distinct(filmy_species)` taxa including varieties) of filmy ferns on Moorea (French Polynesia) and used chlorophyll fluorescence to measure desiccation tolerance and light responses.
 We conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.
-Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima than sporophytes.
-Epiphytes were more tolerant of desiccation than terrestrial species.
+Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima (maximum electron transport rate of photosystem II and light level at 95% of that rate) than sporophytes.
+Epiphytes were more tolerant of desiccation than terrestrial species in both generations.
 Despite their lack of greater physiological tolerances, gametophytes of several species occurred over a wider elevational range than conspecific sporophytes.
@@
 The movement of plants from water to land was one of the most important evolutionary events in the history of the earth, fundamentally altering the global carbon and water cycle and setting the stage for modern terrestrial ecosystems [@Graham1993; @Kenrick1997a].
-The transition from aquatic to terrestrial growth was complex, requiring a suite of evolutionary adaptations to survive in dry habitats bombarded with radiation [@Waters2003].
-The earliest land plants were poikilohydric, \ie, they lacked any mechanisms to regulate water relations [@Raven1999].
-Therefore, one particularly important early adaptation that enabled plants to survive on land was desiccation tolerance (DT), the ability to lose water until reaching equilibrium with atmospheric humidity, then recover upon rewetting [@Mishler1985].
+The transition from aquatic to terrestrial growth was complex, requiring a suite of evolutionary adaptations to survive in deadly dry habitats bombarded with radiation [@Waters2003].
+Lacking mechanisms to regulate water loss, the earliest land plants were likely poikilohydric, \ie, they existed in a state of equilibrium with the humidity of the external environment [@Raven1999].
+Therefore, one particularly important early adaptation that enabled plants to survive on land was desiccation tolerance (DT), the ability to lose all metabolically active water, reach equilibrium with atmospheric humidity, and recover upon rewetting [@Mishler1985].
 
@@
 In extant vascular plants (tracheophytes), DT occurs in certain non-vegetative stages (\eg, seeds and pollen), but has largely been lost in vegetative tissues [with some notable exceptions, the so-called 'resurrection plants'; @Hartung1998; @Kappen2007; @Scott2000].
-Instead, tracheophytes avoid desiccation with adaptations including a waxy cuticle, stomata, and an extensive vascular system.
-Correlated with these differences in functional strategies to cope with water stress is a major difference in life cycle: in most tracheophytes (including all seed plants) the long-lived, dominant phase of the life cycle is the sporophyte (diploid phase), whereas in bryophytes it is the gametophyte (haploid phase).
-Little is known about the details of the transition from desiccation tolerance to avoidance, nor how this took place in conjunction with the switch from a gametophyte to a sporophyte-dominated life cycle; yet, these two steps underpin the radiation of plants onto dry land [@Puttick2018; @Qiu2006; @Wickett2014].
-
-Ferns represent an important, understudied group of tracheophytes for investigating the transition from desiccation tolerance to desiccation avoidance in land plants.
+Instead, the majority of tracheophytes solely rely on avoiding desiccation with adaptations including a waxy cuticle, stomata, roots, and an extensive vascular system.
+Correlated with these differences in functional strategies to cope with water stress is a major difference in life cycle.
+Over time, plants shifted away from wholly gametophyte-dominated life cycle (as in extant bryophytes) to co-independence of the gametophyte and sporophyte (\ie, ferns and lycophytes) to complete loss of the independent gametophyte (\ie, spermatophytes).
+Little is known about the details of the transition from desiccation tolerance to avoidance, nor how this took place in conjunction with the switch from a gametophyte to a sporophyte-dominated life cycle; yet, these two steps underpin the radiation of vascular plants onto dry land [@Puttick2018; @Qiu2006; @Wickett2014].
+
+Ferns represent an important, understudied group of tracheophytes for investigating both the ecology and evolution of desiccation tolerance.
 Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
@@
 Filmy ferns are a large family [\circa{ }430 spp.; @PteridophytePhylogenyGroupI2016] of primarily tropical, leptosporangiate ferns named for their remarkably thin leaf laminae, which are usually only a single cell layer in thickness and lack cuticle or stomata (`r figure("habit")`a--d).
-Filmy fern sporophytes are an exception to the 'typical' fern sporophyte described above because they have reverted to poikilohydry: their thin laminae allow for passive, rapid uptake and loss of moisture, thereby maintaining nearly constant equilibrium with atmospheric humidity [@Raven1999].
+Filmy fern sporophytes are a remarkable exception to the 'typical' homoiohydric fern sporophyte described above because they have reverted to poikilohydry: their thin laminae allow for passive, rapid uptake and loss of moisture, thereby maintaining nearly constant equilibrium with atmospheric humidity [@Raven1999].
 Many species of epiphytic filmy ferns have extremely reduced root systems [@Schneider2000], relying completely upon absorption through leaf laminae for water [@Shreve1911], although terrestrial species possess true roots and use ground water [@Dubuisson2011].
+Thus, the morphological and physiological gap between sporophytes and gametophytes is arguably smaller in filmy ferns compared to any other fern clade.
+This makes filmy ferns a potentially useful group for understanding the transition from a desiccation tolerant, gametophyte-dominated strategy to a desiccation avoiding, sporophyte-dominated one, since the sporophytes and gametophytes of ancestral tracheophytes were also likely initially similar, then gradually diverged by a series of step-wise adaptations [@Bateman1998; @Ligrone2012a].
+
+Filmy fern sporophytes encompass a broad diversity of growth forms (terrestrial, saxicolous, hemi-epiphytic, and holoepiphytic) and morphologies, with rhizome and leaf sizes varying over an order of magnitude [@Dubuisson2013; @Hennequin2008; @Nitta2009].
 As would be expected from their poikilohydric nature, DT of varying degrees has been reported in sporophytes of many filmy fern species (see references below).
-The sporophytes of most filmy ferns are relatively small, with many less than \circa{ }10 cm.
-Thus, the morphological and physiological gap between sporophytes and gametophytes is arguably smaller in filmy ferns compared to any other fern clade.
-This makes filmy ferns a potentially useful group for understanding the transition from a desiccation tolerant, gametophyte-dominated strategy to a desiccation avoiding, sporophyte-dominated one, since the sporophytes and gametophytes of ancestral tracheophytes were likely initially similar, then gradually diverged by a series of step-wise adaptations [@Bateman1998; @Ligrone2012a].
-
-Filmy fern sporophytes encompass a broad diversity of growth forms (terrestrial, saxicolous, hemi-epiphytic, and holoepiphytic) and morphologies, with rhizome and leaf sizes varying over an order of magnitude [@Dubuisson2013; @Hennequin2008; @Nitta2009].
 Filmy fern gametophytes are similarly diverse, with morphotypes including filamentous, ribbon, and a combination of these two types (`r figure("habit")`e--h).
@@
 They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
-A recent study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly [@GarcesCea2014].
+A study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a homoiochlorophyllous, constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly [@GarcesCea2014].
+This was further supported by evidence that induced biomechanical mechanisms likely do not play a role in DT of filmy ferns [@Fallard2018; but see @Mkhize2020].
+Recently, transcriptomics has been applied to elucidate the genes involved at each step of the desiccation-recovery cycle of filmy ferns [@Ostria-Gallardo2020; @Ostria-Gallardo2020a]. 
+In epiphytic *Hymenoglossum cruentum* Cav.
+from Chile, key genes and gene networks were identified related to intracellular mobility and photosynthetic metabolism in the desiccated state, and those related to detoxifying pathways and stabilization of photosystems during rehydration [@Ostria-Gallardo2020a].
+In a comparison of low-trunk and canopy epiphytes, @Ostria-Gallardo2020 identified multiple differentially expressed genes that may be specific to tolerating desiccation in these microhabitats.
 
@@
 Terrestrial and epiphytic species experience different levels of photosynthetic radiation and desiccation, and light intensity varies with height along the host plant in epiphytes [@Benzing1990; @WatkinsJr.2009].
-In a recent study (Testo et al., unpublished), the combined effect of desiccation and light stress was found to have a greater impact on filmy fern physiological performance than either of these stressors alone, suggesting the need to integrate studies of DT with photosynthesis to gain a full picture of physiological adaptation in filmy ferns.
+The combined effect of desiccation and light stress may induce greater oxidative damage in filmy ferns than either of these stressors alone [@Niinemets2018], suggesting the need to integrate studies of DT with photosynthesis to gain a full picture of physiological adaptation in filmy ferns.
 Here, we leverage a recently developed DNA barcoding system [@Nitta2017] to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes:
@@
 3.) DT increases with decreasing environmental water availability.
-4.) Gametophytes with wider ranges have higher DT than those with with narrower ranges.
+4.) Gametophytes with broader elevational ranges have higher DT than those with with narrower ranges.
 
@@
 ### Study system
-
-```{r study-system-stats}
-# Calculate number of all filmy fern species, trichomanoids, and hymenophylloids
-# (not considering numbered taxa)
-
-n_filmy_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
-  n_distinct()
-
-n_trich_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
-  magrittr::extract(., str_detect(., "^Hymeno", negate = TRUE)) %>%
-  n_distinct()
-
-n_hymeno_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
-  magrittr::extract(., str_detect(., "^Hymeno", negate = FALSE)) %>%
-  n_distinct()
-
-```
 
@@
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
-A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Florence; @Murdock2003; @Nitta2011b; @Nitta2017; @Ranker2005] (\figs `r figure_num("habit")`, `r s_figure_num("tree")`; all supplemental figures available in Online Resource 1).
+A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Florence; @Murdock2003; @Nitta2011b; @Nitta2017; @Ranker2005] (\figs{ }`r figure_num("habit")`, `r s_figure_num("tree")`; all supplemental figures available in Online Resource 1).
 The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
@@
 Unlike sporophytes, fern gametophytes are cryptic and cannot be identified to species based on morphology alone. 
-Therefore, we used DNA barcoding to identify filmy fern gametophytes to species as described in @Nitta2017.
+Therefore, we used DNA barcoding to identify filmy fern gametophytes to species following @Nitta2017.
 Briefly, DNA extraction was performed using a modified CTAB protocol [@Doyle1987] or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
@@
 Gametophytes matching > 99.5 % with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched > 99.5 % with multiple reference sequences were not identified and excluded from further analysis.
-For this study, only gametophytes that could be identified as belonging to Hymenophyllaceae were used, and others were excluded from analysis.
+For this study, only gametophytes that could be identified as Hymenophyllaceae were used; others were excluded from analysis.
 
@@
 We used microclimate data collected by @Nitta2020.
-This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) and vapor pressure deficit (VPD) calculated from these two values.
+This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the fern survey sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA), and vapor pressure deficit (VPD) calculated from these two values.
 The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at \circa{ }1.5 m on a tree, one 'terrestrial' datalogger at \circa{ }10 cm next to the tree).
@@
 Mean daily temperature on Moorea is strongly correlated with elevation, dropping \circa{ }6 °C from the lowest surveyed site (201 m) to the highest (1206 m); temperature does not differ significantly between epiphytic and terrestrial habitats [@Nitta2020].
-However, epiphytic habitats tend to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones.
+Epiphytic habitats tend to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones.
 This difference is greatest at low elevations, and smallest at high elevations [@Nitta2020].
@@
 Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
-For sporophytes, 8 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
+For sporophytes, eight individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
 We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
@@
 Samples were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
-Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012 field season for species other than *Callistopteris apiifolia*) or Hobo ProV2 dataloggers logging every 5 min (2012 *Callistopteris apiifolia*, 2013, 2014 field seasons).
+Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012 field season for species other than *Callistopteris apiifolia* (C.
+Presl) Copel.) or Hobo ProV2 dataloggers logging every 5 min (2012 *C.
+apiifolia*, 2013, 2014 field seasons).
 The desiccation chambers were kept in an air-conditioned room. 
-Temperature inside the chambers was mean `mean(filmy_dt_chamber$temp) %>% round(1)` ± SD `r sd(filmy_dt_chamber$temp) %>% round(2)` °C (\figs `r s_figure_num("chamber")`).
+Temperature inside the chambers was mean `r mean(filmy_dt_chamber$temp) %>% round(1)` ± SD `r sd(filmy_dt_chamber$temp) %>% round(2)` °C (`r s_figure("chamber")`) (hereafter, the value following '±' is standard deviation unless otherwise mentioned).
 Salts used for desiccation and their corresponding mean water potentials and approximate relative humidity and VPD are as follows: LiCl (-282 MPa, 18 % RH, 2.45 kPa), Mg(NO~3~)~2~ (-86 MPa, 58 % RH, 1.25 kPa), and NaCl (-38 MPa, 80 % RH, 0.60 kPa).
-Following the desiccation treatment, plants were rewetted by wrapping them in wet tissues using water from a local stream.
+Following the desiccation treatment, plants were rewetted by wrapping them in moist tissues using water from a local stream.
 Plants were allowed to rehydrate, and F~v~/F~m~ was again measured at 0.5 h, 24 h, and 48 h following rewetting.
@@
 
-We measured relative water content (RWC) in a subset of samples by recording the mass of samples prior to the DT test (fresh mass), at each step of the DT test (turgid mass), then after drying them overnight in a drying oven at 65 °C following the DT test (dry mass).
+We measured relative water content (RWC) in a subset of samples by recording the mass of samples prior to the DT test (fresh mass), at each step of the DT test (turgid mass), then following the DT test after drying them overnight in an oven at 65 °C (dry mass).
 Relative water content was calculated as \(RWC = \frac{TM - DM}{FM - DM} \times 100\), where *TM* is turgid mass, *DM* is dry mass, and *FM* is fresh mass.
@@
 We therefore applied the same treatment for all gametophytes (2 d at -86 MPa) in batches of individuals (`r s_table("gameto-batches")`), and verified the species of each individual after the experiment using the DNA barcode approach described above.
-Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
+Thus, gametophytes of some species were sampled many times whereas those of other species were not.
 Relative water content was not calculated for gametophytes, as these were too small to measure accurately with the balances available.
@@
 
-All statistical analyses were done in R v.4.0.2 [@RCoreTeam2020a].
+All statistical analyses were done in R v.`r getRversion()` [@RCoreTeam2020a].
 
 We derived the phylogenetic tree used in this study by extracting Hymenophyllaceae from the tree of @Nitta2017 with the \function{extract.clade} function of the \package{ape} package [@Paradis2004].
-This tree is in accord with previously published, more densely sampled phylogenies [@Dubuisson2003; @Ebihara2007a; @Hennequin2006; @Pryer2001a].
+This tree is in agreement with previously published, more densely sampled phylogenies [@Dubuisson2003; @Ebihara2007a; @Hennequin2006; @Pryer2001a].
 
@@
 λ is a scaling parameter that ranges from zero to one: λ near zero indicates random distribution of trait values across the tree, and λ near one indicates evolution of traits along the phylogeny following BM.
-For *K*, values near one indicate evolution of traits following BM; *K* >1 indicates traits more conserved than expected under BM, and *K* <1 indicates that traits have less phylogenetic signal than expected under BM.
-We tested the significance of *K* by comparing the observed value against values from a null distribution of \num{1000} phylogenies with the traits randomly shuffled across the tips.
+For *K*, values near one indicate evolution of traits following BM; *K* > 1 indicates traits more conserved than expected under BM, and *K* < 1 indicates that traits have less phylogenetic signal than expected under BM.
+We tested the significance of *K* by comparing the observed value against values from a null distribution of 10,000 phylogenies with the traits randomly shuffled across the tips.
 We tested the significance of λ with a log-likelihood test comparing the likelihood of the observed value of λ vs. λ = 0.
-For all further analyses, we used methods that account for phylogenetic history for physiological traits that showed significant phylogenetic signal in at least one of the two metrics, and standard methods otherwise.
+For all further analyses, we used methods that account for phylogenetic history for physiological traits that showed evidence of phylogenetic signal in at least one of the two metrics, and standard methods otherwise.
 
 We used general linear mixed models (GLMMs) to investigate the effect of growth habit and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at -86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the \package{MCMCglmm} package [@Hadfield2010].
-For each of the three response variables (DT, ETR~max~, and PPFD~95%~), we constructed five models: a null model with no fixed effects, or models with either growth habit (high-elevation epiphyte, low-elevation epiphyte, terrestrial or saxicolous), generation (sporophyte vs.
+For each of the three response variables (species mean DT, ETR~max~, and PPFD~95%~), we constructed five models: a null model with no fixed effects, or models with either growth habit (high-elevation epiphyte, low-elevation epiphyte, terrestrial or saxicolous), generation (sporophyte vs.
 gametophyte), growth habit and generation, or growth habit, generation, and their interaction as fixed effects.
-We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix in the case the response variable had significant phylogenetic signal [@Hadfield2010].
+We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix if the response variable showed evidence of phylogenetic signal [@Hadfield2010].
 We used an inverse-Gamma distribution for all priors, and ran analyses for 500,000 iterations, with burnin after 1,000 iterations and thinning every 500 iterations.
@@
 
-To test for a correlation between environmental water availability and DT, we first derived a single maximum VPD for each species by taking the average of the maximum daily VPD at all sites where that species was present.
+To test for a correlation between environmental water availability and DT, we first derived a single maximum VPD for each species by calculating the average of the maximum daily VPD at all sites where that species was present.
 We used maximum VPD (\ie, minimum humidity) values because these may represent a climatic extreme that limits the occurrence of species.
@@
 
-To test whether gametophytes with wider ranges have higher DT, we calculated the extent of gametophyte range beyond sporophytes ("G beyond S") as the total elevational range (m) gametophytes were observed growing beyond sporophytes at either end of the elevational gradient.
+To test whether gametophytes with broader ranges have higher DT, we calculated the extent of gametophyte range beyond sporophytes ("G beyond S") as the total elevational range (m) gametophytes were observed growing beyond sporophytes at either end of the elevational gradient.
 We conducted PGLS with DT as the response variable and G beyond S as the explanatory variable with the \function{pgls} function in the \package{caper} [@Orme2018].
@@
 Datasets used in this study are available on FigShare (https://doi.org/10.6084/m9.figshare.14184572). 
-Code used to analyze the data and generate figures and this manuscript are available on GitHub (https://github.com/joelnitta/moorea-filmies).
+Code used to analyze the data and generate figures and this manuscript are available on GitHub (https://github.com/joelnitta/moorea_filmies).
 A Docker image is available to run the code at https://hub.docker.com/r/joelnitta/moorea_filmies.
@@
 
-In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes including *Callistopteris apiifolia* (C.
-Presl) Copel., *Crepidomanes minutum* var.
+In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes including *C.
+apiifolia*, *C.
+minutum* var.
 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
 Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson (`r s_figure("range")`).
-All of these have sporophytes that are normally only found in cloud forest (\circa{ }500--1200 m), but gametophytes that are distributed to lower elevations.
+All of these have sporophytes that are normally restricted to cloud forest (\circa{ }500--1200 m), but gametophytes that are distributed to lower elevations.
 
@@
 
-Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were mostly similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT when measured with *K* (gametophytes *K* = `r gameto_recover_k`, *P* = `r gameto_recover_k_pval`; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval`), but no significant signal in either metric for PPFD~95%~ or ETR~max~ (`r table("physig")`).
-Results for phylogenetic signal in DT as measured with λ differed slightly; although gametophytes showed significant signal (λ = `r gameto_recover_lambda`, *P* = `r gameto_recover_lambda_pval`) sporophytes did not (λ =`r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`).
+For DT, *K* was near 1 with *P* < 0.1 in both generations (gametophytes *K* = `r gameto_recover_k`, *P* = `r gameto_recover_k_pval`; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval`); λ was near 1 for both generations, but not significant for sporophytes (gametophytes λ = `r gameto_recover_lambda`, *P* = `r gameto_recover_lambda_pval`; sporophytes λ = `r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`).
+There was no evidence of significant phylogenetic signal in photosynthetic parameters (PPFD~95%~ and ETR~max~) in either generation (*P* >> 0.05, `r table("physig")`).
+Therefore, phylogenetic methods were used for analyses involving DT, but not photosynthetic parameters.
 
@@
 
+# Extract list of species with significant p-values for gameto DT more than sporo
+dt_g_more_than_s <-
+t_test_results %>% 
+  filter(est_diff_sporo_gameto < 0, response == "recovery", p_value < 0.05) %>%
+  abbrev_sp() %>%
+  mutate(species = str_replace_all(species, "_", ". ")) %>%
+  pull(species) %>%
+  knitr::combine_words(before = "*")
+
 # Extract p-values for species with gameto DT less than sporo
@@
   filter(est_diff_sporo_gameto > 0, response == "recovery", p_value < 0.05) %>%
+  pull(species) %>%
+  map_chr(~ttest_pval(species_select = ., response_select = "recovery", data = t_test_results)) %>%
+  knitr::combine_words()
+
+# Extract p-values for species with gameto DT more than sporo
+dt_g_more_than_s_pvals <-
+t_test_results %>%
+  filter(est_diff_sporo_gameto < 0, response == "recovery", p_value < 0.05) %>%
   pull(species) %>%
@@
   mutate(across(c(mean, sd), ~scales::percent(., suffix = " %", accuracy = 0.1))) %>%
-  mutate(sum_val = glue("{mean} ± {sd} (*n* = {n})"))
-
-rel_water_table %>% filter(salt == "LiCl", dry_time == "15") %>% pull(sum_val)
+  mutate(sum_val = glue("{mean} ± {sd} (*n* = {n} species)"))
+
+# Gameto DT
+mean_gameto_dt_by_habit <-
+combined_species_means %>%
+  filter(!is.na(recovery_mean), generation == "gametophyte") %>%
+  left_join(filmy_habit, by = "species") %>%
+  arrange(recovery_mean) %>%
+  group_by(habit) %>%
+  summarize(
+    mean = mean(recovery_mean),
+    sd = sd(recovery_mean),
+    n = n(),
+    .groups = "drop"
+  ) %>%
+  mutate(across(c(mean, sd), ~scales::percent(., suffix = " %", accuracy = 0.1))) %>%
+   mutate(sum_val = glue("{mean} ± {sd}, *n* = {n} species"))
+
+gameto_dt_ter_mean <- mean_gameto_dt_by_habit %>%
+  filter(habit == "Terrestrial") %>%
+  pull(sum_val)
 ```
@@
 A general correlation of DT with habitat was observed in sporophytes (`r figure("dt")`a).
-Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats.
-and *C.
-apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 NaCl (more stringent tests were not conducted on *A.
+The two terrestrial species lacked DT, although *C.
+apiifolia* showed slightly greater ability to recover by 48 h compared to *Abrodictyum dentatum* (Bosch) Ebihara & K Iwats, which completely failed to recover from even the gentlest treatment of 2 d at --38 NaCl (more stringent tests were not conducted on *A.
 dentatum* after confirming its lack of DT at this level).
@@
 kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
-All epiphytic species were capable of recovering from desiccation to a certain degree, but ability varied between species.
-Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
-Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.
-pallidum* and *P.
-borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
-Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.
+Nearly all epiphytic species were capable of recovering from desiccation of up to 2 d at --86 MPa.
+*Crepidomanes bipunctatum* (Poir.) Copel., a widespread low-elevation epiphyte, showed the greatest ability to recover from desiccation across all treatments.
+Interestingly, high-elevation epiphytes *Hymenophyllum multifidum* (G.
+Forst.) Sw.
+and *Hymenophyllum polyanthos* (Sw.) Sw.
+could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.
+
+Gametophytes of terrestrial species could not tolerate desiccation (mean 48 h recovery `r gameto_dt_ter_mean`).
+Levels of DT were in general higher for saxicolous and epiphytic gametophytes, but otherwise did not show a clear pattern (`r figure("dt")`b).
+
+Desiccation tolerance of gametophytes tended to be lower than sporophytes (`r figure("combined")`), and was significantly lower in `r dt_g_less_than_s` (*t*-test, *P* = `r dt_g_less_than_s_pvals`, respectively).
+Only `r dt_g_more_than_s` showed significantly higher DT in gametophytes compared to sporophytes (*t*-test, *P* = `r dt_g_more_than_s_pvals`).
 
 Control samples generally did not show any decrease in F~v~/F~m~ throughout the duration of the experiment (`r s_figure("dt-control")`) and maintained RWC near 100% (\figs{ }`r s_figure_num("water-all")`, `r s_figure_num("water-s")`).
-
-Relative water content in the desiccated state ranged from `r rel_water_table %>% filter(salt == "LiCl", dry_time == "15", rec_time == "desiccated") %>% pull(sum_val)` (--282 MPa for 15 d) to `r rel_water_table %>% filter(salt == "NaCl", dry_time == "2", rec_time == "desiccated") %>% pull(sum_val)` (--38 MPa for 2 d), and returned to \circa{ }100 % upon re-wetting in all species sampled (`r s_figure("water-s")`).
-
-DT of gametophytes did not significantly exceed that of sporophytes for any species, and was significantly lower in `r dt_g_less_than_s` (`r figure("combined")`; *t*-test, *P* = `r dt_g_less_than_s_pvals`, respectively).
-
-Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; `r figure("dt")`b).
+Relative water content in the desiccated state ranged from `r rel_water_table %>% filter(salt == "LiCl", dry_time == "15", rec_time == "desiccated") %>% pull(sum_val)` in the --282 MPa, 15 d treatment to `r rel_water_table %>% filter(salt == "NaCl", dry_time == "2", rec_time == "desiccated") %>% pull(sum_val)` in the --38 MPa, 2 d treatment (`r s_figure("water-all")`), and returned to \circa{ }100 % upon re-wetting in all species sampled (\figs{ }`r s_figure_num("water-all")`, `r s_figure_num("water-s")`).
 
 Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (`r s_figure("chamber")`).
-Brief fluctuations to higher humidities were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).
+In a few cases (particularly species with larger fronds, \eg, *C. apiifolia*), the NaCl chamber did not reach equilibrium humidity before 2 d (`r s_figure("chamber")`).
+Brief fluctuations in humidity were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).
 
@@
 # (so mean is the mean of species means, and sd is sd of species means)
-light_overall_means <-
+par_means <-
 combined_species_means %>%
+  filter(!is.na(par_mean)) %>%
   group_by(generation) %>%
   summarize(
-    etr = mean(etr_mean, na.rm = TRUE),
-    etr_sd = sd(etr_mean, na.rm = TRUE),
-    par = mean(par_mean, na.rm = TRUE),
-    par_sd = sd(par_mean, na.rm = TRUE),
+    mean = mean(par_mean),
+    sd = sd(par_mean),
+    n = n(),
     .groups = "drop"
   ) %>%
-  mutate(across(where(is.numeric), ~round_t(.x, 1))) %>%
   mutate(
-    etr_pretty = glue("{etr} ± {etr_sd}"),
-    par_pretty = glue("{par} ± {par_sd}"),
+    mutate(across(c(mean, sd), ~round_t(.x, 1))),
+    sum_val = glue("{mean} ± {sd}"),
   )
 
+etr_means <-
+combined_species_means %>%
+  filter(!is.na(etr_mean)) %>%
+  group_by(generation) %>%
+  summarize(
+    mean = mean(etr_mean),
+    sd = sd(etr_mean),
+    n = n(),
+    .groups = "drop"
+  ) %>%
+  mutate(
+    mutate(across(c(mean, sd), ~round_t(.x, 1))),
+    sum_val = glue("{mean} ± {sd}"),
+  )
+
 # Extract values by response and generation for citing in MS
-mean_par_sporo <- light_overall_means %>% filter(generation == "sporophyte") %>% pull(par_pretty)
-mean_par_gameto <- light_overall_means %>% filter(generation == "gametophyte") %>% pull(par_pretty)
-mean_etr_sporo <- light_overall_means %>% filter(generation == "sporophyte") %>% pull(etr_pretty)
-mean_etr_gameto <- light_overall_means %>% filter(generation == "gametophyte") %>% pull(etr_pretty)
+mean_par_sporo <- par_means %>% filter(generation == "sporophyte") %>% pull(sum_val)
+mean_par_gameto <- par_means %>% filter(generation == "gametophyte") %>% pull(sum_val)
+mean_etr_sporo <- etr_means %>% filter(generation == "sporophyte") %>% pull(sum_val)
+mean_etr_gameto <- etr_means %>% filter(generation == "gametophyte") %>% pull(sum_val)
+mean_par_sporo_n <- par_means %>% filter(generation == "sporophyte") %>% pull(n)
+mean_par_gameto_n <- par_means %>% filter(generation == "gametophyte") %>% pull(n)
+mean_etr_sporo_n <- etr_means %>% filter(generation == "sporophyte") %>% pull(n)
+mean_etr_gameto_n <- etr_means %>% filter(generation == "gametophyte") %>% pull(n)
 ```
@@
 
-Photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (\figs `r figure_num("combined")`, `r s_figure_num("light-curves")`, `r s_figure_num("traits-box")`).
+Photosynthetic optima showed clear differences across generations: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (\figs{ }`r figure_num("combined")`, `r s_figure_num("light-curves")`, `r s_figure_num("traits-box")`).
 Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
-`r mean_par_sporo` µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at `r mean_par_gameto` µmol·m^--2^ ·s^--1^.
-Similarly, sporophytes attained higher rates of electron transport (ETR~max~; `r mean_etr_sporo` µmol·m^--2^ ·s^--1^) than gametophytes (`r mean_etr_gameto` µmol·m^--2^ ·s^--1^).
-
-The GLMM including the effect of generation only was selected as the best-fitting model for ETR~max~, and growth habit + generation was selected for PPFD~95%~ (`r table("glmm")`).
-However, in the GLMM for PPFD~95%~, only the generation parameter was significant (`r table("glmm-pars")`).
+`r mean_par_sporo` µmol·m^--2^ ·s^--1^ (*n* = `r mean_par_sporo_n` species), whereas gametophytes did so at `r mean_par_gameto` µmol·m^--2^ ·s^--1^ (*n* = `r mean_par_gameto_n` species).
+Similarly, sporophytes attained higher rates of electron transport (ETR~max~) (`r mean_etr_sporo` µmol·m^--2^ ·s^--1^, *n* = `r mean_etr_sporo_n` species) than gametophytes (`r mean_etr_gameto` µmol·m^--2^ ·s^--1^, *n* = `r mean_etr_gameto_n` species).
+
+The GLMM including the effect of generation only was selected as the best-fitting model for ETR~max~, and the combined effect of growth habit plus generation and their interaction was selected for PPFD~95%~ (`r table("glmm")`).
 
@@
 
-Desiccation tolerance increased significantly with VPD for sporophytes, but not gametophytes (`r figure("corr")`; sporophytes *P* = `r sporo_vpd_pval`; `r table("pgls")`).
+Desiccation tolerance increased significantly with VPD for sporophytes, but not gametophytes (PGLS, *P* = `r sporo_vpd_pval`; `r figure("corr")`; `r table("pgls")`).
 
@@
 
-Vegetative DT was a key innovation linked to the rise of early land plants, but is largely limited to non-vascular groups such as bryophytes and algae in extant plants.
-Filmy ferns have received considerable attention over the years for their remarkable reversion to vegetative DT in the sporophyte phase [\eg, @GarcesCea2014; @Proctor2003; @Shreve1911].
+Vegetative DT was a key innovation linked to the rise of early land plants, but is largely limited to non-vascular groups such as bryophytes and algae among extant plants.
+Filmy ferns have received considerable attention for their remarkable reversion to vegetative DT in the sporophyte phase [\eg, @GarcesCea2014; @Ostria-Gallardo2020; @Proctor2003; @Shreve1911].
 Yet, no previous study to our knowledge has evaluated parallel physiology of sporophytes and gametophytes of filmy ferns in relation to DT.
@@
 
-We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below -200 MPa (VPD > 2.45 kPa) in some epiphytes.
+We identified a wide variation of DT in filmy fern sporophytes on Moorea, from marked sensitivity in terrestrial species to tolerance of water potentials well below -200 MPa (VPD > 2.45 kPa) in some epiphytes.
 Levels of similarly extreme DT have also been reported in other species of *Hymenophyllum* [@Proctor2003; @Proctor2012] and a general correlation of habitat with DT in filmy ferns has been previously established [@Parra2009; @Proctor2003; @Proctor2012; @Saldana2013; @Shreve1911], but no other studies to our knowledge have documented a similar range of DT values in filmy ferns from a single site.
@@
 -38 MPa when dry for 15 d.
-This may be an adaptation to cloud forest conditions; although cloud forests are frequently moist due to constant fog, dry periods are intense during the times when there is no cloud cover.
+This may be an adaptation to cloud forest conditions; although cloud forests are frequently moist due to constant fog, dry periods are intense during the times when there is no cloud cover (J.
+Nitta, pers.
+obs.).
 The complete lack of DT in terrestrial species is not surprising, considering they have much more direct access to water than epiphytes and therefore are under substantially less selective pressure to maintain DT.
-Since controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (`r s_figure("dt-control")`), it is unlikely that the lack of tolerance seen in terrestrial species is due solely to removal from the soil.
-The significant relationship between average daily maximum VPD and DT in sporophytes supports the specialization of sporophyte DT according to niche, and indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with VPD exceeding \circa{ }`r max_ter_vpd` kPa (`r figure("corr")`).
+Because controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (`r s_figure("dt-control")`), it is unlikely that the lack of tolerance seen in terrestrial species is solely due to removal from the soil.
+The significant relationship between average daily maximum VPD and DT in sporophytes supports the specialization of sporophyte DT according to niche, and indicates that sporophytes of non-tolerant terrestrial and saxicolous species may be unable to occur in habitats with VPD exceeding \circa{ }`r max_ter_vpd` kPa (`r figure("corr")`).
 
@@
 This differs from other fern groups with desiccation tolerant gametophytes but desiccation avoiding sporophytes [@WatkinsJr.2007c].
-In the filmy ferns we studied, DT levels of both sporophytes and gametophytes largely reflected habitat; in general, epiphytic species tended to have higher DT than terrestrial or saxicolous species, regardless of life stage (\figs `r figure_num("combined")`, `r s_figure_num("traits-box")`).
+In the filmy ferns we studied, DT levels of both sporophytes and gametophytes largely reflected habitat; in general, epiphytic species tended to have higher DT than terrestrial species, regardless of life stage (\figs{ }`r figure_num("combined")`, `r s_figure_num("traits-box")`).
 Interestingly, this echoes much earlier field observations (not experimental data) by @Holloway1930 on filmy fern gametophytes in New Zealand: he reported apparent DT in gametophytes of *Hymenophyllum rarum* R.
@@
 @WatkinsJr.2007c also found that DT of fern gametophytes tended to match habitat, with terrestrial species having lower DT than epiphytes.
-However, we found that gametophytes had lower DT than sporophytes regardless of growth habit (\figs `r figure_num("combined")`, `r s_figure_num("traits-box")`).
+However, we found that gametophytes had lower DT than sporophytes regardless of growth habit (\figs{ }`r figure_num("combined")`, `r s_figure_num("traits-box")`).
 At high elevation sites (\ie, cloud forest), epiphytic ferns often occur in a dense matrix of bryophytes (`r figure("habit")`b).
-It is possible that this bryophyte cover helps to retain moisture during brief periods of drought and thus protect the epiphytic fern gametophytes that grow amongst them.
-More detailed studies of the high elevation epiphytic environment and interactions between co-occurring gametophytes of ferns and bryophytes are needed to test this hypothesis.
-
-Although we predicted that species with gametophytes that occur far beyond the range of their sporophytes would have higher DT, this was not the case.
+It is possible that this bryophyte cover helps to retain moisture during brief periods of drought and thus protect the epiphytic fern gametophytes that grow amongst them [@Harrington2021].
+More detailed studies of the high elevation epiphytic environment and interactions between co-occurring gametophytes of ferns and bryophytes are required to test this hypothesis.
+
+Although we predicted that species with gametophytes that occur far beyond the range of their sporophytes would have higher DT, this was not the case (`r figure("corr")`c).
 This is perhaps best illustrated by *C. apiifolia* (`r figure("habit")`d, e, h).
-Sporophytes of this species are usually restricted to moist cloud forest sites from \circa{ }600 m to 1200 m, but we observed gametophytes over a much wider range, down to 200 m.
+Sporophytes of this species are usually restricted to moist cloud forest sites from \circa{ }600 m to 1200 m, but we observed gametophytes over a much wider range, down to 200 m (`r s_figure("range")`).
 A similar distribution pattern has been reported in congeneric *Callistopteris baldwinii* (D. C. Eaton) Copel.
@@
 Since gametophytes of *C.
-apiifolia* are common on Moorea and have a unique morphology [@Nitta2017], we were able to collect enough material of this species for additional tests of DT at 100% RH for 2 d.
-However, it still failed to recover from even this gentle treatment (`r s_figure("capi-water")`).
+apiifolia* are common on Moorea and have a unique morphology [@Nitta2017], we were able to collect sufficient material of this species for additional tests of DT at 100% RH for 2 d.
+However, *C. apiifolia* gametophytes still failed to recover from even this gentle treatment (`r s_figure("capi-water")`).
 This suggests that the widespread gametophytes of *C. apiifolia* do not rely on DT to attain their distribution, but instead must do so via some other mechanism.
@@
 It is likely that the gametophytes of *C.
-apiifolia* do not actually experience the conditions measured with our 'terrestrial' dataloggers, but instead grow within a much more consistently moist boundary layer.
+apiifolia* do not actually experience the conditions measured with our 'terrestrial' dataloggers, but instead grow within a more consistently moist boundary layer.
 Gametophytes of *C. apiifolia* produce gemmae (`r figure("habit")`e, h), and often form dense clonal mats.
@@
 It is possible that there is a tradeoff between growth rate and DT, such that species capable of rapid asexual growth like *C.
-apiifolia* are unable to tolerate desiccation, but slower growing epiphytes such as *Hymenophyllum* spp.
+apiifolia* are less able to tolerate desiccation, but slower growing epiphytes such as *Hymenophyllum* spp.
 can better tolerate it [@Oliver2000].
-Growth rates of tolerant and non-tolerant gametophytes including quantification of gemmae production should be investigated in these species to test this hypothesis.
+Growth rates of tolerant versus non-tolerant gametophytes, including quantification of gemmae production, should be investigated in these species to test this hypothesis.
 
@@
 
-Photosynthetic parameters did not vary significantly between species from different habitats.
-Rather, we observed a clear intergenerational difference, with gametophytes consistently adapted for lower light levels than sporophytes (`r figure("combined")`, `r table("glmm")`).
+We observed a clear intergenerational difference in photosynthetic optima, with gametophytes consistently adapted for lower light levels than sporophytes (`r figure("combined")`, `r table("glmm")`).
 In the only other study to our knowledge that compared photosynthetic rates between gametophytes and sporophytes of a filmy fern, @Johnson2000 also found that gametophytes of *V.
@@
 speciosa* is an extremely deep-shade plant, with gametophytes occupying protected rock crevices that receive less than 1 µmol·m^-2^·s^-1^ PPFD for most of the day [@Johnson2000]; many of the gametophytes sampled in our study occur in more exposed conditions.
-It is surprising, however, that we did not observe any differences between species in different habitats, given that light levels should vary between tree trunks and the forest floor, at least on a \circa{ }1 m scale.
+It is surprising, however, that we did not observe any consistent differences between species in different habitats, given that light levels should vary between tree trunks and the forest floor, at least on a \circa{ }1 m scale.
+Although the GLMM showed an interaction effect between generation and growth habit for PPFD~95%~, the model including only generation was selected for ETR~max~ (`r table("glmm")`).
 It is possible that the actual light levels experienced by gametophytes growing in these habitats may not vary significantly if they occupy protected microsites such as bryophyte mats or crevices in rock or bark.
@@
 Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates [@Parra2009; @Proctor2003; @Proctor2012; @Saldana2013]; however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
-We detected significant phylogenetic signal in DT with Blomberg's *K* (`r table("physig")`), suggesting that phylogeny should be considered when interpreting differences in DT between species.
-In contrast, the lack of phylogenetic signal in photosynthetic traits, coupled with the clear intergenerational differences we observed, suggest that sporophytes and gametophytes are adapted for different light levels regardless of clade or ecological niche.
+We detected phylogenetic signal in DT (`r table("physig")`), suggesting that phylogeny should be considered when interpreting differences in DT between species.
+In contrast, the lack of phylogenetic signal in photosynthetic traits, coupled with the clear intergenerational differences we observed, suggest that sporophytes and gametophytes are adapted for different light levels regardless of clade or growth habit.
 In a more densely sampled study of morphological evolution in the trichomanoid clade, @Dubuisson2013 identified several morphological changes including reduced stele and root systems associated with the evolution of epiphytic growth within the HE subclade (comprising the genera *Polyphlebium*, *Didymoglossum*, *Crepidomanes*, and *Vandenboschia*).
@@
 However, @Dubuisson2013 were unable to determine if these morphological changes had occurred independently, or if there was a single origin at the base of the HE clade followed by multiple losses.
-Increased sampling of both morphological and physiological traits of additional trichomanoid taxa is needed to distinguish between these scenarios.
+Increased sampling of both morphological and physiological traits of additional trichomanoid taxa is needed to distinguish between these competing scenarios.
 
@@
 
-In filmy ferns, both sporophyte and gametophyte generations are poikilohydric; however, we have found here that this does not mean they necessarily share similar physiological optima.
+In filmy ferns, both sporophyte and gametophyte generations are poikilohydric; however, this does not necessarily mean they share similar physiological optima.
 Although filmy fern sporophytes are often of small stature, elevating a leaf even a few cm above the substrate surface means escape from the boundary layer and exposure to substantially greater evapotranspiration.
@@
 life stage (and plant size) on DT.
-Finally, more detailed analyses of the strategies used by each generation to cope with desiccation stress, such as transcriptomics, are needed to better understand the ecology and evolution of filmy ferns as well as the transition from desiccation tolerance to avoidance in vascular plants.
+The recent application of transcriptomics to filmy fern ecophysiology has begun to reveal the genetic mechanisms underlying DT in these plants [@Ostria-Gallardo2020; @Ostria-Gallardo2020a].
+Future studies that apply such methods to gametophytes as well as sporophytes across a variety of growth habits are anticipated to provide even greater insight into the ecology and evolution of filmy ferns as well as the transition from desiccation tolerance to avoidance in vascular plants.
 
@@
 
-Members of the Davis Lab and Jonathan B. Losos provided helpful discussion and comments on drafts.
+Members of the Davis Lab and Jonathan Losos provided helpful discussion and comments on drafts.
 Saad Amer, Ravahere Taputuarai, Tohei Theophilus, and Suzanne Vinnette assisted with fieldwork and experiments.
@@
 phylosig %>%
-  mutate(k_pval_stars = case_when(
-    k_pval < 0.001 ~ "***",
-    k_pval < 0.01 ~ "**",
-    k_pval < 0.05 ~ "*",
-    TRUE ~ ""
-  )) %>%
-  mutate(lambda_pval_stars = case_when(
-    lambda_pval < 0.001 ~ "***",
-    lambda_pval < 0.01 ~ "**",
-    lambda_pval < 0.05 ~ "*",
-    TRUE ~ ""
-  )) %>%
   mutate(
-    kval = ifelse(kval < 0.001, format(kval, digits = 3), round_t(kval, 3)) %>% paste0(., k_pval_stars),
-    lambda = ifelse(lambda < 0.001, format(lambda, digits = 3), round_t(lambda, 3)) %>% paste0(., lambda_pval_stars),
+    kval = ifelse(kval < 0.001, format(kval, digits = 3), round_t(kval, 3)),
+    lambda = ifelse(lambda < 0.001, format(lambda, digits = 3), round_t(lambda, 3)),
+    k_pval = ifelse(k_pval < 0.001, format(k_pval, digits = 3), round_t(k_pval, 3)),
+    lambda_pval = ifelse(lambda_pval < 0.001, format(lambda_pval, digits = 3), round_t(lambda_pval, 3)),
     trait = str_replace_all(trait, "recovery_mean", "Recovery") %>%
@@
   ) %>%
-  select(trait, kval, lambda) %>%
-  kbl(format = "latex", booktabs = TRUE, col.names = c("", "\\textit{K}", "λ"), escape = FALSE) %>%
+  select(trait, kval, k_pval, lambda, lambda_pval) %>%
+  kbl(format = "latex", booktabs = TRUE, col.names = c("", "\\textit{K}","\\textit{P}(\\textit{K})", "λ","\\textit{P}(λ)"), escape = FALSE) %>%
   pack_rows(
@@
 Recovery, recovery of F~v~/F~m~ (%) following 2 d desiccation at –86 MPa; ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached.
-
-Asterisks indicate significance of phylogenetic signal: \*\*, \pval{ }< 0.01, \*, \pval{ }< 0.05.
 
@@
 # Format cell values
+glmm_params_table <-
 glmm_params %>%
@@
   kbl(format = "latex", booktabs = TRUE, escape = FALSE) %>%
-  kable_styling(latex_options = "scale_down") %>%
   # Make significant rows bold
@@
     escape = FALSE)
+
+if(params$output_type == "preprint") glmm_params_table <- kable_styling(glmm_params_table, latex_options = "scale_down")
+
+  glmm_params_table
+
 ```

### 2021-03-11 clean_data.sh
Add shell script to run data cleaning R scripts

@@ -1,0 +1,8 @@
+#!/usr/bin/bash
+
+# Run each data cleaning script in its own R session
+
+Rscript R/clean_chamber_data.R 
+Rscript R/clean_dt_data.R
+Rscript R/clean_lc_data.R
+Rscript R/clean_specimen_data.R

### 2021-03-12 ms/si.Rmd
Update SI
- standardize appearance of species names

@@
 
-**Authors:** Nitta JH; Watkins JE; Holbrook NM; Taputuarai R; Wang T; Davis CC
+**Authors:** Nitta JH; Watkins JE; Holbrook NM; Wang TW; Davis CC
 
@@
   scale_color_manual(values = habit_colors, labels = pretty_text) +
-  scale_y_discrete(labels = pretty_text) +
+  scale_y_discrete(labels = str_abbrev_sp_pretty) +
   labs(x = "Elevation (m)", y = "Species", color = "Growth habit") +
@@
     legend.position = "bottom",
-    axis.text.y = element_text(face = "italic")
+    axis.text.y = element_markdown()
   )
@@
   geom_point(aes(fill = salt, shape = salt), size = 1.25, position=position_dodge(width=0.5)) +
+  scale_y_discrete(labels = str_abbrev_sp_pretty) +
   scale_x_continuous(label = scales::percent) +
@@
     panel.grid.major.y = element_blank(),
+    axis.text.y = element_markdown(),
     legend.position = "bottom"
@@
   facet_grid(vars(dry_time), labeller = as_labeller(c(`2` = "2 days", `15` = "15 days")))
-
-# ggsave(here::here("results/S_fig_water_content_desic.pdf"))
 ```
@@
   rel_water_species_means %>%
-  filter(species %in% species_with_rec)
+  filter(species %in% species_with_rec) %>%
+  # Abbreviate species names to "G. species"
+  mutate(
+    species = str_replace_all(species, "[a-z]+_", ". "),
+    species = glue("*{species}*"))
 
@@
     panel.grid.major.x = element_blank(),
+    strip.text = element_markdown(),
     legend.position = "bottom",
@@
   )
-
-# ggsave(here::here("results/S_fig_water_content_full.png"))
 ```

### 2021-03-12 ms/manuscript.Rmd
Update MS

@@
 Joel H. Nitta  
-Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan  
-Phone: +81-  
-Fax: +81-  
+Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8562, Japan  
+Phone: +81-81-4-7136-3681  
+Fax: +81-4-7136-3680  
 email: joelnitta\@gmail.com
@@
 
-**Number of tables, black-and-white figures, and color figures.**  
+**Number of tables, black-and-white figures, and color figures**  
 4 tables / 0 black-and-white figures / 4 color figures
 
+**Author contributions**  
+Joel H.
+Nitta conceived of and designed the study with input from James E.
+Watkins, N.
+Michele Holbrook, and Charles C.
+Davis.
+Data collection was performed by Joel H.
+Nitta and Tristan W.
+Wang.
+Data analysis was performed by Joel H.
+Nitta.
+The first draft of the manuscript was written by Joel H.
+Nitta, with subsequent comments and edits from the other authors.
+All authors read and approved the final manuscript.
+
 \newpage
@@
 
-^3^Current address: Section of Plant Pathology and Plant-Microbe Biology, Cornell University, Ithaca, New York 14853 USA
-
-^4^Current address: Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan
+^3^Current affiliation: Section of Plant Pathology and Plant-Microbe Biology, Cornell University, Ithaca, New York 14853 USA
+
+^4^Current affiliation: Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8562, Japan
 
@@
 We conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.
-Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima (maximum electron transport rate of photosystem II and light level at 95% of that rate) than sporophytes.
+Gametophytes had similar or less desiccation tolerance (ability to recover from 2 d desiccation at -86 MPa) and lower photosynthetic optima (maximum electron transport rate of photosystem II and light level at 95% of that rate) than sporophytes.
 Epiphytes were more tolerant of desiccation than terrestrial species in both generations.
@@
 Samples were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
-Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012 field season for species other than *Callistopteris apiifolia* (C.
-Presl) Copel.) or Hobo ProV2 dataloggers logging every 5 min (2012 *C.
+Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012 field season for species other than *Callistopteris apiifolia* (Presl) Copel.) or Hobo ProV2 dataloggers logging every 5 min (2012 *C.
 apiifolia*, 2013, 2014 field seasons).
@@
 minutum* var.
-3 and *Hymenophyllum braithwaitei* (Ebihara & K.
-Iwats.), did not occur in the study plots, but each was observed from a single population on Mt.
+3 and *Hymenophyllum braithwaitei* Ebihara & K.
+Iwats., did not occur in the study plots, but each was observed from a single population on Mt.
 Tohiea, at \circa{ }1000 m and near the summit (1206 m), respectively.
@@
 The two terrestrial species lacked DT, although *C.
-apiifolia* showed slightly greater ability to recover by 48 h compared to *Abrodictyum dentatum* (Bosch) Ebihara & K Iwats, which completely failed to recover from even the gentlest treatment of 2 d at --38 NaCl (more stringent tests were not conducted on *A.
+apiifolia* showed slightly greater ability to recover by 48 h compared to *Abrodictyum dentatum* (Bosch) Ebihara & K.
+Iwats., which completely failed to recover from even the gentlest treatment of 2 d at --38 NaCl (more stringent tests were not conducted on *A.
 dentatum* after confirming its lack of DT at this level).
@@
 and *H.
-villosum* Colenso, both high elevation epiphytes with desiccation tolerant sporophytes, and lack of DT in terrestrial filamentous gametophytes of *Polyphlebium colensoi* (Hook.) Ebihara & K.
+villosum* Colenso, both high elevation epiphytes with desiccation tolerant sporophytes, and lack of DT in terrestrial filamentous gametophytes of *Polyphlebium colensoi* (Hook.
+f.) Ebihara & K.
 Iwats., *Abrodictyum strictum* (Menzies ex Hook.
@@
 
-## Figure Captions
+## Figure legends
 <!-- JPR requirements: Figure captions begin with the term Fig.
@@
     x = "Recovery Time (hr)",
-    y = "Recovery (%)",
+    y = "Recovery",
     title = glue("**{plot_part}**")
@@
 make_dt_plot(dt_plot_data, "sporophyte", "a")
-ggsave(here::here("results/ms/fig_1a_sporo_dt.pdf"), width = 6.85, height = 7.5)
+figname <- figure("dt") %>% str_remove_all("[:punct:]| ") %>% paste0("a.pdf")
+ggsave(file = here::here("results/ms", figname), width = 6.85, height = 7.5)
 ```
@@
 make_dt_plot(dt_plot_data, "gametophyte", "b")
-ggsave(here::here("results/ms/fig_1b_gameto_dt.pdf"), width = 6.85, height = 6.5)
+figname <- figure("dt") %>% str_remove_all("[:punct:]| ") %>% paste0("b.pdf")
+ggsave(here::here("results/ms", figname), width = 6.85, height = 6.5, device = "pdf")
 ```
@@
   scale_fill_manual(values = generation_colors, labels = pretty_text) +
-  # theme_bw(base_size = 10) +
   scale_y_discrete(
@@
 
-ggsave(here::here("results/ms/fig_combined.pdf"), width = 6.85, height = 5)
+figname <- figure("combined") %>% str_remove_all("[:punct:]| ") %>% paste0(".pdf")
+ggsave(here::here("results/ms", figname), width = 6.85, height = 5)
 ```
@@
     x = Inf, y = -Inf, 
-    label = glue("*P* = {vpd_sporophyte_pval}<br>*R*<sup>2</sup> = {vpd_sporophyte_adj_r_squared}"),
+    label = glue("*P* = {vpd_sporophyte_pval}<br>adj. *R*<sup>2</sup> = {vpd_sporophyte_adj_r_squared}"),
     hjust = 1, vjust = 0, size = 3, label.colour = "transparent") +
@@
 
-ggsave(here::here("results/ms/fig_corr.pdf"), width = 6.85, height = 5)
+figname <- figure("corr") %>% str_remove_all("[:punct:]| ") %>% paste0(".pdf")
+ggsave(here::here("results/ms", figname), width = 6.85, height = 5)
 ```

### 2021-03-12 _targets.R

@@ -218,9 +218,10 @@ tar_plan(
     ),
   
   # Render MS Word
-  # FIXME: 
+  # FIXME: some fixes need to be made manually for MS submission
   # - change Florence to "in press" (in-line and refs)
   # - remove extra space from bib entries
+  # - add missing table header rows
   ms_docx = latex2docx(
     latex = "results/ms/manuscript.tex",
     docx = "results/ms/manuscript.docx",

### 2021-03-12 R/functions.R
Fix author list in ESM2

@@ -1230,7 +1230,7 @@ write_gameto_times <- function(gameto_indiv_times, file_out) {
   gameto_indiv_times_lines <- c(
     "# Electronic supplementary materials: Online Resource 2",
     "# Title: Intergenerational niche differentiation in filmy ferns (Hymenophyllaceae)",
-    "# Authors: Nitta JH; Watkins JE; Holbrook NM; Taputuarai R; Wang T; Davis CC",
+    "# Authors: Nitta JH; Watkins JE; Holbrook NM; Wang TW; Davis CC",
     "# Journal: Journal of Plant Research",
     "# Corresponding author: Joel H. Nitta (joelnitta@gmail.com)",
     "# Contents: start and end times of desiccation tolerance test on gametophytes of filmy ferns from Moorea French Polynesia",

### 2021-03-12 ms/manuscript.Rmd
Update MS

@@
 Recently, transcriptomics has been applied to elucidate the genes involved at each step of the desiccation-recovery cycle of filmy ferns [@Ostria-Gallardo2020; @Ostria-Gallardo2020a]. 
-In epiphytic *Hymenoglossum cruentum* Cav.
-from Chile, key genes and gene networks were identified related to intracellular mobility and photosynthetic metabolism in the desiccated state, and those related to detoxifying pathways and stabilization of photosystems during rehydration [@Ostria-Gallardo2020a].
+In epiphytic *Hymenoglossum cruentum* Cav., key genes and gene networks were identified related to intracellular mobility and photosynthetic metabolism in the desiccated state, and those related to detoxifying pathways and stabilization of photosystems during rehydration [@Ostria-Gallardo2020a].
 In a comparison of low-trunk and canopy epiphytes, @Ostria-Gallardo2020 identified multiple differentially expressed genes that may be specific to tolerating desiccation in these microhabitats.
@@
 For this study, only gametophytes that could be identified as Hymenophyllaceae were used; others were excluded from analysis.
+
+Fieldwork was done under permits issued by the French Polynesian Government (Délégation à la Recherche) and the Haut-commissariat de la République en Polynésie Francaise (Protocole d’Accueil 2012–2014).
+Voucher specimens were deposited at UC, with duplicates at GH and PAP (abbreviations follow @Thiers2021).
 
@@
 
-\*\*\* NOTE TO REVIEWERS: The dataset and code will not be made public until the paper is accepted.
-While the paper is under review, these links can be used for the dataset (https://figshare.com/s/d6349abf01a3756a5aae) and code (https://www.dropbox.com/sh/h788g1h3sjcqj0j/AACwFWTmL6GxrbLkCfOrXZU9a?dl=0)  \*\*\*
+\*\*\* NOTE TO REVIEWERS: The DOI for the dataset will not be active until the paper is accepted. Until then, this link can be used for the dataset https://figshare.com/s/d6349abf01a3756a5aae  \*\*\*
 

### 2021-03-12 ms/si.Rmd
Convert yield back to 0-1 range after comparison with manually entered values

@@
   # Filter out individuals with low pre-treatment yields
-  filter(!is.na(yield_pre), yield_pre > 400) %>%
+  filter(!is.na(yield_pre), yield_pre > 0.4) %>%
   # Calculate recovery by individual
@@
   # Filter out individuals with low pre-treatment yields
-  filter(!is.na(yield_pre), yield_pre > 400) %>%
+  filter(!is.na(yield_pre), yield_pre > 0.4) %>%
   calculate_recovery() %>%

### 2021-03-12 R/clean_dt_data.R

@@ -792,7 +792,8 @@ filmy_dt_wide <-
   # Remove rows with no yield or weight at all
   filter(!(is.na(yield_manual_entry) & is.na(yield_pam) & is.na(weight))) %>%
   filter(str_detect(note, "exclude", negate = TRUE)) %>%
-  # Convert yield to how it appears on the miniPAM (1000*actual yield)
+  # Convert yield to how it appears on the miniPAM (1000*actual yield) 
+  # to check against manual entries
   mutate(yield_pam = 1000*yield_pam) %>%
   # Check for problems in yield differences between 
   # manual entry and values matched from miniPAM
@@ -816,6 +817,8 @@ filmy_dt_wide <-
     # - use the miniPAM otherwise
     TRUE ~ yield_pam
   )) %>%
+  # Convert yield back to range between 0 and 1
+  mutate(yield = yield/1000) %>%
   select(-yield_manual_entry, -yield_pam, -memory, -note) %>%
   rename(time = date_time) %>%
   # Convert to wide format

### 2021-03-12 _targets.R

@@ -79,7 +79,7 @@ tar_plan(
     # Only include samples that should be presented in the main MS
     filter(section == "main") %>%
     # Filter out individuals with low pre-treatment yields
-    filter(!is.na(yield_pre), yield_pre > 400) %>%
+    filter(!is.na(yield_pre), yield_pre > .4) %>%
     calculate_recovery(),
   
   # - calculate water content during DT test per individual

### 2021-03-13 ms/data_readme.Rmd
Add Rmd to generate data README

---
output: rtf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE, cache = FALSE}
# TO CONVERT TO PLAIN TEXT:
# after rendering to rtf,
# 1. Copy the contents of the rtf file, paste to txt.
# 2. Replace all single line breaks with double line breaks
# (in Atom search-replace with grep: \n for \n\n)
# 3. Hard-wrap all text at 80 characters (in Atom with autoflow: cmd-alt-q)
# (optional: add name after first line)
# 4. Save as "README.txt"

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE,
  warning = FALSE, results = "hide", cache = FALSE)

# Load objects from targets workflow (all paths to data files)
tar_load(gameto_times_2012_file)
tar_load(sporo_dt_times_file)
tar_load(specimens_raw_file)
tar_load(filmy_dt_file)
tar_load(filmy_dt_chamber_file)
tar_load(filmy_habit_file)
tar_load(light_data_file)
```

This README.txt file was generated on `r format(Sys.time(), '%d %B, %Y')`

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\- GENERAL INFORMATION \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

Title of Dataset: Data from: Intergenerational niche differentiation in filmy
ferns (Hymenophyllaceae)

Author Information

Principal Investigator: Joel H. Nitta

Department of Integrated Biosciences, Graduate School of Frontier Sciences,
The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8562, Japan
joelnitta@gmail.com

Associate or Co-investigators: James E. Watkins, N. Michele Holbrook,
Tristan W. Wang, Charles C. Davis

Date of data collection: 2012--2014

Geographic location of data collection: Moorea, French Polynesia

Information about funding sources or sponsorship that supported the
collection of the data: Funding provided in part by the National Science
Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to JHN and
CCD), Setup Funds from Harvard University to CCD, American Society of Plant
Taxonomists (Research Grant for Graduate Students to JHN), Garden Club of
America (Award in Tropical Botany to JHN), Harvard University Herbaria (Fernald
Fieldwork Fellowship to JHN), Society of Systematic Biologists (Graduate Student
Research Award to JHN), and Systematics Association (Systematics Research Fund
to JHN).

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

SHARING/ACCESS INFORMATION

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

Licenses/restrictions placed on the data, or limitations of reuse: CC0 1.0
Universal (CC0 1.0)

Recommended citation for the data: Nitta JH, Watkins JE, Holbrook NM, Wang TW,
Davis CC (2021) Data from: Intergenerational niche differentiation in filmy
ferns (Hymenophyllaceae) https://doi.org/10.6084/m9.figshare.14184572

Citation for and links to publications that cite or use the data: Nitta JH,
Watkins JE, Holbrook NM, Wang TW, Davis CC (2021) Intergenerational niche
differentiation in filmy ferns (Hymenophyllaceae) <<FIXME: ADD JOURNAL>>

Code for analyzing the data is available on github:
https://github.com/joelnitta/moorea_filmies

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

DATA & FILE OVERVIEW

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

File list (filenames, directory structure (for zipped files) and brief
description of all data files):

- 2012_gameto_dt_times.csv: Start and end times of desiccation tolerance (DT)
  test on fern gametophytes during the 2012 field season.
- 2012_sporo_dt_times.csv: Start and end times of desiccation tolerance (DT)
  test on fern sporophytes during the 2012 field season.
- data_raw.tar.gz: Zipped raw data files. For a description, see README.md in
  the zipped archive.
- fern_specimens.csv: Specimen data for ferns collected in French Polynesia
  primarily by Joel H. Nitta.
- filmy_dt.csv: Data from desiccation tolerance test on filmy ferns from
  from Moorea, French Polynesia.
- filmy_dt_chamber.csv: Temperature and humidity inside desiccation chambers during
  desiccation tolerance test.
- filmy_growth_habit.csv: Growth habit of filmy ferns from Moorea,
  French Polynesia.
- filmy_light_curves.csv: Light levels and chlorophyll fluorescence measured
  in filmy ferns from Moorea, French Polynesia with a chlorophyll
  fluorometer.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

METHODOLOGICAL INFORMATION

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

Description of methods used for collection/generation of data:

Desiccation tolerance (DT) and light responses were measured in sporophytes and
gametophytes of filmy ferns (family Hymenophyllaceae) on the island of Moorea,
French Polynesia.

Samples were collected in the field and stored in plastic bags with a small
amount of water to keep them fresh during transport to the lab.

To measure DT, pre-treatment maximum photochemical yield of photosystem II was
measured in fresh plants after a 10 min period of dark-adaptation using a
portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany). Samples were
then transferred to desiccation chambers containing saturated salts at three
different desiccation intensities or a control treatment with moist tissues
(100% RH), and water withheld for either a short (2 d) or long (15 d) interval.
Conditions inside the desiccation chambers were monitored during the experiment
using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging
every 10 min or Hobo ProV2 dataloggers logging every 5 min. Salts used for
desiccation and their corresponding mean water potentials and approximate
relative humidity and VPD are as follows: LiCl (-282 MPa, 18% RH, 2.45 kPa),
Mg(NO3)2 (-86 MPa, 58% RH, 1.25 kPa), and NaCl (-38 MPa, 80% RH, 0.60 kPa).
Following the desiccation treatment, plants were rewetted and yield of
photosystem II was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Eight individuals per treatment were used for sporophytes. No replication by
species was possible for gametophytes due to their cryptic morphology. All
gametophytes were subjected to the same treatment (2 d at -86 MPa). Gametophytes
were later identified to species by DNA barcoding.

Relative water content (RWC) was measured in a subset of samples by recording
the mass of samples prior to the DT test (fresh mass), at each step of the DT
test (turgid mass), then after drying them overnight in a drying oven at 65 °C
following the DT test (dry mass). Relative water content was calculated as RWC =
(TM - DM) / (FM - DM), where TM is turgid mass, DM is dry mass, and FM is fresh
mass. Relative water content was not calculated for gametophytes, as these were
too small to measure accurately with the balances available.

To measure light responses, rapid light response curves were constructed for
each species by measuring photosynthetic yield at gradually increasing levels of
photosynthetically active actinic light (400 nm to 700 nm) with the Light Curve
function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh,
Effeltrich, Germany) as described by the manufacturer.

Fieldwork was done under permits issued by the French Polynesian Government
(Délégation à la Recherche) and the Haut-commissariat de la République en
Polynésie Francaise (Protocole d’Accueil 2012–2014).

fern_specimens.csv, filmy_dt.csv, filmy_dt_chamber.csv, and
filmy_light_curves.csv were generated by processing the raw data files using
scripts available at https://github.com/joelnitta/moorea_filmies (scripts
beginning with "clean_" in the "R/" directory).

For full methods, see Nitta JH, Watkins JE, Holbrook NM, Wang TW, Davis CC (2021) <<FIXME: ADD JOURNAL>>

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

DATA-SPECIFIC INFORMATION

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r gameto-dt-times}
gameto_dt_times <- read_csv(here::here(gameto_times_2012_file))
```

2012_gameto_dt_times.csv: Start and end times of desiccation tolerance (DT)
experiment on fern gametophytes during the 2012 field season. All gametophytes
were desiccated 2 d at -86 MPa. For other field seasons (2013, 2014), start and
end times were recorded during each measurement, so are not included in this
file.

Number of variables: `r ncol(gameto_dt_times)`

Number of cases/rows: `r nrow(gameto_dt_times)`

Variable list:

- date_time: Date and time, formatted YYYY-MM-DD HH-MM-SS.
- event: Start or end of experiment.
- group: Code corresponding to site and growth habit of samples.
- cluster: Number assigned to each group of samples (gametophyte individuals)
that were included together in a single experiment.

Missing data codes: No missing data.

Specialized formats or other abbreviations used: For 'group', 'ter' indicates
terrestrial samples, 'epi' indicates epiphytic samples; number indicates site
as follows: 1 = Three Pines 200m, 2 = Mt. Tohiea 400m, 3 = Mt. Tohiea 600m,  4 =
Mt. Tohiea 800m, 5 = Mt. Tohiea 1000m, 6 = Mt. Tohiea 1170m. For a description
of sites, see Nitta, Joel H.; Meyer, Jean-Yves; Taputuarai, Ravahere; Davis,
Charles C. (2016), Data from: Life cycle matters: DNA barcoding reveals
contrasting community structure between fern sporophytes and gametophytes,
Dryad, Dataset, https://doi.org/10.5061/dryad.df59g

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r sporo-dt-times}
sporo_dt_times <- read_csv(here::here(sporo_dt_times_file))
```

2012_sporo_dt_times.csv: Start and end times of desiccation tolerance (DT) test
on fern sporophytes during the 2012 field season. For other field seasons
(2013, 2014), start and end times were recorded during each measurement, so are
not included in this file.

Number of variables: `r ncol(sporo_dt_times)`

Number of cases/rows: `r nrow(sporo_dt_times)`

Variable list:

- date_time: Date and time, formatted YYYY-MM-DD HH-MM-SS
- event: Start or end of experiment.
- salt: Name of salt used to maintain constant humidity in the desiccation
chamber (see methods above for corresponding relative humidity).
- dry_time: Length (d) of desiccation.
- species: Species included in experiment.

Missing data codes: No missing data.

Specialized formats or other abbreviations used: None.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

data_raw.tar.gz: Zipped raw data files. For a description, see README.md in
the zipped archive. These were used to generate the clean data files with code
available at https://github.com/joelnitta/moorea_filmies (scripts starting with
'clean_' in the 'R' folder).

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r specimens}
specimens <- read_csv(here::here(specimens_raw_file))
```

fern_specimens.csv: Specimen data for ferns collected in French Polynesia
primarily by Joel H. Nitta.

Number of variables: `r ncol(specimens)`

Number of cases/rows: `r nrow(specimens)`

Variable list:

- specimen_id: Unique key assigned to each specimen.
- specimen: Specimen voucher (collector name and number).
- collector: Name of primary collector.
- coll_num: Collection number.
- family: Family.
- genus: Genus.
- specific_epithet: Specific epithet.
- infraspecific_rank: Infraspecific rank.
- infraspecific_name: Infraspecific name.
- certainty: Taxonomic certainty.
- species: Genus and specific epithet.
- taxon: Genus, specific epithet, and any infraspecific names.
- country: Country.
- county: County (or approximate political division).
- locality: General locality.
- site: Name of collection site.
- elevation: Elevation (m).
- latitude: Latitude (decimal degrees).
- longitude: Longitude (decimal degrees).
- observations: Observations about the specimen.
- generation: Gametophyte or sporophyte.
- gameto_net_location: Location of gametophyte sampling grid.
- gameto_square: Grid square where gametophyte was sampled.
- gameto_habit: Gametophyte growth habit.
- date_collected: Collection date.
- other_collectors: Names of other collectors present.
- herbaria: Herbarium code where specimen is deposited.

Missing data codes: Missing data entered as 'NA'.

Specialized formats or other abbreviations used: Herbarium codes follow Index
Herbariorum (http://sweetgum.nybg.org/science/ih/)

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r filmy-dt}
filmy_dt <- read_csv(here::here(filmy_dt_file))
```

filmy_dt.csv: Data from desiccation tolerance test on filmy ferns from from
Moorea, French Polynesia.

Number of variables: `r ncol(filmy_dt)`

Number of cases/rows: `r nrow(filmy_dt)`

Variable list:

- species: Species.
- salt: Name of salt used to maintain constant humidity in the desiccation
chamber (see methods above for corresponding relative humidity).
- dry_time: Length (d) of desiccation.
- individual: Individual code assigned to each sample per species per treatment.
(sporophytes) or collection number (gametophytes).
- generation: Gametophyte or sporophyte.
- dataset: Name of corresponding raw dataset file (sporophytes only).
- yield_pre: Maximum photochemical yield of photosystem II (Fv/Fm) prior to
desiccation treatment.
- yield_30min: Fv/Fm 30 m after rewetting following desiccation treatment.
- yield_24hr: Fv/Fm 24 h after rewetting.
- yield_48hr: Fv/Fm 48 h after rewetting.
- yield_72hr: Fv/Fm 72 h after rewetting (gametophytes only).
- yield_dry: Fv/Fm in the desiccated state.
- weight_pre: Mass (g) prior to desiccation treatment.
- weight_desiccated: Mass (g) in the desiccated state.
- weight_30min: Mass (g) 30 m after rewetting.
- weight_24hr: Mass (g) 24 h after rewetting.
- weight_48hr: Mass (g) 48 h after rewetting.
- weight_dry: Mass (g) after drying overnight at 60 C.
- time_pre: Date and time just before starting desiccation treatment,
formatted YYYY-MM-DD HH-MM-SS.
- time_30min: Date and time 30 m after rewetting.
- time_24hr: Date and time 24 h after rewetting.
- time_48hr: Date and time 48 h after rewetting.
- time_72hr: Date and time 72 h after rewetting (gametophytes only).
- time_dry: Date and time when yield was measured in the desiccated state.

Missing data codes: Missing data entered as 'NA'.

Specialized formats or other abbreviations used: None.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r filmy-dt-chamber}
filmy_dt_chamber <- read_csv(here::here(filmy_dt_chamber_file))
```

filmy_dt_chamber.csv: Temperature and humidity inside desiccation chambers during
desiccation tolerance test, measured using Track-It RH/Temp dataloggers
(Monarch Instrument, Amherst, NH) logging every 10 min or Hobo ProV2 dataloggers
logging every 5 min.

Number of variables: `r ncol(filmy_dt_chamber)`

Number of cases/rows: `r nrow(filmy_dt_chamber)`

Variable list:

- date_time: Date and time, formatted YYYY-MM-DD HH-MM-SS.
- temp: Temperature (degrees Celsius)
- rh: Relative humidity (%)
- salt: Name of salt used to maintain constant humidity in the desiccation
chamber (see methods above for corresponding relative humidity).
- generation: Gametophyte or sporophyte, or "mixed" if samples of both generations
were present in the growth chamber.
- year: Year.
- species: Species.
- file: Name of file containing raw data.
- serial_no: Serial number of datalogger.

Missing data codes: Missing data entered as 'NA'.

Specialized formats or other abbreviations used: None.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r filmy-habit}
filmy_growth_habit <- read_csv(here::here(filmy_habit_file))
```

filmy_growth_habit.csv: Growth habit of filmy ferns from Moorea, French
Polynesia. Growth habit determined by field observation. Species were
categorized into one of four growth habits: terrestrial (growing in soil, with
true roots), saxicolous  (growing on rocks), low-elevation epiphyte (mostly
occurring outside of cloud  forest, i.e., below ca. 500 m), and high-elevation
epiphyte (mostly occurring  within cloud forest, i.e., above ca. 500 m).
Crepidomanes bipunctatum and Crepidomanes  humile were observed growing as both
epiphytes and saxicoles, but are coded as  epiphytes to distinguish them from
exclusively saxicolous species.

Number of variables: `r ncol(filmy_growth_habit)`

Number of cases/rows: `r nrow(filmy_growth_habit)`

Variable list:

- species: Species.
- habit: Growth habit.

Missing data codes: None.

Specialized formats or other abbreviations used: None.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r filmy-light-curves}
filmy_light_curves <- read_csv(here::here(light_data_file))
```

filmy_light_curves.csv: Light levels and chlorophyll fluorescence measured in
filmy ferns from Moorea, French Polynesia with a chlorophyll fluorometer.
Chlorophyll  fluorescence measured with "light curve" function of mini-PAM
chlorophyll fluorometer.

Number of variables: `r ncol(filmy_light_curves)`

Number of cases/rows: `r nrow(filmy_light_curves)`

Variable list:

- type: Type of fluorescence measurement, from mini-PAM raw data.
- no: Number of fluorescence measurement, from mini-PAM raw data (not unique, 
because this file includes measurements from multiple mini-PAM raw data files).
- f: Momentary fluorescence level before application of saturation pulse.
- fm: Maximum fluorescence level of the illuminated sample.
- par: Level of photosynthetically active radiation during fluorescence 
measurement.
- yield: Effective photochemical quantum yield of photosystem II.
- etr: Relative electron transport rate.
- date_time: Date and time of measurement, formatted YYYY-MM-DD HH-MM-SS.
- species: Species.
- individual: Individual code assigned to each sample per species per treatment.
- generation: Gametophyte or sporophyte.
- coll_num: J.H. Nitta collection number
- sporo_site: Name of site where sporophyte sample was collected.
- condition: Condition under which measurements were taken (lab or field).
- date: Date of measurement, formatted YYYY-MM-DD.
- outlier: "TRUE" or "FALSE", indicating whether measurement should be excluded 
from light curve analysis.
- light_id: Unique ID number corresponding to light curve measured from a single 
sample.

Missing data codes: Missing data entered as 'NA'.

Specialized formats or other abbreviations used: For a description of sporophyte
sites, see Nitta, Joel H.; Meyer, Jean-Yves; Taputuarai, Ravahere; Davis,
Charles C. (2016), Data from: Life cycle matters: DNA barcoding reveals
contrasting community structure between fern sporophytes and gametophytes,
Dryad, Dataset, https://doi.org/10.5061/dryad.df59g

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

CHANGE LOG

\-\-\-

2021-03-14

Generate this README file.

### 2021-03-13 _targets.R

@@ -15,6 +15,7 @@ tar_plan(
   moorea_sites = read_csv(
     moorea_sites_file,
     col_types = "cnnn") %>%
+    # remove sites on Tahiti (Mt. Aorai)
     filter(str_detect(site, "Aorai", negate = TRUE)),
   
   # - growth habit
@@ -27,11 +28,13 @@ tar_plan(
   # - phylogenetic tree (time tree, lacks BS values at nodes)
   tar_file(filmy_phy_file, "data/nitta_2017/treepl_Moorea_Tahiti.tre"),
   filmy_phy_no_bs = ape::read.tree(filmy_phy_file) %>%
+    # trim to only filmy ferns
     ape::keep.tip(filmy_species),
   
   # - phylogenetic tree (ML tree with BS values at nodes)
   tar_file(filmy_phy_bs_file, "data/nitta_2017/RAxML_bipartitions.all_broad.reduced"),
   filmy_phy_bs = ape::read.tree(filmy_phy_bs_file) %>%
+    # trim to only filmy ferns
     ape::keep.tip(filmy_phy_no_bs$tip.label),
   
   # - transfer bootstrap values from ML tree to time tree
@@ -59,11 +62,11 @@ tar_plan(
   tar_file(specimens_raw_file, "data/fern_specimens.csv"),
   specimens_raw = read_csv(specimens_raw_file),
   
-  # - gametophyte DT groups
+  # - gametophyte DT times entered manually (2012 only)
   tar_file(gameto_times_2012_file, "data/2012_gameto_dt_times.csv"),
   gameto_time_summary_2012 = load_gameto_time_summary_2012(gameto_times_2012_file),
   
-  # - sporophyte DT times entered manually
+  # - sporophyte DT times entered manually (2012 only)
   tar_file(sporo_dt_times_file, "data/2012_sporo_dt_times.csv"),
   sporo_dt_times = load_sporo_dt_times(sporo_dt_times_file),
   
@@ -131,7 +134,7 @@ tar_plan(
   filmy_lc_model_fitted_data = extract_fitted_lc_data(light_models),
   
   # - extract model parameters: 
-  # critical PAR, and ETR at 95% of estimated max value
+  #  ETR at 95% of estimated max value and PAR at that value
   filmy_lc_model_params = extract_lc_model_params(light_models),
   
   # - calculate mean light curve parameters by species and generation
@@ -143,7 +146,7 @@ tar_plan(
   # - calculate relative water content by species (sporophytes only)
[…]

### 2021-03-13 ms/manuscript.Rmd
Update MS

@@
 The growth habit of each species was scored based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, \ie, below \circa{ }500 m), or high elevation epiphyte (mostly occurring in cloud forest, \ie, above \circa{ }500 m).
+*Crepidomanes bipunctatum* (Poir.) Copel.
+and *C.
+humile* (G.
+Forst.) Bosch were observed growing as both epiphytes and saxicoles, but were treated as epiphytes to distinguish them from exclusively saxicolous species.
 
@@
 Nearly all epiphytic species were capable of recovering from desiccation of up to 2 d at --86 MPa.
-*Crepidomanes bipunctatum* (Poir.) Copel., a widespread low-elevation epiphyte, showed the greatest ability to recover from desiccation across all treatments.
+*Crepidomanes bipunctatum*, a widespread low-elevation epiphyte, showed the greatest ability to recover from desiccation across all treatments.
 Interestingly, high-elevation epiphytes *Hymenophyllum multifidum* (G.

### 2021-03-13 ms/manuscript.Rmd
Add image file for fig 1
Use rmarkdown parameters to print figures or not

@@
     fig_caption: yes
-    keep_tex: yes
+    keep_tex: yes # for converting tex to doc
     latex_engine: xelatex
@@
       in_header: journal-of-plant-research.sty
+# Prevent figures from moving around, suppress automatic "Figure #:" in captions
 header-includes: 
@@
   \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
+  \renewcommand{\caption}{}
 params:
-  output_type: ms_for_word
+  output_type: "ms"
 editor_options: 
@@
 <!-- to move to the next line without starting a new paragraph, 
-add two spaces after last word, e.g., Atsushi Ebihara__ -->
+add two spaces after last word, e.g., Joel Nita__ -->
 
@@
 ## Figure legends
+
 <!-- JPR requirements: Figure captions begin with the term Fig.
@@
 Relationships between recovery and VPD or range of gametophytes beyond sporophytes were analyzed with phylogenetic generalized least squares (PGLS); significant relationships (*P* < 0.05) shown with grey line.
-
-\newpage
 
@@
 
-```{r filmies-fig, fig.width = 6.85, results = "asis", purl = FALSE}
-knitr::include_graphics(here::here("data/figure_1.png"))
-```
-
-**`r figure("habit")`**
-
-\newpage
-
-```{r filmy-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = "", purl = FALSE}
-make_dt_plot(dt_plot_data, "sporophyte", "a")
-figname <- figure("dt") %>% str_remove_all("[:punct:]| ") %>% paste0("a.pdf")
-ggsave(file = here::here("results/ms", figname), width = 6.85, height = 7.5)
-```
-
-**`r figure("dt")`**
-
-\newpage
-
-```{r gameto-dt-plot, fig.width = 6.85, fig.height = 7.5, fig.cap = "", purl = FALSE}
-make_dt_plot(dt_plot_data, "gametophyte", "b")
-figname <- figure("dt") %>% str_remove_all("[:punct:]| ") %>% paste0("b.pdf")
-ggsave(here::here("results/ms", figname), width = 6.85, height = 6.5, device = "pdf")
-```
-
-**`r figure("dt")`** (continued)
-
-\newpage
-
-```{r combined-plot, fig.width = 6.85, fig.height = 5, fig.cap = "", purl = FALSE}
+<!-- From here, use params$output_type to output plots directly for preprint PDF, or save to file for journal submission -->
+
+```{r filmies-print, fig.width = 6.85, results = "asis", fig.cap = figure("habit"), eval = params$output_type == "preprint", purl = FALSE}
+# Print plot if knitting MS to preprint
+knitr::include_graphics(here::here("images/figure_1.png"))
+```
+
+```{r filmy-dt-plot-make, purl = FALSE}
+# Make plot
+filmy_dt_plot <- make_dt_plot(dt_plot_data, "sporophyte", "a")
+```
+
+```{r filmy-dt-plot-save, eval = params$output_type == "ms", purl = FALSE}
+# Save plot to file if knitting MS for submission
+filmy_dt_plot_name <- figure("dt") %>% str_remove_all("[:punct:]| ") %>% paste0("a.pdf")
+ggsave(plot = filmy_dt_plot, file = here::here("results/ms", filmy_dt_plot_name), width = 6.85, height = 7.5)
+```
+
+```{r filmy-dt-plot-print, fig.width = 6.85, fig.height = 7.5, fig.cap = figure("dt"), eval = params$output_type == "preprint", purl = FALSE}
+# Print plot if knitting MS to preprint
+filmy_dt_plot
+```
+
+```{r gameto-dt-plot-make, purl = FALSE}
+# Make plot
+gameto_dt_plot <- make_dt_plot(dt_plot_data, "gametophyte", "b")
+```
+
+```{r gameto-dt-plot-save, eval = params$output_type == "ms", purl = FALSE}
+# Save plot to file if knitting MS for submission
+gameto_dt_plot_name <- figure("dt") %>% str_remove_all("[:punct:]| ") %>% paste0("b.pdf")
+ggsave(plot = gameto_dt_plot, file = here::here("results/ms", gameto_dt_plot_name), width = 6.85, height = 6.5, device = "pdf")
+```
+
+```{r gameo-dt-plot-print, fig.width = 6.85, fig.height = 7.5, fig.cap = paste(figure("dt"), "(continued)"), eval = params$output_type == "preprint", purl = FALSE}
+# Print plot if knitting MS to preprint
+gameto_dt_plot
+```
+
+```{r combined-plot-make, purl = FALSE}
+# Make plot
+
 # Copy data so they can be manipulated for this plot
@@
 
-# Combine plots
-a + b + c + d + plot_layout(nrow = 1, guides = "collect") &
+# Make combined plot
+combined_plot <- a + b + c + d + plot_layout(nrow = 1, guides = "collect") &
   theme(
@@
     legend.margin = margin())
-
-figname <- figure("combined") %>% str_remove_all("[:punct:]| ") %>% paste0(".pdf")
-ggsave(here::here("results/ms", figname), width = 6.85, height = 5)
-```
-
-**`r figure("combined")`**
-
-```{r corr-plot, fig.width = 6.85, fig.height = 5, fig.cap = "", purl = FALSE}
+```
+
+```{r combined-plot-save, eval = params$output_type == "ms", purl = FALSE}
+# Save plot to file if knitting MS for submission
+combined_plot_name <- figure("combined") %>% str_remove_all("[:punct:]| ") %>% paste0(".pdf")
+ggsave(plot = combined_plot, file = here::here("results/ms", combined_plot_name), width = 6.85, height = 5)
+```
+
+```{r combined-plot-print, fig.width = 6.85, fig.height = 5, fig.cap = figure("combined"), eval = params$output_type == "preprint", purl = FALSE}
+# Print plot if knitting MS to preprint
+combined_plot
+```
+
+```{r corr-plot-make, fig.width = 6.85, fig.height = 5, fig.cap = "", purl = FALSE}
+# Make plot
+
 # Combine recovery, VPD, and growth habit into a single dataframe (wide format)
@@
 
-a + b + c + 
+corr_plot <- a + b + c + 
   guide_area() + 
@@
   theme(plot.tag = element_text(face = 'bold'))
-
-figname <- figure("corr") %>% str_remove_all("[:punct:]| ") %>% paste0(".pdf")
-ggsave(here::here("results/ms", figname), width = 6.85, height = 5)
-```
-
-**`r figure("corr")`**
+```
+
+```{r corr-plot-save, eval = params$output_type == "ms", purl = FALSE}
+# Save plot to file if knitting MS for submission
+corr_plot_name <- figure("corr") %>% str_remove_all("[:punct:]| ") %>% paste0(".pdf")
+ggsave(plot = corr_plot, file = here::here("results/ms", corr_plot_name), width = 6.85, height = 5)
+```
+
+```{r corr-plot-print, fig.width = 6.85, fig.height = 5, fig.cap = figure("corr"), eval = params$output_type == "preprint", purl = FALSE}
+# Print plot if knitting MS to preprint
+corr_plot
+```

### 2021-03-13 R/functions.R

@@ -265,8 +265,6 @@ parse_pam <- function (file, exclude_lines = NULL, ret_type = "lc", recalc_yield
 
 #' Parse datalogger data for a desiccation chamber
 #'
-#' FIXME: find the name of this type of datalogger
-#'
 #' @param file Path to datalogger data in CSV format
 #'
 #' @return Dataframe
@@ -983,7 +981,8 @@ calculate_mean_vpd_sporo <- function (community_matrix_raw, filmy_species, moore
     select(-generation) %>%
     left_join(select(traits, species, habit), by = "species") %>%
     assert(not_na, habit) %>%
-    # For sporophytes, treat "saxicolous" as "terrestrial"
+    # For sporophytes, treat "saxicolous" as "terrestrial", since they were collected
+    # on the ground (not a vertical rock surface)
     mutate(habit = case_when(
       str_detect(habit, regex("terr|saxi", ignore_case = TRUE)) ~ "terrestrial",
       str_detect(habit, regex("epi", ignore_case = TRUE)) ~ "epiphytic",

### 2021-03-13 _targets.R

@@ -224,7 +224,8 @@ tar_plan(
   tar_render(
     manuscript_pdf, 
     "ms/manuscript.Rmd", 
-    output_dir = here::here("results/ms")
+    output_dir = here::here("results/ms"),
+    params = list(output_type = "ms")
     ),
   
   # - docx for submission

### 2021-03-13 ms/manuscript.Rmd
Fix 'in press' for Florence
hacky but it works

@@
 Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
-A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Florence; @Murdock2003; @Nitta2011b; @Nitta2017; @Ranker2005] (\figs{ }`r figure_num("habit")`, `r s_figure_num("tree")`; all supplemental figures available in Online Resource 1).
+A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Florence in press; @Murdock2003; @Nitta2011b; @Nitta2017; @Ranker2005] (\figs{ }`r figure_num("habit")`, `r s_figure_num("tree")`; all supplemental figures available in Online Resource 1).
 The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.

### 2021-03-13 _targets.R

@@ -230,7 +230,6 @@ tar_plan(
   
   # - docx for submission
   # FIXME: some fixes need to be made manually for MS submission
-  # - change Florence to "in press" (in-line and refs)
   # - remove extra space from bib entries
   # - add missing table header rows
   ms_docx = latex2docx(

### 2021-03-13 clean_data.sh
Change order of data cleaning scripts

@@ -2,7 +2,10 @@
 
 # Run each data cleaning script in its own R session
 
+# Data files "filmy_growth_habit.csv" and those extracted by R/unzip_data.R must
+# already be present in data/
+
+Rscript R/clean_specimen_data.R
 Rscript R/clean_chamber_data.R 
 Rscript R/clean_dt_data.R
 Rscript R/clean_lc_data.R
-Rscript R/clean_specimen_data.R

### 2021-03-13 ms/manuscript.Rmd
Update MS
- fix various little things

@@
   bookdown::pdf_document2:
-    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
+    citation_package: default # Use pandoc default, otherwise csl can't be used for reference formatting
     toc: no
@@
 
+```{r first-page, results = "asis", eval = params$output_type == "ms", purl = FALSE}
+cat("
 **Corresponding author**  
@@
 Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8562, Japan  
-Phone: +81-81-4-7136-3681  
+Phone: +81-4-7136-3681  
 Fax: +81-4-7136-3680  
-email: joelnitta\@gmail.com
+email: joelnitta\\@gmail.com
 
@@
 
-\newpage
+\\newpage
+")
+```
 
@@
 
-\*\*\* NOTE TO REVIEWERS: The DOI for the dataset will not be active until the paper is accepted. Until then, this link can be used for the dataset https://figshare.com/s/d6349abf01a3756a5aae  \*\*\*
+\*\*\* NOTE TO READER: The DOI for the dataset will not be active until the paper is accepted. Until then, this link can be used for the dataset https://figshare.com/s/d6349abf01a3756a5aae  \*\*\*
 
@@
 In a few cases (particularly species with larger fronds, \eg, *C. apiifolia*), the NaCl chamber did not reach equilibrium humidity before 2 d (`r s_figure("chamber")`).
-Brief fluctuations in humidity were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).
+Brief fluctuations in humidity were observed when chambers were opened to insert or remove samples (`r s_figure("chamber")`).
 

### 2021-03-13 ms/si.Rmd
Format SI slightly differently for preprint

@@
   \newcommand{\elandscape}{\end{landscape}}
+params:
+  output_type: "ms"
 editor_options: 
@@
 
+```{r paper-info, results = "asis", eval = params$output_type == "ms", purl = FALSE}
+# Only print journal name in SI for submission
+cat("
 **Authors:** Nitta JH; Watkins JE; Holbrook NM; Wang TW; Davis CC
@@
 **Corresponding author:** Joel H. Nitta  
-Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan  
-Phone: +81-  
-Fax: +81-  
-email: joelnitta\@gmail.com
+Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8562, Japan  
+Phone: +81-4-7136-3681  
+Fax: +81-4-7136-3680  
+email: joelnitta\\@gmail.com
+")
+```
 
@@
 **`r s_figure("tree")`**. Maximum likelihood ultrametric phylogenetic tree including all filmy fern species from Moorea, French Polynesia.
-Tree obtained by extracting Hymenophyllaceae from the tree of Nitta et al. (2016) with the extract.clade function of the ape package in R (Paradis et al., 2004).
-Bootstrap support values >50% shown at nodes.
+Tree obtained by extracting Hymenophyllaceae from the tree of Nitta et al. (2017) with the ‘extract.clade’ function of the ‘ape’ package (Paradis et al. 2004).
+Bootstrap support values > 50 % shown at nodes.
 Growth habit indicated by colored dots. 
@@
 Recovery was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not recovery of 0%).
+All error bars shown are standard deviation unless otherwise mentioned.
 
@@
 ### Make full sets of light curve plots ###
+
+# Remove one sample that had only outlier points and no good measurements
+light_data_all_mod <-
+light_data_all %>%
+  group_by(light_id) %>%
+  mutate(all_outliers = all(outlier)) %>%
+  ungroup() %>%
+  filter(!all_outliers) %>%
+  select(-all_outliers)
+
 # - sporos in lab
 lc_plots_sporo_lab <- make_facet_lc_plot(
-  light_data = light_data_all %>% filter(generation == "sporophyte", condition == "lab"),
+  light_data = light_data_all_mod %>% filter(generation == "sporophyte", condition == "lab"),
   lc_model_fitted_data = filmy_lc_model_fitted_data %>% filter(generation == "sporophyte", condition == "lab"),
@@
 lc_plots_sporo_field <- make_facet_lc_plot(
-  light_data = light_data_all %>% filter(generation == "sporophyte", condition == "field"),
+  light_data = light_data_all_mod %>% filter(generation == "sporophyte", condition == "field"),
   lc_model_fitted_data = filmy_lc_model_fitted_data %>% filter(generation == "sporophyte", condition == "field"),
@@
 lc_plots_gameto <- make_facet_lc_plot(
-  light_data = light_data_all %>% filter(generation == "gametophyte"),
+  light_data = light_data_all_mod %>% filter(generation == "gametophyte"),
   lc_model_fitted_data = filmy_lc_model_fitted_data %>% filter(generation == "gametophyte"),
@@
 **a**, sporophytes; **b**, gametophytes.
-Title of each subplot indicates species followed by light curve ID number assigned to each sample.
+Title of each subplot indicates species followed by light curve ID number assigned to each sample (light curve IDs match those in filmy_light_curves.csv data file).
 Light curves fit using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (ETR), *x* is photosynthetic photon flux density (PPFD), *A* is the asymptote of the curve, and *k* is a slope parameter.
@@
 
-a + b + c + guide_area() + plot_layout(ncol = 2, guides = 'collect') + plot_annotation(tag_levels = "a")
+a + b + c + guide_area() + 
+  plot_layout(ncol = 2, guides = 'collect') + 
+  plot_annotation(tag_levels = "a") &
+  theme(plot.tag = element_text(face = 'bold'))
 ```

### 2021-03-13 R/functions.R

@@ -1211,11 +1211,12 @@ prepare_gameto_dt_groups <- function(filmy_dt, gameto_time_summary_2012, filmy_s
 #' Data written out as CSV with metadata in comments
 #'
 #' @param gameto_indiv_times Data with gametophyte DT data to include membership in treatment batches
+#' @param ms_type Type of manuscript: preprint or MS for submission to journal
 #' @param file_out Path to write file
 #'
 #' @return Path to output file
 #' 
-write_gameto_times <- function(gameto_indiv_times, file_out) {
+write_gameto_times <- function(gameto_indiv_times, ms_type = "ms", file_out) {
   
   # In order to include metadata as commented lines above CSV, first
   # write out to plain temporary CSV file,
@@ -1226,7 +1227,8 @@ write_gameto_times <- function(gameto_indiv_times, file_out) {
   fs::file_delete(temp_file)
   
   # Add metadata
-  gameto_indiv_times_lines <- c(
+  if(ms_type == "ms") {
+    gameto_indiv_times_lines <- c(
     "# Electronic supplementary materials: Online Resource 2",
     "# Title: Intergenerational niche differentiation in filmy ferns (Hymenophyllaceae)",
     "# Authors: Nitta JH; Watkins JE; Holbrook NM; Wang TW; Davis CC",
@@ -1242,6 +1244,25 @@ write_gameto_times <- function(gameto_indiv_times, file_out) {
     "# --- END COLUMN METADATA ---",
     gameto_indiv_times_lines
   )
+  } else if(ms_type == "preprint") {
+   gameto_indiv_times_lines <- c(
+    "# Electronic supplementary materials: Online Resource 2",
+    "# Title: Intergenerational niche differentiation in filmy ferns (Hymenophyllaceae)",
+    "# Authors: Nitta JH; Watkins JE; Holbrook NM; Wang TW; Davis CC",
+    "# Corresponding author: Joel H. Nitta (joelnitta@gmail.com)",
+    "# Contents: start and end times of desiccation tolerance test on gametophytes of filmy ferns from Moorea French Polynesia",
+    "# --- BEGIN COLUMN METADATA ---",
+    "# species: Species name",
+    "# individual: J.H. Nitta specimen collection number of gametophyte individual",
+    "# pre: start time of desiccation",
+    "# post: end time of desiccation",
[…]

### 2021-03-13 _targets.R

@@ -241,7 +241,7 @@ tar_plan(
     depends = manuscript_pdf
   ),
   
-  # Render SI: figures (ESM 1)
+  # Render SI: figures (ESM 1) for submission
   tar_render(
     si_pdf, 
     "ms/si.Rmd", 
@@ -249,9 +249,23 @@ tar_plan(
     output_dir = here::here("results/si")
     ),
   
-  # Render SI: tables (ESM 2)
+  # Render SI: figures (ESM 1) for pre-print
+  tar_render(
+    si_pdf_bioarxiv, 
+    "ms/si.Rmd", 
+    output_file = "ESM_1_preprint.pdf", 
+    params = list(output_type = "preprint"),
+    output_dir = here::here("results/si")
+  ),
+  
+  # Render SI: tables (ESM 2) for sumbission
   tar_file(
-    gameto_indiv_times_table,
-    write_gameto_times(gameto_indiv_times, "results/si/ESM_2.csv"))
+    gameto_indiv_times_table_ms,
+    write_gameto_times(gameto_indiv_times, ms_type = "ms", "results/si/ESM_2.csv")),
+  
+  # Render SI: tables (ESM 2) for pre-print
+  tar_file(
+    gameto_indiv_times_table_preprint,
+    write_gameto_times(gameto_indiv_times, ms_type = "preprint", "results/si/ESM_2_preprint.csv"))
   
 )

### 2021-03-16 ms/si.Rmd
Update SI
fix various small issues: units on axes, size of font for BS in tree, bold face for part labels, etc

@@
     labs(
-      y = "ETR<sub>max</sub> (µmol m<sup>-2</sup>·s<sup>-1</sup>)", 
+      y = "ETR (µmol m<sup>-2</sup>·s<sup>-1</sup>)", 
       x = "PPFD (µmol m<sup>-2</sup>·s<sup>-1</sup>)", 
@@
 
-**`r s_figure("dt-control")`**. Recovery of controls (samples kept on moist tissues) for desiccation tolerance test of filmy ferns from Moorea, French Polynesia.
+**`r s_figure("dt-control")`**. Recovery of controls (samples kept on moist tissues) during desiccation tolerance test of filmy fern sporophytes from Moorea, French Polynesia.
 
@@
 ggtree(phy) %<+% filmy_habit %<+% bs_tibble +
-  geom_text(aes(label=bootstrap), hjust=-.25, size = 3, color = "grey30") +
+  geom_text(aes(label=bootstrap), hjust=-.25, size = 2, color = "grey10") +
   geom_tippoint(aes(color = habit), position = position_nudge(x = 0, y = 0), size = 4) +
@@
   theme(
-    plot.margin = margin(t = 0, r = 2.2, b = 0, l = 0, unit = "in"),
+    plot.margin = margin(t = 0.1, r = 2.2, b = 0.1, l = 0.5, unit = "in"),
     legend.position = "bottom"
@@
 licl_temp_plot_2012 <- plot_chamber_temp(licl_chamber_data_2012, licl_time_range_2012) +
-  scale_y_continuous(limits = licl_temp_range) +
+  scale_y_continuous(limits = licl_temp_range, label = label_number(suffix = "\u00b0C", accuracy = 1)) +
   labs(title = "2012") +
@@
   plot_layout(ncol = 3, nrow = 3, guides = "collect") +
-  plot_annotation(title = "a") &
-  theme(plot.tag = element_text(face = "bold"))
+  plot_annotation(title = "**a**") &
+  theme(plot.title = element_markdown())
 ```
@@
 mgno3_temp_plot_2012 <- plot_chamber_temp(mgno3_chamber_data_2012, mgno3_time_range_2012) +
-  scale_y_continuous(limits = mgno3_temp_range) +
+  scale_y_continuous(limits = mgno3_temp_range, label = label_number(suffix = "\u00b0C", accuracy = 1)) +
   labs(title = "2012") +
@@
   plot_layout(ncol = 3, nrow = 3, guides = "collect") +
-  plot_annotation(title = "b") &
-  theme(plot.tag = element_text(face = "bold"))
+  plot_annotation(title = "**b**") &
+  theme(plot.title = element_markdown())
 ```
@@
   plot_layout(ncol = 3, nrow = 3, guides = "collect") +
-  plot_annotation(title = "c") &
-  theme(
-    plot.tag = element_text(face = "bold"),
+  plot_annotation(title = "**c**") &
+  theme(
+    plot.title = element_markdown(),
     legend.position = "bottom")
@@
     axis.title.x = element_blank()
-  )
+  ) +
+  blank_y_theme()
 
@@
     axis.title.x = element_blank()
-  ) +
-  blank_y_theme()
+  )
 
@@
 chamber_gameto_2013_rh_plot <-
-  plot_chamber_rh(chamber_gameto_2013, gameto_time_limits_2013)
+  plot_chamber_rh(chamber_gameto_2013, gameto_time_limits_2013) +
+  blank_y_theme()
 
@@
     legend.position = "none",
-    axis.title.x = element_blank()) +
-  blank_y_theme()
+    axis.title.x = element_blank())
 
@@
   labs(title = "2013") +
-  theme(plot.title = element_text(hjust = 0.5, size = 11))
+  theme(plot.title = element_text(hjust = 0.5, size = 11)) +
+  blank_y_theme()
 
@@
     plot.title = element_text(hjust = 0.5, size = 11),
-    axis.title.x = element_blank()) +
-  blank_y_theme()
+    axis.title.x = element_blank())
 
 # Combine into final plot
-chamber_gameto_2013_temp_plot + chamber_gameto_2012_temp_plot +
-  chamber_gameto_2013_rh_plot + chamber_gameto_2012_rh_plot + 
-  dt_times_gameto_2013_plot + dt_times_gameto_2012_plot + 
+chamber_gameto_2012_temp_plot + chamber_gameto_2013_temp_plot +
+  chamber_gameto_2012_rh_plot + chamber_gameto_2013_rh_plot + 
+  dt_times_gameto_2012_plot + dt_times_gameto_2013_plot + 
   plot_layout(ncol = 2, nrow = 3) +
-  plot_annotation(title = "d") &
-  theme(plot.tag = element_text(face = "bold"))
+  plot_annotation(title = "**d**") &
+  theme(plot.title = element_markdown())
 ```
@@
   facet_wrap(vars(habit, species)) +
-  scale_shape_manual(name = "Desiccation Time",
+  scale_shape_manual(name = "Desiccation time",
                      values = c(24, 21),
@@
   xlab("Recovery Time (hr)") +
-  ylab("Recovery (%)") +
+  ylab("Recovery") +
   theme_bw() +
@@
     panel.grid.major.x = element_blank(),
-    strip.text = element_markdown(),
-    axis.title.x = element_blank()
+    strip.text = element_markdown()
   )
@@
 **`r s_figure("dt-control")`**.
-Recovery of controls (samples kept on moist tissues) for desiccation tolerance test of filmy fern sporophytes from Moorea, French Polynesia.
+Recovery of controls (samples kept on moist tissues) during desiccation tolerance test of filmy fern sporophytes from Moorea, French Polynesia.
 Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation treatment (see Methods).
@@
   scale_x_continuous(label = scales::percent) +
-  scale_shape_manual(name = "Desiccation Intensity",
+  scale_shape_manual(name = "Desiccation intensity",
                      values = salt_shapes,
@@
                      labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
-  scale_fill_manual(name = "Desiccation Intensity",
+  scale_fill_manual(name = "Desiccation intensity",
                     values = salt_cols, 
@@
   labs(
-    x = "Relative Water Content",
+    x = "Relative water content",
     y = "Species",
@@
   geom_point(aes(shape = salt, fill = salt), position = pd) +
-  scale_linetype_manual(name = "Desiccation Time",
+  scale_linetype_manual(name = "Desiccation time",
                         values = c("solid", "dashed"),
@@
                         labels = c("15 days", "2 days")) +
-  scale_shape_manual(name = "Desiccation Intensity",
+  scale_shape_manual(name = "Desiccation intensity",
                      values = salt_shapes,
@@
 **`r s_figure("water-s")`**.
-Relative water content (RWC) of sporophytes during desiccation tolerance test.
-RWC was not measured for all combinations of species and desiccation treatments; only species with complete data shown.
+Relative water content (RWC) of sporophytes during desiccation tolerance (DT) test.
+RWC was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not RWC of zero).
+x-axis indicates condition of sample during DT test; times refer to amount of elapsed time since re-wetting.
 
@@
 Blue line indicates maximum modeled relative electron transport rate (ETR~max~).
-Red line indicates PPFD at 95 % saturation of ETR (PPFD~95%~).
+Red line indicates PPFD at 95 % of ETR~max~ (PPFD~95%~).
 
@@
 humile* were observed growing as both epiphytes and saxicoles, but are treated as epiphytes to distinguish them from exclusively saxicolous species.
+Growth habit abbreviations: HE, high elevation epiphyte; LE, low elevation epiphyte; S, saxicolous; T, terrestrial.
 

### 2021-04-28 ms/manuscript.Rmd
Fix subject area
Update MS: change address

@@
 Joel H. Nitta  
-Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8562, Japan  
+Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo 113-0032, Japan  
 Phone: +81-4-7136-3681  
@@
 **Subject area**  
-(1) Ecology/Ecophysiology/Environmental Biology
+(2) Ecology/Ecophysiology/Environmental Biology
 

### 2021-04-28 ms/manuscript.Rmd
Update MS: change title
Update MS: add linear model stats for climate

@@
 
-# Intergenerational niche differentiation in filmy ferns (Hymenophyllaceae)
+# Ecophysiological differentiation between life stages in filmy ferns (Hymenophyllaceae)
 
@@
 The dataset includes `r climate_n_days` days of data from `r climate_n_sites` sites (`r climate_n_loggers` dataloggers).
-Mean daily temperature on Moorea is strongly correlated with elevation, dropping \circa{ }6 °C from the lowest surveyed site (201 m) to the highest (1206 m); temperature does not differ significantly between epiphytic and terrestrial habitats [@Nitta2020].
+Mean daily temperature on Moorea is strongly correlated with elevation, dropping \circa{ }6 °C from the lowest surveyed site (201 m) to the highest (1206 m) (*R*^2^ = 0.87; \pval{ }< 0.001; linear model); temperature does not differ significantly between epiphytic and terrestrial habitats [@Nitta2020].
 Epiphytic habitats tend to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones.

### 2021-04-28 ms/manuscript.Rmd
Update MS: change 'generation' to 'life stage'

@@
 Gametophytes had similar or less desiccation tolerance (ability to recover from 2 d desiccation at -86 MPa) and lower photosynthetic optima (maximum electron transport rate of photosystem II and light level at 95% of that rate) than sporophytes.
-Epiphytes were more tolerant of desiccation than terrestrial species in both generations.
+Epiphytes were more tolerant of desiccation than terrestrial species in both life stages.
 Despite their lack of greater physiological tolerances, gametophytes of several species occurred over a wider elevational range than conspecific sporophytes.
@@
 Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@Pittermann2013; @WatkinsJr.2007c].
-Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.
+Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life stages within a single organism.
 
@@
 Furthermore, the diversity of growth forms and habitats in filmy ferns makes them an ideal group for a comparative study of water-stress strategies (\ie, desiccation tolerance vs.
-desiccation avoidance) and the correlation of these strategies with ecological niche.
+desiccation avoidance) and their association with different ecological niches (\ie, along gradients of temperature and relative humidity).
 
 The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
-In a series of elegant experiments on Jamaican filmy ferns, @Shreve1911 demonstrated that water-use strategies differ between species occupying contrasting habitats (\ie, terrestrial, low elevation epiphyte, and high elevation epiphyte).
+In a series of elegant experiments on Jamaican filmy ferns, @Shreve1911 demonstrated that water-use strategies differ between species with contrasting growth habits including terrestrial, low elevation epiphyte, and high elevation epiphyte.
 Proctor [-@Proctor2003; -@Proctor2012] investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs.
@@
 
-Despite this considerable interest in filmy fern ecophysiology, with the sole exception of @Johnson2000, no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
+Despite this considerable interest in filmy fern ecophysiology, with the sole exception of @Johnson2000, no studies have compared physiology between gametophyte and sporophyte stages, and we are aware of none that have done so across multiple species.
 Furthermore, to our knowledge, no physiological studies of filmy ferns have accounted for their phylogenetic relationships, and we lack an understanding of how physiological strategies have evolved in this clade.
@@
 We used general linear mixed models (GLMMs) to investigate the effect of growth habit and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at -86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the \package{MCMCglmm} package [@Hadfield2010].
-For each of the three response variables (species mean DT, ETR~max~, and PPFD~95%~), we constructed five models: a null model with no fixed effects, or models with either growth habit (high-elevation epiphyte, low-elevation epiphyte, terrestrial or saxicolous), generation (sporophyte vs.
-gametophyte), growth habit and generation, or growth habit, generation, and their interaction as fixed effects.
+For each of the three response variables (species mean DT, ETR~max~, and PPFD~95%~), we constructed five models: a null model with no fixed effects, or models with either growth habit (high-elevation epiphyte, low-elevation epiphyte, terrestrial or saxicolous), life stage (sporophyte vs.
+gametophyte), growth habit and life stage, or growth habit, life stage, and their interaction as fixed effects.
 We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix if the response variable showed evidence of phylogenetic signal [@Hadfield2010].
@@
 
-For DT, *K* was near 1 with *P* < 0.1 in both generations (gametophytes *K* = `r gameto_recover_k`, *P* = `r gameto_recover_k_pval`; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval`); λ was near 1 for both generations, but not significant for sporophytes (gametophytes λ = `r gameto_recover_lambda`, *P* = `r gameto_recover_lambda_pval`; sporophytes λ = `r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`).
-There was no evidence of significant phylogenetic signal in photosynthetic parameters (PPFD~95%~ and ETR~max~) in either generation (*P* >> 0.05, `r table("physig")`).
+For DT, *K* was near 1 with *P* < 0.1 in both life stages (gametophytes *K* = `r gameto_recover_k`, *P* = `r gameto_recover_k_pval`; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval`); λ was near 1 for both life stages, but not significant for sporophytes (gametophytes λ = `r gameto_recover_lambda`, *P* = `r gameto_recover_lambda_pval`; sporophytes λ = `r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`).
+There was no evidence of significant phylogenetic signal in photosynthetic parameters (PPFD~95%~ and ETR~max~) in either life stage (*P* >> 0.05, `r table("physig")`).
 Therefore, phylogenetic methods were used for analyses involving DT, but not photosynthetic parameters.
@@
 
-Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases.
+Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life stages.
 
@@
 
-The GLMM including the combined effects of generation plus growth habit and their interaction was selected as the best-fitting model for DT (`r table("glmm")`). 
+The GLMM including the combined effects of life stage plus growth habit and their interaction was selected as the best-fitting model for DT (`r table("glmm")`). 
 Sporophytes had `r sporo_dt_effect_mean` greater DT (95 % CI `r sporo_dt_effect_lower` to `r sporo_dt_effect_upper`) relative to gametophytes (`r table("glmm-pars")`).
@@
 
-Light responses were measured in sporophytes of `r n_light_species_sporo` species and gametophytes of `r n_light_species_gameto` species, including `r n_light_species_both` species with both life phases.
-
-Photosynthetic optima showed clear differences across generations: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (\figs{ }`r figure_num("combined")`, `r s_figure_num("light-curves")`, `r s_figure_num("traits-box")`).
+Light responses were measured in sporophytes of `r n_light_species_sporo` species and gametophytes of `r n_light_species_gameto` species, including `r n_light_species_both` species with both life stages.
+
+Photosynthetic optima showed clear differences across life stages: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (\figs{ }`r figure_num("combined")`, `r s_figure_num("light-curves")`, `r s_figure_num("traits-box")`).
 Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
@@
 
-The GLMM including the effect of generation only was selected as the best-fitting model for ETR~max~, and the combined effect of growth habit plus generation and their interaction was selected for PPFD~95%~ (`r table("glmm")`).
+The GLMM including the effect of life stage only was selected as the best-fitting model for ETR~max~, and the combined effect of growth habit plus life stage and their interaction was selected for PPFD~95%~ (`r table("glmm")`).
 
@@
 Vegetative DT was a key innovation linked to the rise of early land plants, but is largely limited to non-vascular groups such as bryophytes and algae among extant plants.
-Filmy ferns have received considerable attention for their remarkable reversion to vegetative DT in the sporophyte phase [\eg, @GarcesCea2014; @Ostria-Gallardo2020; @Proctor2003; @Shreve1911].
+Filmy ferns have received considerable attention for their remarkable reversion to vegetative DT in the sporophyte stage [\eg, @GarcesCea2014; @Ostria-Gallardo2020; @Proctor2003; @Shreve1911].
 Yet, no previous study to our knowledge has evaluated parallel physiology of sporophytes and gametophytes of filmy ferns in relation to DT.
-Here, we present the first comparisons of physiological parameters between generations within filmy ferns in a phylogenetic comparative framework.
+Here, we present the first comparisons of physiological parameters between life stages within filmy ferns in a phylogenetic comparative framework.
 
@@
 
-We observed a clear intergenerational difference in photosynthetic optima, with gametophytes consistently adapted for lower light levels than sporophytes (`r figure("combined")`, `r table("glmm")`).
+We observed a clear difference between life stages in photosynthetic optima, with gametophytes consistently adapted for lower light levels than sporophytes (`r figure("combined")`, `r table("glmm")`).
 In the only other study to our knowledge that compared photosynthetic rates between gametophytes and sporophytes of a filmy fern, @Johnson2000 also found that gametophytes of *V.
@@
 It is surprising, however, that we did not observe any consistent differences between species in different habitats, given that light levels should vary between tree trunks and the forest floor, at least on a \circa{ }1 m scale.
-Although the GLMM showed an interaction effect between generation and growth habit for PPFD~95%~, the model including only generation was selected for ETR~max~ (`r table("glmm")`).
+Although the GLMM showed an interaction effect between life stage and growth habit for PPFD~95%~, the model including only life stage was selected for ETR~max~ (`r table("glmm")`).
 It is possible that the actual light levels experienced by gametophytes growing in these habitats may not vary significantly if they occupy protected microsites such as bryophyte mats or crevices in rock or bark.
@@
 We detected phylogenetic signal in DT (`r table("physig")`), suggesting that phylogeny should be considered when interpreting differences in DT between species.
-In contrast, the lack of phylogenetic signal in photosynthetic traits, coupled with the clear intergenerational differences we observed, suggest that sporophytes and gametophytes are adapted for different light levels regardless of clade or growth habit.
+In contrast, the lack of phylogenetic signal in photosynthetic traits, coupled with the clear differences in traits between life stages, suggest that sporophytes and gametophytes are adapted for different light levels regardless of clade or growth habit.
 In a more densely sampled study of morphological evolution in the trichomanoid clade, @Dubuisson2013 identified several morphological changes including reduced stele and root systems associated with the evolution of epiphytic growth within the HE subclade (comprising the genera *Polyphlebium*, *Didymoglossum*, *Crepidomanes*, and *Vandenboschia*).
@@
 
-In filmy ferns, both sporophyte and gametophyte generations are poikilohydric; however, this does not necessarily mean they share similar physiological optima.
+In filmy ferns, both sporophyte and gametophyte life stages are poikilohydric; however, this does not necessarily mean they share similar physiological optima.
 Although filmy fern sporophytes are often of small stature, elevating a leaf even a few cm above the substrate surface means escape from the boundary layer and exposure to substantially greater evapotranspiration.
-It is likely that even in this clade, wherein the physiological divide between sporophyte and gametophyte is probably the smallest amongst ferns, the two generations experience substantially different selective pressures.
+It is likely that even in this clade, wherein the physiological divide between sporophyte and gametophyte is probably the smallest amongst ferns, the two life stages experience substantially different selective pressures.
 Fossils indicate that some of the earliest vascular plants may have had sporophytes and gametophytes that were both branched and small [@Kenrick1997a], but it is unclear at what point selective pressures began to diverge to favor large, complex sporophytes and small, simple gametophytes.
@@
 
-**`r table("physig")`** Results of tests for phylogenetic signal (Blomberg’s *K* and Pagel’s λ) in desiccation tolerance (DT) and photosynthetic parameters of filmy ferns from Moorea, French Polynesia by generation
+**`r table("physig")`** Results of tests for phylogenetic signal (Blomberg’s *K* and Pagel’s λ) in desiccation tolerance (DT) and photosynthetic parameters of filmy ferns from Moorea, French Polynesia by life stage
 
@@
 
-**`r table("glmm")`** Deviance information criterion (DIC) and delta DIC for general linear mixed models (GLMMs) of photosynthetic parameters and desiccation tolerance in response to generation and growth habit in filmy ferns from Moorea, French Polynesia
+**`r table("glmm")`** Deviance information criterion (DIC) and delta DIC for general linear mixed models (GLMMs) of photosynthetic parameters and desiccation tolerance in response to life stage and growth habit in filmy ferns from Moorea, French Polynesia
 
@@
     fixed_effects = case_when(
-      fixed_effects == "both" ~ "Generation + habit",
-      fixed_effects == "intersect" ~ "Generation + habit + Generation:habit",
+      fixed_effects == "both" ~ "Life stage + habit",
+      fixed_effects == "intersect" ~ "Life stage + habit + Life stage:habit",
       TRUE ~ str_to_sentence(fixed_effects)
@@
 
-**`r table("glmm-pars")`** Predictor effects, lower and upper 95% confidence interval (CI), effective sample size, and significance of model parameters for the best general linear mixed models (GLMMs) of photosynthetic parameters and desiccation tolerance in response to generation and growth habit in filmy ferns from Moorea, French Polynesia
+**`r table("glmm-pars")`** Predictor effects, lower and upper 95% confidence interval (CI), effective sample size, and significance of model parameters for the best general linear mixed models (GLMMs) of photosynthetic parameters and desiccation tolerance in response to life stage and growth habit in filmy ferns from Moorea, French Polynesia
 
@@
 
-Effect of generation is relative to gametophytes; effect of growth habit is relative to high elevation epiphytes
+Effect of life stage is relative to gametophytes; effect of growth habit is relative to high elevation epiphytes
 
@@
 *Crepidomanes bipunctatum* and *C. humile* were observed growing as both epiphytes and saxicoles, but are shown as epiphytes to distinguish them from exclusively saxicolous species.
-Asterisk indicates significant difference in means (*P* <0.05) between generations for a particular species (*t*-test); ns indicates no significant difference; lack of ns or asterisk means not enough observations available for that species for *t*-test.
+Asterisk indicates significant difference in means (*P* <0.05) between life stages for a particular species (*t*-test); ns indicates no significant difference; lack of ns or asterisk means not enough observations available for that species for *t*-test.
 Left bar plot shows recovery (%) of chlorophyll fluorescence (F~v~/F~m~) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa). 
@@
   # theme_bw(base_size = 10) +
-  labs(x = "Recovery", fill = "Generation") +
+  labs(x = "Recovery", fill = "Life stage") +
   theme(
@@
   scale_x_continuous(expand = expansion(mult = c(0,0.025))) +
-  labs(x = "ETR<sub>max</sub>", fill = "Generation") +
+  labs(x = "ETR<sub>max</sub>", fill = "Life stage") +
   theme(
@@
   scale_x_continuous(expand = expansion(mult = c(0,0.025))) +
-  labs(x = "PPFD<sub>95%</sub>", fill = "Generation") +
+  labs(x = "PPFD<sub>95%</sub>", fill = "Life stage") +
   theme(

### 2021-05-06 R/clean_specimen_data.R
Specify packages without using source()

@@ -1,7 +1,10 @@
 # Clean specimen collection data
 
 # Load packages
-source("R/packages.R")
+library(readxl)
+library(tidyverse)
+library(janitor)
+library(glue)
 
 # Load functions
 source("R/functions.R")

### 2021-05-07 R/clean_specimen_data.R
Fix missing dates in specimens.csv
from 'NA-NA-NA' to 'NA'

@@ -1,10 +1,7 @@
 # Clean specimen collection data
 
 # Load packages
-library(readxl)
-library(tidyverse)
-library(janitor)
-library(glue)
+source(here::here("R/packages.R"))
 
 # Load functions
 source("R/functions.R")
@@ -118,7 +115,7 @@ fern_specimens_gps_raw <- read_csv("data_raw/specimens.csv", col_types = specime
     year = year(date_collected),
     month = month(date_collected) %>% str_pad(side = "left", pad = "0", width = 2),
     day = day(date_collected) %>% str_pad(side = "left", pad = "0", width = 2),
-    date_collected = paste(year, month, day, sep = "-")
+    date_collected = paste3(year, month, day, sep = "-")
   ) %>%
   # Add taxon
   mutate(taxon = paste3(genus, specific_epithet, infraspecific_name)) %>%

### 2021-05-07 R/functions.R
Update ESM2

@@ -1230,7 +1230,7 @@ write_gameto_times <- function(gameto_indiv_times, ms_type = "ms", file_out) {
   if(ms_type == "ms") {
     gameto_indiv_times_lines <- c(
     "# Electronic supplementary materials: Online Resource 2",
-    "# Title: Intergenerational niche differentiation in filmy ferns (Hymenophyllaceae)",
+    "# Title: Ecophysiological differentiation between life stages in filmy ferns (Hymenophyllaceae)",
     "# Authors: Nitta JH; Watkins JE; Holbrook NM; Wang TW; Davis CC",
     "# Journal: Journal of Plant Research",
     "# Corresponding author: Joel H. Nitta (joelnitta@gmail.com)",
@@ -1247,7 +1247,7 @@ write_gameto_times <- function(gameto_indiv_times, ms_type = "ms", file_out) {
   } else if(ms_type == "preprint") {
    gameto_indiv_times_lines <- c(
     "# Electronic supplementary materials: Online Resource 2",
-    "# Title: Intergenerational niche differentiation in filmy ferns (Hymenophyllaceae)",
+    "# Title: Ecophysiological differentiation between life stages in filmy ferns (Hymenophyllaceae)",
     "# Authors: Nitta JH; Watkins JE; Holbrook NM; Wang TW; Davis CC",
     "# Corresponding author: Joel H. Nitta (joelnitta@gmail.com)",
     "# Contents: start and end times of desiccation tolerance test on gametophytes of filmy ferns from Moorea French Polynesia",

### 2021-05-07 ms/data_readme.Rmd
Change date of data readme creation
Update data README Rmd

@@
 
-Title of Dataset: Data from: Intergenerational niche differentiation in filmy
-ferns (Hymenophyllaceae)
+Title of Dataset: Data from: Ecophysiological differentiation between life stages in filmy ferns (Hymenophyllaceae)
 
@@
 
-Department of Integrated Biosciences, Graduate School of Frontier Sciences,
-The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8562, Japan
+Department of Biological Sciences, Graduate School of Science, The University 
+of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo 113-0032, Japan 
 joelnitta@gmail.com
@@
 Recommended citation for the data: Nitta JH, Watkins JE, Holbrook NM, Wang TW,
-Davis CC (2021) Data from: Intergenerational niche differentiation in filmy
-ferns (Hymenophyllaceae) https://doi.org/10.6084/m9.figshare.14184572
+Davis CC (2021) Data from: Ecophysiological differentiation between life stages 
+in filmy ferns (Hymenophyllaceae) https://doi.org/10.6084/m9.figshare.14184572
 
 Citation for and links to publications that cite or use the data: Nitta JH,
-Watkins JE, Holbrook NM, Wang TW, Davis CC (2021) Intergenerational niche
-differentiation in filmy ferns (Hymenophyllaceae) <<FIXME: ADD JOURNAL>>
+Watkins JE, Holbrook NM, Wang TW, Davis CC (2021) Ecophysiological differentiation 
+between life stages in filmy ferns (Hymenophyllaceae) (Hymenophyllaceae) <<FIXME: ADD JOURNAL>>
 
@@
   fluorometer.
+
+Checksums are 32-byte MD5 hashes generated with tools::md5sum() in R.
 
@@
 
+MD5 checksum: `r tools::md5sum(here::here(gameto_times_2012_file))`
+
 \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
@@
 
+MD5 checksum: `r tools::md5sum(here::here(sporo_dt_times_file))`
+
 \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
@@
 'clean_' in the 'R' folder).
+
+MD5 checksum: `r tools::md5sum(here::here("data/data_raw.tar.gz"))`
 
@@
 
+MD5 checksum: `r tools::md5sum(here::here(specimens_raw_file))`
+
 \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
@@
 
+MD5 checksum: `r tools::md5sum(here::here(filmy_dt_file))`
+
 \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
@@
 chamber (see methods above for corresponding relative humidity).
-- generation: Gametophyte or sporophyte, or "mixed" if samples of both generations
+- generation: Gametophyte or sporophyte, or "mixed" if samples of both life stages
 were present in the growth chamber.
@@
 Specialized formats or other abbreviations used: None.
+
+MD5 checksum: `r tools::md5sum(here::here(filmy_dt_chamber_file))`
 
@@
 Specialized formats or other abbreviations used: None.
+
+MD5 checksum: `r tools::md5sum(here::here(filmy_habit_file))`
 
@@
 
+MD5 checksum: `r tools::md5sum(here::here(light_data_file))`
+
 \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
@@
 
+2021-05-07
+
+- Update specimens.csv in raw data.
+
 \-\-\-
 
-2021-03-14
+2021-05-06
+
+- Add MD5 checksums.
+- Correct missing values in 'date_collected' column of fern_specimens.csv from 'NA-NA-NA'
+to 'NA'.
+- Change title of paper.
+- Change address of first author.
+- Change description of "generation" to "life stage".
+
+\-\-\-
+
+2021-03-13
 

### 2021-05-07 ms/manuscript.Rmd
Update MS
Update MS
fix address
fix note to readers/reviewers
fix generation/life stage

@@
 
-^4^Current affiliation: Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8562, Japan
+^4^Current affiliation: Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo 113-0032, Japan
 
@@
 
-\*\*\* NOTE TO READER: The DOI for the dataset will not be active until the paper is accepted. Until then, this link can be used for the dataset https://figshare.com/s/d6349abf01a3756a5aae  \*\*\*
+\*\*\* NOTE TO `r ifelse(params$output_type == "ms", "EDITOR AND REVIEWERS", "READER")`: The DOI for the dataset will not be active until the paper is accepted.
+Until then, this link can be used for the dataset https://figshare.com/s/d6349abf01a3756a5aae  \*\*\*
 
@@
 Weston Testo provided advice on settings for the DT test.
+Two anonymous reviewers provided feedback that improved the manuscript.
 Staff at the University of California Berkeley's Richard B.
@@
 
+```{r esm-statement, results = "asis", eval = params$output_type == "ms", purl = FALSE}
+# FIXME: add DOI when available
+cat("
 **Electronic supplementary material**
@@
 The online version of this article (DOI) contains supplementary material, which is available to authorized users.
+")
+```
 
@@
     fixed_effects = case_when(
+      fixed_effects == "generation" ~ "Life stage",
       fixed_effects == "both" ~ "Life stage + habit",
@@
   c(
-    "Generation" = "generation", 
+    "Life stage" = "generation", 
     "Predictor" = "predictor",
@@
 (**d**) *Callistopteris apiifolia*, a terrestrial species with sporophytes occurring only in humid cloud forests. *Nitta 3990* (GH).
-(**e**) *C. apiifolia* (ribbon morphotype). Arrow indicates gemmae (asexual propagules) at tip of lobe. *Nitta 2565* (GH). 
+(**e**) *Callistopteris apiifolia* (ribbon morphotype). Arrow indicates gemmae (asexual propagules) at tip of lobe. *Nitta 2565* (GH). 
 (**f**) *Hymenophyllum polyanthos* (ribbon morphotype). *Nitta 3862* (GH).
 (**g**) Filamentous gametophytes of an unidentified species of Hymenophyllaceae. *Nitta 4036* (GH). 
-(**h**) *C. apiifolia*, showing detail of gemmifers (attachment points of gemmae, example indicated by arrow) and branched gemmae. *Nitta 3887* (GH). 
+(**h**) *Callistopteris apiifolia*, showing detail of gemmifers (attachment points of gemmae, example indicated by arrow) and branched gemmae. *Nitta 3887* (GH). 
 Scalebars: (**a**–**c**) 1.0 cm; (**d**) 5.0 cm; (**e**) 5.0 mm; (**f**) 2.0 mm; (**g**, **h**) 1.0 mm. 

### 2021-05-07 ms/si.Rmd
Update SI
remove minor gridlines
fix generation / life stage

@@
 
-**Title:** Intergenerational niche differentiation in filmy ferns (Hymenophyllaceae)
+**Title:** Ecophysiological differentiation between life stages in filmy ferns (Hymenophyllaceae)
 
@@
 **Corresponding author:** Joel H. Nitta  
-Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8562, Japan  
+Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo 113-0032, Japan  
 Phone: +81-4-7136-3681  
@@
   plot_annotation(title = "**a**") &
-  theme(plot.title = element_markdown())
+  theme(plot.title = element_markdown(), panel.grid.minor = element_blank())
 ```
@@
   plot_annotation(title = "**b**") &
-  theme(plot.title = element_markdown())
+  theme(plot.title = element_markdown(), panel.grid.minor = element_blank())
 ```
@@
     plot.title = element_markdown(),
-    legend.position = "bottom")
+    legend.position = "bottom",
+    panel.grid.minor = element_blank()
+  )
 ```
@@
   plot_annotation(title = "**d**") &
-  theme(plot.title = element_markdown())
+  theme(plot.title = element_markdown(), panel.grid.minor = element_blank())
 ```
@@
     panel.grid.major.x = element_blank(),
+    panel.grid.minor = element_blank(),
     strip.text = element_markdown(),
@@
       x = "Growth Habit",
-      fill = "Generation"
+      fill = "Life stage"
     )
@@
   plot_annotation(tag_levels = "a") &
-  theme(plot.tag = element_text(face = 'bold'))
+  theme(plot.tag = element_text(face = 'bold'), panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())
 ```

### 2021-06-02 ms/manuscript.Rmd
Update MS
in response to comments from JPRE editor 2021-05-25

@@
 
-**Keywords**: chlorophyll fluorescence; desiccation tolerance; DNA barcoding; French Polynesia; gametophyte; Hymenophyllaceae
+**Keywords**: Chlorophyll fluorescence; Desiccation tolerance; DNA barcoding; French Polynesia; Gametophyte; Hymenophyllaceae
 
@@
 Filmy fern gametophytes are similarly diverse, with morphotypes including filamentous, ribbon, and a combination of these two types (`r figure("habit")`e--h).
-While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages [@Pittermann2013; @WatkinsJr.2007a; @WatkinsJr.2012], and may be expected to occur in filmy ferns as well.
+While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages [@Pittermann2013; @WatkinsJr.2012; @WatkinsJr.2007a], and may be expected to occur in filmy ferns as well.
 The possibility for varying degrees of vegetative DT in both stages of the life cycle of filmy ferns makes them a potentially useful system to compare how DT is utilized between haploid, non-vascular and diploid, vascular life stages.
@@
 This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca.
-500 µmol·m^-2^·s^-1^.
+500 µmol m^-2^ s^-1^.
 Response curves were fitted using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (RETR), *x* is photosynthetic photon flux density (PPFD), *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded [@Proctor2012].
@@
 The \package{ggplot2} [@Wickham2016], \package{ggtree}, [@Yu2017] and \package{patchwork} (https://github.com/thomasp85/patchwork) packages were used to generate the figures.
-
-\*\*\* NOTE TO `r ifelse(params$output_type == "ms", "EDITOR AND REVIEWERS", "READER")`: The DOI for the dataset will not be active until the paper is accepted.
-Until then, this link can be used for the dataset https://figshare.com/s/d6349abf01a3756a5aae  \*\*\*
 
@@
 Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
-`r mean_par_sporo` µmol·m^--2^ ·s^--1^ (*n* = `r mean_par_sporo_n` species), whereas gametophytes did so at `r mean_par_gameto` µmol·m^--2^ ·s^--1^ (*n* = `r mean_par_gameto_n` species).
-Similarly, sporophytes attained higher rates of electron transport (ETR~max~) (`r mean_etr_sporo` µmol·m^--2^ ·s^--1^, *n* = `r mean_etr_sporo_n` species) than gametophytes (`r mean_etr_gameto` µmol·m^--2^ ·s^--1^, *n* = `r mean_etr_gameto_n` species).
+`r mean_par_sporo` µmol m^--2^ s^--1^ (*n* = `r mean_par_sporo_n` species), whereas gametophytes did so at `r mean_par_gameto` µmol m^--2^ s^--1^ (*n* = `r mean_par_gameto_n` species).
+Similarly, sporophytes attained higher rates of electron transport (ETR~max~) (`r mean_etr_sporo` µmol m^--2^ s^--1^, *n* = `r mean_etr_sporo_n` species) than gametophytes (`r mean_etr_gameto` µmol m^--2^ s^--1^, *n* = `r mean_etr_gameto_n` species).
 
@@
 In the only other study to our knowledge that compared photosynthetic rates between gametophytes and sporophytes of a filmy fern, @Johnson2000 also found that gametophytes of *V.
-speciosa* reached maximum ETR at lower levels than sporophytes: in gametophytes of this species, photosynthesis became saturated at a PPFD of 4--5 µmol·m^-2^·s^-1^ in the field or 30--50 µmol·m^-2^·s^-1^ in the lab vs.
-30--50 µmol·m^-2^ ·s^-1^ for sporophytes (lab measurement only).
-Although we found similar differences in direction between gametophytes and sporophytes, the measurements of @Johnson2000 are considerably lower than what we observed in the field for both Moorean filmy fern gametophytes (PPFD~95%~ \circa{ }40--60 µmol·m^-2^·s^-1^) and sporophytes (PPFD~95%~ \circa{ }130--160 µmol·m^-2^·s^-1^).
+speciosa* reached maximum ETR at lower levels than sporophytes: in gametophytes of this species, photosynthesis became saturated at a PPFD of 4--5 µmol m^-2^ s^-1^ in the field or 30--50 µmol m^-2^ s^-1^ in the lab vs.
+30--50 µmol m^-2^ s^-1^ for sporophytes (lab measurement only).
+Although we found similar differences in direction between gametophytes and sporophytes, the measurements of @Johnson2000 are considerably lower than what we observed in the field for both Moorean filmy fern gametophytes (PPFD~95%~ \circa{ }40--60 µmol m^-2^ s^-1^) and sporophytes (PPFD~95%~ \circa{ }130--160 µmol m^-2^ s^-1^).
 This makes sense in light of the fact that *V.
-speciosa* is an extremely deep-shade plant, with gametophytes occupying protected rock crevices that receive less than 1 µmol·m^-2^·s^-1^ PPFD for most of the day [@Johnson2000]; many of the gametophytes sampled in our study occur in more exposed conditions.
+speciosa* is an extremely deep-shade plant, with gametophytes occupying protected rock crevices that receive less than 1 µmol m^-2^ s^-1^ PPFD for most of the day [@Johnson2000]; many of the gametophytes sampled in our study occur in more exposed conditions.
 It is surprising, however, that we did not observe any consistent differences between species in different habitats, given that light levels should vary between tree trunks and the forest floor, at least on a \circa{ }1 m scale.
@@
 Left bar plot shows recovery (%) of chlorophyll fluorescence (F~v~/F~m~) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa). 
-Middle bar plot shows maximum relative electron transport rate of photosystem II (ETR~max~) (µmol e-·m^-2^·s^-1^). 
-Right bar plot shows photosynthetic photon flux density at 95% saturation of RETR (PPFD~95%~) (µmol quanta·m^-2^·s^-1^).
+Middle bar plot shows maximum relative electron transport rate of photosystem II (ETR~max~) (µmol e- m^-2^ s^-1^). 
+Right bar plot shows photosynthetic photon flux density at 95% saturation of RETR (PPFD~95%~) (µmol quanta m^-2^ s^-1^).
 Absence of a bar indicates measurements not available for gametophytes or sporophytes of that species.

### 2021-06-09 ms/manuscript.Rmd
Change 'd' to 'days'

@@
 We conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.
-Gametophytes had similar or less desiccation tolerance (ability to recover from 2 d desiccation at -86 MPa) and lower photosynthetic optima (maximum electron transport rate of photosystem II and light level at 95% of that rate) than sporophytes.
+Gametophytes had similar or less desiccation tolerance (ability to recover from 2 days desiccation at -86 MPa) and lower photosynthetic optima (maximum electron transport rate of photosystem II and light level at 95% of that rate) than sporophytes.
 Epiphytes were more tolerant of desiccation than terrestrial species in both life stages.
@@
 Any samples with pre-treatment F~v~/F~m~ < 0.4 were excluded from analysis, as typical values of F~v~/F~m~ in healthy plants tend to be \circa{ }0.8 [@Bjorkman1987].
-Samples were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
+Samples were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 days) or long (15 days) interval [@Testo2013].
 Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012 field season for species other than *Callistopteris apiifolia* (Presl) Copel.) or Hobo ProV2 dataloggers logging every 5 min (2012 *C.
@@
 Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible.
-We therefore applied the same treatment for all gametophytes (2 d at -86 MPa) in batches of individuals (`r s_table("gameto-batches")`), and verified the species of each individual after the experiment using the DNA barcode approach described above.
+We therefore applied the same treatment for all gametophytes (2 days at -86 MPa) in batches of individuals (`r s_table("gameto-batches")`), and verified the species of each individual after the experiment using the DNA barcode approach described above.
 Thus, gametophytes of some species were sampled many times whereas those of other species were not.
@@
 
-We used general linear mixed models (GLMMs) to investigate the effect of growth habit and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at -86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the \package{MCMCglmm} package [@Hadfield2010].
+We used general linear mixed models (GLMMs) to investigate the effect of growth habit and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at -86 MPa for 2 days) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the \package{MCMCglmm} package [@Hadfield2010].
 For each of the three response variables (species mean DT, ETR~max~, and PPFD~95%~), we constructed five models: a null model with no fixed effects, or models with either growth habit (high-elevation epiphyte, low-elevation epiphyte, terrestrial or saxicolous), life stage (sporophyte vs.
@@
 apiifolia* showed slightly greater ability to recover by 48 h compared to *Abrodictyum dentatum* (Bosch) Ebihara & K.
-Iwats., which completely failed to recover from even the gentlest treatment of 2 d at --38 NaCl (more stringent tests were not conducted on *A.
+Iwats., which completely failed to recover from even the gentlest treatment of 2 days at --38 NaCl (more stringent tests were not conducted on *A.
 dentatum* after confirming its lack of DT at this level).
@@
 Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C.
-kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
-Nearly all epiphytic species were capable of recovering from desiccation of up to 2 d at --86 MPa.
+kurzii* was able to recover from desiccation up to --86 MPa for 2 days but not more.
+Nearly all epiphytic species were capable of recovering from desiccation of up to 2 days at --86 MPa.
 *Crepidomanes bipunctatum*, a widespread low-elevation epiphyte, showed the greatest ability to recover from desiccation across all treatments.
@@
 and *Hymenophyllum polyanthos* (Sw.) Sw.
-could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.
+could recover from desiccation at --86 or --282 MPa for 15 days, but not --38 MPa.
 
@@
 Control samples generally did not show any decrease in F~v~/F~m~ throughout the duration of the experiment (`r s_figure("dt-control")`) and maintained RWC near 100% (\figs{ }`r s_figure_num("water-all")`, `r s_figure_num("water-s")`).
-Relative water content in the desiccated state ranged from `r rel_water_table %>% filter(salt == "LiCl", dry_time == "15", rec_time == "desiccated") %>% pull(sum_val)` in the --282 MPa, 15 d treatment to `r rel_water_table %>% filter(salt == "NaCl", dry_time == "2", rec_time == "desiccated") %>% pull(sum_val)` in the --38 MPa, 2 d treatment (`r s_figure("water-all")`), and returned to \circa{ }100 % upon re-wetting in all species sampled (\figs{ }`r s_figure_num("water-all")`, `r s_figure_num("water-s")`).
+Relative water content in the desiccated state ranged from `r rel_water_table %>% filter(salt == "LiCl", dry_time == "15", rec_time == "desiccated") %>% pull(sum_val)` in the --282 MPa, 15 days treatment to `r rel_water_table %>% filter(salt == "NaCl", dry_time == "2", rec_time == "desiccated") %>% pull(sum_val)` in the --38 MPa, 2 days treatment (`r s_figure("water-all")`), and returned to \circa{ }100 % upon re-wetting in all species sampled (\figs{ }`r s_figure_num("water-all")`, `r s_figure_num("water-s")`).
 
 Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (`r s_figure("chamber")`).
-In a few cases (particularly species with larger fronds, \eg, *C. apiifolia*), the NaCl chamber did not reach equilibrium humidity before 2 d (`r s_figure("chamber")`).
+In a few cases (particularly species with larger fronds, \eg, *C. apiifolia*), the NaCl chamber did not reach equilibrium humidity before 2 days (`r s_figure("chamber")`).
 Brief fluctuations in humidity were observed when chambers were opened to insert or remove samples (`r s_figure("chamber")`).
@@
 moderate desiccation, with higher recovery values at -282 and -86 MPa vs.
--38 MPa when dry for 15 d.
+-38 MPa when dry for 15 days.
 This may be an adaptation to cloud forest conditions; although cloud forests are frequently moist due to constant fog, dry periods are intense during the times when there is no cloud cover (J.
@@
 Since gametophytes of *C.
-apiifolia* are common on Moorea and have a unique morphology [@Nitta2017], we were able to collect sufficient material of this species for additional tests of DT at 100% RH for 2 d.
+apiifolia* are common on Moorea and have a unique morphology [@Nitta2017], we were able to collect sufficient material of this species for additional tests of DT at 100% RH for 2 days.
 However, *C. apiifolia* gametophytes still failed to recover from even this gentle treatment (`r s_figure("capi-water")`).
@@
 
-Recovery, recovery of F~v~/F~m~ (%) following 2 d desiccation at –86 MPa; ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached.
+Recovery, recovery of F~v~/F~m~ (%) following 2 days desiccation at –86 MPa; ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached.
 
@@
 
-ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached; Recovery, recovery of F~v~/F~m~ (%) following 2 d desiccation at –86 MPa
+ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached; Recovery, recovery of F~v~/F~m~ (%) following 2 days desiccation at –86 MPa
 
@@
 
-**`r figure("corr")`** Recovery of F~v~/F~m~ (%) following 2 d desiccation at –86 MPa plotted against environmental vapor pressure deficit (VPD; kPa) (**a**, **b**) or range of gametophytes beyond sporophytes (m) (**c**) for gametophytes (**a**, **c**) or sporophytes (**b**).
+**`r figure("corr")`** Recovery of F~v~/F~m~ (%) following 2 days desiccation at –86 MPa plotted against environmental vapor pressure deficit (VPD; kPa) (**a**, **b**) or range of gametophytes beyond sporophytes (m) (**c**) for gametophytes (**a**, **c**) or sporophytes (**b**).
 Species means ± 1 SD are shown. 

### 2021-06-10 ms/manuscript.Rmd
Update MS

@@
 Correlated with these differences in functional strategies to cope with water stress is a major difference in life cycle.
-Over time, plants shifted away from wholly gametophyte-dominated life cycle (as in extant bryophytes) to co-independence of the gametophyte and sporophyte (\ie, ferns and lycophytes) to complete loss of the independent gametophyte (\ie, spermatophytes).
+Over time, plants shifted away from a wholly gametophyte-dominated life cycle (as in extant bryophytes) to co-independence of the gametophyte and sporophyte (\ie, ferns and lycophytes) to complete loss of the independent gametophyte (\ie, spermatophytes).
 Little is known about the details of the transition from desiccation tolerance to avoidance, nor how this took place in conjunction with the switch from a gametophyte to a sporophyte-dominated life cycle; yet, these two steps underpin the radiation of vascular plants onto dry land [@Puttick2018; @Qiu2006; @Wickett2014].
@@
 Filmy fern gametophytes are similarly diverse, with morphotypes including filamentous, ribbon, and a combination of these two types (`r figure("habit")`e--h).
-While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages [@Pittermann2013; @WatkinsJr.2012; @WatkinsJr.2007a], and may be expected to occur in filmy ferns as well.
+While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages [@Pittermann2013; @WatkinsJr.2012; @WatkinsJr.2007c], and may be expected to occur in filmy ferns as well.
 The possibility for varying degrees of vegetative DT in both stages of the life cycle of filmy ferns makes them a potentially useful system to compare how DT is utilized between haploid, non-vascular and diploid, vascular life stages.
@@
 Unlike sporophytes, fern gametophytes are cryptic and cannot be identified to species based on morphology alone. 
-Therefore, we used DNA barcoding to identify filmy fern gametophytes to species following @Nitta2017.
+Therefore, we used DNA barcoding to identify filmy fern gametophytes to species as described in @Nitta2017.
 Briefly, DNA extraction was performed using a modified CTAB protocol [@Doyle1987] or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
-Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following @Nitta2017.
+Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols as described in @Nitta2017.
 For each marker, a reference library was constructed from sporophyte sequences using the \function{makeblastDB} command in BLAST [@Altschul1997].
@@
 
-Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400 nm to 700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer as described by the manufacturer.
+Rapid light response curves were constructed by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400 nm to 700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer as described by the manufacturer.
 Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
@@
 Rotui 830 m site were excluded because this site is much more exposed than the other high elevation sites, and appears as a significant outlier in the microclimatic data [@Nitta2020].
-Since DT showed significant phylogenetic signal (see Results), we used phylogenetic generalized least squares (PGLS) [@Freckleton2002] to test for a relationship between VPD and DT while taking phylogeny into account, using species' means.
+Since DT showed significant phylogenetic signal (see Results), we used phylogenetic generalized least squares (PGLS) [@Freckleton2002] to test for a relationship between VPD and DT while taking phylogeny into account with the \function{pgls} function in the \package{caper} package [@Orme2018].
 
 To test whether gametophytes with broader ranges have higher DT, we calculated the extent of gametophyte range beyond sporophytes ("G beyond S") as the total elevational range (m) gametophytes were observed growing beyond sporophytes at either end of the elevational gradient.
-We conducted PGLS with DT as the response variable and G beyond S as the explanatory variable with the \function{pgls} function in the \package{caper} [@Orme2018].
+We conducted PGLS with DT as the response variable and G beyond S as the explanatory variable with the \function{pgls} function as above.
 
 Datasets used in this study are available on FigShare (https://doi.org/10.6084/m9.figshare.14184572). 
-Code used to analyze the data and generate figures and this manuscript are available on GitHub (https://github.com/joelnitta/moorea_filmies).
+Code used to analyze the data and generate figures and this manuscript is available on GitHub (https://github.com/joelnitta/moorea_filmies).
 A Docker image is available to run the code at https://hub.docker.com/r/joelnitta/moorea_filmies.
@@
 1 and 2.
+These `r n_distinct(range_comparison$species)` filmy fern taxa were treated as species for data analysis.
 
@@
 apiifolia* showed slightly greater ability to recover by 48 h compared to *Abrodictyum dentatum* (Bosch) Ebihara & K.
-Iwats., which completely failed to recover from even the gentlest treatment of 2 days at --38 NaCl (more stringent tests were not conducted on *A.
+Iwats., which completely failed to recover from even the gentlest treatment of 2 days at --38 MPa (more stringent tests were not conducted on *A.
 dentatum* after confirming its lack of DT at this level).
@@
 This suggests that the widespread gametophytes of *C. apiifolia* do not rely on DT to attain their distribution, but instead must do so via some other mechanism.
-Although we measured terrestrial microhabitats by positioning sensors a few cm above the ground, these sensors are still outside of the boundary layer.
+Although we measured terrestrial microhabitats by positioning sensors \circa{ }10 cm above the ground, these sensors are still outside of the boundary layer.
 It is likely that the gametophytes of *C.
@@
 Weston Testo provided advice on settings for the DT test.
-Two anonymous reviewers provided feedback that improved the manuscript.
+James M. R. Brock and an anonymous reviewer provided feedback that improved the manuscript.
 Staff at the University of California Berkeley's Richard B.
@@
 
-Recovery, recovery of F~v~/F~m~ (%) following 2 days desiccation at –86 MPa; ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached.
+**Recovery** recovery of F~v~/F~m~ (%) following 2 days desiccation at –86 MPa, **ETR~max~** maximum relative electron transport rate, **PPFD~95%~** light level at which 95% of ETR~max~ is reached.
 
@@
 
-ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached; Recovery, recovery of F~v~/F~m~ (%) following 2 days desiccation at –86 MPa
-
 Best model (that with the lowest DIC) for each response variable in bold
+
+**ETR~max~** maximum relative electron transport rate, **PPFD~95%~** light level at which 95% of ETR~max~ is reached, **Recovery** recovery of F~v~/F~m~ (%) following 2 days desiccation at –86 MPa
 
@@
 
-ETR~max~, maximum relative electron transport rate; PPFD~95%~, light level at which 95% of ETR~max~ is reached; Recovery, recovery of F~v~/F~m~ (%) following 2 days desiccation at –86 MPa
-
 Effect of life stage is relative to gametophytes; effect of growth habit is relative to high elevation epiphytes
@@
 *n* = 8 samples per species/treatment unless otherwise indicated.
-Growth habit indicated in parentheses after each species: T, terrestrial; S, saxicolous; LE, low elevation epiphyte; HE, high elevation epiphyte
+Growth habit indicated in parentheses after each species: **T** terrestrial, **S** saxicolous, **LE** low elevation epiphyte, **HE** high elevation epiphyte
 
 **`r figure("combined")`** Phylogeny of filmy ferns from Moorea, French Polynesia with comparison of physiological parameters (mean ± 1 SD) between sporophytes (dark grey bars) and gametophytes (light grey bars).
-Phylogeny adapted from Nitta et al. (2016); only species with physiological data shown.
+Phylogeny adapted from @Nitta2017; only species with physiological data shown.
 Bootstrap support >50% from 100 ML bootstraps indicated at nodes.

### 2021-06-10 ms/manuscript.Rmd
Update MS
various small changes after receiving proofs

@@
 
-Comparisons of DT, ETR~max~, and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a two-sided *t*-test for these.
+Comparisons of DT, ETR~max~, and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a two-sided *t* test for these.
 
@@
 
-Desiccation tolerance of gametophytes tended to be lower than sporophytes (`r figure("combined")`), and was significantly lower in `r dt_g_less_than_s` (*t*-test, *P* = `r dt_g_less_than_s_pvals`, respectively).
-Only `r dt_g_more_than_s` showed significantly higher DT in gametophytes compared to sporophytes (*t*-test, *P* = `r dt_g_more_than_s_pvals`).
+Desiccation tolerance of gametophytes tended to be lower than sporophytes (`r figure("combined")`), and was significantly lower in `r dt_g_less_than_s` (*t* test, *P* = `r dt_g_less_than_s_pvals`, respectively).
+Only `r dt_g_more_than_s` showed significantly higher DT in gametophytes compared to sporophytes (*t* test, *P* = `r dt_g_more_than_s_pvals`).
 
@@
 
-The GLMM including the effect of life stage only was selected as the best-fitting model for ETR~max~, and the combined effect of growth habit plus life stage and their interaction was selected for PPFD~95%~ (`r table("glmm")`).
+The GLMM including the effect of life stage only was selected as the best-fitting model for ETR~max~, and the combined effects of growth habit plus life stage and their interaction was selected for PPFD~95%~ (`r table("glmm")`).
 
@@
 ```{r esm-statement, results = "asis", eval = params$output_type == "ms", purl = FALSE}
-# FIXME: add DOI when available
 cat("
-**Electronic supplementary material**
-
-The online version of this article (DOI) contains supplementary material, which is available to authorized users.
+**Supplementary Information** The online version contains supplementary material available at https://doi.org/10.1007/s10265-021-01318-z.
 ")
@@
 Bootstrap support >50% from 100 ML bootstraps indicated at nodes.
-Bold letters indicate major clades (H = hymenophylloid clade, T = trichomanoid clade).
+Bold letters indicate major clades (**H** hymenophylloid clade, **T** trichomanoid clade).
 Growth habit indicated by colored symbols (blue = low elevation epiphyte, green = high elevation epiphyte, yellow = saxicolous, red = terrestrial).
 *Crepidomanes bipunctatum* and *C. humile* were observed growing as both epiphytes and saxicoles, but are shown as epiphytes to distinguish them from exclusively saxicolous species.
-Asterisk indicates significant difference in means (*P* <0.05) between life stages for a particular species (*t*-test); ns indicates no significant difference; lack of ns or asterisk means not enough observations available for that species for *t*-test.
+Asterisks indicates significant difference in means (\*\*\* = *P* < 0.001, \*\* = *P* < 0.01, \* = *P* < 0.05) between life stages for a particular species (*t* test); ns indicates no significant difference; lack of ns or asterisk means not enough observations available for that species for *t* test.
 Left bar plot shows recovery (%) of chlorophyll fluorescence (F~v~/F~m~) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa). 
@@
 
-**`r figure("corr")`** Recovery of F~v~/F~m~ (%) following 2 days desiccation at –86 MPa plotted against environmental vapor pressure deficit (VPD; kPa) (**a**, **b**) or range of gametophytes beyond sporophytes (m) (**c**) for gametophytes (**a**, **c**) or sporophytes (**b**).
+**`r figure("corr")`** Recovery of F~v~/F~m~ (%) following 2 days desiccation at –86 MPa plotted against environmental vapor pressure deficit (VPD; kPa) (**a**, **b**) or range of gametophytes beyond sporophytes (m) (**c**) for sporophytes (**a**) or gametophytes (**b**, **c**).
 Species means ± 1 SD are shown. 
@@
 
+# Make tibble of clade labels to plot on tree
+clade_labels <-
+tibble(
+  node = c(
+    getMRCA(phy, c("Crepidomanes_bipunctatum", "Callistopteris_apiifolia")),
+    getMRCA(phy, c("Hymenophyllum_multifidum", "Hymenophyllum_pallidum"))
+  ),
+  clade_label = c(
+    "T", # trichomanoid clade
+    "H" # hymenophylloid clade
+  )
+)
+
 # Sort species by position in phylogeny
@@
 # Part A (left): tree
-a <- ggtree(phy) %<+% habit +
+a <- ggtree(phy) %<+% habit %<+% clade_labels +
+  geom_nodelab(
+    aes(label = clade_label), 
+    hjust = 1, vjust = 0, geom = "label", 
+    label.padding = unit(0.4, "lines"),
+    label.size = 0,
+    size = 4,
+    fill = "transparent",
+    fontface = "bold"
+    ) +
   geom_tippoint(aes(color = habit), position = position_nudge(x = 14, y = 0), size = 3.0) +
@@
 
+# a: sporophytes recovery vs. vpd
 a <- ggplot(plot_data, aes(x = vpd_sporophyte_mean, y = recovery_sporophyte_mean, color = habit)) +
@@
 
+# b: gametophytes recovery vs. vpd
 b <- ggplot(plot_data, aes(x = vpd_gametophyte_mean, y = recovery_gametophyte_mean, color = habit)) +
@@
 
+# c: gametophytes recovery vs G beyond S
 c <- ggplot(plot_data, aes(x = gameto_beyond_sporo, y = recovery_gametophyte_mean, color = habit)) +

### 2021-06-10 ms/si.Rmd
Update MS
various small changes after receiving proofs

@@
     labs(
-      y = "ETR (µmol m<sup>-2</sup>·s<sup>-1</sup>)", 
-      x = "PPFD (µmol m<sup>-2</sup>·s<sup>-1</sup>)", 
+      y = "ETR (µmol m<sup>-2</sup> s<sup>-1</sup>)", 
+      x = "PPFD (µmol m<sup>-2</sup> s<sup>-1</sup>)", 
       alpha = "Outlier") +
@@
 
-**`r s_figure("water-all")`**. Relative water content (RWC) of sporophytes after desiccation by duration of desiccation treatment (2 or 15 days).
+**`r s_figure("water-all")`**. Relative water content (RWC) of sporophytes in desiccated state by duration of desiccation treatment (2 or 15 days).
 
@@
     panel.grid.minor.y = element_blank(),
-    panel.grid.major.y = element_blank(),
+    # panel.grid.major.y = element_blank(),
     axis.text.y = element_markdown(),
@@
 **`r s_figure("water-all")`**.
-Relative water content (RWC) of sporophytes after desiccation by duration of desiccation treatment (2 or 15 days).
+Relative water content (RWC) of sporophytes in desiccated state (after desiccation but before recovery) by duration of desiccation treatment (2 or 15 days).
 RWC was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not RWC of zero).
@@
 **a**, sporophytes; **b**, gametophytes.
-Title of each subplot indicates species followed by light curve ID number assigned to each sample (light curve IDs match those in filmy_light_curves.csv data file).
+Title of each subplot indicates species followed by light curve ID number assigned to each sample (light curve IDs match those in filmy_light_curves.csv data file available at https://doi.org/10.6084/m9.figshare.14184572).
 Light curves fit using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (ETR), *x* is photosynthetic photon flux density (PPFD), *A* is the asymptote of the curve, and *k* is a slope parameter.
@@
 Each point represents one species; dark line is median, lower and upper edges correspond to first and third quartiles, and whiskers extend to the most extreme value no further than 1.5 times the inter-quartile range.
-**a**, ETR~max~ (µmol electrons m^--2^ ·s^--1^); **b** PPFD~95%~ (µmol photons m^--2^ ·s^--1^); **c**, recovery (%) of chlorophyll fluorescence (F~v~/F~m~) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa).
+**a**, ETR~max~ (µmol electrons m^--2^ s^--1^); **b** PPFD~95%~ (µmol photons m^--2^ s^--1^); **c**, recovery (%) of chlorophyll fluorescence (F~v~/F~m~) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa).
 *Crepidomanes bipunctatum* and *C.
